### 2019-01-03 paper/paper.Rmd
add paper before new year intro re-write.

---
title: "The interactions of rational pragmatic agents provide a framework for understanding efficient language structure and use"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Author 1} \\ \texttt{bpeloqui@stanford.edu} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Author 2} \\ \texttt{author1@university.edu} \\ Department of Psychology \\ Some University}

abstract: 
    We discuss a framework for studying the distributional properties of linguistic systems as emerging from in-the-moment interactions of speakers and listeners. Our work takes Zipfian notions of lexicon-level efficiency as a starting point, connecting these ideas to Gricean notions of conversational-level efficiency. To do so, we begin by deriving an objective function for measuring the communicative efficiency of linguistic systems and then examining the behavior of this objective in a series of simulations focusing on the communicative function of ambiguity in language.

keywords:
    "Communicaive efficiency, Rational Speech Act theory, computational modeling, information theory, agent-based simulation"

output: cogsci2016::cogsci_paper
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
```

# Introduction
Why do natural languages look they way they do? Functionalist approaches to language science bring a utility-based perspective to this question, typically reframing it in terms of information-theoretic principles of efficiency -- how does having property $x$ make using language $\ell$ more or less useful for communication? This family of theories has described a range of phenomena in language-design (compositionality, ambiguity, dependency length) and language-use (redundancy and reduction) in terms of communicative efficiency. Consider two examples of such theories: Piantadosi et al. (2012) argued that efficiently *designed* linguistic systems will contain ambiguous lexical items when communication is contextualized. If words with multiple meanings can be disambiguated in context, less costly forms can be re-used efficiently.  From the language *use* perspective, Levy & Jaeger (2007) showed that speakers are sensitive to spikes and troughs in the information rate of their utterances. The authors showed that speakers modulate use of optional material according to the surprisal of oncoming material -- an optimal strategy for efficient information transfer from an information-theoretic analysis.\par

While Piantadosi et al. (2012) and Levy & Jaeger (2007) provide two recent examples of functionalist theory, these frameworks are largely in line with earlier work by Zipf (1949). Zipf's Principle of Least Effort is a domain-general framework for describing human behavior in terms of effort minimization. Zipf claimed that humans have a preference to minimize "the average rate of work-expenditure over time" (Zipf, 1949, p.6). In terms of language behavior, he described a relation between a word's frequency in corpora and its length (number of characters) in terms of the principle of least effort. Intuitively, if effort is related to a word’s length (perhaps via articulatory effort) then, holding all else constant, a language with shorter words should be more efficient than a language with longer words, what Zipf called the *Law of Abbreviation*. Indeed this relation between frequency and length is well-documented in natural languages (Kanwal et al 2017).\par

Perhaps even more important than the actual predictions made by Zipf’s Principle of Least Effort is the insight that we can analyse the structure of language systems in terms *effort* or *cost* along competing dimensions.  With such a formulation we can begin to ask questions about what optimal languages might look like under different cost structures. For example, Zipf noted that purely from the standpoint of speaker effort, what Zipf called "Speaker's Economy," an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of just a single, low-cost word. Given such a language, the full set of potential meanings would be conveyed using only that word, i.e. $\ell_{speaker}^*$ would be fully ambiguous and all possible meanings would need to be disambiguated by a listener. Conversely, from the standpoint of listener effort, what Zipf called "Auditor's Economy," an optimal language $\ell_{listener}^*$ would bijectively map all possible meanings to distinct words, eliminating a listener's need to disambiguate. Under this analysis, speaker effort is related to production cost and listener effort to understanding (disambiguation) cost. Clearly natural languages fall between the two extremes of $\ell_{speaker}^*$ and $\ell_{listener}^*$. Zipf's insight was that the particular distributional properties we observe in natural languages emerge from the competition of these forces -- the pressure to jointly minimize speaker and listener effort.\par

Zipf’s principle of least effort and its speaker and listener instantations were intended as a framework for explaining all of natural language. However, Zipf’s primary unit of analysis was at the level of languages themselves (i.e. the vocabulary). Subsequent work by Horn (1984), highlighted the importance of Zipf’s principles for explaining phenomena at the level of a single conversation. Horn suggested a direct link between Zipf’s Speaker and Listener economy and, what was at the time, more recent work on conversational pragmatics by Grice (1975). Horn highlighted that the interaction of Zipf’s forces were “largely responsible for generating Grice’s conversational maxims and the schema for pragmatic inference derived therefrom.” Put differently, system-level efficiency we see in languages is deeply related to local-level efficiency during the in-the-moment interactions of speakers and listeners. \par

In this work we present an initial step toward formalizing this connection first highlighted by Horn. That is, we present a framework for evaluating the aggregate, distributional properties of linguistic systems as emerging from the accumulation of many in-the-moment interactions of pragmatic speaker and listener agents. We are exploring the questions - to what extent can we connect the Gricean and Zipfian programs in a single framework? To do so, we adopt a simulation-based approach, modeling communication as reference games in which we can vary the types of agents and languages present. This framework allows us to investigate a range of functionalist theories broadly under the categories of efficient language design and language use. \par

For this particular project we focus on an analysis of the communicative function of ambiguity in language following work by Piantadosi et al. (2012). We adopt this particular linguistic phenomena to demonstrate the extent to which we can explore both the design- and use- based functionalist ideas. We focus on three questions -- (1) when is it desirable to have a language with lexical ambiguity? (2) Under what circumstances do speakers and listeners use ambiguity efficiently? (3) Assuming Gricean agents what is the impact of increasing pragmatics on efficiency?\par

We begin with a high-level introduction to the modelling framework introducing the basic ingredients we will need to represent language as repeated reference games. Following this introduction we derive a simple objective function for measuring the efficiency of linguistics systems in this setting. Subsequently, we move on to three case-studies examining the questions posed above, framing results in terms of our efficiency measure.

# Exploring efficient language- design and use in rational pragmatic agents

## Simulation set-up.

### Reference games
We take as our unit of analysis the basic referential communication game similar to those described by WittgensteIn (CITATION). Both speakers and listeners are aware of a set of objects $M$, which will refer to as *meanings* and are knowledgeable about the set of possible signals $U$ (*utterances*) that could be used to refer to a given meaning. Utterances may have different relative costs, operationalized via a prior over utterances $p(u)$. Similarly, meanings differ in the degree to which they need to be talked about. This is operationalized as a prior over meanings $p(m)$. Note that the prior over meanings are analogous to the *need probabilities* assumed in previous work (Regier, CITATION). We consider a set of contexts $C$ which describe different need probability distributions over our set of meanings $p(m|c)$.

### Languages
A language $\ell$ defines the set of semantic mappings between utterance and meanings. For example, in a world with three utterances $U = \{u_1, u_2, u_3\} \text{and three meanings }M = \{m_1, m_2, m_3\}$ the boolean matrix

\begin{center}
\begin{tabular}{ c | c c c } 
& $m_1$ & $m_2$ & $m_3$ \\
\hline
$u_1$ & 1 & 1 & 0 \\
$u_2$ & 0 & 1 & 0 \\
$u_3$ & 0 & 0 & 1 \\
\end{tabular}
\end{center}

describes the literal semantics of $\ell$. E.g. the language describes semantic mappings $[[u_1]]\ = \{m_1, m_2\}, [[u_2]]\ = \{m_2\}, [[u_3]]\ = \{m_3\}$.

### Speakers and listeners
A *speaker agent* defines a conditional distribution over utterances, mapping from intended meanings $M$ to utterances $U$ using a particular semantic mapping given by $\ell$. That is, a speaker defines $P_{speaker}(u|m;\ell)$. We will use $S(u|m;\ell)$ as short-hand throughout.  A *listener agent* defines a conditional distribution over meanings, mapping from utterances $U$ to meanings $M$ using a particular semantic mapping given by $\ell$. We will use $L(m|u; \ell)$ as shorthand. Note that both speakers and listeners can induce joint distributions over the set of all signaling events $E$, although, importantly, these distributions may differ:
$$P_{speaker}(u, m; \ell) = S(u|m; \ell)p(m)$$
$$P_{listener}(u, m; \ell) = L(m|u; \ell)p(u)$$

In general, we would like to consider the efficiency of a system $\ell$ in terms of these joint distributions.

# Zipfian objective for linguistic system efficiency

## Metrics for optimal linguistic systems in the reference game setting

To assess the optimality of a linguistic system we need some measure of goodness. In other words an objective function, which, as a function of the particular linguistic system $\ell$ and the speaker $S$ and listener $L$ agents returns a measure of the system efficiency ( $f(S, L, \ell) \rightarrow \mathbb{R}$).  For our purposes we’d like to tie this objective to fundamental principles of speaker and listener effort suggested by Zipf. Additionally, we will take the fundamental unit of analysis as the reference game event. \par
We now derive a simple objective function from Zipfian first principles. Zipf proposed that the particular distributional properties found in natural language emerge as a result of competing speaker and listener pressures. We operationalize this in equation (1) -- the efficiency of a linguistic system $\ell$ being used by speaker and listener agents $S$ and $L$ is sum of the expected speaker and listener effort to communicate.
\begin{equation}
\text{Efficiency}(\ell) = \mathbb{E}_{e \in E}[\text{speaker effort}] + \\ \mathbb{E}_{e \in E}[\text{listener effort}]
\end{equation}

Let speaker effort be the log probability of a particular utterance. $log_2(p(u))$ E.g. the number of bits needed to encode the utterance $u$.

$$\text{speaker effort} = -log_2(p(u))$$

Let listener effort be the log probability a listener disambiguates an intended meaning $m$ given an utterance $u$ E.g. the number of guesses a listener would need to discover the intended meaning $m$ given an utterance $u$.
$$\text{listener effort} = -log_2(L(m|u; \ell))$$

Rewriting (1) we now have
\begin{equation}
= \mathbb{E}_{e \in E}[-log_2(p(u))] + \\
\mathbb{E}_{e \in E}[-log_2(L(m|u; \ell))]
\end{equation}
In general, these expectations are each taken over all possible communicative events $e\in E$ weighted by the probability of a particular event $p(e)$. Recall that is the set of all utterance, meaning pairs $<u, m> = e \in E$.

\begin{equation}
 = \sum_{e \in E}p(e)[-log_2(p(u))] + \sum_{e \in E}p(e)[-log_2(L(m|u; \ell))]
\end{equation}

We assume that the particular joint distribution over $u, m$ pairs follows from a simple generative model. First, some meaning is sampled with probability $p(m)$. Our speaker attempts to convey this intended meaning to a speaker by encoding it in the utterance $u$ via the conditional $S(u|m; \ell)$. Combining these terms leads to the *speaker's joint distribution over events* which we can write as $P_{Speaker}(u, m; \ell) = S(u | m; \ell)p(m)$.

\begin{equation}
\begin{split}
  = \sum_{u, m}p_{speaker}(u, m; \ell)[-log_2(p(u))] + \\ 
   \sum_{u, m}p_{speaker}(u, m; \ell)[-log_2(L(m|u; \ell))]
\end{split}
\end{equation}

Simplifying we arrive at (5):
\begin{equation}
\begin{split}
 = \sum_{u, m}P_{speaker}(u, m; \ell)[-log_2(L(m|u; \ell)p(u))]
 \end{split}
\end{equation}

Note that $L(m|u; \ell)p(u)$ is the listener-based joint distribution over all communicative events ($P_{listener}(u, m; \ell)$).
\begin{equation}
  = \sum_{u, m}P_{speaker}(u, m; \ell)[-log_2(P_{listener}(u, m; \ell))]
\end{equation}
This is the simply the cross-entropy between the speaker an listener joint distributions.
\begin{equation}
\begin{split}
  = \mathbb{E}_{P_{speaker}}[-log_2(P_{listener})] \\
  = H_{cross}(P_{speaker}, P_{listener})
\end{split}
\end{equation}
From a mathematical standpoint this objective is intuitive. An optimal linguistic system being used by a speaker and listener agent will induce a particular set of form-meaning mappings that are both *close* between speaker and listener agents and *peaky* in that they assign non-uniform probability mass to the valid form-meaning mappings.

# Three case-studies on efficient language use and design
We now have a metric for exploring questions of langauge optimality in the reference game setting (the speaker-listener cross-entropy). In the following sections we examine the behavior of the objective with respect to the communicative function of ambiguity. In particular, we examine three questions. The first is a question of optimal design -- given a set of possible languages $L$ in what circumstances does the optimal languages $\ell^*$ contain ambiguity? The second is a question of optimal *use* -- give a particular language that theoretically allows for the use of ambiguious material under what circumstances will ambiguous forms be used efficiently?

## Experiment 1:  Rational-pragmatic speakers prefer languages with ambiguous items

Piantadosi et al. (2012) proposed that a language with ambiguous items is strictly more efficient than a language with out, *when ambiguous items can be disambiguated with context*. Formally, they authors highlighted the fact that if we measure the efficiency of a linguistic system in terms of entropy, we have the basic result from (information theory citation) that the conditional entropy of an ensemble is strictly less that the unconditional entropy, when the information we condition on is informative ($H(X|C) < H(X)$).\par

Piantadosi et al. (2012) provided empirical evidence for this claim, showing (SAY WHAT PIANTADOSI DID). We consider their basic arguemnt in the reference game setting. Our approach procceds as follows, we consider the full-space of valid languages (semantic mappings) between forms and utterances in which $|U|=|M|=4$. A language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assigment). For example, a one word language that is fully ambiguous would map a single form to all possible meanings. We represent that via the boolean matrix $\ell_1$ in which $[[u_1]] = \{m_1,m_2, m_3 \}, [[u_2]]=[[u_3]]=\{\}$

\begin{center}
\begin{tabular}{ c | c c c} 
& $m_1$ & $m_2$ & $m_3$ \\
\hline
$u_1$ & 1 & 1 & 1 \\
$u_2$ & 0 & 0 & 0 \\
$u_3$ & 0 & 0 & 0 \\
\end{tabular}
\end{center}

Note that $\mathbb{E}[\text{speaker effort}]$ can be arbitrarily low for such a language, as it is directly proportional to $p(u_1)$. That is in the single utterance language, $\mathbb{E}[\text{speaker effort}] \rightarrow 0$ as $p(u_1) \rightarrow 1$. However, $\mathbb{E}[\text{listener effort}]$ is likely to be high as the only available form ($u_1$) is full ambiguous between possible meanings.\par
By contrast consider a language $\ell_2$:

\begin{center}
\begin{tabular}{ c | c c c} 
& $m_1$ & $m_2$ & $m_3$ \\
\hline
$u_1$ & 1 & 0 & 0 \\
$u_2$ & 0 & 1 & 0 \\
$u_3$ & 0 & 0 & 1 \\
\end{tabular}
\end{center}

In this case $\mathbb{E}[\text{listener effort}]=0$ -- each form has a one-to-one mapping to a given meaning. However speaker effort will likely be greater than in $\ell_1$ so long as $p(u_1) < p(u_2)$ and $p(u_1) < p(u_3)$.

Finally, consider a langauge $\ell_3$:

\begin{center}
\begin{tabular}{ c | c c c} 
& $m_1$ & $m_2$ & $m_3$ \\
\hline
$u_1$ & 1 & 1 & 0 \\
$u_2$ & 0 & 0 & 0 \\
$u_3$ & 0 & 0 & 1 \\
\end{tabular}
\end{center}

Our third language contains an ambiguous lexical item - $u_1$ can be used to refer to either $m_1$ or $m_2$. This language represents a trade-off between $\ell_1$ and $\ell_2$. From the speaker perspective $\ell_3$ likely requires more effort from the speaker than $\ell_1$, however it likely requires less than $\ell_2$ (so long as $p(u_1)$ is cheap). From the listener perspective $\ell_3$ requires more effort than $\ell_3$ which requires no disambiguation, but less than $\ell_1$. \par
As this small set of languages demonstrate, the efficiency of a particular language depends on both utterances costs $P(u)$ and need probabilities $P(M)$. E.g. If $u_1$ is actually a costly utterance ($p(u_1)$ is close to 0) then $\ell_1$ is inefficient both for the speaker and listener. For this reason, in the following simulations, we randomly sample utterance costs $P(u)$ and need probabilities $P(m)$. Our simulations procced as follows. For $n=100$ simulations we enumerate all valid langauges ($4\times 4$ boolean matrices) as well as a set of need probabilities ($p(m)$) and utterance costs ($p(u)$).

Given this set-up we can find the optimal language:

$$\ell^* = \text{argmin}_{\ell \in L} H_{Cross}(P_{Speaker}(u, m; \ell), P_{Listener}(u, m; \ell))$$

Critical to Piantadosi's claim, however is the fact that ambiguous langauges should be preferred *when* context is disambiguating. For that reason we compare the optimal langauge $\ell^*$ as the amount of contextual information increases. In our case a "context" describes a conditional distribution over meanings. For example, take a setting with two contexts $c_1$ and $c_2$. The first context dictates a set of need probabilites which differs from the second ($p(m|c_1) \neq p(m|c_2)$). If ambiguity is useful when context is disambiguating we should expect the likelihood that an optimal langauge $\ell^*$ contains an ambiguous item to increase as the number of contexts increases.

### Simulation set-up. \newline

$L = \text{set of all valid languages with} |U|=|M|=4$\par
$p(m) = \text{Dirichlet}(ones)$\par
$S(u|m; \ell) = \text{S1 rsa speaker}$\par
$L(m|u; \ell) = \text{L0 rsa listener}$\par

### Results
We run $n=100$ simulations in four condtions. Our first is a *single-context* condition ( $|C|=1$) -- there is a only a single context describing $p(m|c_1)$. Our second condition contains two-contexts  ($|C| = 2$) -- we consider efficiency under both $p(m|c_1)$ as well as $p(m|c_2)$. The third and fourth condition correspond accordingly with $|C|=3$ and $|C| = 4$, respectively. In each condition we consider a random set of utterances costs $P(u)$ and need probabilities $p(m|c)$ parametrized as Dirichlet distributions with $\alpha=1$. Piantadosi et al. (2012) argued that ambiguity is an efficient property of contextualized langauge use. That is, it is useful to assign words multiple meanings if those words can be disambiguated in context. Formalized in our setting, this would mean that we expect $\ell*$ to be more likely to contain ambiguous mappings as the number of contexts increases.par
Figure 1. plots the proportion of optimal languages under $H_{cross}(P_{Speaker}, P_{Listener})$ for each condition. We find that as the number of contexts increases, so does the probability that optimal language $\ell*$ contains ambiguity. For comparison we also include the optimal language under an objective that only considers speaker or listener effort. In line, with Zipf's predictions, if langauges are designed only to minimize speaker effort then optimal languages will assign all meanings to a single, low-cost utterance. Likewise, if langauges are designed only to minimize speaker listener effor then ambiguity should alwasy be avoided.

```{r fig1, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=4, fig.height=4, fig.cap = "Optimal languages are more likely to contain ambiguous items as amount of contextual information increases. Vertical axis shows the proportion of optimal languages that contain ambiguity. Horizontal axis shows the number of contexts in each condition (1-4). Red-line represents the optimal langauge under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only consideration."}
img <- png::readPNG("figs/fig1.png")
grid::grid.raster(img)
```

Piantadosi et al. (2012) framed their theory in terms of conditional entropy. That is, $H(X|C) < H(X)$ when $C$ provides information about $X$. Put differently, when $I(X, C)$ is non-zero. In our setting this would indicate that as the amount of contextual information increases, the difference between the conditional and unconditional measures of information should increase. That is, $H_{cross}(P_{Speaker}, P_{Listener}|C_1) - H_{cross}(P_{Speaker}, P_{Listener}) < H_{cross}(P_{Speaker}, P_{Listener}|C_2) - H_{cross}(P_{Speaker}, P_{Listener})$ when $C_1$ contains more information than $C_2$. We can consider this very metric in our conditions. Figure 2 plots $H_{cross}(P_{Speaker}, P_{Listener}|C_1) - H_{cross}(P_{Speaker}, P_{Listener})$ in each of our conditions. As the number of contexts increases, the difference in efficiency when context is disambiguating increaess as well.
```{r fig2, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2.5, fig.height=3, fig.cap = "The gains in efficinecy increase as the amount of contextual information increases. Vertical axis shows the difference in cross-entropy for when context is informative. Horizontal axis shows each of the four conditions. Note that the difference for condition 1 is zero as thered is only a single context." }
img <- png::readPNG("figs/fig2.png")
grid::grid.raster(img)
```

### Summary
Piantadosi et al. (2012) argued that it is useful to re-use low-cost linguistic forms for multiple meanings when they can be disambiguated in context. Using our speaker-listener cross-entropy measure of efficiency we showed that optimal langauges are more likely to have ambiguous items when context is informative. Further, we showed that the impact of being able to disambiguate langauge (having access to $p(m|c_i)$) is increasingly efficienct as the amount of common-ground (context) increases.

## Experiment 2: Rational-pragmatic speakers use ambiguity efficiently
Our first experiment made a fairly direct test of the communicative function of ambiguity proposed by Piantadosi et al. (2012). Sampling a set of need probabilities and utterance costs we explored the space of possible languages, examining the properties of the language $\ell*$ which minimized our objective. Results indicated that ambiguity is an efficient property when context is informative. This experiment, however, assumed that our speaker and listener agents could disambiguate items context. That is, in our four conditions both agents had access to these conditional distributions $p(m|c_1), \dots, p(m|c_4)$. More often than not in day-to-day language use, speakers and listeners may not have perfect knowledge of the particular set of need probabilities $p(m)$ being used. Put differently, if ambiguity is only efficient when it can be disambiguated in context, speakers should avoid using ambiguous items if they know their listener may not be able to disambiguate the item.\par
In experiment two we examine speaker behavior in a scenario in which the listener does not know the exact set of need probabilites a priori. That is, the listener has uncertainty over $p(m)$. Over the course of a discourse $D$ the listener tries to infer the context along with the particular intended meaning of a given utterance. That is, we consider a listener model $L(c, m|u;D)$ where $c$ is the particular context (or *topic*) being discussed and $D$ is the set of previous utterances the speaker has made. Likewise we consider a speaker model $S(u|m, c, D)$ who chooses an utterance based on an intended meaning $m$, the particular context of conversation $c$, while also considering the history of their utterance $D$. If speakers are behaving efficiently they should only use ambiguous items when they can be disambiguated in context. That is, if the listener agents is able to correctly infer the topic of conversation as the conversation progresses $D\rightarrow\inf$ then the speaker should be more likely to use ambiguous items later in discourse.

### Simulation set-up \newline

$\ell = \text{current language known to both speaker and listener}$\par
$D = \text{History of previous utterances by speaker}$\par
$S(u|m; \ell) = \text{S1 rsa speaker}$\par
$L(m|u; \ell) = \text{L0 rsa listener}$\par
$P(u) = \text{utterance costs in which } p(u_1) < p(u_2) < p(u_3) < p(u_4)$\par
$c = \text{current topic, which specifies need probabilities}$\par
$P(m|c) = \text{set of need probabilites specified by } c$\par

### Results

Fig3
```{r fig3, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2.5, fig.height=3, fig.cap = "R plot" }
img <- png::readPNG("figs/fig3.png")
grid::grid.raster(img)
```

Fig4
```{r fig4, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2.5, fig.height=3, fig.cap = "R plot" }
img <- png::readPNG("figs/fig4.png")
grid::grid.raster(img)
```

### Summary

## Experiment 3: “More” pragmatic leads to more efficiency

### Simulation set-up

### Results

Fig5
```{r fig5, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2.5, fig.height=3, fig.cap = "R plot" }
img <- png::readPNG("figs/fig5.png")
grid::grid.raster(img)
```

### Summary

# Discussion

# Conclusion

# General Formatting Instructions

For general information about authoring in markdown, see **[here](http://rmarkdown.rstudio.com/authoring_basics.html).**

For standard spoken papers and standard posters, the entire 
contribution (including figures, references, everything) can be 
no longer than six pages. For abstract posters, the entire contribution 
can be no longer than one page. For symposia, the entire contribution 
can be no longer than two pages.

The text of the paper should be formatted in two columns with an
overall width of 7 inches (17.8 cm) and length of 9.25 inches (23.5
cm), with 0.25 inches between the columns. Leave two line spaces
between the last author listed and the text of the paper. The left
margin should be 0.75 inches and the top margin should be 1 inch.
\textbf{The right and bottom margins will depend on whether you use
U.S. letter or A4 paper, so you must be sure to measure the width of
the printed text.} Use 10 point Times Roman with 12 point vertical
spacing, unless otherwise specified.

The title should be in 14 point, bold, and centered. The title should
be formatted with initial caps (the first letter of content words
capitalized and the rest lower case). Each author's name should appear
on a separate line, 11 point bold, and centered, with the author's
email address in parentheses. Under each author's name list the
author's affiliation and postal address in ordinary 10 point  type.

Indent the first line of each paragraph by 1/8~inch (except for the
first paragraph of a new section). Do not add extra vertical space
between paragraphs.

# First-Level Headings

First level headings should be in 12 point , initial caps, bold and
centered. Leave one line space above the heading and 1/4~line space
below the heading.

## Second-Level Headings

Second level headings should be 11 point , initial caps, bold, and
flush left. Leave one line space above the heading and 1/4~ line
space below the heading.

### Third-Level Headings

Third-level headings should be 10 point , initial caps, bold, and flush
left. Leave one line space above the heading, but no space after the
heading.

# Formalities, Footnotes, and Floats

Use standard APA citation format. Citations within the text should
include the author's last name and year. If the authors' names are
included in the sentence, place only the year in parentheses, as in
[-@NewellSimon1972a], but otherwise place the entire reference in
parentheses with the authors and year separated by a comma
[@NewellSimon1972a]. List multiple references alphabetically and
separate them by semicolons [@ChalnickBillman1988a; @NewellSimon1972a]. 
Use the et. al. construction only after listing all the authors to a
publication in an earlier reference and for citations with four or
more authors.

For more information on citations in RMarkdown, see **[here](http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html#citations).**

## Footnotes

Indicate footnotes with a number\footnote{Sample of the first
footnote.} in the text. Place the footnotes in 9 point type at the
bottom of the page on which they appear. Precede the footnote with a
horizontal rule.\footnote{Sample of the second footnote.} You can also use 
markdown formatting to include footnotes using this syntax [^1].

[^1]: Sample of a markdown footnote.

## Figures

All artwork must be very dark for purposes of reproduction and should
not be hand drawn. Number figures sequentially, placing the figure
number and caption, in 10 point, after the figure with one line space
above the caption and one line space below it. If necessary, leave extra white space at
the bottom of the page to avoid splitting the figure and figure
caption. You may float figures to the top or bottom of a column, or
set wide figures across both columns.

## Two-column images

You can read local images using png package for example and plot 
it like a regular plot using grid.raster from the grid package. 
With this method you have full control of the size of your image. **Note: Image must be in .png file format for the readPNG function to work.**

You might want to display a wide figure across both columns. To do this, you change the `fig.env` chunk option to `figure*`. To align the image in the center of the page, set `fig.align` option to `center`. To format the width of your caption text, you set the `num.cols.cap` option to `2`.

```{r 2-col-image, fig.env = "figure*", fig.pos = "h", fig.width=4, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns. And the caption text is limited to 0.8 of the width of the document."}
img <- png::readPNG("figs/walrus.png")
grid::grid.raster(img)
```

## One-column images

Single column is the default option, but if you want set it explicitly, set `fig.env` to `figure`. Notice that the `num.cols` option for the caption width is set to `1`.

```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
img <- png::readPNG("figs/lab_logo_stanford.png")
grid::grid.raster(img)
```

## R Plots

You can use R chunks directly to plot graphs. And you can use latex floats in the
fig.pos chunk option to have more control over the location of your plot on the page. For more information on latex placement specifiers see **[here](https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions)**

```{r plot, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2, fig.height=2, fig.cap = "R plot" }
x <- 0:100
y <- 2 * (x + rnorm(length(x), sd = 3) + 3)

ggplot2::ggplot(data = data.frame(x, y), 
       aes(x = x, y = y)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

## Tables

Number tables consecutively; place the table number and title (in
10 point) above the table with one line space above the caption and
one line space below it, as in Table 1. You may float
tables to the top or bottom of a column, set wide tables across both
columns.

You can use the xtable function in the xtable package.

```{r xtable, results="asis"}
n <- 100
x <- rnorm(n)
y <- 2*x + rnorm(n)
out <- lm(y ~ x)

tab1 <- xtable::xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
                      caption = "This table prints across one column.")

print(tab1, type="latex", comment = F, table.placement = "H")
```

# Acknowledgements

Place acknowledgments (including funding information) in a section at
the end of the paper.

# References

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent

### 2019-01-03 paper/paper.Rmd
updates to intro -- now much more focus on Zipf, probably need to carry this throughout.

@@
 # Introduction
+
 Why do natural languages look they way they do?
-Functionalist approaches to language science bring a utility-based perspective to this question, typically reframing it in terms of information-theoretic principles of efficiency -- how does having property $x$ make using language $\ell$ more or less useful for communication?
-This family of theories has described a range of phenomena in language-design (compositionality, ambiguity, dependency length) and language-use (redundancy and reduction) in terms of communicative efficiency.
-Consider two examples of such theories: Piantadosi et al.
-(2012) argued that efficiently *designed* linguistic systems will contain ambiguous lexical items when communication is contextualized.
-If words with multiple meanings can be disambiguated in context, less costly forms can be re-used efficiently.
-From the language *use* perspective, Levy & Jaeger (2007) showed that speakers are sensitive to spikes and troughs in the information rate of their utterances.
-The authors showed that speakers modulate use of optional material according to the surprisal of oncoming material -- an optimal strategy for efficient information transfer from an information-theoretic analysis.\par
+While Zipf (1935) presented his “Principle of Least effort” as a domain general framework for understanding human behavior, he took language as his central case-study.
+He proposed that distributional properties found in natural language were evidence of speaker-listener utility maximization or, equivalently, effort minimization.
+In his own words, “we are arguing that people do in fact act with a maximum economy of effort, and that therefore in the process of speaking-listening they will automatically minimize the expenditure of effort.” Evidence supporting this hypothesis has been realized in a set of well-known findings, largely falling under what is now known as Zipf’s Law or Zipfian distributions -- one of a family of discrete power law distributions.
+A range of empirical phenomena in social and physical sciences are well-described as Zipfian, from the relationship between word rank and frequency (CITATION) to the populations of U.S.
+cities (CITATION) Stanford stats class 306a).\par
 
-While Piantadosi et al.
-(2012) and Levy & Jaeger (2007) provide two recent examples of functionalist theory, these frameworks are largely in line with earlier work by Zipf (1949).
-Zipf's Principle of Least Effort is a domain-general framework for describing human behavior in terms of effort minimization.
-Zipf claimed that humans have a preference to minimize "the average rate of work-expenditure over time" (Zipf, 1949, p.6).
-In terms of language behavior, he described a relation between a word's frequency in corpora and its length (number of characters) in terms of the principle of least effort.
-Intuitively, if effort is related to a word’s length (perhaps via articulatory effort) then, holding all else constant, a language with shorter words should be more efficient than a language with longer words, what Zipf called the *Law of Abbreviation*.
-Indeed this relation between frequency and length is well-documented in natural languages (Kanwal et al 2017).\par
-
-Perhaps even more important than the actual predictions made by Zipf’s Principle of Least Effort is the insight that we can analyse the structure of language systems in terms *effort* or *cost* along competing dimensions.
-With such a formulation we can begin to ask questions about what optimal languages might look like under different cost structures.
+Language-structure and use provide a particularly fascinating case-study under an efficiency-based analysis as speaker- and listener-effort is asymmetric.
+Simply put, in the act of communicating with another person, what is effortful as a speaker is likely different from what is effortful as a listener.
 For example, Zipf noted that purely from the standpoint of speaker effort, what Zipf called "Speaker's Economy," an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of just a single, low-cost word.
@@
 Clearly natural languages fall between the two extremes of $\ell_{speaker}^*$ and $\ell_{listener}^*$.
-Zipf's insight was that the particular distributional properties we observe in natural languages emerge from the competition of these forces -- the pressure to jointly minimize speaker and listener effort.\par
+Zipf’s Law of Meaning proposes that the particular relation between a word’s frequency and denotation size emerge from the competition of these forces -- the pressure to jointly minimize speaker and listener effort.\par
 
-Zipf’s principle of least effort and its speaker and listener instantations were intended as a framework for explaining all of natural language.
-However, Zipf’s primary unit of analysis was at the level of languages themselves (i.e.
-the vocabulary).
-Subsequent work by Horn (1984), highlighted the importance of Zipf’s principles for explaining phenomena at the level of a single conversation.
+A range of more recent approaches to language science have distinctly Zipfian roots (functionalist theories).
+In terms of our opening question, “why do languages look the way they do?”, a functionalist approach would formalize the question in terms of information-theoretic principles of efficiency -- how does having property $x$ make using language $\ell$ more or less useful for communication?
+This family of theories has described a range of phenomena in language-design (compositionality, ambiguity, dependency length) and language-use (redundancy and reduction) in terms of communicative efficiency.
+Of particular interest to our current work, Piantadosi et al.
+(2012) argued that efficiently *designed* linguistic systems will contain ambiguous lexical items when communication is contextualized.
+If words with multiple meanings can be disambiguated in context, less costly forms can be re-used efficiently.
+They showed in a series of analyses that … [WHAT PIANTADOSI DID HERE!!!!].
+\par
+
+Zipf’s principle of least effort and it’s associated speaker and listener instantations were intended as a framework for explaining all of natural language.
+Evidence for the theory was largely derived from vocabulary-level analysis.
+Implicit in Zipf’s (1935) argument is that aggregate, distributional-level effects emerge from optimization during local interactions between speaker-listeners.
+While Functionalist projects have provided evidence that speaker’s attempt to optimize their output at a variety of levels -- phonologically (morphological contraction; Bell et al., 2003), syntactically (Levy & Jaeger, 2007; Jaeger, 2010) and over even over discourse (Genzel & Charniak, 2002), Zipf was less precise in terms of his exact level of analysis -- however his example of optimal speaker and listener language in which a single word $u$ can express some meaning $m$$[[u]]\neq \emptyset$.
+
+A half century after Zipf, the linguist Lawrence Horn (1984), highlighted the importance of Zipf’s principles for explaining conversation-level phenomena.
 Horn suggested a direct link between Zipf’s Speaker and Listener economy and, what was at the time, more recent work on conversational pragmatics by Grice (1975).
@@
 
-In this work we present an initial step toward formalizing this connection first highlighted by Horn.
-That is, we present a framework for evaluating the aggregate, distributional properties of linguistic systems as emerging from the accumulation of many in-the-moment interactions of pragmatic speaker and listener agents.
-We are exploring the questions - to what extent can we connect the Gricean and Zipfian programs in a single framework?
-To do so, we adopt a simulation-based approach, modeling communication as reference games in which we can vary the types of agents and languages present.
+In this work we present a step toward formalizing this connection highlighted by Horn -- formalizing the theoretical connection between Zipfian and Grice, and in doing so, deriving aggregate, distributional level properties from local, in-the-moment interactions of speaker-listeners.
+To gain traction on this project we adopt a simulation-based approach, modeling communication as reference games in which we can vary the types of agents and languages present.
 This framework allows us to investigate a range of functionalist theories broadly under the categories of efficient language design and language use.
@@
 We focus on three questions -- (1) when is it desirable to have a language with lexical ambiguity?
-(2) Under what circumstances do speakers and listeners use ambiguity efficiently?
-(3) Assuming Gricean agents what is the impact of increasing pragmatics on efficiency?\par
+(2) Under what circumstances do speakers and listeners use ambiguity efficiently?\par
 
@@
 Following this introduction we derive a simple objective function for measuring the efficiency of linguistics systems in this setting.
-Subsequently, we move on to three case-studies examining the questions posed above, framing results in terms of our efficiency measure.
+Subsequently, we move on to two case-studies examining the questions posed above, framing results in terms of our efficiency measure.
 
@@
 \par
-We now derive a simple objective function from Zipfian first principles.
+
+We now derive an objective function from Zipfian first principles.
 Zipf proposed that the particular distributional properties found in natural language emerge as a result of competing speaker and listener pressures.
-We operationalize this in equation (1) -- the efficiency of a linguistic system $\ell$ being used by speaker and listener agents $S$ and $L$ is sum of the expected speaker and listener effort to communicate.
+We operationalize this in equation (1) -- the efficiency of a linguistic system $\ell$ being used by speaker and listener agents $S$ and $L$ is sum of the expected speaker and listener effort to communicate over all possible communicative events.
+
 \begin{equation}
-\text{Efficiency}(\ell) = \mathbb{E}_{e \in E}[\text{speaker effort}] + \\ \mathbb{E}_{e \in E}[\text{listener effort}]
+\text{Efficiency}(S, L, \ell) = \mathbb{E}_{e \in E}[\text{speaker effort}] + \\ \mathbb{E}_{e \in E}[\text{listener effort}]
 \end{equation}
@@
 \begin{equation}
-= \mathbb{E}_{e \in E}[-log_2(p(u))] + \\
+\begin{split}
+\text{Efficiency}(S, L, \ell) = \mathbb{E}_{e \in E}[-log_2(p(u))] + \\
 \mathbb{E}_{e \in E}[-log_2(L(m|u; \ell))]
+\end{split}
 \end{equation}
@@
 
-# Three case-studies on efficient language use and design
-We now have a metric for exploring questions of langauge optimality in the reference game setting (the speaker-listener cross-entropy).
+# Two case-studies on efficient language use and design
+We now have a metric for exploring questions of language optimality in the reference game setting (the speaker-listener cross-entropy).
 In the following sections we examine the behavior of the objective with respect to the communicative function of ambiguity.
@@
 The first is a question of optimal design -- given a set of possible languages $L$ in what circumstances does the optimal languages $\ell^*$ contain ambiguity?
-The second is a question of optimal *use* -- give a particular language that theoretically allows for the use of ambiguious material under what circumstances will ambiguous forms be used efficiently?
+The second is a question of optimal *use* -- give a particular language that theoretically allows for the use of ambiguous material under what circumstances will ambiguous forms be used efficiently?
 
@@
 (2012) provided empirical evidence for this claim, showing (SAY WHAT PIANTADOSI DID).
-We consider their basic arguemnt in the reference game setting.
-Our approach procceds as follows, we consider the full-space of valid languages (semantic mappings) between forms and utterances in which $|U|=|M|=4$.
-A language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assigment).
+We consider their basic argument in the reference game setting examining the full-space of valid (footnote here for “valid” we have the constraint that all meanings must be expressible) languages (semantic mappings) between forms and utterances in which $|U|=|M|=4$.
+A language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment).
 For example, a one word language that is fully ambiguous would map a single form to all possible meanings.
@@
 That is, it is useful to assign words multiple meanings if those words can be disambiguated in context.
-Formalized in our setting, this would mean that we expect $\ell*$ to be more likely to contain ambiguous mappings as the number of contexts increases.par
+Formalized in our setting, this would mean that we expect $\ell*$ to be more likely to contain ambiguous mappings as the number of contexts increases.\par
 Figure 1.

### 2019-01-08 paper/paper.Rmd
paper updates, still very rough

@@
 keywords:
-    "Communicaive efficiency, Rational Speech Act theory, computational modeling, information theory, agent-based simulation"
+    "Communicative efficiency, Rational Speech Act theory, computational modeling, information theory, agent-based simulation"
 
@@
 Why do natural languages look they way they do?
-While Zipf (1935) presented his “Principle of Least effort” as a domain general framework for understanding human behavior, he took language as his central case-study.
-He proposed that distributional properties found in natural language were evidence of speaker-listener utility maximization or, equivalently, effort minimization.
-In his own words, “we are arguing that people do in fact act with a maximum economy of effort, and that therefore in the process of speaking-listening they will automatically minimize the expenditure of effort.” Evidence supporting this hypothesis has been realized in a set of well-known findings, largely falling under what is now known as Zipf’s Law or Zipfian distributions -- one of a family of discrete power law distributions.
-A range of empirical phenomena in social and physical sciences are well-described as Zipfian, from the relationship between word rank and frequency (CITATION) to the populations of U.S.
-cities (CITATION) Stanford stats class 306a).\par
+While Zipf (1935) presented his “Principle of Least Effort” as a domain general framework for understanding human behavior, he took language as his central case-study.
+He proposed that distributional properties found in natural language were evidence of speaker-listener effort minimization.
+In his own words, “we are arguing that people do in fact act with a maximum economy of effort, and that therefore in the process of speaking-listening they will automatically minimize the expenditure of effort.” Evidence for this claim was largely derived at the level of the lexicon.
+Zipf argued that the particular relationship between a word’s frequency and its rank, length, and denotation size could be explained as emergent phenomena from speaker-listener effort minimization.
+\par
 
-Language-structure and use provide a particularly fascinating case-study under an efficiency-based analysis as speaker- and listener-effort is asymmetric.
+Zipf articulated what is now considered a *functionalist* approach to language science -- analyzing language structure and use in terms of efficiency.
+Such an approach might reframe our opening question as follows -- how does having property $x$ make using language $\ell$ more or less useful for communication?
+This reframing provides an opportunity to study both language *structure* (as Zipf primarily did) or *use* in terms of efficiency.
+For example, Regier et al.
+(201X) showed that languages appear organize semantic domains (form-meaning mappings) in maximally efficient ways.
+Likewise, Piantadosi et al.
+(2012) argued that lexical ambiguity is a useful property of any communication system when communication is contextualized.
+We will return to Piantadosi’s argument later in the paper.\par
+
+As the opening question indicates, language has both *structure* and *users*.
+Either of these dimensions are available to analysis.
+For example, both Regier (CITATION) and Piantadosi, et al (2012) assumed that language users were approximately Bayesian.
+In particular, Piantadosi needed to assume that language users could *use* context during conversation to derive their particular argument.
+\par
+
+// Maybe don’t include par below...
+Other work has focused more directly on efficient use of linguistic properties.
+For example, Levy & Jaeger (2007) showed speakers will choose to include an optional language items when oncoming material is high-surprisal.
+They argued this behavior indicates an efficient strategy under an information-theoretic analysis, following from Shannon’s Source Coding Theorem.
+Such efficient use of redundancy has also been described at the morphological level (CITATION) and across multiple languages (CITATION).
+Similar theories have described efficient production behavior at the level of discourse.\par
+
+There is little coincidence that Zipf introduced his Principle of Least Effort with an analysis of language.
+Language-structure and -use provide a particularly fascinating case-study under an efficiency-based analysis as speaker- and listener-effort is asymmetric.
 Simply put, in the act of communicating with another person, what is effortful as a speaker is likely different from what is effortful as a listener.
@@
 Clearly natural languages fall between the two extremes of $\ell_{speaker}^*$ and $\ell_{listener}^*$.
-Zipf’s Law of Meaning proposes that the particular relation between a word’s frequency and denotation size emerge from the competition of these forces -- the pressure to jointly minimize speaker and listener effort.\par
+Zipf proposed that the particular lexicon-level properties he observed emerged from the competition of these forces -- the pressure to jointly minimize speaker and listener effort.\par
 
-A range of more recent approaches to language science have distinctly Zipfian roots (functionalist theories).
-In terms of our opening question, “why do languages look the way they do?”, a functionalist approach would formalize the question in terms of information-theoretic principles of efficiency -- how does having property $x$ make using language $\ell$ more or less useful for communication?
-This family of theories has described a range of phenomena in language-design (compositionality, ambiguity, dependency length) and language-use (redundancy and reduction) in terms of communicative efficiency.
-Of particular interest to our current work, Piantadosi et al.
-(2012) argued that efficiently *designed* linguistic systems will contain ambiguous lexical items when communication is contextualized.
-If words with multiple meanings can be disambiguated in context, less costly forms can be re-used efficiently.
-They showed in a series of analyses that … [WHAT PIANTADOSI DID HERE!!!!].
-\par
-
-Zipf’s principle of least effort and it’s associated speaker and listener instantations were intended as a framework for explaining all of natural language.
-Evidence for the theory was largely derived from vocabulary-level analysis.
 Implicit in Zipf’s (1935) argument is that aggregate, distributional-level effects emerge from optimization during local interactions between speaker-listeners.
-While Functionalist projects have provided evidence that speaker’s attempt to optimize their output at a variety of levels -- phonologically (morphological contraction; Bell et al., 2003), syntactically (Levy & Jaeger, 2007; Jaeger, 2010) and over even over discourse (Genzel & Charniak, 2002), Zipf was less precise in terms of his exact level of analysis -- however his example of optimal speaker and listener language in which a single word $u$ can express some meaning $m$$[[u]]\neq \emptyset$.
+However, Zipf did not formalize ``local interaction” beyond the example of a optimal speaker and listener languages.
+In that example, he described a reference game setting; similar to those described by Wittgenstein (CITATION).
+In this scenario, speakers and listeners are aware of a set of objects $M$, which will refer to as *meanings* and are knowledgeable about the set of possible signals $U$ (*utterances*) that could be used by a speaker to refer to a given meaning.
+Utterances may differ in their cost, the number of meanings to which they refer, and the objects may also vary in the degree to which they need to be described in a given context.
+Subsequent functionalist projects examining efficient language structure (Piantadosi, Regier) have also adopted the reference game setting as the primary unit of analysis.
 
@@
 
-In this work we present a step toward formalizing this connection highlighted by Horn -- formalizing the theoretical connection between Zipfian and Grice, and in doing so, deriving aggregate, distributional level properties from local, in-the-moment interactions of speaker-listeners.
+In this work we present a step toward formalizing this connection highlighted by Horn -- formalizing the theoretical connection between Zipf and Grice, and in doing so, deriving aggregate, distributional level properties from local, in-the-moment interactions of speaker-listeners.
 To gain traction on this project we adopt a simulation-based approach, modeling communication as reference games in which we can vary the types of agents and languages present.
@@
 
-For this particular project we focus on an analysis of the communicative function of ambiguity in language following work by Piantadosi et al.
+As a proof of concept, we focus on an analysis of the communicative function of ambiguity in language following work by Piantadosi et al.
 (2012).
 We adopt this particular linguistic phenomena to demonstrate the extent to which we can explore both the design- and use- based functionalist ideas.
-We focus on three questions -- (1) when is it desirable to have a language with lexical ambiguity?
-(2) Under what circumstances do speakers and listeners use ambiguity efficiently?\par
+We focus on the question of efficient language structure and efficient language use.
+In the former, we ask ''when is it desirable to have a language with lexical ambiguity?” In the latter we ask ``under what circumstances do speakers and listeners use ambiguity efficiently?” \par
 
 We begin with a high-level introduction to the modelling framework introducing the basic ingredients we will need to represent language as repeated reference games.
-Following this introduction we derive a simple objective function for measuring the efficiency of linguistics systems in this setting.
+Following this introduction we derive a simple objective function for measuring the efficiency of linguistic systems in this setting.
 Subsequently, we move on to two case-studies examining the questions posed above, framing results in terms of our efficiency measure.
@@
 
-To assess the optimality of a linguistic system we need some measure of goodness.
-In other words an objective function, which, as a function of the particular linguistic system $\ell$ and the speaker $S$ and listener $L$ agents returns a measure of the system efficiency ( $f(S, L, \ell) \rightarrow \mathbb{R}$).
-For our purposes we’d like to tie this objective to fundamental principles of speaker and listener effort suggested by Zipf.
+Given a linguistic system $\ell$ used by a speaker $S$ and listener $L$ we need some measure of efficiency.
+For our present purposes, we’d like to tie this objective to fundamental principles of speaker and listener effort suggested by Zipf.
+To this end, we define an objective function, which as a function of a particular linguistic system $\ell$, speaker $S$ and listener $L$ returns some measure of efficiency (i.e.
+$f(S, L, \ell) \rightarrow \mathbb{R}$).
 Additionally, we will take the fundamental unit of analysis as the reference game event.
@@
 
-We now derive an objective function from Zipfian first principles.
 Zipf proposed that the particular distributional properties found in natural language emerge as a result of competing speaker and listener pressures.
@@
 
-# General Formatting Instructions
-
-For general information about authoring in markdown, see **[here](http://rmarkdown.rstudio.com/authoring_basics.html).**
-
-For standard spoken papers and standard posters, the entire 
-contribution (including figures, references, everything) can be 
-no longer than six pages. For abstract posters, the entire contribution 
-can be no longer than one page. For symposia, the entire contribution 
-can be no longer than two pages.
-
-The text of the paper should be formatted in two columns with an
-overall width of 7 inches (17.8 cm) and length of 9.25 inches (23.5
-cm), with 0.25 inches between the columns. Leave two line spaces
-between the last author listed and the text of the paper. The left
-margin should be 0.75 inches and the top margin should be 1 inch.
-\textbf{The right and bottom margins will depend on whether you use
-U.S. letter or A4 paper, so you must be sure to measure the width of
-the printed text.} Use 10 point Times Roman with 12 point vertical
-spacing, unless otherwise specified.
-
-The title should be in 14 point, bold, and centered. The title should
-be formatted with initial caps (the first letter of content words
-capitalized and the rest lower case). Each author's name should appear
-on a separate line, 11 point bold, and centered, with the author's
-email address in parentheses. Under each author's name list the
-author's affiliation and postal address in ordinary 10 point  type.
-
-Indent the first line of each paragraph by 1/8~inch (except for the
-first paragraph of a new section). Do not add extra vertical space
-between paragraphs.
-
-# First-Level Headings
-
-First level headings should be in 12 point , initial caps, bold and
-centered. Leave one line space above the heading and 1/4~line space
-below the heading.
-
-## Second-Level Headings
-
-Second level headings should be 11 point , initial caps, bold, and
-flush left. Leave one line space above the heading and 1/4~ line
-space below the heading.
-
-### Third-Level Headings
-
-Third-level headings should be 10 point , initial caps, bold, and flush
-left. Leave one line space above the heading, but no space after the
-heading.
-
-# Formalities, Footnotes, and Floats
-
-Use standard APA citation format. Citations within the text should
-include the author's last name and year. If the authors' names are
-included in the sentence, place only the year in parentheses, as in
-[-@NewellSimon1972a], but otherwise place the entire reference in
-parentheses with the authors and year separated by a comma
-[@NewellSimon1972a]. List multiple references alphabetically and
-separate them by semicolons [@ChalnickBillman1988a; @NewellSimon1972a]. 
-Use the et. al. construction only after listing all the authors to a
-publication in an earlier reference and for citations with four or
-more authors.
-
-For more information on citations in RMarkdown, see **[here](http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html#citations).**
-
-## Footnotes
-
-Indicate footnotes with a number\footnote{Sample of the first
-footnote.} in the text. Place the footnotes in 9 point type at the
-bottom of the page on which they appear. Precede the footnote with a
-horizontal rule.\footnote{Sample of the second footnote.} You can also use 
-markdown formatting to include footnotes using this syntax [^1].
-
-[^1]: Sample of a markdown footnote.
-
-## Figures
-
-All artwork must be very dark for purposes of reproduction and should
-not be hand drawn. Number figures sequentially, placing the figure
-number and caption, in 10 point, after the figure with one line space
-above the caption and one line space below it. If necessary, leave extra white space at
-the bottom of the page to avoid splitting the figure and figure
-caption. You may float figures to the top or bottom of a column, or
-set wide figures across both columns.
-
-## Two-column images
-
-You can read local images using png package for example and plot 
-it like a regular plot using grid.raster from the grid package. 
-With this method you have full control of the size of your image. **Note: Image must be in .png file format for the readPNG function to work.**
-
-You might want to display a wide figure across both columns.
-To do this, you change the `fig.env` chunk option to `figure*`.
-To align the image in the center of the page, set `fig.align` option to `center`.
-To format the width of your caption text, you set the `num.cols.cap` option to `2`.
-
-```{r 2-col-image, fig.env = "figure*", fig.pos = "h", fig.width=4, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns.
-And the caption text is limited to 0.8 of the width of the document."}
-img <- png::readPNG("figs/walrus.png")
-grid::grid.raster(img)
-```
-
-## One-column images
-
-Single column is the default option, but if you want set it explicitly, set `fig.env` to `figure`. Notice that the `num.cols` option for the caption width is set to `1`.
-
-```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
-img <- png::readPNG("figs/lab_logo_stanford.png")
-grid::grid.raster(img)
-```
-
-## R Plots
-
-You can use R chunks directly to plot graphs. And you can use latex floats in the
-fig.pos chunk option to have more control over the location of your plot on the page.
-For more information on latex placement specifiers see **[here](https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions)**
-
-```{r plot, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2, fig.height=2, fig.cap = "R plot" }
-x <- 0:100
-y <- 2 * (x + rnorm(length(x), sd = 3) + 3)
-
-ggplot2::ggplot(data = data.frame(x, y), 
-       aes(x = x, y = y)) + 
-  geom_point() + 
-  geom_smooth(method = "lm")
-```
-
-## Tables
-
-Number tables consecutively; place the table number and title (in
-10 point) above the table with one line space above the caption and
-one line space below it, as in Table 1. You may float
-tables to the top or bottom of a column, set wide tables across both
-columns.
-
-You can use the xtable function in the xtable package.
-
-```{r xtable, results="asis"}
-n <- 100
-x <- rnorm(n)
-y <- 2*x + rnorm(n)
-out <- lm(y ~ x)
-
-tab1 <- xtable::xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
-                      caption = "This table prints across one column.")
-
-print(tab1, type="latex", comment = F, table.placement = "H")
-```
-
 # Acknowledgements

### 2019-01-09 paper/paper.Rmd
additional updates to paper, including more prelim writing for discourse section

@@
 Zipf articulated what is now considered a *functionalist* approach to language science -- analyzing language structure and use in terms of efficiency.
-Such an approach might reframe our opening question as follows -- how does having property $x$ make using language $\ell$ more or less useful for communication?
+Such an approach might reframe our opening question as follows -- how does having \textbf{property x} make using language $\ell$ more or less useful for communication?
 This reframing provides an opportunity to study both language *structure* (as Zipf primarily did) or *use* in terms of efficiency.
 For example, Regier et al.
-(201X) showed that languages appear organize semantic domains (form-meaning mappings) in maximally efficient ways.
+(201X) showed that languages appear to organize semantic domains (form-meaning mappings) in maximally efficient ways.
 Likewise, Piantadosi et al.
-(2012) argued that lexical ambiguity is a useful property of any communication system when communication is contextualized.
+(2012) argued that lexical ambiguity is an efficient property of any communication system when communication is contextualized.
 We will return to Piantadosi’s argument later in the paper.\par
 
-As the opening question indicates, language has both *structure* and *users*.
-Either of these dimensions are available to analysis.
-For example, both Regier (CITATION) and Piantadosi, et al (2012) assumed that language users were approximately Bayesian.
-In particular, Piantadosi needed to assume that language users could *use* context during conversation to derive their particular argument.
-\par
-
-// Maybe don’t include par below...
-Other work has focused more directly on efficient use of linguistic properties.
-For example, Levy & Jaeger (2007) showed speakers will choose to include an optional language items when oncoming material is high-surprisal.
-They argued this behavior indicates an efficient strategy under an information-theoretic analysis, following from Shannon’s Source Coding Theorem.
-Such efficient use of redundancy has also been described at the morphological level (CITATION) and across multiple languages (CITATION).
-Similar theories have described efficient production behavior at the level of discourse.\par
-
-There is little coincidence that Zipf introduced his Principle of Least Effort with an analysis of language.
-Language-structure and -use provide a particularly fascinating case-study under an efficiency-based analysis as speaker- and listener-effort is asymmetric.
-Simply put, in the act of communicating with another person, what is effortful as a speaker is likely different from what is effortful as a listener.
-For example, Zipf noted that purely from the standpoint of speaker effort, what Zipf called "Speaker's Economy," an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of just a single, low-cost word.
+Zipf’s original work and subsequent functionalist projects describe a varied set of linguistic properties and human behaviors.
+Common to many is a characterization of a fundamental effort-asymmetry underlying everyday communication.
+Simply put, what is effortful as a speaker is likely different from what is effortful as a listener.
+Zipf characterized this as follows -- purely from the standpoint of speaker effort, what Zipf called "Speaker's Economy," an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of just a single, low-cost word.
 Given such a language, the full set of potential meanings would be conveyed using only that word, i.e.
 $\ell_{speaker}^*$ would be fully ambiguous and all possible meanings would need to be disambiguated by a listener.
-Conversely, from the standpoint of listener effort, what Zipf called "Auditor's Economy," an optimal language $\ell_{listener}^*$ would bijectively map all possible meanings to distinct words, eliminating a listener's need to disambiguate.
-Under this analysis, speaker effort is related to production cost and listener effort to understanding (disambiguation) cost.
+From the standpoint of listener effort, what Zipf called "Auditor's Economy," an optimal language $\ell_{listener}^*$ would bijectively map all possible meanings to distinct words, eliminating a listener's need to disambiguate.
+Here the asymmetry is operationalized by different costs -- speaker effort is related to *production cost* and listener effort to *understanding or disambiguation cost*.
 Clearly natural languages fall between the two extremes of $\ell_{speaker}^*$ and $\ell_{listener}^*$.
@@
 
-Implicit in Zipf’s (1935) argument is that aggregate, distributional-level effects emerge from optimization during local interactions between speaker-listeners.
-However, Zipf did not formalize ``local interaction” beyond the example of a optimal speaker and listener languages.
-In that example, he described a reference game setting; similar to those described by Wittgenstein (CITATION).
-In this scenario, speakers and listeners are aware of a set of objects $M$, which will refer to as *meanings* and are knowledgeable about the set of possible signals $U$ (*utterances*) that could be used by a speaker to refer to a given meaning.
-Utterances may differ in their cost, the number of meanings to which they refer, and the objects may also vary in the degree to which they need to be described in a given context.
-Subsequent functionalist projects examining efficient language structure (Piantadosi, Regier) have also adopted the reference game setting as the primary unit of analysis.
+This formulation begs the question -- at what level of analysis do local, in-the-moment interactions of speakers and listeners lead to optimized structure at the level of the lexicon?
+The example given by Zipf (1935) appears to describe local interaction in terms of a reference game.
+Speakers intend to refer to some object in the world $m$.
+They choose some utterance $u$ to transmit this intended meaning, $u \rightarrow m$.
+The listener attempts to reconstruct this intended meaning given the transmitted utterance $m \rightarrow u$.
+Other functionalist projects have also assumed this basic reference game setting (Regier, et al.
+(201X); Piantadosi et al.
+(2012); Kirby, Smith & Folds (201X), OTHERS) and this simplification of the communicative act has proven productive in both theoretical (Ferrer-i-cancho, et al.,  201X), simulation-based (iterated learning stuff) and empirical explorations of efficient language structure and use (Some of Robert’s work here).
 
+This move to study efficient language use in terms of reference games has a strong theoretical underpinnings as well.
 A half century after Zipf, the linguist Lawrence Horn (1984), highlighted the importance of Zipf’s principles for explaining conversation-level phenomena.
 Horn suggested a direct link between Zipf’s Speaker and Listener economy and, what was at the time, more recent work on conversational pragmatics by Grice (1975).
-Horn highlighted that the interaction of Zipf’s forces were “largely responsible for generating Grice’s conversational maxims and the schema for pragmatic inference derived therefrom.” Put differently, system-level efficiency we see in languages is deeply related to local-level efficiency during the in-the-moment interactions of speakers and listeners.
+Horn highlighted that the interaction of Zipf’s forces were “largely responsible for generating Grice’s conversational maxims and the schema for pragmatic inference derived therefrom.” Put differently, system-level efficiency we see in languages is deeply related to local-level efficiency during the in-the-moment interactions of rational pragmatic speakers and listeners.
 \par
 
-In this work we present a step toward formalizing this connection highlighted by Horn -- formalizing the theoretical connection between Zipf and Grice, and in doing so, deriving aggregate, distributional level properties from local, in-the-moment interactions of speaker-listeners.
-To gain traction on this project we adopt a simulation-based approach, modeling communication as reference games in which we can vary the types of agents and languages present.
-This framework allows us to investigate a range of functionalist theories broadly under the categories of efficient language design and language use.
-\par
-
-As a proof of concept, we focus on an analysis of the communicative function of ambiguity in language following work by Piantadosi et al.
-(2012).
-We adopt this particular linguistic phenomena to demonstrate the extent to which we can explore both the design- and use- based functionalist ideas.
-We focus on the question of efficient language structure and efficient language use.
-In the former, we ask ''when is it desirable to have a language with lexical ambiguity?” In the latter we ask ``under what circumstances do speakers and listeners use ambiguity efficiently?” \par
+In this work, we present a step toward formalizing Horn’s observation -- connecting Zipfian notions of efficient language structure to Gricean notions of rational conversation.
+To do so minimally requires three basic ingredients -- (1) a language property we’d like to explain (2) a framework for describing in-the-moment interactions of speaker-listeners and (3) some measure of linguistic efficiency.
+For the current project we focus on the communicative function of ambiguity for our property (1).
+To model local, conversational interactions (2) we adopt the Rational Speech-act framework.
+Finally, (3) we derive an objective function for measuring language efficiency in the reference game setting.
+Given this framework we can ask the question, “how does having \textbf{lexical ambiguity} make using language $\ell$ more or less useful for communication”?
+This project is meant to demonstrate the ways in which a range of functionalist theories might make use of such a framework -- replacing *lexical ambiguity* with a range of interesting behaviors and properties.\par
 
@@
 ### Reference games
-We take as our unit of analysis the basic referential communication game similar to those described by WittgensteIn (CITATION).
-Both speakers and listeners are aware of a set of objects $M$, which will refer to as *meanings* and are knowledgeable about the set of possible signals $U$ (*utterances*) that could be used to refer to a given meaning.
+Zipf’s canonical (1935) example of optimal speaker- and listener-languages took the form of a reference game.
+We adopt that formulation here, assuming the referential communication, similar to those described by Wittgenstein, as our basic unit of analysis.
+In this setting, speakers and listeners are aware of a set of objects $M$, which will refer to as *meanings* and are knowledgeable about the set of possible signals $U$ (*utterances*) that could be used to refer to a given meaning.
 Utterances may have different relative costs, operationalized via a prior over utterances $p(u)$.
-Similarly, meanings differ in the degree to which they need to be talked about.
-This is operationalized as a prior over meanings $p(m)$.
+Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $p(m)$.
 Note that the prior over meanings are analogous to the *need probabilities* assumed in previous work (Regier, CITATION).
@@
 A language $\ell$ defines the set of semantic mappings between utterance and meanings.
-For example, in a world with three utterances $U = \{u_1, u_2, u_3\} \text{and three meanings }M = \{m_1, m_2, m_3\}$ the boolean matrix
+For example, in a world with three utterances $U = \{u_1, u_2, u_3\} \text{ and three meanings }M = \{m_1, m_2, m_3\}$ the boolean matrix
 
@@
 
-describes the literal semantics of $\ell$. E.g. the language describes semantic mappings $[[u_1]]\ = \{m_1, m_2\}, [[u_2]]\ = \{m_2\}, [[u_3]]\ = \{m_3\}$.
+describes the literal semantics of $\ell$. E.g. the language describes semantic mappings $[\![u_1]\!]\ = \{m_1, m_2\}, [\![u_2]\!]\ = \{m_2\}, [\![u_3]\!]\ = \{m_3\}$.
 
 ### Speakers and listeners
-A *speaker agent* defines a conditional distribution over utterances, mapping from intended meanings $M$ to utterances $U$ using a particular semantic mapping given by $\ell$.
+We adopt the Rational Speech-act framework (Frank & Goodman, 2012; Goodman & Frank, 2016) to operationalize our rational speaker-listeners.
+Under this setting, a *speaker agent* defines a conditional distribution over utterances, mapping from intended meanings $M$ to utterances $U$ using a particular semantic mapping given by $\ell$.
 That is, a speaker defines $P_{speaker}(u|m;\ell)$.
@@
 
-Given a linguistic system $\ell$ used by a speaker $S$ and listener $L$ we need some measure of efficiency.
-For our present purposes, we’d like to tie this objective to fundamental principles of speaker and listener effort suggested by Zipf.
-To this end, we define an objective function, which as a function of a particular linguistic system $\ell$, speaker $S$ and listener $L$ returns some measure of efficiency (i.e.
-$f(S, L, \ell) \rightarrow \mathbb{R}$).
-Additionally, we will take the fundamental unit of analysis as the reference game event.
-\par
-
 Zipf proposed that the particular distributional properties found in natural language emerge as a result of competing speaker and listener pressures.
@@
 
-Let speaker effort be the log probability of a particular utterance. $log_2(p(u))$ E.g. the number of bits needed to encode the utterance $u$.
+Let speaker effort be the negative log probability (surprisal) of a particular utterance. Intuitively, the number of bits needed to encode the utterance $u$.
 
@@
 
-Let listener effort be the log probability a listener disambiguates an intended meaning $m$ given an utterance $u$ E.g.
-the number of guesses a listener would need to discover the intended meaning $m$ given an utterance $u$.
+Let listener effort be the negative log probability a listener disambiguates an intended meaning $m$ given an utterance $u$ Intuitively, the number of guesses a listener would need to discover the intended meaning $m$ given an utterance $u$.
 $$\text{listener effort} = -log_2(L(m|u; \ell))$$
@@
 From a mathematical standpoint this objective is intuitive.
-An optimal linguistic system being used by a speaker and listener agent will induce a particular set of form-meaning mappings that are both *close* between speaker and listener agents and *peaky* in that they assign non-uniform probability mass to the valid form-meaning mappings.
+An optimal linguistic system being used by a speaker and listener agent will induce a particular set of form-meaning mappings that are both *close* between speaker and listener agents and *peaky* in that they assign non-uniform probability mass to the valid form-meaning mappings.\par
+
+In terms of Zipf’s initial example about an optimal speaker or listener language we might consider just the first or second term.
+Hypothetically, we might expect the optimal speaker language $\ell_{speaker}*$ to minimize:
+
+\begin{equation}
+\ell_{speaker}* = argmin_{\ell\in L}\mathbb{E}_{P_{speaker}(u, m; \ell)}(p(u))
+\end{equation}
+
+Likewise we might expect the optimal listener language $\ell_{listener}*$ to minimize:
+
+\begin{equation}
+\ell_{listener}* = argmin_{\ell\in L}\mathbb{E}_{P_{speaker}(u, m; \ell)}(L(m|u;\ell)
+\end{equation}
 
@@
 In the following sections we examine the behavior of the objective with respect to the communicative function of ambiguity.
-In particular, we examine three questions.
-The first is a question of optimal design -- given a set of possible languages $L$ in what circumstances does the optimal languages $\ell^*$ contain ambiguity?
-The second is a question of optimal *use* -- give a particular language that theoretically allows for the use of ambiguous material under what circumstances will ambiguous forms be used efficiently?
 
-## Experiment 1:  Rational-pragmatic speakers prefer languages with ambiguous items
+Ambiguity is ubiquitous in natural language.
+At a minimum, the listener task in everyday communication is marked by the frequent need to handle lexical ambiguity (words often have multiple meanings) as well as syntactic ambiguity (sentences often have multiple parses), in addition to many other instances of ambiguity (Wasow, Perfors & Beaver, 201X).
+Chomsky (2002) famously claimed that the presence of ambiguity in natural language provides evidence that language has not been optimized for communication, stating “If you want to make sure that we never misunderstand one another, for that purpose language is not well designed, because you have such properties as ambiguity.” Such an analysis is proximal to the Zipfian optimal Listener language, in which the listener is never required to disambiguate a speaker’s utterance.\par
 
-Piantadosi et al.
-(2012) proposed that a language with ambiguous items is strictly more efficient than a language with out, *when ambiguous items can be disambiguated with context*.
-Formally, they authors highlighted the fact that if we measure the efficiency of a linguistic system in terms of entropy, we have the basic result from (information theory citation) that the conditional entropy of an ensemble is strictly less that the unconditional entropy, when the information we condition on is informative ($H(X|C) < H(X)$).\par
+Piantadosi et al, (2012) essentially argue the opposite.
+They claim that ambiguity is an *efficient* property of any communication system in which *communication is contextualized*.
+We provide the heart of their argument here, however we refer to Piantados et al.
+(2012) for more detail:
 
-Piantadosi et al.
-(2012) provided empirical evidence for this claim, showing (SAY WHAT PIANTADOSI DID).
-We consider their basic argument in the reference game setting examining the full-space of valid (footnote here for “valid” we have the constraint that all meanings must be expressible) languages (semantic mappings) between forms and utterances in which $|U|=|M|=4$.
-A language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment).
-For example, a one word language that is fully ambiguous would map a single form to all possible meanings.
-We represent that via the boolean matrix $\ell_1$ in which $[[u_1]] = \{m_1,m_2, m_3 \}, [[u_2]]=[[u_3]]=\{\}$
+*In an unambiguous linguistic system, $m_1$ is mapped to $\ell_1 \in L$ and $m_2$ is mapped to $\ell_2 \in L$, with $\ell_1 \neq \ell_2$.
+Suppose that $\ell_1$ is easier than $\ell_2$...
+If $\ell_1$ is easier than $\ell_2$, and $m_1$ and $m_2$ are well-disambiguated by context, then we can always create a linguistic system which is easier overall by mapping $m_1$ and $m_2$ both to $\ell_1$.
+This costs the same in terms of effort every time we communicate $m_1$, but saves effort every time we communicate $m_2$*\par
 
-\begin{center}
-\begin{tabular}{ c | c c c} 
-& $m_1$ & $m_2$ & $m_3$ \\
-\hline
-$u_1$ & 1 & 1 & 1 \\
-$u_2$ & 0 & 0 & 0 \\
-$u_3$ & 0 & 0 & 0 \\
-\end{tabular}
-\end{center}
+In words, the Piantadosi argument might be summarised as -- it is useful to have a language that re-uses low-cost material (has ambiguity) so long as the cost of disambiguating the material is low (so long as the context is informative with respect to the disambiguation).
 
-Note that $\mathbb{E}[\text{speaker effort}]$ can be arbitrarily low for such a language, as it is directly proportional to $p(u_1)$.
-That is in the single utterance language, $\mathbb{E}[\text{speaker effort}] \rightarrow 0$ as $p(u_1) \rightarrow 1$.
-However, $\mathbb{E}[\text{listener effort}]$ is likely to be high as the only available form ($u_1$) is full ambiguous between possible meanings.\par
-By contrast consider a language $\ell_2$:
+In the following experiments we inspect two aspects of Piantadosi’s claim.
+In the first, we examine the efficient language *structure* aspect of the question exploring when the optimal linguistic system (under our cross-entropy objective) is most likely to contain ambiguous lexical items.
+In the second experiment we explore an efficient language *use* aspect of the question -- at what point in a conversation is it useful for a speaker to use ambiguous lexical material?
 
-\begin{center}
-\begin{tabular}{ c | c c c} 
-& $m_1$ & $m_2$ & $m_3$ \\
-\hline
-$u_1$ & 1 & 0 & 0 \\
-$u_2$ & 0 & 1 & 0 \\
-$u_3$ & 0 & 0 & 1 \\
-\end{tabular}
-\end{center}
+## Experiment 1:  Optimal languages contain ambiguity when context is informative
 
-In this case $\mathbb{E}[\text{listener effort}]=0$ -- each form has a one-to-one mapping to a given meaning.
-However speaker effort will likely be greater than in $\ell_1$ so long as $p(u_1) < p(u_2)$ and $p(u_1) < p(u_3)$.
+The argument outlined by Piantadosi et al.
+(2012) is clearly compatible with our current reference game setting with just a single addition -- context.
+The authors argue that mapping both $m_1$ and $m_2$ to $l_1$ is useful when they can disambiguated *in context*.
+In our case we consider a context $c$ to specify a re-ordering of the referent need probabilities.
+That is, $p(m|c_1) \neq p(m|c_2)$.
+In the two context case.
 
-Finally, consider a langauge $\ell_3$:
+This in turn, leads to an update to our linguistic efficiency objective as we now would like to consider the average speaker-listener effort over all contextualized communicative events. That is,
 
-\begin{center}
-\begin{tabular}{ c | c c c} 
-& $m_1$ & $m_2$ & $m_3$ \\
-\hline
-$u_1$ & 1 & 1 & 0 \\
-$u_2$ & 0 & 0 & 0 \\
-$u_3$ & 0 & 0 & 1 \\
-\end{tabular}
-\end{center}
+\begin{equation}
+\sum_{c\in C}p(c)\sum_{u, m}S(u|m, c;\ell)p(m|c)-log[L(m|u,c;\ell)p(u)] = \\ \mathbb{E}_{c\sim P(c)[\mathbb{E}_{P_speaker(u, m, c; \ell)}[-log(P_listener(u, m, c; \ell))]}]
+\end{equation}
 
-Our third language contains an ambiguous lexical item - $u_1$ can be used to refer to either $m_1$ or $m_2$.
-This language represents a trade-off between $\ell_1$ and $\ell_2$.
-From the speaker perspective $\ell_3$ likely requires more effort from the speaker than $\ell_1$, however it likely requires less than $\ell_2$ (so long as $p(u_1)$ is cheap).
-From the listener perspective $\ell_3$ requires more effort than $\ell_3$ which requires no disambiguation, but less than $\ell_1$.
-\par
-As this small set of languages demonstrate, the efficiency of a particular language depends on both utterances costs $P(u)$ and need probabilities $P(M)$.
-E.g.
-If $u_1$ is actually a costly utterance ($p(u_1)$ is close to 0) then $\ell_1$ is inefficient both for the speaker and listener.
-For this reason, in the following simulations, we randomly sample utterance costs $P(u)$ and need probabilities $P(m)$.
-Our simulations procced as follows.
-For $n=100$ simulations we enumerate all valid langauges ($4\times 4$ boolean matrices) as well as a set of need probabilities ($p(m)$) and utterance costs ($p(u)$).
+Note that in the case that $|C| =1$, our objective simplifies to our original equation (7).
+With this update we can compare linguistics systems $L$ while varying the degree to which context contains useful information by varying the size of $|C|$.\par
 
-Given this set-up we can find the optimal language:
+### Simulation set-up
 
-$$\ell^* = \text{argmin}_{\ell \in L} H_{Cross}(P_{Speaker}(u, m; \ell), P_{Listener}(u, m; \ell))$$
+#### Basic ingredients
+We conduct $N=400$ simulations.
+For each simulation we enumerate the set of *valid* languages in which $|U|=|M|=4$.
+Note that a language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment).
+For a given simulation the goal is to find the language $\ell^*$ which minimizers our cross-entropy objective.\par
 
-Critical to Piantadosi's claim, however is the fact that ambiguous langauges should be preferred *when* context is disambiguating.
-For that reason we compare the optimal langauge $\ell^*$ as the amount of contextual information increases.
-In our case a "context" describes a conditional distribution over meanings.
-For example, take a setting with two contexts $c_1$ and $c_2$.
-The first context dictates a set of need probabilites which differs from the second ($p(m|c_1) \neq p(m|c_2)$).
-If ambiguity is useful when context is disambiguating we should expect the likelihood that an optimal langauge $\ell^*$ contains an ambiguous item to increase as the number of contexts increases.
+Recall that language efficiency is both a function of the particular semantic mappings induced by that language, the Speaker and Listener agents, as well as the utterance and meaning priors.
+Therefore, we randomly generate priors of these values such that $p(m) \sim \text{Dir}(1, |M|)$ and $p(u) \sim \text{Dir}(1, |U|)$ where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.
 
-### Simulation set-up. \newline
-
-$L = \text{set of all valid languages with} |U|=|M|=4$\par
-$p(m) = \text{Dirichlet}(ones)$\par
-$S(u|m; \ell) = \text{S1 rsa speaker}$\par
-$L(m|u; \ell) = \text{L0 rsa listener}$\par
-
-### Results
-We run $n=100$ simulations in four condtions.
+#### Context
+Given that we’d like to assess the impact of *context* on our objective we consider four conditions with $n=100$ each.
 Our first is a *single-context* condition ( $|C|=1$) -- there is a only a single context describing $p(m|c_1)$.
@@
 The third and fourth condition correspond accordingly with $|C|=3$ and $|C| = 4$, respectively.
-In each condition we consider a random set of utterances costs $P(u)$ and need probabilities $p(m|c)$ parametrized as Dirichlet distributions with $\alpha=1$.
-Piantadosi et al.
-(2012) argued that ambiguity is an efficient property of contextualized langauge use.
-That is, it is useful to assign words multiple meanings if those words can be disambiguated in context.
-Formalized in our setting, this would mean that we expect $\ell*$ to be more likely to contain ambiguous mappings as the number of contexts increases.\par
+
+#### Baselines
+For comparison, we also examine properties of optimal languages under two  additional objectives.
+Zipf (1935) proposed that a language optimized for speaker-ease should consist of a single word.
+We operationalize this using the first half of equation (1).
+
+\begin{equation}
+\text{Speaker-only objective} = \mathbb{E}_{u, m\sim P_{speaker}}[\text{speaker effort}] = \sum_{u, m}P_{speaker}(u, m; \ell)[-log(p(u))]
+\end{equation}
+
+Zipf also proposed that the optimal listener language should bijectively map all words to unique meanings. We operationalize this using the second half of equation (1).
+\begin{equation}
+\text{Listener-only objective} = \mathbb{E}_{u, m\sim P_{speaker}}[\text{listener effort}] = \sum_{u, m}P_{speaker}(u, m; \ell)[-log(L(m|u; \ell)]
+\end{equation}
+
+#### Hypotheses
+This leads to the following set of hypotheses: (1) The probability that an optimal language contains ambiguity under our Cross-Entropy objective should increase with the number of contexts (as $|C|\rightarrow\inf$).
+(2) The optimal speaker language under our Speaker-only objective should always map all meanings to the single, lowest cost utterance.
+(3) The optimal listener language should never contain ambiguous material.
+
+### Results
 Figure 1.
@@
 For comparison we also include the optimal language under an objective that only considers speaker or listener effort.
-In line, with Zipf's predictions, if langauges are designed only to minimize speaker effort then optimal languages will assign all meanings to a single, low-cost utterance.
-Likewise, if langauges are designed only to minimize speaker listener effor then ambiguity should alwasy be avoided.
-
-```{r fig1, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=4, fig.height=4, fig.cap = "Optimal languages are more likely to contain ambiguous items as amount of contextual information increases.
-Vertical axis shows the proportion of optimal languages that contain ambiguity.
-Horizontal axis shows the number of contexts in each condition (1-4).
-Red-line represents the optimal langauge under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only consideration."}
-img <- png::readPNG("figs/fig1.png")
-grid::grid.raster(img)
-```
+In line, with Zipf's predictions, if languages are designed only to minimize speaker effort then optimal languages will assign all meanings to a single, low-cost utterance.
+Likewise, if languages are designed only to minimize speaker listener effort then ambiguity should always be avoided.\par
 
@@
 Figure 2 plots $H_{cross}(P_{Speaker}, P_{Listener}|C_1) - H_{cross}(P_{Speaker}, P_{Listener})$ in each of our conditions.
-As the number of contexts increases, the difference in efficiency when context is disambiguating increaess as well.
-```{r fig2, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2.5, fig.height=3, fig.cap = "The gains in efficinecy increase as the amount of contextual information increases.
-Vertical axis shows the difference in cross-entropy for when context is informative.
-Horizontal axis shows each of the four conditions.
-Note that the difference for condition 1 is zero as thered is only a single context." }
-img <- png::readPNG("figs/fig2.png")
-grid::grid.raster(img)
-```
+As the number of contexts increases, the difference in efficiency when context is disambiguating increaess as well.\par
 
@@
 Using our speaker-listener cross-entropy measure of efficiency we showed that optimal langauges are more likely to have ambiguous items when context is informative.
-Further, we showed that the impact of being able to disambiguate langauge (having access to $p(m|c_i)$) is increasingly efficienct as the amount of common-ground (context) increases.
+Further, we showed that the impact of being able to disambiguate langauge (having access to $p(m|c_i)$) is increasingly efficienct as the amount of common-ground (context) increases.\par
 
@@
 Put differently, if ambiguity is only efficient when it can be disambiguated in context, speakers should avoid using ambiguous items if they know their listener may not be able to disambiguate the item.\par
+
 In experiment two we examine speaker behavior in a scenario in which the listener does not know the exact set of need probabilites a priori.
@@
 
-Fig3
-```{r fig3, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2.5, fig.height=3, fig.cap = "R plot" }
-img <- png::readPNG("figs/fig3.png")
-grid::grid.raster(img)
-```
-
-Fig4
-```{r fig4, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2.5, fig.height=3, fig.cap = "R plot" }
-img <- png::readPNG("figs/fig4.png")
-grid::grid.raster(img)
-```
-
 ### Summary
-
-## Experiment 3: “More” pragmatic leads to more efficiency
-
-### Simulation set-up
-
-### Results
-
-Fig5
-```{r fig5, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2.5, fig.height=3, fig.cap = "R plot" }
-img <- png::readPNG("figs/fig5.png")
-grid::grid.raster(img)
-```
-
-### Summary
-
-# Discussion
 
@@
 
+```{r fig1, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=4, fig.height=4, fig.cap = "Optimal languages are more likely to contain ambiguous items as amount of contextual information increases.
+Vertical axis shows the proportion of optimal languages that contain ambiguity.
+Horizontal axis shows the number of contexts in each condition (1-4).
+Red-line represents the optimal langauge under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only consideration."}
+img <- png::readPNG("figs/fig1.png")
+grid::grid.raster(img)
+```
+
+```{r fig2, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2.5, fig.height=3, fig.cap = "The gains in efficinecy increase as the amount of contextual information increases.
+Vertical axis shows the difference in cross-entropy for when context is informative.
+Horizontal axis shows each of the four conditions.
+Note that the difference for condition 1 is zero as thered is only a single context." }
+img <- png::readPNG("figs/fig2.png")
+grid::grid.raster(img)
+```
+
+```{r fig3, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2.5, fig.height=3, fig.cap = "R plot" }
+img <- png::readPNG("figs/fig3.png")
+grid::grid.raster(img)
+```
+
+```{r fig4, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2.5, fig.height=3, fig.cap = "R plot" }
+img <- png::readPNG("figs/fig4.png")
+grid::grid.raster(img)
+```
+
+```{r fig5, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2.5, fig.height=3, fig.cap = "R plot" }
+img <- png::readPNG("figs/fig5.png")
+grid::grid.raster(img)
+```
+
 ```{r}

### 2019-01-14 paper/paper.Rmd
fixed latex errors, updated RD -- need to work on experiment two section.

@@
 In his own words, “we are arguing that people do in fact act with a maximum economy of effort, and that therefore in the process of speaking-listening they will automatically minimize the expenditure of effort.” Evidence for this claim was largely derived at the level of the lexicon.
-Zipf argued that the particular relationship between a word’s frequency and its rank, length, and denotation size could be explained as emergent phenomena from speaker-listener effort minimization.
+Zipf argued that the particular relationship between a word’s frequency and its rank, length, and denotation size could be explained as an emergent property of speaker-listener effort minimization.
 \par
@@
 
-This move to study efficient language use in terms of reference games has a strong theoretical underpinnings as well.
+This move to study efficient language use in terms of reference games has clear theoretical underpinnings as well.
 A half century after Zipf, the linguist Lawrence Horn (1984), highlighted the importance of Zipf’s principles for explaining conversation-level phenomena.
@@
 To do so minimally requires three basic ingredients -- (1) a language property we’d like to explain (2) a framework for describing in-the-moment interactions of speaker-listeners and (3) some measure of linguistic efficiency.
-For the current project we focus on the communicative function of ambiguity for our property (1).
-To model local, conversational interactions (2) we adopt the Rational Speech-act framework.
+For the current project we focus on the “communicative function of ambiguity” for our property -- translating the functionalist theory pursued by Piantadosi et al.
+(2012) into our framework.
+To model local, conversational interactions (2) we adopt the Rational Speech-act framework (RSA) (Frank & Goodman, 2012; Goodman & Frank, 2016).
 Finally, (3) we derive an objective function for measuring language efficiency in the reference game setting.
-Given this framework we can ask the question, “how does having \textbf{lexical ambiguity} make using language $\ell$ more or less useful for communication”?
-This project is meant to demonstrate the ways in which a range of functionalist theories might make use of such a framework -- replacing *lexical ambiguity* with a range of interesting behaviors and properties.\par
+Given this framework we can ask the question, “how does having \textit{lexical ambiguity} make using language $\ell$ more or less useful for communication”?
+This project is meant to demonstrate the ways in which a range of functionalist theories might make use of such a framework -- in theory replacing \textit{lexical ambiguity} with a range of behaviors and properties identified by functionalist approaches to language science.\par
 
@@
 
-## Simulation set-up.
-
 ### Reference games
-Zipf’s canonical (1935) example of optimal speaker- and listener-languages took the form of a reference game.
-We adopt that formulation here, assuming the referential communication, similar to those described by Wittgenstein, as our basic unit of analysis.
+Zipf’s canonical (1935) example of optimal speaker- and listener-languages took the form of a reference game (similar to those described by Wittgenstein).
+We adopt that formulation here, assuming these communication games as our basic unit of analysis.
 In this setting, speakers and listeners are aware of a set of objects $M$, which will refer to as *meanings* and are knowledgeable about the set of possible signals $U$ (*utterances*) that could be used to refer to a given meaning.
@@
 We consider a set of contexts $C$ which describe different need probability distributions over our set of meanings $p(m|c)$.
+Finally, we consider a set of communicative events $e \in E$ where $<u, m> = e$ is a tuple of utterance, meanings pairs.
+The set $E$ all possible utterance-meaning mappings.
 
@@
 We adopt the Rational Speech-act framework (Frank & Goodman, 2012; Goodman & Frank, 2016) to operationalize our rational speaker-listeners.
-Under this setting, a *speaker agent* defines a conditional distribution over utterances, mapping from intended meanings $M$ to utterances $U$ using a particular semantic mapping given by $\ell$.
+Under this setting, a *speaker agent* defines a conditional distribution over utterances, mapping from intended meanings $M$ to utterances $U$ using $\ell$.
 That is, a speaker defines $P_{speaker}(u|m;\ell)$.
 We will use $S(u|m;\ell)$ as short-hand throughout.
-A *listener agent* defines a conditional distribution over meanings, mapping from utterances $U$ to meanings $M$ using a particular semantic mapping given by $\ell$.
+A *listener agent* defines a conditional distribution over meanings, mapping from utterances $U$ to meanings $M$ using $\ell$.
 We will use $L(m|u; \ell)$ as shorthand.
@@
 In general, we would like to consider the efficiency of a system $\ell$ in terms of these joint distributions.
+For a more detailed description of the RSA framework we refer the reader to Goodman & Frank (2016).
 
@@
 
-## Metrics for optimal linguistic systems in the reference game setting
-
 Zipf proposed that the particular distributional properties found in natural language emerge as a result of competing speaker and listener pressures.
-We operationalize this in equation (1) -- the efficiency of a linguistic system $\ell$ being used by speaker and listener agents $S$ and $L$ is sum of the expected speaker and listener effort to communicate over all possible communicative events.
+We operationalize this in equation (1) -- the efficiency of a linguistic system $\ell$ being used by speaker and listener agents $S$ and $L$ is the sum of the expected speaker and listener effort to communicate over all possible communicative events $E$.
 
@@
 
-Rewriting (1) we now have
+Rewriting (1) we have
 \begin{equation}
@@
 \end{equation}
-In general, these expectations are each taken over all possible communicative events $e\in E$ weighted by the probability of a particular event $p(e)$.
-Recall that is the set of all utterance, meaning pairs $<u, m> = e \in E$.
+In general, these expectations are each taken over the set of possible communicative events $e\in E$ weighted by their probability, $p(e)$.
+Recall this is the set of all utterance, meaning pairs $<u, m> = e \in E$.
 
@@
 
-We assume that the particular joint distribution over $u, m$ pairs follows from a simple generative model.
+We assume that the particular joint distribution over utterance-meaning $<u, m> = e$ pairs follows from a simple generative model.
 First, some meaning is sampled with probability $p(m)$.
-Our speaker attempts to convey this intended meaning to a speaker by encoding it in the utterance $u$ via the conditional $S(u|m; \ell)$.
-Combining these terms leads to the *speaker's joint distribution over events* which we can write as $P_{Speaker}(u, m; \ell) = S(u | m; \ell)p(m)$.
+Our speaker attempts to convey this intended meaning to a listener by encoding it in the utterance $u$ via the conditional $S(u|m; \ell)$.
+Combining these terms leads to the *speaker's joint distribution over events* which we can write as $P(e) = P_{speaker}(u, m; \ell) = S(u | m; \ell)p(m)$.
 
@@
 \end{equation}
-From a mathematical standpoint this objective is intuitive.
-An optimal linguistic system being used by a speaker and listener agent will induce a particular set of form-meaning mappings that are both *close* between speaker and listener agents and *peaky* in that they assign non-uniform probability mass to the valid form-meaning mappings.\par
+From an information-theoretic perspective this objective is intuitive.
+Cross-entropy gives us a measure of dissimilarity between two distributions -- informally, the the average length of communicating an event from one distribution with the code from another.
+In our current setting, the optimal linguistic system being used by a speaker and listener agent will induce a particular set of form-meaning mappings which brings the speaker and listener joint distributions over communicative events as *close* together as possible.\par
 
-In terms of Zipf’s initial example about an optimal speaker or listener language we might consider just the first or second term.
-Hypothetically, we might expect the optimal speaker language $\ell_{speaker}*$ to minimize:
+## Optimal speaker and listener languages in the reference game setting
+In terms of Zipf’s initial example about an optimal speaker or listener language we might consider just the first or second term of equation (1).
+That is, given our formulation we should expect optimal speaker languages $\ell_{speaker}*$ to minimize:
 
@@
 
-Likewise we might expect the optimal listener language $\ell_{listener}*$ to minimize:
+Likewise we might expect optimal listener languages $\ell_{listener}*$ to minimize:
 
 \begin{equation}
-\ell_{listener}* = argmin_{\ell\in L}\mathbb{E}_{P_{speaker}(u, m; \ell)}(L(m|u;\ell)
+\ell_{listener}* = argmin_{\ell\in L}\mathbb{E}_{P_{speaker}(u, m; \ell)}(L(m|u;\ell))
 \end{equation}
@@
 We now have a metric for exploring questions of language optimality in the reference game setting (the speaker-listener cross-entropy).
-In the following sections we examine the behavior of the objective with respect to the communicative function of ambiguity.
+In the following sections we examine the behavior of this objective with respect to the communicative function of ambiguity.
 
@@
 At a minimum, the listener task in everyday communication is marked by the frequent need to handle lexical ambiguity (words often have multiple meanings) as well as syntactic ambiguity (sentences often have multiple parses), in addition to many other instances of ambiguity (Wasow, Perfors & Beaver, 201X).
-Chomsky (2002) famously claimed that the presence of ambiguity in natural language provides evidence that language has not been optimized for communication, stating “If you want to make sure that we never misunderstand one another, for that purpose language is not well designed, because you have such properties as ambiguity.” Such an analysis is proximal to the Zipfian optimal Listener language, in which the listener is never required to disambiguate a speaker’s utterance.\par
+Chomsky (2002) famously claimed that the presence of ambiguity in natural language provides evidence that language has not been optimized for communication, stating “If you want to make sure that we never misunderstand one another, for that purpose language is not well designed, because you have such properties as ambiguity.” Note that this analysis is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$ we described earlier and define in terms of equation (9).\par
 
-Piantadosi et al, (2012) essentially argue the opposite.
+In their paper, “The communicative function of ambiguity,” Piantadosi et al.
+(2012) essentially argue the opposite.
 They claim that ambiguity is an *efficient* property of any communication system in which *communication is contextualized*.
-We provide the heart of their argument here, however we refer to Piantados et al.
+We provide the heart of their argument here, however we refer to Piantadosi et al.
 (2012) for more detail:
@@
 
-In words, the Piantadosi argument might be summarised as -- it is useful to have a language that re-uses low-cost material (has ambiguity) so long as the cost of disambiguating the material is low (so long as the context is informative with respect to the disambiguation).
+In words, the Piantadosi argument might be summarised as -- it is useful to have a language that re-uses low-cost material (has ambiguity) so long as the cost of disambiguating the material is low.
+They argue that disambiguation cost is low when context if informative.
 
-In the following experiments we inspect two aspects of Piantadosi’s claim.
-In the first, we examine the efficient language *structure* aspect of the question exploring when the optimal linguistic system (under our cross-entropy objective) is most likely to contain ambiguous lexical items.
+In the following experiments we explore two aspects of Piantadosi’s claim.
+First, we examine the efficient language *structure* aspect of the question exploring when the optimal linguistic system (under our cross-entropy objective) is most likely to contain ambiguous lexical items.
 In the second experiment we explore an efficient language *use* aspect of the question -- at what point in a conversation is it useful for a speaker to use ambiguous lexical material?
@@
 \begin{equation}
+\begin{split}
 \sum_{c\in C}p(c)\sum_{u, m}S(u|m, c;\ell)p(m|c)-log[L(m|u,c;\ell)p(u)] = \\ \mathbb{E}_{c\sim P(c)[\mathbb{E}_{P_speaker(u, m, c; \ell)}[-log(P_listener(u, m, c; \ell))]}]
+\end{split}
 \end{equation}
@@
 
-Recall that language efficiency is both a function of the particular semantic mappings induced by that language, the Speaker and Listener agents, as well as the utterance and meaning priors.
-Therefore, we randomly generate priors of these values such that $p(m) \sim \text{Dir}(1, |M|)$ and $p(u) \sim \text{Dir}(1, |U|)$ where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.
+Recall that language efficiency is both a function of the particular semantic mappings induced by that language, the speaker and listener agents, as well as the utterance and meaning priors.
+Rather than assume particular structure for our utterance and meaning prior distributions, for each simulation we generate $p(m) \sim \text{Dir}(1, |M|)$ and $p(u) \sim \text{Dir}(1, |U|)$ where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.
 
 #### Context
-Given that we’d like to assess the impact of *context* on our objective we consider four conditions with $n=100$ each.
+Following the argument given by Piantadosi et al.
+(2012) we want to assess the impact of *context* on our objective.
+To do so we consider four conditions with $n=100$ simulations each.
 Our first is a *single-context* condition ( $|C|=1$) -- there is a only a single context describing $p(m|c_1)$.
@@
 Zipf (1935) proposed that a language optimized for speaker-ease should consist of a single word.
-We operationalize this using the first half of equation (1).
+We operationalize this using the *first half* of equation (1).
 
 \begin{equation}
-\text{Speaker-only objective} = \mathbb{E}_{u, m\sim P_{speaker}}[\text{speaker effort}] = \sum_{u, m}P_{speaker}(u, m; \ell)[-log(p(u))]
+\begin{split}
+\text{Speaker-only objective} = \mathbb{E}_{u, m\sim P_{speaker}}[\text{speaker effort}] \\= \sum_{u, m}P_{speaker}(u, m; \ell)[-log(p(u))]
+\end{split}
 \end{equation}
 
-Zipf also proposed that the optimal listener language should bijectively map all words to unique meanings. We operationalize this using the second half of equation (1).
+Zipf also proposed that the optimal listener language should bijectively map all words to unique meanings. We operationalize this using the *second half* of equation (1).
 \begin{equation}
-\text{Listener-only objective} = \mathbb{E}_{u, m\sim P_{speaker}}[\text{listener effort}] = \sum_{u, m}P_{speaker}(u, m; \ell)[-log(L(m|u; \ell)]
+\begin{split}
+\text{Listener-only objective} = \mathbb{E}_{u, m\sim P_{speaker}}[\text{listener effort}] \\ = \sum_{u, m}P_{speaker}(u, m; \ell)[-log(L(m|u; \ell)]
+\end{split}
 \end{equation}
@@
 plots the proportion of optimal languages under $H_{cross}(P_{Speaker}, P_{Listener})$ for each condition.
-We find that as the number of contexts increases, so does the probability that optimal language $\ell*$ contains ambiguity.
-For comparison we also include the optimal language under an objective that only considers speaker or listener effort.
+We find that as the number of contexts increases, so does the probability that the optimal language $\ell*$ contains ambiguity.
+For comparison we also include the optimal language under an objective that only considers speaker or listener effort (equations 11 and 12).
 In line, with Zipf's predictions, if languages are designed only to minimize speaker effort then optimal languages will assign all meanings to a single, low-cost utterance.
@@
 
+#### More context leads to larger efficiency gains
 Piantadosi et al.
-(2012) framed their theory in terms of conditional entropy.
-That is, $H(X|C) < H(X)$ when $C$ provides information about $X$.
-Put differently, when $I(X, C)$ is non-zero.
-In our setting this would indicate that as the amount of contextual information increases, the difference between the conditional and unconditional measures of information should increase.
-That is, $H_{cross}(P_{Speaker}, P_{Listener}|C_1) - H_{cross}(P_{Speaker}, P_{Listener}) < H_{cross}(P_{Speaker}, P_{Listener}|C_2) - H_{cross}(P_{Speaker}, P_{Listener})$ when $C_1$ contains more information than $C_2$.
-We can consider this very metric in our conditions.
-Figure 2 plots $H_{cross}(P_{Speaker}, P_{Listener}|C_1) - H_{cross}(P_{Speaker}, P_{Listener})$ in each of our conditions.
-As the number of contexts increases, the difference in efficiency when context is disambiguating increaess as well.\par
+(2012) framed their theory in terms of conditional entropy following the relation between conditional and unconditional entropy.
+That is, $H(X|C) < H(X)$ when $C$ provides information about $X$ (CITATION).
+In our setting, since knowing the particular context $c_i$ is always informative with respect the distribution over meanings ($p(m|c_i)$) this should mean that as the amount of contextual information increases, the difference between the conditional and unconditional measures of information should increase.
+In other words, $H_{cross}(P_{Speaker}, P_{Listener}|C_1) - H_{cross}(P_{Speaker}, P_{Listener}) = 0 < H_{cross}(P_{Speaker}, P_{Listener}|C_2) - H_{cross}(P_{Speaker}, P_{Listener})$ where $C_1 = |C| = 1$ and $C_2 = |C| = 2$.
+Figure 2 shows that as the number of contexts increases (the amount of information contained in the common ground, the difference in efficiency also increase.
 
@@
 (2012) argued that it is useful to re-use low-cost linguistic forms for multiple meanings when they can be disambiguated in context.
-Using our speaker-listener cross-entropy measure of efficiency we showed that optimal langauges are more likely to have ambiguous items when context is informative.
-Further, we showed that the impact of being able to disambiguate langauge (having access to $p(m|c_i)$) is increasingly efficienct as the amount of common-ground (context) increases.\par
+Using our speaker-listener cross-entropy measure of efficiency we showed that optimal languages are more likely to have ambiguous items when context is informative.
+Further, we showed that the impact of being able to disambiguate language (having access to $p(m|c_i)$) is increasingly efficient as the amount of common-ground (context) increases.\par
 
 ## Experiment 2: Rational-pragmatic speakers use ambiguity efficiently
-Our first experiment made a fairly direct test of the communicative function of ambiguity proposed by Piantadosi et al.
+
+Our first experiment made a fairly direct test of the communicative function of ambiguity proposed theory presented by Piantadosi et al.
 (2012).
-Sampling a set of need probabilities and utterance costs we explored the space of possible languages, examining the properties of the language $\ell*$ which minimized our objective.
+Sampling a set of need probabilities and utterance costs we explored the space of possible languages, examining the properties of the language $\ell^*$ which minimized our objective.
 Results indicated that ambiguity is an efficient property when context is informative.
-This experiment, however, assumed that our speaker and listener agents could disambiguate items context.
-That is, in our four conditions both agents had access to these conditional distributions $p(m|c_1), \dots, p(m|c_4)$.
-More often than not in day-to-day language use, speakers and listeners may not have perfect knowledge of the particular set of need probabilities $p(m)$ being used.
-Put differently, if ambiguity is only efficient when it can be disambiguated in context, speakers should avoid using ambiguous items if they know their listener may not be able to disambiguate the item.\par
+This experiment, however, assumed that our speaker and listener agents could always disambiguate items context.
+That is, in our four conditions both agents had perfect knowledge of the conditional distributions $p(m|c_1), \dots, p(m|c_4)$.
+Consider the scenario in which a reader is beginning a news article.
+While they may have some knowledge about the article’s topic (perhaps from the title), they may not have complete knowledge of its contents, including the persons or events involved.
+In this scenario, using an ambiguous pronoun early in the article lead to misunderstanding.
+Using such ambiguity is only warranted when the appropriate person or event is salient in the context.\par
 
-In experiment two we examine speaker behavior in a scenario in which the listener does not know the exact set of need probabilites a priori.
-That is, the listener has uncertainty over $p(m)$.
-Over the course of a discourse $D$ the listener tries to infer the context along with the particular intended meaning of a given utterance.
-That is, we consider a listener model $L(c, m|u;D)$ where $c$ is the particular context (or *topic*) being discussed and $D$ is the set of previous utterances the speaker has made.
-Likewise we consider a speaker model $S(u|m, c, D)$ who chooses an utterance based on an intended meaning $m$, the particular context of conversation $c$, while also considering the history of their utterance $D$.
-If speakers are behaving efficiently they should only use ambiguous items when they can be disambiguated in context.
-That is, if the listener agents is able to correctly infer the topic of conversation as the conversation progresses $D\rightarrow\inf$ then the speaker should be more likely to use ambiguous items later in discourse.
+We consider such a scenario in experiment two -- analyzing efficient *use* of ambiguity over discourse.
+In this scenario, we assume a single language $\ell$ which contains ambiguous items, known to both the speaker and listener.
+The topic of conversation $t\in T$ defines the particular meaning distribution at hand -- $p(m|T=t)$.
+This conditional distribution is known only to the speaker a priori.
+Over the course of a discourse, $D$ the listener tries to infer both the current topic as well as the meanings.
+That is we consider a listener model $L(m, t|u,D;\ell)$ and speaker model $S(u|m,t,D;\ell)$.
+Note that both the speaker and listener can track the history of previous utterances $D$.\par
 
-### Simulation set-up \newline
+### Hypothesis
+In this setting, if speakers are *using ambiguity efficiently* they should use the ambiguous material only when they know the listener can disambiguate the items in context.
+Since listeners accumulate information about the current topic distribution ($p(m|t)$) over as the discourse progresses, we should expect to see more frequent use of ambiguous items later in the discourse.
 
-$\ell = \text{current language known to both speaker and listener}$\par
-$D = \text{History of previous utterances by speaker}$\par
-$S(u|m; \ell) = \text{S1 rsa speaker}$\par
-$L(m|u; \ell) = \text{L0 rsa listener}$\par
-$P(u) = \text{utterance costs in which } p(u_1) < p(u_2) < p(u_3) < p(u_4)$\par
-$c = \text{current topic, which specifies need probabilities}$\par
-$P(m|c) = \text{set of need probabilites specified by } c$\par
+### Basic ingredients
+
+We conduct $n=50$ simulations, generating discourses of length $|D|=50$.  We consider a single language $\ell$ with $|U|=6$ and $|M|=4$ specified by the boolean matrix:
+
+\begin{center}
+\begin{tabular}{ c | c c c c} 
+& $m_1$ & $m_2$ & $m_3$ & $m_4$ \\
+\hline
+$u_1$ & 1 & 0 & 0 & 0 \\
+$u_2$ & 0 & 1 & 0 & 0 \\
+$u_3$ & 0 & 0 & 1 & 0 \\
+$u_4$ & 0 & 0 & 0 & 1 \\
+$u_5$ & 1 & 1 & 0 & 0 \\
+$u_6$ & 0 & 0 & 1 & 1 \\
+\end{tabular}
+\end{center}
+
+We assume that $p(u_5) = p(u_6) > p(u_{1,...,4})$.
+That is, the two ambiguous utterances are less costly than the non-ambiguous utterances.
+Over the discourse both agents have complete recall of the set of possible utterances by the speaker $D$.
+Note that the listener can attempt to infer the current topic distribution $p(C=c|D)$ via $p(C=c|D) \propto \prod_{i=0}^|D|S(u_i|m_i)p(m_i|c)$.
 
@@
 
+The left facet of figure 3 shows the empirical probability that the speaker used an ambiguous utterance as a function of discourse position.
+The right facet shows the listener belief about the current topic $c \in C$ given the discourse history ($p(c|D)$).
+Note that the trajectory of these items are close -- the speaker is more likely to use ambiguous material only when the listener is able to disambiguate that material.
+Figure 4 shows the listener belief about the current context.
+Additionally we can track the behavior of our cross-entropy objective and it’s component speaker and listener elements over the discourse.
+Overall we see declines in the speaker-listener cross-entropy.
+These are derived primarily from the listener learning about the current context of conversation (i.e.
+the current $c$ defining $p(m|c)$), however we also see declines in speaker effort, as they are able to use loss-costly material later in the discourse.
+
 ### Summary
+
+These demonstrated the efficient *use* of ambiguous material over a discourse, in line with the argument of Piantadosi (2012).
+While the speaker may prefer to use the less costly, ambiguous utterances at all points throughout the discourse, using the ambiguous items early leads to higher listener-level cost.
+Only once the context is known to the listener does it makes sense to use the lower-cost material.
+
+# General Discussion
+
+How might the competing pressures of speakers and listeners give rise to the distributional forms found in natural language?
+Zipf (1935) proposed that the asymmetry between speaker and listener costs gives rise to a range of properties the the level of the lexicon.
+Subsequent functionalist approaches to language science have provided additional evidence that various aspects of language structure and language use appear to be optimized for efficiency.
+But, it is not clear at what level this optimization occurs.
+In Zipf’s famous example of an optimal speaker and listener language he appears indicate that such optimization occurs in in-the-moment local interactions between speakers and listeners.
+We explore this idea in our current project, following theoretical work by Lawrence Horn (1984) who proposed the speaker-listener pressures identified by Zipf can also be interpreted in terms of the maxims governing rational conversation proposed by Grice (1975).
+To so, we explored the degree to which the interactions of rational pragmatic agents can provide a framework for understanding efficient language structure and use.
+We focused on an argument on the communicative function of ambiguity put forth by Piantadosi.
+We derive a novel speaker-listener cross-entropy objective for measuring the efficiency of a language in a reference game setting showing that optimal languages are more likely to contain ambiguous material when context is informative.
+Further we show that rational pragmatic agents will *use* ambiguous material efficiently, only when such use is supported by context.
+The intent is that this project provides a demonstration for the types of functionalist theories that can be explored within this framework.
+Among other language properties open to this type of analysis are work examining semantic typology (CITATIONS) and  the emergence of compositionality (CITATIONS), as well as efficient examples of language use including the efficient use of reduction and redundancy at the level  of syntax (Levy & Jaeger, 2007) and discourse (Genzel & Charniak, 2002).
 

### 2019-01-15 paper/paper.Rmd
updated plots

@@
 library(grid)
+library(gridExtra)
 library(ggplot2)
@@
 ### Speakers and listeners
-We adopt the Rational Speech-act framework (Frank & Goodman, 2012; Goodman & Frank, 2016) to operationalize our rational speaker-listeners.
-Under this setting, a *speaker agent* defines a conditional distribution over utterances, mapping from intended meanings $M$ to utterances $U$ using $\ell$.
+The Rational Speech-act framework (RSA) is computational-level theory of pragmatic language use.
+This framework has proven productive in describing a range of phenomena in natural language understanding and generation including hyperbole, metaphor, scalar and ad-hoc implicature, among others (CITATIONS).
+RSA can largely be understood as a formalization of essential Gricean pragmatic principles -- interlocutors reasoning about one another and their context.
+For this reason, we adopt RSA as our representational framework to operationalize rational speaker-listeners.
+For a more extensive description of RSA as well as its application we refer readers to Goodman & Frank (2016), limiting the current discussion to essential, high-level details of the framework.
+\par
+
+An RSA *speaker agent* defines a conditional distribution over utterances, mapping from intended meanings $M$ to utterances $U$ using $\ell$.
 That is, a speaker defines $P_{speaker}(u|m;\ell)$.
@@
 In general, we would like to consider the efficiency of a system $\ell$ in terms of these joint distributions.
-For a more detailed description of the RSA framework we refer the reader to Goodman & Frank (2016).
 
@@
 
+```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Left facet -- optimal languages are more likely to contain ambiguous items as amount of contextual information increases.
+Vertical axis shows the proportion of optimal languages that contain ambiguity.
+Horizontal axis shows the number of contexts in each condition (1-4).
+Red-line represents the optimal langauge under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only consideration.
+Right facet -- Gains in efficiency increase with the amount of contextual information (as the number of contexts increases).
+Vertical axis shows the difference in cross-entropy when the number of contexts is 1 and when the number of contexts is greater than 1.
+Horizontal axis shows each of the four conditions (the four context sizes).
+Note that the difference for condition 1 is zero as there is only a single context."}
+img1 <- grid::rasterGrob(as.raster(png::readPNG("figs/fig1.png")))
+img2 <- grid::rasterGrob(as.raster(png::readPNG("figs/fig2.png")))
+gridExtra::grid.arrange(grobs=list(img1, img2), ncol=2)
+```
+
 ### Simulation set-up
@@
 Further, we showed that the impact of being able to disambiguate language (having access to $p(m|c_i)$) is increasingly efficient as the amount of common-ground (context) increases.\par
+
+```{r plot-optimal-use, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Left facet shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position.
+As the discourse proceeds, the speaker is more likely to use an ambiguous term.
+Right face shows the probability the listener assigns to the current topic of conversation.
+That is, P(c|D) which is inferred by reasoning about the discourse history."}
+img3 <- grid::rasterGrob(as.raster(png::readPNG("figs/fig3.png")))
+img4 <- grid::rasterGrob(as.raster(png::readPNG("figs/fig4.png")))
+gridExtra::grid.arrange(grobs=list(img3, img4), ncol=2)
+```
 
@@
 
-```{r fig1, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=4, fig.height=4, fig.cap = "Optimal languages are more likely to contain ambiguous items as amount of contextual information increases.
-Vertical axis shows the proportion of optimal languages that contain ambiguity.
-Horizontal axis shows the number of contexts in each condition (1-4).
-Red-line represents the optimal langauge under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only consideration."}
-img <- png::readPNG("figs/fig1.png")
-grid::grid.raster(img)
-```
-
-```{r fig2, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2.5, fig.height=3, fig.cap = "The gains in efficinecy increase as the amount of contextual information increases.
-Vertical axis shows the difference in cross-entropy for when context is informative.
-Horizontal axis shows each of the four conditions.
-Note that the difference for condition 1 is zero as thered is only a single context." }
-img <- png::readPNG("figs/fig2.png")
-grid::grid.raster(img)
-```
-
-```{r fig3, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2.5, fig.height=3, fig.cap = "R plot" }
-img <- png::readPNG("figs/fig3.png")
-grid::grid.raster(img)
-```
-
-```{r fig4, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2.5, fig.height=3, fig.cap = "R plot" }
-img <- png::readPNG("figs/fig4.png")
-grid::grid.raster(img)
-```
-
-```{r fig5, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2.5, fig.height=3, fig.cap = "R plot" }
-img <- png::readPNG("figs/fig5.png")
-grid::grid.raster(img)
-```
-
 ```{r}

### 2019-01-17 paper/paper.Rmd
first changes from MF reading

@@
 abstract: 
-    We discuss a framework for studying the distributional properties of linguistic systems as emerging from in-the-moment interactions of speakers and listeners.
+    We discuss a framework for studying how the distributional properties of linguistic systems emerge from in-the-moment interactions of speakers and listeners.
 Our work takes Zipfian notions of lexicon-level efficiency as a starting point, connecting these ideas to Gricean notions of conversational-level efficiency.
 To do so, we begin by deriving an objective function for measuring the communicative efficiency of linguistic systems and then examining the behavior of this objective in a series of simulations focusing on the communicative function of ambiguity in language.
+These simulations suggest that rational pragmatic agents will produce communicatively efficient systems.
 
@@
 
-Why do natural languages look they way they do?
-While Zipf (1935) presented his “Principle of Least Effort” as a domain general framework for understanding human behavior, he took language as his central case-study.
-He proposed that distributional properties found in natural language were evidence of speaker-listener effort minimization.
+Why do languages look they way they do?
+While Zipf (1935) proposed that distributional properties found in natural language were evidence of speaker-listener effort minimization.
 In his own words, “we are arguing that people do in fact act with a maximum economy of effort, and that therefore in the process of speaking-listening they will automatically minimize the expenditure of effort.” Evidence for this claim was largely derived at the level of the lexicon.
@@
 (2012) argued that lexical ambiguity is an efficient property of any communication system when communication is contextualized.
-We will return to Piantadosi’s argument later in the paper.\par
 
@@
 Common to many is a characterization of a fundamental effort-asymmetry underlying everyday communication.
-Simply put, what is effortful as a speaker is likely different from what is effortful as a listener.
-Zipf characterized this as follows -- purely from the standpoint of speaker effort, what Zipf called "Speaker's Economy," an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of just a single, low-cost word.
+Simply put, what is effortful for a speaker is likely different from what is effortful for a listener.
+Zipf characterized this as follows -- purely from the standpoint of speaker effort, an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of just a single, low-cost word.
 Given such a language, the full set of potential meanings would be conveyed using only that word, i.e.
 $\ell_{speaker}^*$ would be fully ambiguous and all possible meanings would need to be disambiguated by a listener.
-From the standpoint of listener effort, what Zipf called "Auditor's Economy," an optimal language $\ell_{listener}^*$ would bijectively map all possible meanings to distinct words, eliminating a listener's need to disambiguate.
-Here the asymmetry is operationalized by different costs -- speaker effort is related to *production cost* and listener effort to *understanding or disambiguation cost*.
+From the standpoint of listener effort, an optimal language $\ell_{listener}^*$ should map all possible meanings to distinct words, removing a listener's need to disambiguate.
+In this example, speaker effort is related to *production cost* and listener effort to *understanding or disambiguation cost*.
 Clearly natural languages fall between the two extremes of $\ell_{speaker}^*$ and $\ell_{listener}^*$.
-Zipf proposed that the particular lexicon-level properties he observed emerged from the competition of these forces -- the pressure to jointly minimize speaker and listener effort.\par
+Zipf proposed that the particular lexicon-level properties he observed were a result of these competing forces -- the pressure to jointly minimize speaker and listener effort.\par
 
-This formulation begs the question -- at what level of analysis do local, in-the-moment interactions of speakers and listeners lead to optimized structure at the level of the lexicon?
-The example given by Zipf (1935) appears to describe local interaction in terms of a reference game.
+But at what level of does local, in-the-moment interaction lead to optimized structure at the level of the lexicon?
+The example given by Zipf (1935) appears to describe local interaction in terms of a \textit{reference game}.
 Speakers intend to refer to some object in the world $m$.
@@
 This move to study efficient language use in terms of reference games has clear theoretical underpinnings as well.
-A half century after Zipf, the linguist Lawrence Horn (1984), highlighted the importance of Zipf’s principles for explaining conversation-level phenomena.
-Horn suggested a direct link between Zipf’s Speaker and Listener economy and, what was at the time, more recent work on conversational pragmatics by Grice (1975).
-Horn highlighted that the interaction of Zipf’s forces were “largely responsible for generating Grice’s conversational maxims and the schema for pragmatic inference derived therefrom.” Put differently, system-level efficiency we see in languages is deeply related to local-level efficiency during the in-the-moment interactions of rational pragmatic speakers and listeners.
+The linguist Lawrence Horn (1984), highlighted the importance of Zipf’s principles for explaining conversation-level phenomena.
+Horn suggested a direct link between Zipf’s Speaker and Listener economy and work on conversational pragmatics by Grice (1975).
+Horn highlighted that the interaction of Zipf’s forces were “largely responsible for generating Grice’s conversational maxims and the schema for pragmatic inference derived therefrom.” Put differently, system-level efficiency we see in languages is deeply related to local-level efficiency during the in-the-moment interactions of rational pragmatic speakers and listeners at the level of conversation.
 \par
 
-In this work, we present a step toward formalizing Horn’s observation -- connecting Zipfian notions of efficient language structure to Gricean notions of rational conversation.
-To do so minimally requires three basic ingredients -- (1) a language property we’d like to explain (2) a framework for describing in-the-moment interactions of speaker-listeners and (3) some measure of linguistic efficiency.
+In this work, we present a step toward formalizing the framing presented above  -- connecting Zipfian notions of efficient language structure to Gricean notions of rational conversation.
+To do so minimally requires three basic ingredients -- (1) a language property we’d like to explain, (2) a framework for describing in-the-moment interactions of speaker-listeners, and (3) some measure of linguistic efficiency.
 For the current project we focus on the “communicative function of ambiguity” for our property -- translating the functionalist theory pursued by Piantadosi et al.
@@
 Finally, we consider a set of communicative events $e \in E$ where $<u, m> = e$ is a tuple of utterance, meanings pairs.
-The set $E$ all possible utterance-meaning mappings.
+The set, $E$, of all possible utterance-meaning mappings.
 
@@
 The Rational Speech-act framework (RSA) is computational-level theory of pragmatic language use.
-This framework has proven productive in describing a range of phenomena in natural language understanding and generation including hyperbole, metaphor, scalar and ad-hoc implicature, among others (CITATIONS).
 RSA can largely be understood as a formalization of essential Gricean pragmatic principles -- interlocutors reasoning about one another and their context.
@@
 
-## Optimal speaker and listener languages in the reference game setting
-In terms of Zipf’s initial example about an optimal speaker or listener language we might consider just the first or second term of equation (1).
-That is, given our formulation we should expect optimal speaker languages $\ell_{speaker}*$ to minimize:
-
-\begin{equation}
-\ell_{speaker}* = argmin_{\ell\in L}\mathbb{E}_{P_{speaker}(u, m; \ell)}(p(u))
-\end{equation}
-
-Likewise we might expect optimal listener languages $\ell_{listener}*$ to minimize:
-
-\begin{equation}
-\ell_{listener}* = argmin_{\ell\in L}\mathbb{E}_{P_{speaker}(u, m; \ell)}(L(m|u;\ell))
-\end{equation}
-
-# Two case-studies on efficient language use and design
-We now have a metric for exploring questions of language optimality in the reference game setting (the speaker-listener cross-entropy).
-In the following sections we examine the behavior of this objective with respect to the communicative function of ambiguity.
+# Simulations
 
@@
 They claim that ambiguity is an *efficient* property of any communication system in which *communication is contextualized*.
-We provide the heart of their argument here, however we refer to Piantadosi et al.
-(2012) for more detail:
-
-*In an unambiguous linguistic system, $m_1$ is mapped to $\ell_1 \in L$ and $m_2$ is mapped to $\ell_2 \in L$, with $\ell_1 \neq \ell_2$.
-Suppose that $\ell_1$ is easier than $\ell_2$...
-If $\ell_1$ is easier than $\ell_2$, and $m_1$ and $m_2$ are well-disambiguated by context, then we can always create a linguistic system which is easier overall by mapping $m_1$ and $m_2$ both to $\ell_1$.
-This costs the same in terms of effort every time we communicate $m_1$, but saves effort every time we communicate $m_2$*\par
-
-In words, the Piantadosi argument might be summarised as -- it is useful to have a language that re-uses low-cost material (has ambiguity) so long as the cost of disambiguating the material is low.
+Simply put, it is useful to have a language that re-uses low-cost material (has ambiguity) so long as the cost of disambiguating the material is low.
 They argue that disambiguation cost is low when context if informative.
@@
 
-## Experiment 1:  Optimal languages contain ambiguity when context is informative
+## Simulation 1:  Optimal languages contain ambiguity when context is informative
 
@@
 In our case we consider a context $c$ to specify a re-ordering of the referent need probabilities.
-That is, $p(m|c_1) \neq p(m|c_2)$.
-In the two context case.
+That is, $p(m|c_1) \neq p(m|c_2)$, when there are two contexts $|C| = \{c_1, c_2\}$.
 
@@
 \begin{split}
-\sum_{c\in C}p(c)\sum_{u, m}S(u|m, c;\ell)p(m|c)-log[L(m|u,c;\ell)p(u)] = \\ \mathbb{E}_{c\sim P(c)[\mathbb{E}_{P_speaker(u, m, c; \ell)}[-log(P_listener(u, m, c; \ell))]}]
+\sum_{c\in C}p(c)\sum_{u, m}S(u|m, c;\ell)p(m|c)-log[L(m|u,c;\ell)p(u)] = \\ \mathbb{E}_{c\sim P(c)}[\mathbb{E}_{P_{speaker}(u, m, c; \ell)}[-log(P_{listener}(u, m, c; \ell))]]
 \end{split}
@@
 Note that the difference for condition 1 is zero as there is only a single context."}
-img1 <- grid::rasterGrob(as.raster(png::readPNG("figs/fig1.png")))
-img2 <- grid::rasterGrob(as.raster(png::readPNG("figs/fig2.png")))
-gridExtra::grid.arrange(grobs=list(img1, img2), ncol=2)
+
+img <- png::readPNG("figs/fig1.png")
+grid::grid.raster(img)
 ```
@@
 For comparison, we also examine properties of optimal languages under two  additional objectives.
-Zipf (1935) proposed that a language optimized for speaker-ease should consist of a single word.
+Zipf (1935) proposed that the optimal speaker language should consist of a single low cost word $\ell_{speaker}^*$.
 We operationalize this using the *first half* of equation (1).
@@
 \begin{equation}
-\begin{split}
-\text{Speaker-only objective} = \mathbb{E}_{u, m\sim P_{speaker}}[\text{speaker effort}] \\= \sum_{u, m}P_{speaker}(u, m; \ell)[-log(p(u))]
-\end{split}
+\ell_{speaker}^* = argmin_{\ell\in L}\mathbb{E}_{P_{speaker}(u, m; \ell)}(p(u))
 \end{equation}
 
-Zipf also proposed that the optimal listener language should bijectively map all words to unique meanings. We operationalize this using the *second half* of equation (1).
+The optimal listener language $\ell_{listener}^*$, by contrast, should bijectively map all words to unique meanings. We operationalize this using the *second half* of equation (1).
+
 \begin{equation}
-\begin{split}
-\text{Listener-only objective} = \mathbb{E}_{u, m\sim P_{speaker}}[\text{listener effort}] \\ = \sum_{u, m}P_{speaker}(u, m; \ell)[-log(L(m|u; \ell)]
-\end{split}
+\ell_{listener}^* = argmin_{\ell\in L}\mathbb{E}_{P_{speaker}(u, m; \ell)}(L(m|u;\ell))
 \end{equation}
@@
 
-## Experiment 2: Rational-pragmatic speakers use ambiguity efficiently
+## Simulation 2: Rational-pragmatic speakers use ambiguity efficiently
 
@@
 (2012).
-Sampling a set of need probabilities and utterance costs we explored the space of possible languages, examining the properties of the language $\ell^*$ which minimized our objective.
+Sampling a set of need probabilities and utterance costs, we explored the space of possible languages, examining the properties of the language $\ell^*$ which minimized our objective.
 Results indicated that ambiguity is an efficient property when context is informative.
-This experiment, however, assumed that our speaker and listener agents could always disambiguate items context.
+This simulation, however, assumed that our speaker and listener agents could always disambiguate items context.
 That is, in our four conditions both agents had perfect knowledge of the conditional distributions $p(m|c_1), \dots, p(m|c_4)$.
@@
 
-We consider such a scenario in experiment two -- analyzing efficient *use* of ambiguity over discourse.
+We consider such a scenario in simulation two -- analyzing efficient *use* of ambiguity over discourse.
 In this scenario, we assume a single language $\ell$ which contains ambiguous items, known to both the speaker and listener.
-The topic of conversation $t\in T$ defines the particular meaning distribution at hand -- $p(m|T=t)$.
-This conditional distribution is known only to the speaker a priori.
-Over the course of a discourse, $D$ the listener tries to infer both the current topic as well as the meanings.
-That is we consider a listener model $L(m, t|u,D;\ell)$ and speaker model $S(u|m,t,D;\ell)$.
+Crucially, we do not assume that a listener has access the conditional distributions $p(m|c)$.
+Rather, we assume that the listener has knowledge of the set of possible context $C = \{c_1, \dots, c_k\}$, but does not know which one is currently being used by the speaker.
+In this setting, one can thing of the “contexts” as analogous to “topics” of conversation, where a “topic” specifies a particular seed of need probabilities $p(m|c)$.\par
+
+Over the course of a discourse, $D$ the listener tries to infer both the current topic, $c$, as well as the particular meaning of a given utterance.
+That is we consider a listener model $L(m, c|u,D;\ell)$ and speaker model $S(u|m,c,D;\ell)$.
 Note that both the speaker and listener can track the history of previous utterances $D$.\par
@@
 ### Hypothesis
-In this setting, if speakers are *using ambiguity efficiently* they should use the ambiguous material only when they know the listener can disambiguate the items in context.
+In this setting, if speakers are *using ambiguity efficiently* they should only use the ambiguous material when they know the listener can disambiguate the items in context.
 Since listeners accumulate information about the current topic distribution ($p(m|t)$) over as the discourse progresses, we should expect to see more frequent use of ambiguous items later in the discourse.
@@
 
-We conduct $n=50$ simulations, generating discourses of length $|D|=50$.  We consider a single language $\ell$ with $|U|=6$ and $|M|=4$ specified by the boolean matrix:
+We conduct $n=200$ simulations, generating discourses of length $|D|=30$ utterances.  We consider a single language $\ell$ with $|U|=6$ and $|M|=4$ specified by the boolean matrix:
 
@@
 That is, the two ambiguous utterances are less costly than the non-ambiguous utterances.
-Over the discourse both agents have complete recall of the set of possible utterances by the speaker $D$.
-Note that the listener can attempt to infer the current topic distribution $p(C=c|D)$ via $p(C=c|D) \propto \prod_{i=0}^|D|S(u_i|m_i)p(m_i|c)$.
+In this setting, it is less costly for the speaker to use the ambiguous utterances.
+However, if the listener cannot disambiguate them, then they will incur a disambiguation cose.
+We assume that over the discourse both agents have complete recall of the set of possible utterances by the speaker $D$.
+This means that the listener can infer the current topic distribution $p(C=c|D)$ via $p(C=c|D) \propto \prod_{i=0}^|D|S(u_i|m_i)p(m_i|c)$.
+In this way, an efficient speaker should should avoid ambiguity \textit{early} in the discourse when the listener is unsure of the context, only using it later once they are confident the listener knows the context $p(m|C=c)$.
 
@@
 
-The left facet of figure 3 shows the empirical probability that the speaker used an ambiguous utterance as a function of discourse position.
+The left facet of figure 3 shows the empirical probability that an RSA speaker used an ambiguous utterance as a function of discourse position.
 The right facet shows the listener belief about the current topic $c \in C$ given the discourse history ($p(c|D)$).
@@
 Figure 4 shows the listener belief about the current context.
-Additionally we can track the behavior of our cross-entropy objective and it’s component speaker and listener elements over the discourse.
+Additionally, we can track the behavior of our cross-entropy objective and it’s component speaker and listener elements over the discourse.
 Overall we see declines in the speaker-listener cross-entropy.
@@
 
-These demonstrated the efficient *use* of ambiguous material over a discourse, in line with the argument of Piantadosi (2012).
-While the speaker may prefer to use the less costly, ambiguous utterances at all points throughout the discourse, using the ambiguous items early leads to higher listener-level cost.
-Only once the context is known to the listener does it makes sense to use the lower-cost material.
+These simulations demonstrated the efficient *use* of ambiguous material over a discourse, in an alternative instantiation of the argument outlined by Piantadosi et al.
+(2012) -- Only once the context is known to the listener does it makes sense to use the lower-cost material.
 

### 2019-01-17 ambiguity/simulations.py
add logging, big code clean-up, add non-uniform contex priors (priors generated from dirichlet)

@@ -2,12 +2,15 @@
 
 from collections import defaultdict
 import pandas as pd
+import logging
 import numpy as np
 import tqdm
 
 from .matrix import *
 from .objectives import *
 from .utils import *
+
+logging.getLogger().setLevel(logging.INFO)
 
 
 class Simulation:
@@ -60,6 +63,8 @@ class OrderedSimulation(Simulation):
             meaning_order_fn_name,
             speaker_k=1,
             listener_k=1,
+            speaker_alpha=1,
+            listener_alpha=1,
             verbose=False):
         all_results = []
         pbar = tqdm.tqdm(total=n_sims * verbose)
@@ -74,6 +79,9 @@ class OrderedSimulation(Simulation):
             N_MEANINGS = 4
             p_utterances = np.random.dirichlet([utterance_order_fn(i) for i in range(N_UTTERANCES)])
             p_meanings = np.random.dirichlet([meaning_order_fn(i) for i in range(N_MEANINGS)])
+            p_cs = np.random.dirichlet(
+                [1. for _ in range(n_contexts)])
+            p_cs = p_cs.reshape((n_contexts, 1))
             # Populate contexts
             contexts = [p_meanings]
 
@@ -83,8 +91,10 @@ class OrderedSimulation(Simulation):
                 # new_p_m = np.random.dirichlet([meaning_order_fn(k) for k in range(N_MEANINGS)])
                 contexts.append(np.roll(contexts[0], i+1))
             # Baseline model (context isn't informative so we take)
-            # p(m) = \sum_c p(m, c)p(c)
+            # p(m) = \sum_c p(m|c)p(c)
             p_m_no_context = np.array(contexts)
+            p_m_no_context *= p_cs
+
             p_m_no_context = \
                 np.sum(p_m_no_context, axis=0) / np.sum(p_m_no_context)
 
@@ -96,84 +106,93 @@ class OrderedSimulation(Simulation):
             for idxs, m in all_matrices:
                 for context_id, context in enumerate(contexts):
                     p_meanings_ = context
-                    # Pragmatic speaker / listener
-                    m_speaker = col_multiply(speaker(m, p_utterances, p_meanings_, 1., speaker_k), p_meanings_)
-                    m_listener = row_multiply(listener(m, p_utterances, p_meanings_, 1., listener_k), p_utterances)
-                    # non-Pragmatic speaker / listener
[…]

### 2019-01-17 ambiguity/config.py
change sims to 500

@@ -5,17 +5,19 @@ from simulations import VaryContextSimulations, RecursiveSimulation
 CONTEXT_RUN_PARAMS = {
     'simulator': VaryContextSimulations,
     'run_params': {
-        'n_sims': 100,
+        'n_sims': 500,
         'utterance_order_fn': lambda x: 1.,
         'utterance_order_fn_name': 'uniform',
         'meaning_order_fn': lambda x: 1.,
         'meaning_order_fn_name': 'uniform',
+        'speaker_alpha': 1.,
+        'listener_alpha': 1.,
         'speaker_k': 1,
         'listener_k': 1,
         'verbose': False
     },
     # utterance_order_fn_name-meaning_order_fn_name-speaker_k-listener_k
-    'out_file_name': 'uniform_uniform_1_1_100.csv'
+    'out_file_name': 'uniform_uniform_1_1_500.csv'
 }
 
 

### 2019-01-19 ambiguity/simulations.py
minor clean-up

@@ -86,11 +86,12 @@ class OrderedSimulation(Simulation):
             contexts = [p_meanings]
 
             for i in range(1, n_contexts):
-                # new_p_m populates a totally new prior over meanigns.
                 # Currently not using...
+                # new_p_m populates a totally new prior over meanigns.
                 # new_p_m = np.random.dirichlet([meaning_order_fn(k) for k in range(N_MEANINGS)])
                 contexts.append(np.roll(contexts[0], i+1))
-            # Baseline model (context isn't informative so we take)
+            # Baseline model
+            # (context isn't informative so we take)
             # p(m) = \sum_c p(m|c)p(c)
             p_m_no_context = np.array(contexts)
             p_m_no_context *= p_cs
@@ -113,32 +114,23 @@ class OrderedSimulation(Simulation):
                                                      speaker_alpha,
                                                      speaker_k),
                                              p_meanings_)
+
                     # P_{listener}(u, m | c)
                     m_listener = row_multiply(listener(m, p_utterances,
                                                        p_meanings_,
                                                        listener_alpha,
                                                        listener_k),
                                               p_utterances)
-                    # p(u)
+                    # p(u) term
                     # in \sum_{u, m}P_s(u, m)*log(p(u))
                     m_simple_speaker_costs = row_multiply(m, p_utterances)
-                    # L(m|u)
+
+                    # L(m|u) term
                     # in \sum_{u, m}P_s(u, m)*log(L(m|u))
                     m_simple_listener_costs = listener(m, p_utterances,
                                                        p_meanings_,
                                                        listener_alpha,
                                                        listener_k)
-
-                    # Note we assume uniform prior over contexts
-                    # thus (1 / len(n_contexts))
-                    # p_c = (1. / n_contexts)
-
-                    # TODO (20190116) Do I need to do this normalization below?
-                    # Normalise to ensure valid probability distributions.
-                    # (handle languages that have assigned
-                    # 0 to particular utterance.)
-
-
 
                     # Safe-keeping for clean-use below
                     args = [('ce', m_speaker, m_listener, ce),
@@ -149,8 +141,6 @@ class OrderedSimulation(Simulation):
                              m_simple_speaker_costs, ce),
                             ('listener_entropy', m_speaker,
                              m_simple_listener_costs, ce)]
-                            # ('base_speaker_listener_ce', m_speaker_no_context,
-                            #  m_listener_no_context, ce)]
 
                     # Record objectives
                     for arg in args:

### 2019-01-19 paper/paper.Rmd
updates to figures and add reference game.

@@
 ---
-title: "The interactions of rational pragmatic agents provide a framework for understanding efficient language structure and use"
+title: "Pragmatic interactions lead to efficient language structure and use"
 bibliography: library.bib
@@
 author-information: > 
-    \author{{\large \bf Author 1} \\ \texttt{bpeloqui@stanford.edu} \\ Department of Psychology \\ Stanford University
-    \And {\large \bf Author 2} \\ \texttt{author1@university.edu} \\ Department of Psychology \\ Some University}
+    \author{{\large \bf Benjamin N. Peloquin} \\ \texttt{bpeloqui@stanford.edu} \\ Department of Psychology \\ Stanford University
+    \And {\large \bf Noah D. Goodman} \\ \texttt{ngoodman@stanford.edu} \\ Department of Computer Science \\ Stanford University
+    \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@university.edu} \\ Department of Psychology \\ Stanford University}
 
@@
 Why do languages look they way they do?
-While Zipf (1935) proposed that distributional properties found in natural language were evidence of speaker-listener effort minimization.
+Zipf (1935) proposed that distributional properties found in natural language were evidence of speaker-listener effort minimization.
 In his own words, “we are arguing that people do in fact act with a maximum economy of effort, and that therefore in the process of speaking-listening they will automatically minimize the expenditure of effort.” Evidence for this claim was largely derived at the level of the lexicon.
-Zipf argued that the particular relationship between a word’s frequency and its rank, length, and denotation size could be explained as an emergent property of speaker-listener effort minimization.
+Zipf argued that the particular relationship between a word’s frequency and its rank, length, and denotation size could be explained as an emergent properties of speaker-listener effort minimization.
 \par
@@
 Common to many is a characterization of a fundamental effort-asymmetry underlying everyday communication.
-Simply put, what is effortful for a speaker is likely different from what is effortful for a listener.
+Simply put, what is “hard” for a speaker is likely different from what is “hard” for a listener.
 Zipf characterized this as follows -- purely from the standpoint of speaker effort, an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of just a single, low-cost word.
@@
 The linguist Lawrence Horn (1984), highlighted the importance of Zipf’s principles for explaining conversation-level phenomena.
-Horn suggested a direct link between Zipf’s Speaker and Listener economy and work on conversational pragmatics by Grice (1975).
-Horn highlighted that the interaction of Zipf’s forces were “largely responsible for generating Grice’s conversational maxims and the schema for pragmatic inference derived therefrom.” Put differently, system-level efficiency we see in languages is deeply related to local-level efficiency during the in-the-moment interactions of rational pragmatic speakers and listeners at the level of conversation.
+Horn suggested a link between Zipf’s conception of speaker- and listener-effort and ideas related to conversational pragmatics by Grice (1975), arguing that the interaction of Zipf’s forces were “largely responsible for generating Grice’s conversational maxims and the schema for pragmatic inference derived therefrom.” Put differently, the system-level efficiency we see in languages is fundamentally related to local-level efficiency during the in-the-moment interactions of rational, pragmatic speakers and listeners at the level of conversation.
 \par
@@
 Finally, (3) we derive an objective function for measuring language efficiency in the reference game setting.
-Given this framework we can ask the question, “how does having \textit{lexical ambiguity} make using language $\ell$ more or less useful for communication”?
-This project is meant to demonstrate the ways in which a range of functionalist theories might make use of such a framework -- in theory replacing \textit{lexical ambiguity} with a range of behaviors and properties identified by functionalist approaches to language science.\par
+In terms of our opening question, we are exploring the question, “how does having \textit{lexical ambiguity} make using language $\ell$ more or less useful for communication”?
+In more general terms, this project is meant to demonstrate the ways in which a range of functionalist theories might make use of such a framework -- e.g.
+we could replace \textit{lexical ambiguity} with a range of behaviors and properties identified by functionalist approaches to language science.\par
 
@@
 Following this introduction we derive a simple objective function for measuring the efficiency of linguistic systems in this setting.
-Subsequently, we move on to two case-studies examining the questions posed above, framing results in terms of our efficiency measure.
+We move on to two case-studies examining the questions posed above, examining the efficiency of ambiguity from a language design and language use perspective.
 
 # Exploring efficient language- design and use in rational pragmatic agents
+
+```{r plot-reference-game, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2.5, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.cap = "An example reference game with associated literal semantics (in our terminology 'language')."}
+img <- png::readPNG("figs/game1.png")
+grid::grid.raster(img)
+```
 
@@
 Note that the prior over meanings are analogous to the *need probabilities* assumed in previous work (Regier, CITATION).
-We consider a set of contexts $C$ which describe different need probability distributions over our set of meanings $p(m|c)$.
-Finally, we consider a set of communicative events $e \in E$ where $<u, m> = e$ is a tuple of utterance, meanings pairs.
-The set, $E$, of all possible utterance-meaning mappings.
+We consider a set of contexts $C$ which describe different need probability states, formalized as different conditional distributions over meanings $p(m|c)$.
+Finally, we consider a set of communicative events $e \in E$ where $<u, m> = e$ is a tuple of utterance, referent pairs.
 
@@
 The Rational Speech-act framework (RSA) is computational-level theory of pragmatic language use.
-RSA can largely be understood as a formalization of essential Gricean pragmatic principles -- interlocutors reasoning about one another and their context.
-For this reason, we adopt RSA as our representational framework to operationalize rational speaker-listeners.
-For a more extensive description of RSA as well as its application we refer readers to Goodman & Frank (2016), limiting the current discussion to essential, high-level details of the framework.
+RSA can largely be understood as a formalization of essential Gricean pragmatic principles -- interlocutors reason about one another and their shared context.
+For this reason, we adopt RSA as our representational framework to operationalize Gricean (rational and pragmatic) speaker-listeners.
+For a more extensive description of RSA as well as its applications we refer readers to Goodman & Frank (2016), limiting the current discussion to essential, high-level details of the framework.
 \par
@@
 \begin{split}
-  = \sum_{u, m}p_{speaker}(u, m; \ell)[-log_2(p(u))] + \\ 
-   \sum_{u, m}p_{speaker}(u, m; \ell)[-log_2(L(m|u; \ell))]
+  = \sum_{u, m}P_{speaker}(u, m; \ell)[-log_2(p(u))] + \\ 
+   \sum_{u, m}P_{speaker}(u, m; \ell)[-log_2(L(m|u; \ell))]
 \end{split}
@@
 From an information-theoretic perspective this objective is intuitive.
-Cross-entropy gives us a measure of dissimilarity between two distributions -- informally, the the average length of communicating an event from one distribution with the code from another.
-In our current setting, the optimal linguistic system being used by a speaker and listener agent will induce a particular set of form-meaning mappings which brings the speaker and listener joint distributions over communicative events as *close* together as possible.\par
+Cross-entropy gives us a measure of dissimilarity between two distributions -- the average number of bits required to communicate under one distribution, given that the “true” distribution differs.
+In our case, this is the difference between the joint distribution assumed by the speaker $P_{speaker}$ and listener $P_{listener}$.
+A good language $\ell$ used by a set of a pair of speaker-listeners will have properties which minimize this objective.\par
 
@@
 Ambiguity is ubiquitous in natural language.
-At a minimum, the listener task in everyday communication is marked by the frequent need to handle lexical ambiguity (words often have multiple meanings) as well as syntactic ambiguity (sentences often have multiple parses), in addition to many other instances of ambiguity (Wasow, Perfors & Beaver, 201X).
+At a minimum, the listener task in everyday communication is marked by the frequent need to handle lexical ambiguity (words often have multiple meanings) as well as syntactic ambiguity (sentences often have multiple parses), in addition to other instances of ambiguity (Wasow, Perfors & Beaver, 201X).
 Chomsky (2002) famously claimed that the presence of ambiguity in natural language provides evidence that language has not been optimized for communication, stating “If you want to make sure that we never misunderstand one another, for that purpose language is not well designed, because you have such properties as ambiguity.” Note that this analysis is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$ we described earlier and define in terms of equation (9).\par
 
-In their paper, “The communicative function of ambiguity,” Piantadosi et al.
-(2012) essentially argue the opposite.
+Piantadosi et al.
+(2012) argue just the opposite.
 They claim that ambiguity is an *efficient* property of any communication system in which *communication is contextualized*.
 Simply put, it is useful to have a language that re-uses low-cost material (has ambiguity) so long as the cost of disambiguating the material is low.
-They argue that disambiguation cost is low when context if informative.
+Context can provide useful information for disambiguation.\par
 
 In the following experiments we explore two aspects of Piantadosi’s claim.
-First, we examine the efficient language *structure* aspect of the question exploring when the optimal linguistic system (under our cross-entropy objective) is most likely to contain ambiguous lexical items.
+First, we examine the efficient language *structure* aspect of their claim, exploring when the optimal linguistic system $\ell^*$ (under our cross-entropy objective) is most likely to contain ambiguous lexical items.
 In the second experiment we explore an efficient language *use* aspect of the question -- at what point in a conversation is it useful for a speaker to use ambiguous lexical material?
@@
 The argument outlined by Piantadosi et al.
-(2012) is clearly compatible with our current reference game setting with just a single addition -- context.
-The authors argue that mapping both $m_1$ and $m_2$ to $l_1$ is useful when they can disambiguated *in context*.
-In our case we consider a context $c$ to specify a re-ordering of the referent need probabilities.
+(2012) is compatible with our current reference game setting with just a single addition -- context.
+The authors argue that mapping both $m_1$ and $m_2$ to $l_1$ is useful when they can be disambiguated *in context*.
+In our case we consider a context $c$ to specify a unique ordering of the need probabilities.
 That is, $p(m|c_1) \neq p(m|c_2)$, when there are two contexts $|C| = \{c_1, c_2\}$.
@@
 Note that in the case that $|C| =1$, our objective simplifies to our original equation (7).
-With this update we can compare linguistics systems $L$ while varying the degree to which context contains useful information by varying the size of $|C|$.\par
-
-```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Left facet -- optimal languages are more likely to contain ambiguous items as amount of contextual information increases.
-Vertical axis shows the proportion of optimal languages that contain ambiguity.
-Horizontal axis shows the number of contexts in each condition (1-4).
-Red-line represents the optimal langauge under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only consideration.
-Right facet -- Gains in efficiency increase with the amount of contextual information (as the number of contexts increases).
-Vertical axis shows the difference in cross-entropy when the number of contexts is 1 and when the number of contexts is greater than 1.
-Horizontal axis shows each of the four conditions (the four context sizes).
-Note that the difference for condition 1 is zero as there is only a single context."}
-
-img <- png::readPNG("figs/fig1.png")
-grid::grid.raster(img)
-```
+With this update we can compare languages $L$ while varying the degree to which context contains useful information, varying the number of contexts (e.g.
+the size of $|C|$).\par
 
@@
 #### Basic ingredients
-We conduct $N=400$ simulations.
+We conduct $N=2000$ simulations.
 For each simulation we enumerate the set of *valid* languages in which $|U|=|M|=4$.
@@
 
-Recall that language efficiency is both a function of the particular semantic mappings induced by that language, the speaker and listener agents, as well as the utterance and meaning priors.
-Rather than assume particular structure for our utterance and meaning prior distributions, for each simulation we generate $p(m) \sim \text{Dir}(1, |M|)$ and $p(u) \sim \text{Dir}(1, |U|)$ where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.
+Recall that language efficiency is both a function of the particular semantic mappings induced by that language, the speaker and listener agents, as well as the utterance ($P(U)$), meaning ($P(M)$), and context priors ($P(C)$).
+Rather than assume particular structure for our utterance and meaning prior distributions, for each simulation we generate $P(M) \sim \text{Dir}(1, |M|)$, $P(U) \sim \text{Dir}(1, |U|)$ and $P(C) \sim \text{Dir}(1, |C|)$ where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.
 
@@
 (2012) we want to assess the impact of *context* on our objective.
-To do so we consider four conditions with $n=100$ simulations each.
+To do so we consider four conditions with $n=500$ simulations each (that is, 500 unique sets of $\{P(U), P(M), P(C)\}$.
 Our first is a *single-context* condition ( $|C|=1$) -- there is a only a single context describing $p(m|c_1)$.
@@
 #### Hypotheses
-This leads to the following set of hypotheses: (1) The probability that an optimal language contains ambiguity under our Cross-Entropy objective should increase with the number of contexts (as $|C|\rightarrow\inf$).
-(2) The optimal speaker language under our Speaker-only objective should always map all meanings to the single, lowest cost utterance.
+Piantadosi’s claim leads to the following set of hypotheses: (1) The probability that an optimal language contains ambiguity under our Cross-Entropy objective should increase with the number of contexts (as $|C|\rightarrow\inf$).
+(2) The optimal speaker language under our “Speaker-only” objective should always map all meanings to the single, lowest cost utterance.
 (3) The optimal listener language should never contain ambiguous material.
@@
 Figure 1.
-plots the proportion of optimal languages under $H_{cross}(P_{Speaker}, P_{Listener})$ for each condition.
-We find that as the number of contexts increases, so does the probability that the optimal language $\ell*$ contains ambiguity.
-For comparison we also include the optimal language under an objective that only considers speaker or listener effort (equations 11 and 12).
-In line, with Zipf's predictions, if languages are designed only to minimize speaker effort then optimal languages will assign all meanings to a single, low-cost utterance.
-Likewise, if languages are designed only to minimize speaker listener effort then ambiguity should always be avoided.\par
+plots the proportion of optimal languages under $H_{cross}(P_{Speaker}, P_{Listener}; \ell)$ for each context condition.
+The red line indicates that as the number of contexts increases, so does the probability that the optimal language $\ell*$ under our speaker-listener cross-entropy objective contains ambiguity.
+We also plot the proportion of optimal speaker-only and listener-only languages that contain ambiguity.
+In line, with Zipf's predictions, if languages are designed only to minimize speaker effort then optimal languages will assign all meanings to a single utterance (full ambiguous).
+Likewise, if languages are designed only to minimize speaker listener effort then ambiguity should always be avoided (bijective form-meaning mappings).\par
 
@@
 Piantadosi et al.
-(2012) framed their theory in terms of conditional entropy following the relation between conditional and unconditional entropy.
-That is, $H(X|C) < H(X)$ when $C$ provides information about $X$ (CITATION).
-In our setting, since knowing the particular context $c_i$ is always informative with respect the distribution over meanings ($p(m|c_i)$) this should mean that as the amount of contextual information increases, the difference between the conditional and unconditional measures of information should increase.
-In other words, $H_{cross}(P_{Speaker}, P_{Listener}|C_1) - H_{cross}(P_{Speaker}, P_{Listener}) = 0 < H_{cross}(P_{Speaker}, P_{Listener}|C_2) - H_{cross}(P_{Speaker}, P_{Listener})$ where $C_1 = |C| = 1$ and $C_2 = |C| = 2$.
-Figure 2 shows that as the number of contexts increases (the amount of information contained in the common ground, the difference in efficiency also increase.
+(2012) framed their theory in terms of properties of conditional entropy.
+That is,  $H(X|C) \leq H(X)$.
+This inequality is strict when $C$ provides information about $X$ (Cover & Thomas, XXXX).
+Intuitively, this means that they amount of uncertainty we have about $X$ decreases when $C$ tells us something about $X$.
+In our setting, knowing the particular context $c_i$ is always informative with respect the distribution over meanings ($p(m|c_i)$).
+This means that if we have $|C_1|=1$ and $|C_2|=2$ then we expect $H_{cross}(P_{Speaker}, P_{Listener}|C_2) < H_{cross}(P_{Speaker}, P_{Listener}|C_1)$.
+To examine this in our simulations, we look at the difference in efficiency scores, comparing pragmatic agents who have access the conditional distributions $P(m|c)$ an non-pragmatic agents who don’t and simply consider $P(m) =
+
+Additionally, the difference between these two quantities should increase with the difference in the size of $|C_i|$.
+E.g.
+if we also have $|C_3|=3$ then  .
+In our case, we can examine the difference between the unconditional and conditional cross-entropy measures as $|C|$ increases from $|C|=1$ in condition 1 to $|C|=4$ in condition 4.
+The right facet of figure one shows the difference between conditional and unconditional corr-entropy measures as the number of contexts $|C|$ increases.
+
+```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=8, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Left facet -- optimal languages are more likely to contain ambiguous items as amount of contextual information increases.
+Vertical axis shows the proportion of optimal languages that contain ambiguity.
+Horizontal axis shows the number of contexts in each condition (1-4).
+Red-line represents the optimal langauge under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only consideration.
+Right facet -- Gains in efficiency increase with the amount of contextual information (as the number of contexts increases).
+Vertical axis shows the difference in cross-entropy when the number of contexts is 1 and when the number of contexts is greater than 1.
+Horizontal axis shows each of the four conditions (the four context sizes).
+Note that the difference for condition 1 is zero as there is only a single context."}
+
+img <- png::readPNG("figs/fig1.png")
+grid::grid.raster(img)
+```
 
@@
 Further, we showed that the impact of being able to disambiguate language (having access to $p(m|c_i)$) is increasingly efficient as the amount of common-ground (context) increases.\par
-
-```{r plot-optimal-use, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Left facet shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position.
-As the discourse proceeds, the speaker is more likely to use an ambiguous term.
-Right face shows the probability the listener assigns to the current topic of conversation.
-That is, P(c|D) which is inferred by reasoning about the discourse history."}
-img3 <- grid::rasterGrob(as.raster(png::readPNG("figs/fig3.png")))
-img4 <- grid::rasterGrob(as.raster(png::readPNG("figs/fig4.png")))
-gridExtra::grid.arrange(grobs=list(img3, img4), ncol=2)
-```
 
@@
 This simulation, however, assumed that our speaker and listener agents could always disambiguate items context.
-That is, in our four conditions both agents had perfect knowledge of the conditional distributions $p(m|c_1), \dots, p(m|c_4)$.
-Consider the scenario in which a reader is beginning a news article.
+That is, in our four conditions both agents had perfect knowledge of the conditional distributions $p(m|c_i)$.\par
+However, in day to day communication we often don’t have perfect knowledge of the current context of a conversation.
+We can consider this notion of “context” as analogous to knowledge of the *topic* of conversation.
+Consider the scenario in which a reader is beginning a newspaper article.
 While they may have some knowledge about the article’s topic (perhaps from the title), they may not have complete knowledge of its contents, including the persons or events involved.
-In this scenario, using an ambiguous pronoun early in the article lead to misunderstanding.
+In this scenario, using an ambiguous pronoun early in the article lead to misunderstanding as there isn’t sufficient context to disambiguate it.
 Using such ambiguity is only warranted when the appropriate person or event is salient in the context.\par
@@
 In this scenario, we assume a single language $\ell$ which contains ambiguous items, known to both the speaker and listener.
-Crucially, we do not assume that a listener has access the conditional distributions $p(m|c)$.
-Rather, we assume that the listener has knowledge of the set of possible context $C = \{c_1, \dots, c_k\}$, but does not know which one is currently being used by the speaker.
-In this setting, one can thing of the “contexts” as analogous to “topics” of conversation, where a “topic” specifies a particular seed of need probabilities $p(m|c)$.\par
+Crucially, we do not assume that a listener has access the conditional distributions $p(m|c)$ a priori.
+Rather, we assume that the listener has knowledge of the set of possible contexts $C = \{c_1, \dots, c_k\}$, but does not know which one is currently being used by the speaker.\par
 
@@
 Note that both the speaker and listener can track the history of previous utterances $D$.\par
+
+```{r plot-optimal-use, fig.env = "figure*", fig.pos = "h", fig.width=8, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Left facet (A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position.
+As the discourse proceeds, the speaker is more likely to use an ambiguous term.
+Right face (B) shows the probability the listener assigns to the current topic of conversation."}
+img <- png::readPNG("figs/fig2.png")
+grid::grid.raster(img)
+```
 
@@
 We assume that over the discourse both agents have complete recall of the set of possible utterances by the speaker $D$.
-This means that the listener can infer the current topic distribution $p(C=c|D)$ via $p(C=c|D) \propto \prod_{i=0}^|D|S(u_i|m_i)p(m_i|c)$.
-In this way, an efficient speaker should should avoid ambiguity \textit{early} in the discourse when the listener is unsure of the context, only using it later once they are confident the listener knows the context $p(m|C=c)$.
+This means that the listener can infer the current topic distribution $p(C=c|D)$ via $p(C=c|D) \propto \prod_{i=0}^{|D|}S(u_i|m_i)p(m_i|c)$.
+In this way, an efficient speaker should avoid ambiguity \textit{early} in the discourse when the listener is unsure of the context, only using it later once they are confident the listener knows the context $p(m|C=c)$.
 
@@
 
-The left facet of figure 3 shows the empirical probability that an RSA speaker used an ambiguous utterance as a function of discourse position.
-The right facet shows the listener belief about the current topic $c \in C$ given the discourse history ($p(c|D)$).
+Figure 2, plot A shows the empirical probability that an RSA speaker used an ambiguous utterance as a function of discourse position.
+While at the start of the discourse the speaker is unlikely to use the utterance we see consistent use later, once the listener is sure of the particular context $c_i$ this speaker is using.\par
+
+The right facet of figure 2 shows the listener belief about the current topic $c \in C$ given the discourse history ($p(c|D)$).
 Note that the trajectory of these items are close -- the speaker is more likely to use ambiguous material only when the listener is able to disambiguate that material.
@@
 These are derived primarily from the listener learning about the current context of conversation (i.e.
-the current $c$ defining $p(m|c)$), however we also see declines in speaker effort, as they are able to use loss-costly material later in the discourse.
+the current $c$ defining $p(m|c)$), however we also see declines in speaker effort, as they are able to use loss-costly material later in the discourse.\par
 
@@
 How might the competing pressures of speakers and listeners give rise to the distributional forms found in natural language?
-Zipf (1935) proposed that the asymmetry between speaker and listener costs gives rise to a range of properties the the level of the lexicon.
+Zipf (1935) proposed that the asymmetry between speaker and listener costs gives rise to a range of properties at the level of the lexicon.
 Subsequent functionalist approaches to language science have provided additional evidence that various aspects of language structure and language use appear to be optimized for efficiency.
@@
 
-# Conclusion
-
-# Acknowledgements
-
-Place acknowledgments (including funding information) in a section at
-the end of the paper.
-
 # References

### 2019-01-21 paper/paper.Rmd
minor clean-up improvments and clarifications, removed conditional efficiency plots / section for space and narrative flow.

@@
 In his own words, “we are arguing that people do in fact act with a maximum economy of effort, and that therefore in the process of speaking-listening they will automatically minimize the expenditure of effort.” Evidence for this claim was largely derived at the level of the lexicon.
-Zipf argued that the particular relationship between a word’s frequency and its rank, length, and denotation size could be explained as an emergent properties of speaker-listener effort minimization.
+Zipf argued that the particular relationship between a word’s frequency and its rank, length, and denotation size could be explained as an emergent property of speaker-listener effort minimization.
 \par
@@
 $\ell_{speaker}^*$ would be fully ambiguous and all possible meanings would need to be disambiguated by a listener.
-From the standpoint of listener effort, an optimal language $\ell_{listener}^*$ should map all possible meanings to distinct words, removing a listener's need to disambiguate.
+From the standpoint of listener effort, an optimal language $\ell_{listener}^*$ would map all possible meanings to distinct words, removing a listener's need to disambiguate.
 In this example, speaker effort is related to *production cost* and listener effort to *understanding or disambiguation cost*.
@@
 The linguist Lawrence Horn (1984), highlighted the importance of Zipf’s principles for explaining conversation-level phenomena.
-Horn suggested a link between Zipf’s conception of speaker- and listener-effort and ideas related to conversational pragmatics by Grice (1975), arguing that the interaction of Zipf’s forces were “largely responsible for generating Grice’s conversational maxims and the schema for pragmatic inference derived therefrom.” Put differently, the system-level efficiency we see in languages is fundamentally related to local-level efficiency during the in-the-moment interactions of rational, pragmatic speakers and listeners at the level of conversation.
-\par
+Horn suggested a link between Zipf’s conception of speaker- and listener-effort and ideas related to conversational pragmatics by Grice (1975), arguing that the interaction of Zipf’s forces were “largely responsible for generating Grice’s conversational maxims and the schema for pragmatic inference derived therefrom.” Put differently, the system-level efficiencies we see in language are fundamentally related to the local-level efficiencies that occur during rational, pragmatic conversation.\par
 
@@
 To model local, conversational interactions (2) we adopt the Rational Speech-act framework (RSA) (Frank & Goodman, 2012; Goodman & Frank, 2016).
-Finally, (3) we derive an objective function for measuring language efficiency in the reference game setting.
-In terms of our opening question, we are exploring the question, “how does having \textit{lexical ambiguity} make using language $\ell$ more or less useful for communication”?
-In more general terms, this project is meant to demonstrate the ways in which a range of functionalist theories might make use of such a framework -- e.g.
-we could replace \textit{lexical ambiguity} with a range of behaviors and properties identified by functionalist approaches to language science.\par
+Finally, (3) we derive an objective function for measuring language efficiency in the reference game setting.\par
 
+The current work examines the question, “how does having \textit{lexical ambiguity} make using language $\ell$ more or less useful for communication”?
 We begin with a high-level introduction to the modelling framework introducing the basic ingredients we will need to represent language as repeated reference games.
@@
 
-```{r plot-reference-game, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2.5, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.cap = "An example reference game with associated literal semantics (in our terminology 'language')."}
+```{r plot-reference-game, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2.5, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.cap = "An example reference game with associated literal semantics (in our terminology a ‘language’)."}
 img <- png::readPNG("figs/game1.png")
@@
 We adopt that formulation here, assuming these communication games as our basic unit of analysis.
-In this setting, speakers and listeners are aware of a set of objects $M$, which will refer to as *meanings* and are knowledgeable about the set of possible signals $U$ (*utterances*) that could be used to refer to a given meaning.
-Utterances may have different relative costs, operationalized via a prior over utterances $p(u)$.
-Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $p(m)$.
+In this setting, speakers and listeners are aware of a set of objects $M$, which we will refer to as *meanings* and are knowledgeable about the set of possible signals $U$ (*utterances*) that can be used to refer to a given meaning.
+Utterances may have different relative costs, operationalized via a prior over utterances $P(U)$.
+Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$.
 Note that the prior over meanings are analogous to the *need probabilities* assumed in previous work (Regier, CITATION).
-We consider a set of contexts $C$ which describe different need probability states, formalized as different conditional distributions over meanings $p(m|c)$.
+We consider a set of contexts $C$ with an associated prior $P(C)$.
+Importantly, contexts describe different need probability states, formalized as different conditional distributions over meanings e.g.
+$p(M|C=c_i)$.
 Finally, we consider a set of communicative events $e \in E$ where $<u, m> = e$ is a tuple of utterance, referent pairs.
@@
 First, some meaning is sampled with probability $p(m)$.
-Our speaker attempts to convey this intended meaning to a listener by encoding it in the utterance $u$ via the conditional $S(u|m; \ell)$.
-Combining these terms leads to the *speaker's joint distribution over events* which we can write as $P(e) = P_{speaker}(u, m; \ell) = S(u | m; \ell)p(m)$.
+Our speaker attempts to convey this intended meaning to a listener via an utterance $u$ by sampling from the speaker conditional distribution $S(u|m; \ell)$.
+Combining these terms leads to the *speaker's joint distribution over events* which we can write as $P(e) = S(u | m; \ell)p(m) = P_{speaker}(u, m; \ell)$.
 
@@
 Simply put, it is useful to have a language that re-uses low-cost material (has ambiguity) so long as the cost of disambiguating the material is low.
-Context can provide useful information for disambiguation.\par
+In particular, context (or common ground) can provide useful information for disambiguation.
+Say we have two objects in the world $m_1$ and $m_2$ and two languages $\ell_1$ and $\ell_2$.
+In language $\ell_1$ the low-cost utterance $u_1$ can be used to refer to both $m1$ and $m_2$ (i.e.
+$[[u_1]]_{\ell_1} = \{m_1, m_2 \}$.
+Language $\ell_2$, by contrast, defines a mapping such that the low cost utterance $[[u_1]]_{\ell_2} = \{m_1\}$ and a higher cost utterance $[[u_2]]_{\ell_2} = \{m_2 \}$.
+While it is cheaper for a speaker to use $\ell_1$ (because she can always use a low-cost utterance), it is more difficult for a listener (because $u_1$ is ambiguous).
+However, if the meaning $u_1$ can be disambiguated *in context* then the listener will not incur disambiguation cost and the speaker can always use the low-cost utterance.
+This a basic argument for the contextual efficiency of polysemy.
+If I use the word “bat” at a baseball game a listener will likely infer I am referring to the baseball sense of the word.
+If I use the word “bat” in a dark cave, a listener will likely infer I am referring to the animal.
 
@@
 (2012) is compatible with our current reference game setting with just a single addition -- context.
-The authors argue that mapping both $m_1$ and $m_2$ to $l_1$ is useful when they can be disambiguated *in context*.
+The authors argue that mapping two meanings $m_1$ and $m_2$ to a single utterance $u_1$ is useful when they can be disambiguated *in context*.
 In our case we consider a context $c$ to specify a unique ordering of the need probabilities.
@@
 Note that in the case that $|C| =1$, our objective simplifies to our original equation (7).
-With this update we can compare languages $L$ while varying the degree to which context contains useful information, varying the number of contexts (e.g.
-the size of $|C|$).\par
+With this update we can compare languages $L$ while varying the degree to which context contains useful information, varying the number of contexts (the size of $|C|$).\par
 
@@
 
-Recall that language efficiency is both a function of the particular semantic mappings induced by that language, the speaker and listener agents, as well as the utterance ($P(U)$), meaning ($P(M)$), and context priors ($P(C)$).
-Rather than assume particular structure for our utterance and meaning prior distributions, for each simulation we generate $P(M) \sim \text{Dir}(1, |M|)$, $P(U) \sim \text{Dir}(1, |U|)$ and $P(C) \sim \text{Dir}(1, |C|)$ where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.
+Recall that language efficiency is both a function of the particular semantic mappings induced by that language, the speaker and listener agents ($S$ and $L$), as well as the utterance ($P(U)$), meaning ($P(M)$), and context priors ($P(C)$).
+Rather than assume particular structure for our utterance, meaning and context prior distributions, for each simulation we generate $P(U) \sim \text{Dir}(1, |U|)$, $P(M) \sim \text{Dir}(1, |M|)$, and $P(C) \sim \text{Dir}(1, |C|)$ where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.
 
@@
 To do so we consider four conditions with $n=500$ simulations each (that is, 500 unique sets of $\{P(U), P(M), P(C)\}$.
-Our first is a *single-context* condition ( $|C|=1$) -- there is a only a single context describing $p(m|c_1)$.
-Our second condition contains two-contexts  ($|C| = 2$) -- we consider efficiency under both $p(m|c_1)$ as well as $p(m|c_2)$.
+Our first is a *single-context* condition ( $|C|=1$) -- there is a only a single context describing $P(M)$.
+Our second condition contains two-contexts  ($|C| = 2$) -- we consider efficiency under both $P(M|C=c_1)$ as well as $P(M|C=c_2)$.
 The third and fourth condition correspond accordingly with $|C|=3$ and $|C| = 4$, respectively.
@@
 For comparison, we also examine properties of optimal languages under two additional objectives.
-Zipf (1935) proposed that the optimal speaker language should consist of a single low cost word $\ell_{speaker}^*$.
+Zipf (1935) proposed that the optimal speaker language $\ell_{speaker}^*$ should only optimize speaker effort.
 We operationalize this using the *first half* of equation (1).
@@
 
-The optimal listener language $\ell_{listener}^*$, by contrast, should bijectively map all words to unique meanings. We operationalize this using the *second half* of equation (1).
+The optimal listener language $\ell_{listener}^*$, by contrast, should only optimize listener effort. We operationalize this using the *second half* of equation (1).
 
@@
 plots the proportion of optimal languages under $H_{cross}(P_{Speaker}, P_{Listener}; \ell)$ for each context condition.
-The red line indicates that as the number of contexts increases, so does the probability that the optimal language $\ell*$ under our speaker-listener cross-entropy objective contains ambiguity.
-We also plot the proportion of optimal speaker-only and listener-only languages that contain ambiguity.
+The red line indicates that as the number of contexts increases, so does the probability that the optimal language $\ell^*_{cross}$ under our speaker-listener cross-entropy objective contains ambiguity.
+We also plot the proportion of optimal speaker-only $\ell^*_{speaker}$ and listener-only $\ell^*_{listener}$ languages that contain ambiguity.
 In line, with Zipf's predictions, if languages are designed only to minimize speaker effort then optimal languages will assign all meanings to a single utterance (full ambiguous).
@@
 
-#### More context leads to larger efficiency gains
-Piantadosi et al.
-(2012) framed their theory in terms of properties of conditional entropy.
-That is,  $H(X|C) \leq H(X)$.
-This inequality is strict when $C$ provides information about $X$ (Cover & Thomas, XXXX).
-Intuitively, this means that they amount of uncertainty we have about $X$ decreases when $C$ tells us something about $X$.
-In our setting, knowing the particular context $c_i$ is always informative with respect the distribution over meanings ($p(m|c_i)$).
-This means that if we have $|C_1|=1$ and $|C_2|=2$ then we expect $H_{cross}(P_{Speaker}, P_{Listener}|C_2) < H_{cross}(P_{Speaker}, P_{Listener}|C_1)$.
-To examine this in our simulations, we look at the difference in efficiency scores, comparing pragmatic agents who have access the conditional distributions $P(m|c)$ an non-pragmatic agents who don’t and simply consider $P(m) =
-
-Additionally, the difference between these two quantities should increase with the difference in the size of $|C_i|$.
-E.g.
-if we also have $|C_3|=3$ then  .
-In our case, we can examine the difference between the unconditional and conditional cross-entropy measures as $|C|$ increases from $|C|=1$ in condition 1 to $|C|=4$ in condition 4.
-The right facet of figure one shows the difference between conditional and unconditional corr-entropy measures as the number of contexts $|C|$ increases.
-
-```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=8, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Left facet -- optimal languages are more likely to contain ambiguous items as amount of contextual information increases.
+```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=8, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Optimal languages are more likely to contain ambiguous items as amount of contextual information increases.
 Vertical axis shows the proportion of optimal languages that contain ambiguity.
 Horizontal axis shows the number of contexts in each condition (1-4).
-Red-line represents the optimal langauge under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only consideration.
-Right facet -- Gains in efficiency increase with the amount of contextual information (as the number of contexts increases).
-Vertical axis shows the difference in cross-entropy when the number of contexts is 1 and when the number of contexts is greater than 1.
-Horizontal axis shows each of the four conditions (the four context sizes).
-Note that the difference for condition 1 is zero as there is only a single context."}
+Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only consideration."}
 
@@
 (2012) argued that it is useful to re-use low-cost linguistic forms for multiple meanings when they can be disambiguated in context.
-Using our speaker-listener cross-entropy measure of efficiency we showed that optimal languages are more likely to have ambiguous items when context is informative.
-Further, we showed that the impact of being able to disambiguate language (having access to $p(m|c_i)$) is increasingly efficient as the amount of common-ground (context) increases.\par
+We simulated this effect in a reference game setting, examining the probability that an optimal language will contain an word with an ambiguous mapping as a function of context size.
+Using our speaker-listener cross-entropy measure of efficiency we showed that optimal languages are more likely to have ambiguous items when context is informative, but not under objectives which emphasizes speaker- or listener-effort alone.\par
 
@@
 
-Our first experiment made a fairly direct test of the communicative function of ambiguity proposed theory presented by Piantadosi et al.
+Our first experiment made a fairly direct test of the communicative function of ambiguity theory presented by Piantadosi et al.
 (2012).
@@
 This simulation, however, assumed that our speaker and listener agents could always disambiguate items context.
-That is, in our four conditions both agents had perfect knowledge of the conditional distributions $p(m|c_i)$.\par
-However, in day to day communication we often don’t have perfect knowledge of the current context of a conversation.
-We can consider this notion of “context” as analogous to knowledge of the *topic* of conversation.
+That is, in our four conditions both agents had perfect knowledge of the conditional distributions $P(M|C_i)$.\par
+
+This is a strong assumption and likely unrealistic for much of day-to-day communication.
+In talking with another person, we often \textit{don’t have perfect knowledge of the current context of a conversation}.
+We can consider this notion of “context” as analogous to a *topic* of conversation.
 Consider the scenario in which a reader is beginning a newspaper article.
 While they may have some knowledge about the article’s topic (perhaps from the title), they may not have complete knowledge of its contents, including the persons or events involved.
-In this scenario, using an ambiguous pronoun early in the article lead to misunderstanding as there isn’t sufficient context to disambiguate it.
-Using such ambiguity is only warranted when the appropriate person or event is salient in the context.\par
+In this scenario, using an ambiguous pronoun early in the article lead to misunderstanding as there isn’t sufficient context for disambiguation.
+Using such ambiguity is only warranted when the appropriate person or event is salient in the context.
+\par
 
 We consider such a scenario in simulation two -- analyzing efficient *use* of ambiguity over discourse.
-In this scenario, we assume a single language $\ell$ which contains ambiguous items, known to both the speaker and listener.
-Crucially, we do not assume that a listener has access the conditional distributions $p(m|c)$ a priori.
-Rather, we assume that the listener has knowledge of the set of possible contexts $C = \{c_1, \dots, c_k\}$, but does not know which one is currently being used by the speaker.\par
+In this scenario, we assume a single language $\ell$ which contains both ambiguous and unambiguous utterances.
+The ambiguous utterances are less costly than the unambiguous utterances.
+Crucially, we do not assume that a listener has access the conditional distributions $p(M|C=c)$ a priori.
+Rather, we assume that the listener has knowledge of the set of possible contexts $C = \{c_1, \dots, c_k\}$, but \textit{does not know which one is currently being used by the speaker}.
+In this setting, a speaker minimizing her production effort would like to use the low-cost material as frequently as possible.
+However, this will incur a disambiguation cost to the listener if they are uncertain of the current context $c_{current}$ (topic of conversation).\par
 
-Over the course of a discourse, $D$ the listener tries to infer both the current topic, $c$, as well as the particular meaning of a given utterance.
+Over the course of a discourse, $D$ the listener tries to infer both the current topic, $c_{current}$, as well as the particular meaning $m$ of a given utterance $u$.
 That is we consider a listener model $L(m, c|u,D;\ell)$ and speaker model $S(u|m,c,D;\ell)$.
-Note that both the speaker and listener can track the history of previous utterances $D$.\par
+Note that both the speaker and listener can track the history of previous utterances $D$.
+Importantly, a listener can attempt to infer the current topic of conversation $c_{current}$ using the discourse history $D$.\par
 
@@
 In this setting, if speakers are *using ambiguity efficiently* they should only use the ambiguous material when they know the listener can disambiguate the items in context.
-Since listeners accumulate information about the current topic distribution ($p(m|t)$) over as the discourse progresses, we should expect to see more frequent use of ambiguous items later in the discourse.
+Since listeners accumulate information about the current topic distribution ($P(C=c_{current}|D)$) over as the discourse progresses, we should expect to see more frequent use of ambiguous items later in the discourse.
 
@@
 
-We conduct $n=200$ simulations, generating discourses of length $|D|=30$ utterances.  We consider a single language $\ell$ with $|U|=6$ and $|M|=4$ specified by the boolean matrix:
+We conduct $N=200$ simulations, generating discourses of length $|D|=30$ utterances.  We consider a single language $\ell$ with $|U|=6$ and $|M|=4$ specified by the boolean matrix:
 
@@
 
-We assume that $p(u_5) = p(u_6) > p(u_{1,...,4})$.
+We assume that $p(u_5) = p(u_6) > p(u_{1})=\dots=p(u_{4})$.
 That is, the two ambiguous utterances are less costly than the non-ambiguous utterances.
@@
 We assume that over the discourse both agents have complete recall of the set of possible utterances by the speaker $D$.
-This means that the listener can infer the current topic distribution $p(C=c|D)$ via $p(C=c|D) \propto \prod_{i=0}^{|D|}S(u_i|m_i)p(m_i|c)$.
-In this way, an efficient speaker should avoid ambiguity \textit{early} in the discourse when the listener is unsure of the context, only using it later once they are confident the listener knows the context $p(m|C=c)$.
+This means that the listener can infer the current topic distribution $p(C=c|D)$ via $p(C=c|D) \propto \prod_{i=0}^{|D|}S(u_i|m_i)p(m_i|c)p(c)$.
+In this way, an efficient speaker should avoid ambiguity \textit{early} in the discourse when the listener is unsure of the context, only using it later once they are confident the listener knows the current context $p(C=c_{current}|D)$.
 
@@
 
-The right facet of figure 2 shows the listener belief about the current topic $c \in C$ given the discourse history ($p(c|D)$).
-Note that the trajectory of these items are close -- the speaker is more likely to use ambiguous material only when the listener is able to disambiguate that material.
-Figure 4 shows the listener belief about the current context.
-Additionally, we can track the behavior of our cross-entropy objective and it’s component speaker and listener elements over the discourse.
-Overall we see declines in the speaker-listener cross-entropy.
-These are derived primarily from the listener learning about the current context of conversation (i.e.
-the current $c$ defining $p(m|c)$), however we also see declines in speaker effort, as they are able to use loss-costly material later in the discourse.\par
+The right facet of figure 2 shows the listener belief about the current topic $c_{current} \in C$ given the discourse history ($p(C=c_{current}|D)$).
+Note that the trajectories in plot A and B are related -- the speaker is more likely to use ambiguous material only when the listener is able to disambiguate that material.
 
@@
 
-These simulations demonstrated the efficient *use* of ambiguous material over a discourse, in an alternative instantiation of the argument outlined by Piantadosi et al.
-(2012) -- Only once the context is known to the listener does it makes sense to use the lower-cost material.
+These simulations explored the efficient *use* of ambiguous material over a discourse, in an alternative instantiation of the argument outlined by Piantadosi et al.
+(2012).
+In the first set of simulations we varied the set of languages, examining the properties of languages which minimized our cross-entropy objective under different context conditions.
+In the current simulation we assumed that the listener did not know the current context (toptic) a priori, but had to infer it from the discourse.
+We examined speaker production behavior, tracking use of low-cost, but ambiguous material over the discourse.
+We found that early in the discourse our rational speaker avoided using ambiguous material, only using the material later when the context was known to the listener.
 
@@
 
-How might the competing pressures of speakers and listeners give rise to the distributional forms found in natural language?
+How do the competing pressures of speakers and listeners give rise to the distributional forms found in natural language?
 Zipf (1935) proposed that the asymmetry between speaker and listener costs gives rise to a range of properties at the level of the lexicon.
 Subsequent functionalist approaches to language science have provided additional evidence that various aspects of language structure and language use appear to be optimized for efficiency.
-But, it is not clear at what level this optimization occurs.
-In Zipf’s famous example of an optimal speaker and listener language he appears indicate that such optimization occurs in in-the-moment local interactions between speakers and listeners.
-We explore this idea in our current project, following theoretical work by Lawrence Horn (1984) who proposed the speaker-listener pressures identified by Zipf can also be interpreted in terms of the maxims governing rational conversation proposed by Grice (1975).
+In Zipf’s famous example of an optimal speaker and listener language he appears indicate that such optimization occurs in in-the-moment, local interactions between speakers and listeners.
+We explore this idea in our current project, following theoretical work by Horn (1984) who proposed the speaker-listener pressures identified by Zipf can also be interpreted in terms of the maxims governing rational conversation proposed by Grice (1975).
 To so, we explored the degree to which the interactions of rational pragmatic agents can provide a framework for understanding efficient language structure and use.
-We focused on an argument on the communicative function of ambiguity put forth by Piantadosi.
+We focused on an argument on the communicative function of ambiguity put forth by Piantadosi et al.
+(2012).
 We derive a novel speaker-listener cross-entropy objective for measuring the efficiency of a language in a reference game setting showing that optimal languages are more likely to contain ambiguous material when context is informative.
@@
 The intent is that this project provides a demonstration for the types of functionalist theories that can be explored within this framework.
-Among other language properties open to this type of analysis are work examining semantic typology (CITATIONS) and  the emergence of compositionality (CITATIONS), as well as efficient examples of language use including the efficient use of reduction and redundancy at the level  of syntax (Levy & Jaeger, 2007) and discourse (Genzel & Charniak, 2002).
+Among other language properties open to this type of analysis are work examining semantic typology (CITATIONS), the emergence of compositionality (CITATIONS) and convention formation, as well as efficient examples of language use including the efficient use of reduction and redundancy (Levy & Jaeger, 2007) and discourse (Genzel & Charniak, 2002).
 
 # References 
+Chomsky (2002)
+Cover & Thomas, XXXX)
+Ferrer-i-cancho, et al.,  201X
+Frank & Goodman (2012)
+Genzel & Charniak (2002)
+Goodman & Frank (2016)
+Grice (1957)
+Horn (1984)
+Kibry, Smith & Folds (201X)
+Levy & Jaeger (2007)
+Piantadosi et al. (2012)
+Regier, et al.
+Wasow, Perfors & Beaver (chomsky201X)
+Zipf (1935)
 

### 2019-01-22 ambiguity/config.py
general clean-up

@@ -16,6 +16,7 @@ CONTEXT_RUN_PARAMS = {
         'listener_k': 1,
         'verbose': False
     },
+    # fp pattern below
     # utterance_order_fn_name-meaning_order_fn_name-speaker_k-listener_k
     'out_file_name': 'uniform_uniform_1_1_500.csv'
 }

### 2019-01-22 ambiguity/language_constructs.py

@@ -1,6 +1,5 @@
 """language_constructs.py"""
 
-import itertools as it
 from .utils import *
 
 

### 2019-01-22 ambiguity/objectives.py

@@ -13,11 +13,6 @@ class Objective:
     def __init__(self):
         pass
 
-    @staticmethod
-    def _check_matrix(m):
-        return True
-        #return np.abs(np.sum(m) - 1) < 1e-3
-
     def compute_cost(self, m1, m2):
         raise NotImplementedError()
 

### 2019-01-22 ambiguity/utils.py
add matrix generation upto and utterances filter

@@ -28,6 +28,31 @@ def all_meanings_available_filter(M):
     return all(np.sum(M, axis=0) > 0.)
 
 
+def all_utterances_used_filter(M):
+    """
+
+    Parameters
+    ----------
+    M: np.array
+        A boolean matrix.
+
+    Returns
+    -------
+    bool
+        True if all meanings can be talked about (each col as some val.)
+
+    """
+    return all(np.sum(M, axis=1) > 0.)
+
+
+def all_utterances_meanings_used_filter(M):
+    """Combined version .
+
+    """
+    return all_meanings_available_filter(M) and all_utterances_used_filter(M)
+
+
+
 def is_valid_matrix(M, filter_fn):
     """Filter on boolean matrices.
 
@@ -74,6 +99,14 @@ def generate_all_boolean_matrices(m, n, n_true):
     for idxs in all_subsets:
         matrices.append((idxs, idxs2matrix(idxs, m, n)))
     return matrices
+
+
+def generate_all_boolean_matrices_upto(m, n, n_true):
+    """Generate all M x N boolean matrices from n to n_true ones."""
+    langs = []
+    for curr_n_true in range(n, n_true + 1):
+        langs.extend(generate_all_boolean_matrices(m, n, curr_n_true))
+    return langs
 
 
 def matrix2idxs(M):

### 2019-01-22 ambiguity/simulations.py
generate all languages 'upto' and ENSURE THAT SPEAKER/LISTENER JOINTS ARE VALID JOINT DISTRIBUTIOSN YES :)

@@ -81,13 +81,16 @@ class OrderedSimulation(Simulation):
             p_meanings = np.random.dirichlet([meaning_order_fn(i) for i in range(N_MEANINGS)])
             p_cs = np.random.dirichlet(
                 [1. for _ in range(n_contexts)])
+            # Uniform prior over contexts (NOT USING)
+            # p_cs = np.array([1. for _ in range(n_contexts)])
+            # p_cs = p_cs / np.sum(p_cs)
             p_cs = p_cs.reshape((n_contexts, 1))
             # Populate contexts
             contexts = [p_meanings]
 
             for i in range(1, n_contexts):
                 # Currently not using...
-                # new_p_m populates a totally new prior over meanigns.
+                # new_p_m populates a totally new prior over meanings.
                 # new_p_m = np.random.dirichlet([meaning_order_fn(k) for k in range(N_MEANINGS)])
                 contexts.append(np.roll(contexts[0], i+1))
             # Baseline model
@@ -95,13 +98,13 @@ class OrderedSimulation(Simulation):
             # p(m) = \sum_c p(m|c)p(c)
             p_m_no_context = np.array(contexts)
             p_m_no_context *= p_cs
-
             p_m_no_context = \
                 np.sum(p_m_no_context, axis=0) / np.sum(p_m_no_context)
 
-            # All matrices
+            # Generate all valid matrices (as defined by filter)
             all_matrices = \
-                [(idxs, m) for (idxs, m) in generate_all_boolean_matrices(N_UTTERANCES, N_MEANINGS, N_MEANINGS) if all_meanings_available_filter(m)]
+                [(idxs, m) for (idxs, m) in generate_all_boolean_matrices_upto(N_UTTERANCES, N_MEANINGS, N_MEANINGS+1) \
+                 if all_utterances_meanings_used_filter(m)]
 
             d_results = defaultdict(dict)
             for idxs, m in all_matrices:
@@ -121,6 +124,7 @@ class OrderedSimulation(Simulation):
                                                        listener_alpha,
                                                        listener_k),
                                               p_utterances)
+
                     # p(u) term
                     # in \sum_{u, m}P_s(u, m)*log(p(u))
                     m_simple_speaker_costs = row_multiply(m, p_utterances)
@@ -131,6 +135,12 @@ class OrderedSimulation(Simulation):
                                                        p_meanings_,
                                                        listener_alpha,
                                                        listener_k)
+
+                    # Require m_speaker and m_listener are valid
[…]

### 2019-01-22 ambiguity/tests.py
add old tests, this should be updated

@@ -1,0 +1,62 @@
+"""tests.py"""
+
+import numpy as np
+import unittest
+
+from .language_constructs import Semantics, Language, RSAAgent, CrossEntropy
+from .matrix import rsa_speaker_forward, rsa_listener_forward, row_multiply, \
+    col_multiply, \
+    rsa_listener, rsa_speaker, col_normalize, row_normalize
+from .objectives import CrossEntropy as CE
+from .utils import idxs2matrix, matrix2dict
+
+
+def test_language2matrix():
+    pass
+
+
+def test_rsa_speaker():
+    N_UTTERANCES = 4
+    N_MEANINGS = 4
+    idxs = (0, 5, 10, 15)
+    m = idxs2matrix(idxs, N_UTTERANCES, N_MEANINGS)
+    assert rsa_speaker_forward(m, 1) == np.array(m)
+
+def test_matrix_formalism():
+
+    p_utterances = [0.1, 0.3, 0.6]
+    p_meanings = [0.1, 0.3, 0.6]
+    l0 = idxs2matrix([0, 3, 4, 8], 3, 3)
+
+    # Non-matrix formulation
+    utterances = ['a', 'b', 'c']
+    meanings = [1, 2, 3]
+    d = matrix2dict(l0, utterances, meanings)
+    sems = Semantics(d)
+    language = Language(utterances, meanings, p_utterances, p_meanings, sems)
+    agent = RSAAgent(language)
+    cost1 = CrossEntropy(agent, agent, language).compute_cost()
+
+    import pdb;
+    pdb.set_trace();
+
+    listener = rsa_listener(l0, p_meanings, p_utterances)
+    speaker = rsa_speaker(l0, p_meanings)
+    # Matrix formulation
+    m_l0 = row_multiply(l0, p_meanings)
+    m_listener = col_multiply(rsa_listener_forward(m_l0, 1), p_utterances)
+    m_speaker = row_multiply(rsa_speaker_forward(m_l0, 1), p_meanings)
+    ce_matrix_cost = CE()
+    cost2 = ce_matrix_cost.compute_cost(m_listener, m_speaker)
+
+
+
+
+
+
+if __name__ == '__main__':
+    test_matrix_formalism()
+
+
+
+

### 2019-01-22 ambiguity/agents.py
org changes and commenting

@@ -1,4 +1,4 @@
-"""matrix.py
+"""agents.py
 
 Define all experiments in terms of matrix operations.
 

### 2019-01-22 ambiguity/simulations.py

@@ -1,12 +1,9 @@
 """simulations.py"""
 
-from collections import defaultdict
 import pandas as pd
-import logging
-import numpy as np
 import tqdm
 
-from .matrix import *
+from .agents import *
 from .objectives import *
 from .utils import *
 
@@ -323,6 +320,8 @@ class VaryContextSimulations(Simulation):
         return d_results
 
 
+# Below is old code that needs updating and is not used in the
+# paper "Pragmatic interactions lead to efficient language structure and use"
 class RecursiveSimulation(Simulation):
     """Simulation focuses on benefits of pragmatic recursion.
 

### 2019-01-25 paper/paper.Rmd
new draft for MF/NDG with partially complete supplementary materials

@@
 
-Why do languages look they way they do?
-Zipf (1935) proposed that distributional properties found in natural language were evidence of speaker-listener effort minimization.
+Why do languages look the way they do?
+Zipf [-@Zipf1949a] proposed that distributional properties found in natural language were evidence of speaker-listener effort minimization.
 In his own words, “we are arguing that people do in fact act with a maximum economy of effort, and that therefore in the process of speaking-listening they will automatically minimize the expenditure of effort.” Evidence for this claim was largely derived at the level of the lexicon.
@@
 For example, Regier et al.
-(201X) showed that languages appear to organize semantic domains (form-meaning mappings) in maximally efficient ways.
+[-@RegierKempKay2015a] showed that languages appear to organize semantic domains (form-meaning mappings) in maximally efficient ways.
 Likewise, Piantadosi et al.
-(2012) argued that lexical ambiguity is an efficient property of any communication system when communication is contextualized.
+[-@Piantadosi2011a] argued that lexical ambiguity is an efficient property of any communication system when communication is contextualized.\par
 
@@
 
-But at what level of does local, in-the-moment interaction lead to optimized structure at the level of the lexicon?
-The example given by Zipf (1935) appears to describe local interaction in terms of a \textit{reference game}.
+But how does such optimization take place?
+The example given by Zipf [-@Zipf1949a], he appears to describe local, communicative interactions in terms of a \textit{reference game}.
 Speakers intend to refer to some object in the world $m$.
 They choose some utterance $u$ to transmit this intended meaning, $u \rightarrow m$.
-The listener attempts to reconstruct this intended meaning given the transmitted utterance $m \rightarrow u$.
-Other functionalist projects have also assumed this basic reference game setting (Regier, et al.
-(201X); Piantadosi et al.
-(2012); Kirby, Smith & Folds (201X), OTHERS) and this simplification of the communicative act has proven productive in both theoretical (Ferrer-i-cancho, et al.,  201X), simulation-based (iterated learning stuff) and empirical explorations of efficient language structure and use (Some of Robert’s work here).
+The listener attempts to reconstruct this intended meaning given the transmitted utterance, $m \rightarrow u$.
+Other functionalist projects have also assumed this basic reference game setting [@RegierKempKay2015a; @Piantadosi2011a] and this simplification of the communicative act has proven productive in theoretical [@FerreriCancho2018a], simulation-based [@KirbyGriffithsSmith2014a] and empirical explorations [@HawkinsFrankeSmithGoodman2018a] of efficient language structure and use.\par
 
 This move to study efficient language use in terms of reference games has clear theoretical underpinnings as well.
-The linguist Lawrence Horn (1984), highlighted the importance of Zipf’s principles for explaining conversation-level phenomena.
-Horn suggested a link between Zipf’s conception of speaker- and listener-effort and ideas related to conversational pragmatics by Grice (1975), arguing that the interaction of Zipf’s forces were “largely responsible for generating Grice’s conversational maxims and the schema for pragmatic inference derived therefrom.” Put differently, the system-level efficiencies we see in language are fundamentally related to the local-level efficiencies that occur during rational, pragmatic conversation.\par
+The linguist Lawrence Horn highlighted the importance of Zipf’s principles for explaining conversation-level phenomena.
+Horn suggested a link between Zipf’s conception of speaker- and listener-effort and ideas related to conversational pragmatics by the philosopher Paul Grice, arguing that the interaction of Zipf’s forces were “largely responsible for generating Grice’s conversational maxims and the schema for pragmatic inference derived therefrom” [@Horn1984a].
+Put differently, the system-level efficiencies we see in language are fundamentally related to the local-level efficiencies that occur during rational, pragmatic conversation.\par
 
-In this work, we present a step toward formalizing the framing presented above  -- connecting Zipfian notions of efficient language structure to Gricean notions of rational conversation.
+In this work, we present a step toward formalizing the framing presented above  -- connecting Zipfian notions of efficient language structure to Gricean notions of rational conversation [@Grice1975a].
 To do so minimally requires three basic ingredients -- (1) a language property we’d like to explain, (2) a framework for describing in-the-moment interactions of speaker-listeners, and (3) some measure of linguistic efficiency.
-For the current project we focus on the “communicative function of ambiguity” for our property -- translating the functionalist theory pursued by Piantadosi et al.
-(2012) into our framework.
-To model local, conversational interactions (2) we adopt the Rational Speech-act framework (RSA) (Frank & Goodman, 2012; Goodman & Frank, 2016).
+For the current project we focus on the “communicative function of ambiguity” for our property -- translating the functionalist theory pursued by Piantadosi, et al.
+[-@Piantadosi2011a] into our framework.
+To model local, conversational interactions (2) we adopt the Rational Speech-act framework (RSA) [@FrankGoodman2012a; @GoodmanFrank2016a].
 Finally, (3) we derive an objective function for measuring language efficiency in the reference game setting.\par
@@
 Following this introduction we derive a simple objective function for measuring the efficiency of linguistic systems in this setting.
-We move on to two case-studies examining the questions posed above, examining the efficiency of ambiguity from a language design and language use perspective.
+We move on to two case-studies exploring the questions posed above -- examining the efficiency of ambiguity from a language design and language use perspective.\par
 
-# Exploring efficient language- design and use in rational pragmatic agents
+# Exploring efficient language design and use in rational pragmatic agents
 
-```{r plot-reference-game, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2.5, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.cap = "An example reference game with associated literal semantics (in our terminology a ‘language’)."}
+```{r plot-reference-game, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2.5, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.cap = "An example reference game with associated literal semantics (in our terminology a ``language'')."}
 img <- png::readPNG("figs/game1.png")
@@
 ### Reference games
-Zipf’s canonical (1935) example of optimal speaker- and listener-languages took the form of a reference game (similar to those described by Wittgenstein).
+Zipf’s example of optimal speaker- and listener-languages took the form of a reference game (similar to those described by Wittgenstein).
 We adopt that formulation here, assuming these communication games as our basic unit of analysis.
-In this setting, speakers and listeners are aware of a set of objects $M$, which we will refer to as *meanings* and are knowledgeable about the set of possible signals $U$ (*utterances*) that can be used to refer to a given meaning.
+In this setting, speakers and listeners are aware of a set of objects $M$, which we will refer to as *meanings* and are knowledgeable about the set of possible signals $U$ (*utterances*) that can be used to refer to a given meaning (see Figure 1 for an example reference game).
 Utterances may have different relative costs, operationalized via a prior over utterances $P(U)$.
 Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$.
-Note that the prior over meanings are analogous to the *need probabilities* assumed in previous work (Regier, CITATION).
+Note that the prior over meanings are analogous to the *need probabilities* assumed in previous work Regier, Kemp & Kay [-@RegierKempKay2015a].
 We consider a set of contexts $C$ with an associated prior $P(C)$.
-Importantly, contexts describe different need probability states, formalized as different conditional distributions over meanings e.g.
+Contexts describe different orderings of the underlying meaning prior, formalized as different conditional distributions over meanings e.g.
 $p(M|C=c_i)$.
-Finally, we consider a set of communicative events $e \in E$ where $<u, m> = e$ is a tuple of utterance, referent pairs.
+Finally, we consider a set of communicative events $e \in E$ where $<u, m> = e$ is a tuple of utterance, meaning pairs.\par
 
@@
 
-describes the literal semantics of $\ell$. E.g. the language describes semantic mappings $[\![u_1]\!]\ = \{m_1, m_2\}, [\![u_2]\!]\ = \{m_2\}, [\![u_3]\!]\ = \{m_3\}$.
+describes the literal semantics of $\ell$.
 
@@
 RSA can largely be understood as a formalization of essential Gricean pragmatic principles -- interlocutors reason about one another and their shared context.
-For this reason, we adopt RSA as our representational framework to operationalize Gricean (rational and pragmatic) speaker-listeners.
-For a more extensive description of RSA as well as its applications we refer readers to Goodman & Frank (2016), limiting the current discussion to essential, high-level details of the framework.
-\par
+For this reason, we adopt RSA as our representational framework to operationalize Gricean (rational and pragmatic) speaker-listeners [@Grice1975a].
+We refer the reader to Section 1 of our supplementary materials for a more detailed description of basic RSA speaker-listener definitions, and to Goodman & Frank [-@GoodmanFrank2016a] for an overview of its appliations.
+We limit the current discussion to essential, high-level details of the framework.\par
 
@@
 We will use $L(m|u; \ell)$ as shorthand.
-Note that both speakers and listeners can induce joint distributions over the set of all signaling events $E$, although, importantly, these distributions may differ:
+Note that both speakers and listeners can induce joint distributions over the set of all events $E$, although, importantly, these distributions may differ:
 $$P_{speaker}(u, m; \ell) = S(u|m; \ell)p(m)$$
@@
 
-In general, we would like to consider the efficiency of a system $\ell$ in terms of these joint distributions.
-
 # Zipfian objective for linguistic system efficiency
 
-Zipf proposed that the particular distributional properties found in natural language emerge as a result of competing speaker and listener pressures.
-We operationalize this in equation (1) -- the efficiency of a linguistic system $\ell$ being used by speaker and listener agents $S$ and $L$ is the sum of the expected speaker and listener effort to communicate over all possible communicative events $E$.
+Zipf (1949) proposed that the particular distributional properties found in natural language emerge from competing speaker and listener pressures.
+We operationalize this in equation (1) -- the efficiency of a linguistic system $\ell$ being used by speaker and listener agents $S$ and $L$ is the sum of the expected speaker and listener effort to communicate over all possible communicative events $E$.\par
 
 \begin{equation}
-\text{Efficiency}(S, L, \ell) = \mathbb{E}_{e \in E}[\text{speaker effort}] + \\ \mathbb{E}_{e \in E}[\text{listener effort}]
+  \text{Efficiency}(S, L, \ell) = \mathbb{E}_{e \sim P(E)}[\text{speaker effort}] + \mathbb{E}_{e \sim P(E)}[\text{listener effort}]
 \end{equation}
 
-Let speaker effort be the negative log probability (surprisal) of a particular utterance. Intuitively, the number of bits needed to encode the utterance $u$.
-
-$$\text{speaker effort} = -log_2(p(u))$$
-
-Let listener effort be the negative log probability a listener disambiguates an intended meaning $m$ given an utterance $u$ Intuitively, the number of guesses a listener would need to discover the intended meaning $m$ given an utterance $u$.
-$$\text{listener effort} = -log_2(L(m|u; \ell))$$
-
-Rewriting (1) we have
-\begin{equation}
-\begin{split}
-\text{Efficiency}(S, L, \ell) = \mathbb{E}_{e \in E}[-log_2(p(u))] + \\
-\mathbb{E}_{e \in E}[-log_2(L(m|u; \ell))]
-\end{split}
-\end{equation}
-In general, these expectations are each taken over the set of possible communicative events $e\in E$ weighted by their probability, $p(e)$.
-Recall this is the set of all utterance, meaning pairs $<u, m> = e \in E$.
-
-\begin{equation}
- = \sum_{e \in E}p(e)[-log_2(p(u))] + \sum_{e \in E}p(e)[-log_2(L(m|u; \ell))]
-\end{equation}
-
-We assume that the particular joint distribution over utterance-meaning $<u, m> = e$ pairs follows from a simple generative model.
-First, some meaning is sampled with probability $p(m)$.
-Our speaker attempts to convey this intended meaning to a listener via an utterance $u$ by sampling from the speaker conditional distribution $S(u|m; \ell)$.
-Combining these terms leads to the *speaker's joint distribution over events* which we can write as $P(e) = S(u | m; \ell)p(m) = P_{speaker}(u, m; \ell)$.
+We assume that speaker effort is the negative log probability of an utterance $-log_2(p(U))$, listener effort is the negative log probability of the listener's conditional $-log_2(L(M|U))$, and that events $e = <u, m>$ are sampled according the to following generative model -- some object occurs in the world with probability $p(M=m)$, the speaker attempts to refer to that object by sampling from her conditional distribution $S(u|m)$ (i.e.
+$e \sim p(m)S(u|m)$).
+Using these basic ingredients we arrive at the following objective:
 
@@
 \begin{split}
-  = \sum_{u, m}P_{speaker}(u, m; \ell)[-log_2(p(u))] + \\ 
-   \sum_{u, m}P_{speaker}(u, m; \ell)[-log_2(L(m|u; \ell))]
+  = \sum_{u, m}P_{speaker}(u, m; \ell)[-log_2(P_{listener}(u, m; \ell))]
 \end{split}
 \end{equation}
-
-Simplifying we arrive at (5):
-\begin{equation}
-\begin{split}
- = \sum_{u, m}P_{speaker}(u, m; \ell)[-log_2(L(m|u; \ell)p(u))]
- \end{split}
-\end{equation}
-
-Note that $L(m|u; \ell)p(u)$ is the listener-based joint distribution over all communicative events ($P_{listener}(u, m; \ell)$).
-\begin{equation}
-  = \sum_{u, m}P_{speaker}(u, m; \ell)[-log_2(P_{listener}(u, m; \ell))]
-\end{equation}
-This is the simply the cross-entropy between the speaker an listener joint distributions.
+This is the simply the cross-entropy between the speaker and listener joint distributions.
 \begin{equation}
@@
 In our case, this is the difference between the joint distribution assumed by the speaker $P_{speaker}$ and listener $P_{listener}$.
-A good language $\ell$ used by a set of a pair of speaker-listeners will have properties which minimize this objective.\par
+An "efficient" language $\ell$ used by a set of a pair of speaker-listeners will have properties which minimize this objective.
+For a complete derivation and its componetns we refer the reader to Section 2.1 of our supplementary materials.
+\par
 
-# Simulations
+# Simulating the communicative function of ambiguity
 
 Ambiguity is ubiquitous in natural language.
-At a minimum, the listener task in everyday communication is marked by the frequent need to handle lexical ambiguity (words often have multiple meanings) as well as syntactic ambiguity (sentences often have multiple parses), in addition to other instances of ambiguity (Wasow, Perfors & Beaver, 201X).
-Chomsky (2002) famously claimed that the presence of ambiguity in natural language provides evidence that language has not been optimized for communication, stating “If you want to make sure that we never misunderstand one another, for that purpose language is not well designed, because you have such properties as ambiguity.” Note that this analysis is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$ we described earlier and define in terms of equation (9).\par
+The task of understanding language is marked by a frequent need to handle lexical ambiguity (words often have multiple meanings) as well as syntactic ambiguity (sentences often have multiple parses), in addition to many other instances of ambiguity [@WasowPerforsBeaver2005a].
+Chomsky [-@Chomsky2002a] famously claimed that the presence of ambiguity in natural language provides evidence that language has not been optimized for communication, stating “If you want to make sure that we never misunderstand one another, for that purpose language is not well designed, because you have such properties as ambiguity.” Note that this analysis is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$ we described earlier and define in terms of equation 2 in section 2.2 for our supplementary materials.\par
 
 Piantadosi et al.
-(2012) argue just the opposite.
-They claim that ambiguity is an *efficient* property of any communication system in which *communication is contextualized*.
+[-@Piantadosi2011a] argue just the opposite, claiming that ambiguity is an *efficient* property of any communication system in which *communication is contextualized*.
 Simply put, it is useful to have a language that re-uses low-cost material (has ambiguity) so long as the cost of disambiguating the material is low.
-In particular, context (or common ground) can provide useful information for disambiguation.
-Say we have two objects in the world $m_1$ and $m_2$ and two languages $\ell_1$ and $\ell_2$.
-In language $\ell_1$ the low-cost utterance $u_1$ can be used to refer to both $m1$ and $m_2$ (i.e.
-$[[u_1]]_{\ell_1} = \{m_1, m_2 \}$.
-Language $\ell_2$, by contrast, defines a mapping such that the low cost utterance $[[u_1]]_{\ell_2} = \{m_1\}$ and a higher cost utterance $[[u_2]]_{\ell_2} = \{m_2 \}$.
+In particular, context (or common ground) can provide useful information for disambiguation.\par
+
+We provide a toy example of their argument here.
+Say we have two objects in the world $m_1$ and $m_2$, and two languages $\ell_1$ and $\ell_2$.
+In language $\ell_1$, the low-cost utterance $u_1$ can be used to refer to both $m1$ and $m_2$ (i.e.
+$[[u_1]]_{\ell_1} = \{m_1, m_2 \}$).
+With language $\ell_2$, however, the low cost utterance $u_1$ can only refer to $m_1$ and a higher-cost utterance $u_2$ can refer to $m_2$ (i.e.
+$[[u_1]]_{\ell_2} = \{m_1\}$ and $[[u_2]]_{\ell_2} = \{m_2 \}$).
 While it is cheaper for a speaker to use $\ell_1$ (because she can always use a low-cost utterance), it is more difficult for a listener (because $u_1$ is ambiguous).
-However, if the meaning $u_1$ can be disambiguated *in context* then the listener will not incur disambiguation cost and the speaker can always use the low-cost utterance.
-This a basic argument for the contextual efficiency of polysemy.
-If I use the word “bat” at a baseball game a listener will likely infer I am referring to the baseball sense of the word.
-If I use the word “bat” in a dark cave, a listener will likely infer I am referring to the animal.
+If context isn't disambiguating then we might prefer $\ell_2$ to $\ell_1$.
+But, if context is disambiguating then the speaker can use $u_1$ to refer to either $m_1$ or $m_2$ and $\ell_1$ should be preferred.\par
 
 In the following experiments we explore two aspects of Piantadosi’s claim.
-First, we examine the efficient language *structure* aspect of their claim, exploring when the optimal linguistic system $\ell^*$ (under our cross-entropy objective) is most likely to contain ambiguous lexical items.
-In the second experiment we explore an efficient language *use* aspect of the question -- at what point in a conversation is it useful for a speaker to use ambiguous lexical material?
+First, we examine the efficient language *structure* aspect of their claim, exploring when the optimal linguistic system $\ell^*$ is most likely to contain ambiguous lexical items.
+In the second experiment, we explore an efficient language *use* aspect of the claim -- at what point in a conversation is it useful for a speaker to use ambiguous lexical material?\par
 
-## Simulation 1:  Optimal languages contain ambiguity when context is informative
-
-The argument outlined by Piantadosi et al.
-(2012) is compatible with our current reference game setting with just a single addition -- context.
-The authors argue that mapping two meanings $m_1$ and $m_2$ to a single utterance $u_1$ is useful when they can be disambiguated *in context*.
-In our case we consider a context $c$ to specify a unique ordering of the need probabilities.
-That is, $p(m|c_1) \neq p(m|c_2)$, when there are two contexts $|C| = \{c_1, c_2\}$.
-
-This in turn, leads to an update to our linguistic efficiency objective as we now would like to consider the average speaker-listener effort over all contextualized communicative events. That is,
-
-\begin{equation}
-\begin{split}
-\sum_{c\in C}p(c)\sum_{u, m}S(u|m, c;\ell)p(m|c)-log[L(m|u,c;\ell)p(u)] = \\ \mathbb{E}_{c\sim P(c)}[\mathbb{E}_{P_{speaker}(u, m, c; \ell)}[-log(P_{listener}(u, m, c; \ell))]]
-\end{split}
-\end{equation}
-
-Note that in the case that $|C| =1$, our objective simplifies to our original equation (7).
-With this update we can compare languages $L$ while varying the degree to which context contains useful information, varying the number of contexts (the size of $|C|$).\par
-
-### Simulation set-up
-
-#### Basic ingredients
-We conduct $N=2000$ simulations.
-For each simulation we enumerate the set of *valid* languages in which $|U|=|M|=4$.
-Note that a language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment).
-For a given simulation the goal is to find the language $\ell^*$ which minimizers our cross-entropy objective.\par
-
-Recall that language efficiency is both a function of the particular semantic mappings induced by that language, the speaker and listener agents ($S$ and $L$), as well as the utterance ($P(U)$), meaning ($P(M)$), and context priors ($P(C)$).
-Rather than assume particular structure for our utterance, meaning and context prior distributions, for each simulation we generate $P(U) \sim \text{Dir}(1, |U|)$, $P(M) \sim \text{Dir}(1, |M|)$, and $P(C) \sim \text{Dir}(1, |C|)$ where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.
-
-#### Context
-Following the argument given by Piantadosi et al.
-(2012) we want to assess the impact of *context* on our objective.
-To do so we consider four conditions with $n=500$ simulations each (that is, 500 unique sets of $\{P(U), P(M), P(C)\}$.
-Our first is a *single-context* condition ( $|C|=1$) -- there is a only a single context describing $P(M)$.
-Our second condition contains two-contexts  ($|C| = 2$) -- we consider efficiency under both $P(M|C=c_1)$ as well as $P(M|C=c_2)$.
-The third and fourth condition correspond accordingly with $|C|=3$ and $|C| = 4$, respectively.
-
-#### Baselines
-For comparison, we also examine properties of optimal languages under two additional objectives.
-Zipf (1935) proposed that the optimal speaker language $\ell_{speaker}^*$ should only optimize speaker effort.
-We operationalize this using the *first half* of equation (1).
-
-\begin{equation}
-\ell_{speaker}^* = argmin_{\ell\in L}\mathbb{E}_{P_{speaker}(u, m; \ell)}(p(u))
-\end{equation}
-
-The optimal listener language $\ell_{listener}^*$, by contrast, should only optimize listener effort. We operationalize this using the *second half* of equation (1).
-
-\begin{equation}
-\ell_{listener}^* = argmin_{\ell\in L}\mathbb{E}_{P_{speaker}(u, m; \ell)}(L(m|u;\ell))
-\end{equation}
-
-#### Hypotheses
-Piantadosi’s claim leads to the following set of hypotheses: (1) The probability that an optimal language contains ambiguity under our Cross-Entropy objective should increase with the number of contexts (as $|C|\rightarrow\inf$).
-(2) The optimal speaker language under our “Speaker-only” objective should always map all meanings to the single, lowest cost utterance.
-(3) The optimal listener language should never contain ambiguous material.
-
-### Results
-Figure 1.
-plots the proportion of optimal languages under $H_{cross}(P_{Speaker}, P_{Listener}; \ell)$ for each context condition.
-The red line indicates that as the number of contexts increases, so does the probability that the optimal language $\ell^*_{cross}$ under our speaker-listener cross-entropy objective contains ambiguity.
-We also plot the proportion of optimal speaker-only $\ell^*_{speaker}$ and listener-only $\ell^*_{listener}$ languages that contain ambiguity.
-In line, with Zipf's predictions, if languages are designed only to minimize speaker effort then optimal languages will assign all meanings to a single utterance (full ambiguous).
-Likewise, if languages are designed only to minimize speaker listener effort then ambiguity should always be avoided (bijective form-meaning mappings).\par
-
-```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=8, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Optimal languages are more likely to contain ambiguous items as amount of contextual information increases.
+```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=8, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Optimal languages are more likely to contain ambiguous items as the amount of contextual information increases.
 Vertical axis shows the proportion of optimal languages that contain ambiguity.
@@
 
-### Summary
+# Simulation 1:  Optimal languages contain ambiguity when context is informative
+
+The argument outlined by Piantados et al.
+[-@Piantadosi2011a] is compatible with our current reference game setting with an addition -- context.
+The authors argue that mapping two meanings $m_1$ and $m_2$ to a single utterance $u_1$ is useful when they can be disambiguated *in context*.
+In our case we consider a context $c_i \in C$ to specify a unique ordering of the need probabilities.
+That is, $p(m|c_1) \neq p(m|c_2)$, when there are two contexts $|C| = \{c_1, c_2\}$.\par
+
+This in turn, leads to an update to our linguistic efficiency objective as we now would like to consider the average speaker-listener effort over all contextualized communicative events (section 3 of supplementary materials).\par
+
+## Simulation set-up
+
+We conduct $N=2000$ simulations.
+For each simulation we enumerate the set of *valid* languages in which $|U|=|M|=4$.
+Note that a language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment) and each form maps to at least one meaning (every column has some non-zero assignment).
+For a given simulation the goal is to find the language $\ell^*$ which minimizes our cross-entropy objective.\par
+
+Recall that language efficiency is both a function of the particular semantic mappings induced by that language, the speaker and listener agents ($S$ and $L$), as well as the utterance ($P(U)$), meaning ($P(M)$), and context priors ($P(C)$).
+Rather than assume particular structure for our utterance, meaning, and context prior distributions, for each simulation we generate $P(U) \sim \text{Dir}(1, |U|)$, $P(M) \sim \text{Dir}(1, |M|)$, and $P(C) \sim \text{Dir}(1, |C|)$ where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.\par
+
+### Context
+
+Following the argument given by Piantadosi et al.
+[-@Piantadosi2011a], we want to assess the impact of *context* on our objective.
+To do so we consider four conditions with $n=500$ simulations each (that is, 500 unique sets of $\{P(U), P(M), P(C)\}$.
+Our first is a *single-context* condition ($|C|=1$) -- there is a only a single context describing $P(M)$.
+Our second condition contains two-contexts  ($|C| = 2$) -- we consider efficiency under both $P(M|C=c_1)$ as well as $P(M|C=c_2)$.
+The third and fourth condition correspond accordingly with $|C|=3$ and $|C| = 4$, respectively.\par
+
+### Baselines
+
+For comparison, we also examine properties of optimal languages under two additional objectives.
+Zipf [-@Zipf1949a] proposed that the optimal speaker language $\ell_{speaker}^*$ would only optimize speaker effort and the optimal listener language $\ell_{listener}^*$ would only optimize listener effort.
+We define these objectives using the first and second half of equation 1 and with greater detail in the supplementary materials (Section 2.2).\par
+
+## Hypotheses
+
+Piantadosi’s claim leads to the following set of hypotheses: (1) The probability that an optimal language contains ambiguity under our Cross-Entropy objective should increase with the number of contexts (as $|C|\rightarrow\inf$).
+(2) The optimal speaker language under our “Speaker-only” objective should always contain ambiguity (maximally re-using the lowest-cost utterance).
+(3) The optimal listener language should never contain ambiguous material.
+
+## Results
+
+Figure 1.
+plots the proportion of optimal languages under each objective, as a function of context condition.
+The red line shows that as the number of contexts increases, so does the probability that the optimal language $\ell^*_{cross}$ under our speaker-listener cross-entropy objective contains ambiguity.
+We also plot the proportion of optimal speaker-only $\ell^*_{speaker}$ and listener-only $\ell^*_{listener}$ languages that contain ambiguity.
+In line with Zipf's predictions, if languages are designed only to minimize speaker effort then optimal languages will assign lower-cost forms multiple meanings (all of the speaker-optimal languages contained ambiguity).
+Likewise, if languages are designed only to minimize listener effort then ambiguity should always be avoided (none of the listener-optimal languages contained ambiguity).\par
+
+## Summary
+
 Piantadosi et al.
-(2012) argued that it is useful to re-use low-cost linguistic forms for multiple meanings when they can be disambiguated in context.
-We simulated this effect in a reference game setting, examining the probability that an optimal language will contain an word with an ambiguous mapping as a function of context size.
-Using our speaker-listener cross-entropy measure of efficiency we showed that optimal languages are more likely to have ambiguous items when context is informative, but not under objectives which emphasizes speaker- or listener-effort alone.\par
+[-@Piantadosi2011a] argued that it is useful to re-use low-cost linguistic forms for multiple meanings when they can be disambiguated in context.
+We explored the degree to which this this property emerged during communication between rational, pragmatic agents, showing that as contextual information increased, the optimal languages under our Cross-Entropy objective were more likely to have ambiguous lexical items.
+Importantly, we did not see this effect under alternative objectives which only emphasized speaker- or listener effort alone.\par
 
-## Simulation 2: Rational-pragmatic speakers use ambiguity efficiently
+# Simulation 2: Rational, pragmatic speakers use ambiguity efficiently
 
-Our first experiment made a fairly direct test of the communicative function of ambiguity theory presented by Piantadosi et al.
-(2012).
-Sampling a set of need probabilities and utterance costs, we explored the space of possible languages, examining the properties of the language $\ell^*$ which minimized our objective.
+Our first experiment made a fairly direct test of the communicative function of ambiguity theory.
+Sampling meaning, utterance, and context priors, we examined properties of the language $\ell^*$ which minimized our objective.
 Results indicated that ambiguity is an efficient property when context is informative.
 This simulation, however, assumed that our speaker and listener agents could always disambiguate items context.
-That is, in our four conditions both agents had perfect knowledge of the conditional distributions $P(M|C_i)$.\par
+That is, in our four conditions both agents had perfect knowledge of the relevant conditional distributions ($P(M|C_i)$).\par
 
-This is a strong assumption and likely unrealistic for much of day-to-day communication.
-In talking with another person, we often \textit{don’t have perfect knowledge of the current context of a conversation}.
-We can consider this notion of “context” as analogous to a *topic* of conversation.
-Consider the scenario in which a reader is beginning a newspaper article.
+This may be a strong assumption for describing much of day-to-day communication -- we seldom begin a conversation with perfect knowledge of the current context of a conversation.
+Under Piantadosi's framework, if context is not informative with respect to meaning at time position $i$ in a discourse, but is informative at position $i+k$, then the speaker should avoid using ambigous material at position $i$, only using it at a future point, $k$ steps later in the conversation.
+In this toy example we can consider “context” as analogous to a “topic” of conversation.
+Intuitvely, we can imagine a scenario in which a reader is beginning a newspaper article.
 While they may have some knowledge about the article’s topic (perhaps from the title), they may not have complete knowledge of its contents, including the persons or events involved.
-In this scenario, using an ambiguous pronoun early in the article lead to misunderstanding as there isn’t sufficient context for disambiguation.
-Using such ambiguity is only warranted when the appropriate person or event is salient in the context.
-\par
+In this setting, using an ambiguous pronoun early (position $i$) in the article can lead to misunderstanding without sufficient context to support disambiguation.
+But if by position $i+k$ there is sufficient contextual information it may be efficient for the writer to use ambiguous material.\par
 
-We consider such a scenario in simulation two -- analyzing efficient *use* of ambiguity over discourse.
-In this scenario, we assume a single language $\ell$ which contains both ambiguous and unambiguous utterances.
+##  Simulation set-up
+
+We consider such a scenario in simulation two -- analyzing efficient *use* of ambiguity over discourse in a reference game setting.
+Unlike our first simulation, we now assume a single language $\ell$, which contains both ambiguous and unambiguous utterances.
 The ambiguous utterances are less costly than the unambiguous utterances.
-Crucially, we do not assume that a listener has access the conditional distributions $p(M|C=c)$ a priori.
-Rather, we assume that the listener has knowledge of the set of possible contexts $C = \{c_1, \dots, c_k\}$, but \textit{does not know which one is currently being used by the speaker}.
-In this setting, a speaker minimizing her production effort would like to use the low-cost material as frequently as possible.
+Crucially, we do not assume that the listener knows the particular topic ($c_{current}$) of the conversation \textit{a priori}.
+Rather, we assume that the listener has knowledge of the set of possible topics $C = \{c_1, \dots, c_k\}$, but \textit{does not know which one is currently being used by the speaker}.
+Formally, this means the listener does not have access to the correct conditional distribution over meanings $P(M|C=c_{current})$ at the start of the discourse.
+In this setting, a speaker minimizing production effort would like to use the low-cost material as frequently as possible.
 However, this will incur a disambiguation cost to the listener if they are uncertain of the current context $c_{current}$ (topic of conversation).\par
@@
 Over the course of a discourse, $D$ the listener tries to infer both the current topic, $c_{current}$, as well as the particular meaning $m$ of a given utterance $u$.
-That is we consider a listener model $L(m, c|u,D;\ell)$ and speaker model $S(u|m,c,D;\ell)$.
+That is we consider a modified listener model $L(m, c|u,D;\ell)$ and speaker model $S(u|m,c,D;\ell)$.
 Note that both the speaker and listener can track the history of previous utterances $D$.
-Importantly, a listener can attempt to infer the current topic of conversation $c_{current}$ using the discourse history $D$.\par
+Importantly, a listener can attempt to infer the current topic of conversation $c_{current}$ using the discourse history $D$.
+Section 3 of the supplementary materials provides more detail on the speaker and listener models.\par
 
-```{r plot-optimal-use, fig.env = "figure*", fig.pos = "h", fig.width=8, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Left facet (A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position.
-As the discourse proceeds, the speaker is more likely to use an ambiguous term.
-Right face (B) shows the probability the listener assigns to the current topic of conversation."}
-img <- png::readPNG("figs/fig2.png")
+```{r plot-optimal-use, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Left facet (A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position.
+As the discourse proceeds, only our Full Pragmatic speaker is more likely to use an ambiguous term.
+Plot B tracks the Cross-Entropy objective under our three speaker models.
+The objective declines in all cases as the listener infers the correct topic distribution.
+However the pragmatic speaker arrives as the lowest value, do to reduction in both speaker and listener effort.
+Plot C tracks speaker effort across the three models.
+Only the Full Pragmatic speaker shows declines in effort across the discourse as they increasinlgy use ambiguous items as the listener becomes more sure of the topic of conversation."}
+img <- png::readPNG("figs/discourse_grid_plot.png")
 grid::grid.raster(img)
@@
 
-### Hypothesis
-In this setting, if speakers are *using ambiguity efficiently* they should only use the ambiguous material when they know the listener can disambiguate the items in context.
-Since listeners accumulate information about the current topic distribution ($P(C=c_{current}|D)$) over as the discourse progresses, we should expect to see more frequent use of ambiguous items later in the discourse.
-
-### Basic ingredients
-
-We conduct $N=200$ simulations, generating discourses of length $|D|=30$ utterances.  We consider a single language $\ell$ with $|U|=6$ and $|M|=4$ specified by the boolean matrix:
+We conduct $N=600$ simulations, generating discourses of length $|D|=30$ utterances with three different speaker models ($n=200$ each).
+We consider a single language $\ell$ with $|U|=6$ and $|M|=4$ specified by the boolean matrix:
 
@@
 
-We assume that $p(u_5) = p(u_6) > p(u_{1})=\dots=p(u_{4})$.
-That is, the two ambiguous utterances are less costly than the non-ambiguous utterances.
-In this setting, it is less costly for the speaker to use the ambiguous utterances.
-However, if the listener cannot disambiguate them, then they will incur a disambiguation cose.
-We assume that over the discourse both agents have complete recall of the set of possible utterances by the speaker $D$.
-This means that the listener can infer the current topic distribution $p(C=c|D)$ via $p(C=c|D) \propto \prod_{i=0}^{|D|}S(u_i|m_i)p(m_i|c)p(c)$.
-In this way, an efficient speaker should avoid ambiguity \textit{early} in the discourse when the listener is unsure of the context, only using it later once they are confident the listener knows the current context $p(C=c_{current}|D)$.
+We assume that $p(u_5) = p(u_6) > p(u_{1})=\dots=p(u_{4})$. That is, the two ambiguous utterances ($u_5$ and $u_6$) are less costly than the non-ambiguous utterances.\par
 
-### Results
+## Speaker agents
+We consider three types of speaker models.
+In our \textit{Full pragmatics} model, we consider a speaker agent who reasons about her listener and also has complete recall of the set of utterances in the discourse $D$.
+This speaker believes that the listener may not know the current topic $c_{current}$ at the start of the discrouse, but can infer it with enough data.
+In this way, contextual information shared by the speaker and listener may increase over the discourse (as the listeners estimate of the current topic ($P(C=c|D)$) comes in line with the actual topic known to the speaker).
+We compare two baselines models.
+The first, a \textit{Partial pragmatics} baseline describes a speaker who reasons about a listener, but assumes they have no access to discourse history.
+The second is a \textit{No pragmatics} speaker who does not consider a listener at all, but produces utterances according to the underlying langauge semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$).
+We provide detailed descriptions of each model in the supplementary materials Section 3.\par
 
-Figure 2, plot A shows the empirical probability that an RSA speaker used an ambiguous utterance as a function of discourse position.
-While at the start of the discourse the speaker is unlikely to use the utterance we see consistent use later, once the listener is sure of the particular context $c_i$ this speaker is using.\par
+## Hypotheses
+Under Piantadosi's claim it is efficient to use low-cost, ambiguous material when context is disambiguating.
+In the current setting, this corresponds to speaker behavior in which ambiguouous forms are avoided early in the discourse, but prefered later in the discourse once the listener is confident of the topic.
+We should expect this to be reflected in our \textit{Full pragmatics} model.
+By contrast, our \textit{Partial pragmatics} speaker does not believe that there is sufficient contextual information for the listener to disambiguate her utterances.
+This should lead to total  avoidance of ambiguous material over the discourse.
+Finally, our \textit{No pragmatics} speaker should use low-cost material greedily since she does not consider a listener's need to disambiguate whatsoever.\par
 
-The right facet of figure 2 shows the listener belief about the current topic $c_{current} \in C$ given the discourse history ($p(C=c_{current}|D)$).
-Note that the trajectories in plot A and B are related -- the speaker is more likely to use ambiguous material only when the listener is able to disambiguate that material.
+## Results
 
-### Summary
+Figure 2, plot A shows the empirical probability that a speaker uses an ambiguous utterance as a function of discourse position.
+Note that the \text{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse, the \text{Partial pragmatic} baseline avoids ambiguous utterances entirely (reasonining that the utterances cannout be disambiguated), but the \text{Full pragmatic} model avoids ambiguous material at first, but employing them gradually increasingly as the discourse proceeds.
+Figure 2, plot B tracks the speaker-listener Cross-Entropy objective.
+Note that the objective decreases for all three models.
+This decrease in all three models is primarily driven by the listener updating his belief about the actual topic ($P(C=c_{current}|D)$).
+Of particular importance is the value of the objective at the end of the discourse (position $i=30$).
+The difference in the Cross-Entropy between the \text{Full Pragmatic} and \text{Partial Pragmatic} models at the end of the discourse is driven by the reduction in speaker costs described in Figure 2, plot C.
+That is, while speaker effort remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines, effort declines for the \text{Full Pragmatic} model as they are increasingly able to use ambiguous material later in the discourse.\par
 
-These simulations explored the efficient *use* of ambiguous material over a discourse, in an alternative instantiation of the argument outlined by Piantadosi et al.
-(2012).
+## Summary
+
+These simulations explored the efficient *use* of ambiguity over a discourse, in an alternative instantiation of the argument outlined by Piantadosi et al.
+[-@Piantadosi2011a].
 In the first set of simulations we varied the set of languages, examining the properties of languages which minimized our cross-entropy objective under different context conditions.
-In the current simulation we assumed that the listener did not know the current context (toptic) a priori, but had to infer it from the discourse.
+In the current simulation we assumed that the listener did not know the current context (topic) \textit{a priori}, but could infer it from the discourse.
 We examined speaker production behavior, tracking use of low-cost, but ambiguous material over the discourse.
-We found that early in the discourse our rational speaker avoided using ambiguous material, only using the material later when the context was known to the listener.
+We found that early in the discourse our \textit{Full pragmatic} speaker avoided using ambiguous material, only using the material later when the context was known to the listener.
+This is consistent with Piantaodi's claim -- it is useful to use low-cost material, but only when context is disambiguating.
 
@@
 How do the competing pressures of speakers and listeners give rise to the distributional forms found in natural language?
-Zipf (1935) proposed that the asymmetry between speaker and listener costs gives rise to a range of properties at the level of the lexicon.
-Subsequent functionalist approaches to language science have provided additional evidence that various aspects of language structure and language use appear to be optimized for efficiency.
+Zipf [-@Zipf1949a] proposed that the asymmetry between speaker and listener costs gives rise to a range of properties at the level of the lexicon.
+Subsequent functionalist approaches to language science have provided additional evidence that aspects of language structure and language use appear to be optimized for efficiency.
 In Zipf’s famous example of an optimal speaker and listener language he appears indicate that such optimization occurs in in-the-moment, local interactions between speakers and listeners.
-We explore this idea in our current project, following theoretical work by Horn (1984) who proposed the speaker-listener pressures identified by Zipf can also be interpreted in terms of the maxims governing rational conversation proposed by Grice (1975).
-To so, we explored the degree to which the interactions of rational pragmatic agents can provide a framework for understanding efficient language structure and use.
+We explore this idea in our current project, following theoretical work by Horn [-@Horn1984a] who proposed the speaker-listener pressures identified by Zipf can also be interpreted in terms of the maxims governing rational conversation proposed by [@Grice1975a].
+To so, we explored the interactions of rational pragmatic agents as a framework for understanding efficient language structure and use.
 We focused on an argument on the communicative function of ambiguity put forth by Piantadosi et al.
-(2012).
+[-@Piantadosi2011a].
 We derive a novel speaker-listener cross-entropy objective for measuring the efficiency of a language in a reference game setting showing that optimal languages are more likely to contain ambiguous material when context is informative.
@@
 The intent is that this project provides a demonstration for the types of functionalist theories that can be explored within this framework.
-Among other language properties open to this type of analysis are work examining semantic typology (CITATIONS), the emergence of compositionality (CITATIONS) and convention formation, as well as efficient examples of language use including the efficient use of reduction and redundancy (Levy & Jaeger, 2007) and discourse (Genzel & Charniak, 2002).
+Among other language properties open to this type of analysis are work examining semantic typology [@RegierKempKay2015a], the emergence of compositionality [@SmithKirbyBrighton2003a] and convention formation, as well as efficient examples of language use including the efficient use of reduction and redundancy [@LevyJaeger2007a; @GenzelCharniak2002a].
 
 # References
-Chomsky (2002)
-Cover & Thomas, XXXX)
-Ferrer-i-cancho, et al.,  201X
-Frank & Goodman (2012)
-Genzel & Charniak (2002)
-Goodman & Frank (2016)
-Grice (1957)
-Horn (1984)
-Kibry, Smith & Folds (201X)
-Levy & Jaeger (2007)
-Piantadosi et al. (2012)
-Regier, et al.
-Wasow, Perfors & Beaver (chomsky201X)
-Zipf (1935)
 

### 2019-01-28 paper/paper.Rmd
updated paper and supplementary materials after M Hahn comments.

@@
 Utterances may have different relative costs, operationalized via a prior over utterances $P(U)$.
-Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$.
-Note that the prior over meanings are analogous to the *need probabilities* assumed in previous work Regier, Kemp & Kay [-@RegierKempKay2015a].
+Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$\footnote{The prior over meanings are analogous to the \textit{need probabilities} assumed in previous work Regier, Kemp \& Kay (2015).}.
 We consider a set of contexts $C$ with an associated prior $P(C)$.
@@
 A language $\ell$ defines the set of semantic mappings between utterance and meanings.
-For example, in a world with three utterances $U = \{u_1, u_2, u_3\} \text{ and three meanings }M = \{m_1, m_2, m_3\}$ the boolean matrix
-
-\begin{center}
-\begin{tabular}{ c | c c c } 
-& $m_1$ & $m_2$ & $m_3$ \\
-\hline
-$u_1$ & 1 & 1 & 0 \\
-$u_2$ & 0 & 1 & 0 \\
-$u_3$ & 0 & 0 & 1 \\
-\end{tabular}
-\end{center}
-
-describes the literal semantics of $\ell$.
+For example, figure 1 contains four utterances $U = \{\text{blue}, \text{green}, \text{square}, \text{circle}\} \text{ and three meanings }M = \{\text{green-square}, \text{blue-square}, \text{green-circle}\}$.
+The boolean matrix describes the literal semantics of the language $\ell$.
+Note that will consider language to "contain ambiguous items" if there is at least one utterance $u \in U$ which can apply to multiple meanings (i.e.
+$|[[u_i]]| > 1$).
+For example, in Figure 1 both the words "blue" and "green" are ambigous so we would say that $\ell$ contains ambiguity.
 
@@
 \begin{equation}
-  \text{Efficiency}(S, L, \ell) = \mathbb{E}_{e \sim P(E)}[\text{speaker effort}] + \mathbb{E}_{e \sim P(E)}[\text{listener effort}]
+\begin{split}
+  \text{Efficiency}(S, L, \ell) = \mathbb{E}_{e \sim P(E)}[\text{speaker effort}] \\+ \mathbb{E}_{e \sim P(E)}[\text{listener effort}]
+\end{split}
 \end{equation}
 
-We assume that speaker effort is the negative log probability of an utterance $-log_2(p(U))$, listener effort is the negative log probability of the listener's conditional $-log_2(L(M|U))$, and that events $e = <u, m>$ are sampled according the to following generative model -- some object occurs in the world with probability $p(M=m)$, the speaker attempts to refer to that object by sampling from her conditional distribution $S(u|m)$ (i.e.
+We assume that speaker effort is the negative log probability of an utterance (surprisal) -- intuitively, the number of bits needed to encode the utterance $u$.
+This particular formalization of speaker-cost is general enough to accommodate a range of instantiations that might in theory be related to production difficult via articulation effort, cognitive effort related to lexical access, etc [@BennettGoodman2015a].\par
+
+$$\text{speaker effort} = -log_2(p(u))$$
+
+We assume listener effort is the negative log probability of the listener's conditional $-log_2(L(M|U))$.
+This operationalization of listener effort is intuitively related to existing work in sentence processing in which word difficulty during comprehension is proportional to surprisal [@Hale2001a; @Levy2008a].
+
+$$\text{listener effort} = -log_2(L(m|u; \ell))$$
+
+Finally, we assume that events $e = <u, m>$ are sampled according the to following generative model -- some object occurs in the world with probability $p(M=m)$, the speaker attempts to refer to that object by sampling from her conditional distribution $S(u|m)$ (i.e.
 $e \sim p(m)S(u|m)$).
@@
   = \mathbb{E}_{P_{speaker}}[-log_2(P_{listener})] \\
-  = H_{cross}(P_{speaker}, P_{listener})
+  = H_{cross}(P_{speaker}, P_{listener}; \ell)
 \end{split}
@@
 An "efficient" language $\ell$ used by a set of a pair of speaker-listeners will have properties which minimize this objective.
-For a complete derivation and its componetns we refer the reader to Section 2.1 of our supplementary materials.
-\par
+For a complete derivation and its components we refer the reader to Section 2.1 of our supplementary materials.\par
 
@@
 Note that a language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment) and each form maps to at least one meaning (every column has some non-zero assignment).
-For a given simulation the goal is to find the language $\ell^*$ which minimizes our cross-entropy objective.\par
+For a given simulation the goal is to find the language $\ell^*$ which minimizes our cross-entropy objective and then check to see if that language contains ambiguity.\par
 

### 2019-01-28 paper/paper.Rmd
updates from MFs first comments (looks like its mostly intro and he didnt get to simulations

@@
 ---
-title: "Pragmatic interactions lead to efficient language structure and use"
+title: "Interactions of rational, pragmatic agents lead to efficient language structure and use"
 bibliography: library.bib
@@
 abstract: 
-    We discuss a framework for studying how the distributional properties of linguistic systems emerge from in-the-moment interactions of speakers and listeners.
+    Languages display a diverse set of distributional regularities such as the relation between a word's frequency and rank in a corpus, the distribution of dependency lengths, or the presence of lexical properties such as ambiguity.
+We discuss a framework for studying how these properties emerge from in-the-moment interactions of rational, pragmatic speakers and listeners.
 Our work takes Zipfian notions of lexicon-level efficiency as a starting point, connecting these ideas to Gricean notions of conversational-level efficiency.
-To do so, we begin by deriving an objective function for measuring the communicative efficiency of linguistic systems and then examining the behavior of this objective in a series of simulations focusing on the communicative function of ambiguity in language.
-These simulations suggest that rational pragmatic agents will produce communicatively efficient systems.
+To do so, we derive an objective function for measuring the communicative efficiency of linguistic systems and then examining the behavior of this objective in a series of simulations focusing on the communicative function of ambiguity in language.
+These simulations suggest that rational pragmatic agents will produce communicatively efficient systems and that interactions between such agents provide a framework for examining efficient properties of language more broadly.
 
@@
 Zipf [-@Zipf1949a] proposed that distributional properties found in natural language were evidence of speaker-listener effort minimization.
-In his own words, “we are arguing that people do in fact act with a maximum economy of effort, and that therefore in the process of speaking-listening they will automatically minimize the expenditure of effort.” Evidence for this claim was largely derived at the level of the lexicon.
+In his own words, “we are arguing that people do in fact act with a maximum economy of effort, and that therefore in the process of speaking-listening they will automatically minimize the expenditure of effort.” Evidence for this claim has been largely derived at the level of the lexicon.
 Zipf argued that the particular relationship between a word’s frequency and its rank, length, and denotation size could be explained as an emergent property of speaker-listener effort minimization.
@@
 But how does such optimization take place?
-The example given by Zipf [-@Zipf1949a], he appears to describe local, communicative interactions in terms of a \textit{reference game}.
+The example given by Zipf [-@Zipf1949a], describes local, communicative interactions in terms of a \textit{reference game}.
 Speakers intend to refer to some object in the world $m$.
@@
 
-This move to study efficient language use in terms of reference games has clear theoretical underpinnings as well.
-The linguist Lawrence Horn highlighted the importance of Zipf’s principles for explaining conversation-level phenomena.
-Horn suggested a link between Zipf’s conception of speaker- and listener-effort and ideas related to conversational pragmatics by the philosopher Paul Grice, arguing that the interaction of Zipf’s forces were “largely responsible for generating Grice’s conversational maxims and the schema for pragmatic inference derived therefrom” [@Horn1984a].
-Put differently, the system-level efficiencies we see in language are fundamentally related to the local-level efficiencies that occur during rational, pragmatic conversation.\par
+Zipf's conception of speaker and listener effort may be connected to speakers' pragmatic reasoning in the moment.
+Under a Gricean treatment of pragmatics, speakers and listeners are assumed to follow a set of conversational maxims in which they cooperate to transfer information (Grice, 1975).
+These maxims appear to emerge from efficiency concerns, however (Horn, 1984).\par
 
@@
 The Rational Speech-act framework (RSA) is computational-level theory of pragmatic language use.
-RSA can largely be understood as a formalization of essential Gricean pragmatic principles -- interlocutors reason about one another and their shared context.
+RSA can be understood as a formalization of essential Gricean pragmatic principles -- interlocutors reason about one another and their shared context.
 For this reason, we adopt RSA as our representational framework to operationalize Gricean (rational and pragmatic) speaker-listeners [@Grice1975a].
@@
 Zipf (1949) proposed that the particular distributional properties found in natural language emerge from competing speaker and listener pressures.
-We operationalize this in equation (1) -- the efficiency of a linguistic system $\ell$ being used by speaker and listener agents $S$ and $L$ is the sum of the expected speaker and listener effort to communicate over all possible communicative events $E$.\par
+We operationalize this objective in equation (1) -- the efficiency of a linguistic system $\ell$ being used by speaker and listener agents $S$ and $L$ is the sum of the expected speaker and listener effort to communicate over all possible communicative events $E$.\par
 
@@
 \end{equation}
-This is the simply the cross-entropy between the speaker and listener joint distributions.
+This equation is the simply the cross-entropy between the speaker and listener joint distributions.
 \begin{equation}
@@
 \end{equation}
-From an information-theoretic perspective this objective is intuitive.
-Cross-entropy gives us a measure of dissimilarity between two distributions -- the average number of bits required to communicate under one distribution, given that the “true” distribution differs.
+From an information-theoretic perspective this objective is intuitive: cross-entropy gives us a measure of dissimilarity between two distributions -- the average number of bits required to communicate under one distribution, given that the “true” distribution differs.
 In our case, this is the difference between the joint distribution assumed by the speaker $P_{speaker}$ and listener $P_{listener}$.
-An "efficient" language $\ell$ used by a set of a pair of speaker-listeners will have properties which minimize this objective.
+In other words, an "efficient" language $\ell$ minimizes the distance between what speakers and listener think.
 For a complete derivation and its components we refer the reader to Section 2.1 of our supplementary materials.\par

### 2019-01-29 paper/paper.Rmd
updates after most of MF / MWu comments

@@
 ---
-title: "Interactions of rational, pragmatic agents lead to efficient language structure and use"
+title: "The interactions of rational, pragmatic agents  \n lead to efficient language structure and use"
 bibliography: library.bib
@@
 We begin with a high-level introduction to the modelling framework introducing the basic ingredients we will need to represent language as repeated reference games.
-Following this introduction we derive a simple objective function for measuring the efficiency of linguistic systems in this setting.
-We move on to two case-studies exploring the questions posed above -- examining the efficiency of ambiguity from a language design and language use perspective.\par
+We derive anobjective function for measuring the efficiency of linguistic systems in this setting.
+We move on to two case-studies exploring the questions posed above -- examining the efficiency of ambiguity from a language design and language use perspective.
+The contributions of this work is to show how the reference game framework, in combination with formal modles of communication, can be used to study questions at a range of different timescales, from language use in-the-moment to language change across historical time.\par
 
@@
 ### Reference games
-Zipf’s example of optimal speaker- and listener-languages took the form of a reference game (similar to those described by Wittgenstein).
+Zipf’s example of optimal speaker- and listener-languages took the form of a reference game.
 We adopt that formulation here, assuming these communication games as our basic unit of analysis.
-In this setting, speakers and listeners are aware of a set of objects $M$, which we will refer to as *meanings* and are knowledgeable about the set of possible signals $U$ (*utterances*) that can be used to refer to a given meaning (see Figure 1 for an example reference game).
+In this setting, speakers and listeners are aware of a set of objects $M$, which we will refer to as *meanings* and are knowledgeable about the set of possible signals $U$ (*utterances*) that can be used to refer to a given meaning (see Figure 1).
 Utterances may have different relative costs, operationalized via a prior over utterances $P(U)$.
@@
 Contexts describe different orderings of the underlying meaning prior, formalized as different conditional distributions over meanings e.g.
-$p(M|C=c_i)$.
+$p(M|C=c)$.
 Finally, we consider a set of communicative events $e \in E$ where $<u, m> = e$ is a tuple of utterance, meaning pairs.\par
@@
 A language $\ell$ defines the set of semantic mappings between utterance and meanings.
-For example, figure 1 contains four utterances $U = \{\text{blue}, \text{green}, \text{square}, \text{circle}\} \text{ and three meanings }M = \{\text{green-square}, \text{blue-square}, \text{green-circle}\}$.
+For example, Figure 1 contains four utterances $U = \{\text{blue}, \text{green}, \text{square}, \text{circle}\} \text{ and three meanings }M = \{\text{green-square}, \text{blue-square}, \text{green-circle}\}$.
 The boolean matrix describes the literal semantics of the language $\ell$.
 Note that will consider language to "contain ambiguous items" if there is at least one utterance $u \in U$ which can apply to multiple meanings (i.e.
-$|[[u_i]]| > 1$).
+$|[[u_i]]| > 1$)\footnote{We use double brackets $[[\dots]]$ to represent denotation}.
 For example, in Figure 1 both the words "blue" and "green" are ambigous so we would say that $\ell$ contains ambiguity.
@@
 The Rational Speech-act framework (RSA) is computational-level theory of pragmatic language use.
-RSA can be understood as a formalization of essential Gricean pragmatic principles -- interlocutors reason about one another and their shared context.
+RSA can be understood as a formalization of essential Gricean pragmatic principles -- agents reason about one another and their shared context.
 For this reason, we adopt RSA as our representational framework to operationalize Gricean (rational and pragmatic) speaker-listeners [@Grice1975a].
-We refer the reader to Section 1 of our supplementary materials for a more detailed description of basic RSA speaker-listener definitions, and to Goodman & Frank [-@GoodmanFrank2016a] for an overview of its appliations.
-We limit the current discussion to essential, high-level details of the framework.\par
+We refer the reader to Section 1 of our supplementary materials for a more detailed description of basic RSA speaker-listener definitions, and to Goodman & Frank [-@GoodmanFrank2016a] for an overview of its appliations.\par
 
@@
 
-Ambiguity is ubiquitous in natural language.
-The task of understanding language is marked by a frequent need to handle lexical ambiguity (words often have multiple meanings) as well as syntactic ambiguity (sentences often have multiple parses), in addition to many other instances of ambiguity [@WasowPerforsBeaver2005a].
-Chomsky [-@Chomsky2002a] famously claimed that the presence of ambiguity in natural language provides evidence that language has not been optimized for communication, stating “If you want to make sure that we never misunderstand one another, for that purpose language is not well designed, because you have such properties as ambiguity.” Note that this analysis is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$ we described earlier and define in terms of equation 2 in section 2.2 for our supplementary materials.\par
+The task of understanding language is marked by a frequent need to handle various forms of ambiguity; lexical, syntactic, among others [@WasowPerforsBeaver2005a].
+The ubiquity of this property, however, has been argued to provide evidence that langauges are not been optimized for communication\footnote{Chomsky (2002) famously claimed "If you want to make sure that we never misunderstand one another, for that purpose language is not well designed, because you have such properties as ambiguity." Note that this analysis is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$ we described earlier and define in terms of equation 2 in SI 2.2 .} [@Chomsky2002a].
 
@@
 Say we have two objects in the world $m_1$ and $m_2$, and two languages $\ell_1$ and $\ell_2$.
-In language $\ell_1$, the low-cost utterance $u_1$ can be used to refer to both $m1$ and $m_2$ (i.e.
+In language $\ell_1$, the low-cost utterance $u_1$ can be used to refer to both $m_1$ and $m_2$ (i.e.
 $[[u_1]]_{\ell_1} = \{m_1, m_2 \}$).
@@
 
-The argument outlined by Piantados et al.
-[-@Piantadosi2011a] is compatible with our current reference game setting with an addition -- context.
+We adpot the framework above, and (following Piantadosi et al.
+2011), add \textit{context} ($C$).
 The authors argue that mapping two meanings $m_1$ and $m_2$ to a single utterance $u_1$ is useful when they can be disambiguated *in context*.
-In our case we consider a context $c_i \in C$ to specify a unique ordering of the need probabilities.
-That is, $p(m|c_1) \neq p(m|c_2)$, when there are two contexts $|C| = \{c_1, c_2\}$.\par
+In our case we consider a context $c_i \in C$ to specify a permutation of the meaning probabilities.
+That is, $p(m|c_1) \neq p(m|c_2)$, when there are two contexts $|C| = \{c_1, c_2\}$.This in turn, leads to an update to our linguistic efficiency objective as we now would like to consider the average speaker-listener effort over all contextualized communicative events (see section 3 SI).\par
 
-This in turn, leads to an update to our linguistic efficiency objective as we now would like to consider the average speaker-listener effort over all contextualized communicative events (section 3 of supplementary materials).\par
+We proceed by generating languages with different amounts of contextual support (varying the size of $|C|$).
+We search the space of languages, asessing if the ones which minimize our objective contain ambiguity.
+If context leads to more efficient commnication, then optimal languages should be more likley to be ambiguous as the amount of context increases.\par
 
@@
 We conduct $N=2000$ simulations.
-For each simulation we enumerate the set of *valid* languages in which $|U|=|M|=4$.
-Note that a language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment) and each form maps to at least one meaning (every column has some non-zero assignment).
+For each simulation we enumerate the set of *valid* languages in which $|U|=|M|=4$ (recall that $U$ is our set of utterances and $M$ our set of meanings).
+Note that a language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment) and each form maps to at least one meaning (every column has some non-zero assignment)\footnote{For the current set of simulations we consider languages in which there are four or five true values -- in a "four" language ambiguity is not possible, but in a "five" language there are ambiguous mappings.}.
 For a given simulation the goal is to find the language $\ell^*$ which minimizes our cross-entropy objective and then check to see if that language contains ambiguity.\par
@@
 Zipf [-@Zipf1949a] proposed that the optimal speaker language $\ell_{speaker}^*$ would only optimize speaker effort and the optimal listener language $\ell_{listener}^*$ would only optimize listener effort.
-We define these objectives using the first and second half of equation 1 and with greater detail in the supplementary materials (Section 2.2).\par
+We define these objectives using the first and second half of equation 1 (see SI section 2.2.
+for more detail).\par
 
-## Hypotheses
-
-Piantadosi’s claim leads to the following set of hypotheses: (1) The probability that an optimal language contains ambiguity under our Cross-Entropy objective should increase with the number of contexts (as $|C|\rightarrow\inf$).
-(2) The optimal speaker language under our “Speaker-only” objective should always contain ambiguity (maximally re-using the lowest-cost utterance).
-(3) The optimal listener language should never contain ambiguous material.
-
-## Results
+## Results and Discussion
 
@@
 
-## Summary
+We explored the degree to which ambiguity improved efficiency during communication between rational, pragmatic agents.
+We showed that as contextual information increased, the optimal languages under our Cross-Entropy objective were more likely to have ambiguous lexical items.
+Importantly, we did not see this effect under alternative objectives which only emphasized speaker- or listener effort alone.\par
 
-Piantadosi et al.
-[-@Piantadosi2011a] argued that it is useful to re-use low-cost linguistic forms for multiple meanings when they can be disambiguated in context.
-We explored the degree to which this this property emerged during communication between rational, pragmatic agents, showing that as contextual information increased, the optimal languages under our Cross-Entropy objective were more likely to have ambiguous lexical items.
-Importantly, we did not see this effect under alternative objectives which only emphasized speaker- or listener effort alone.\par
+While our results indicate that ambiguity is an efficient property of contextualized language these simulations assumed that speaker-listener agents could always disambiguate items in context.
+That is, in our four conditions both agents had perfect knowledge of the relevant conditional distributions ($P(M|C)$).
+This assumption may be too strong for describing much of day-to-day communication -- we seldom begin a conversation with perfect knowledge of the current context of a conversation.
+We explore a case in which the listener has imperfect knowledge of context at the start of the conversation in Simulation 2.\par
 
@@
 
-Our first experiment made a fairly direct test of the communicative function of ambiguity theory.
-Sampling meaning, utterance, and context priors, we examined properties of the language $\ell^*$ which minimized our objective.
-Results indicated that ambiguity is an efficient property when context is informative.
-This simulation, however, assumed that our speaker and listener agents could always disambiguate items context.
-That is, in our four conditions both agents had perfect knowledge of the relevant conditional distributions ($P(M|C_i)$).\par
-
-This may be a strong assumption for describing much of day-to-day communication -- we seldom begin a conversation with perfect knowledge of the current context of a conversation.
-Under Piantadosi's framework, if context is not informative with respect to meaning at time position $i$ in a discourse, but is informative at position $i+k$, then the speaker should avoid using ambigous material at position $i$, only using it at a future point, $k$ steps later in the conversation.
+We shift time-scales in Simulation 2 -- instead of examining the objective landscape of a space of languages, we now consider a single language that contains ambiguity and examine how that property is used over a discourse.
+If context is not informative with respect to meaning early in a discourse (position $i$), but is at later positions at position (position $i+k$), then the speaker should avoid using ambigous material early $i$, only using it $k$ steps later in the conversation.
 In this toy example we can consider “context” as analogous to a “topic” of conversation.
-Intuitvely, we can imagine a scenario in which a reader is beginning a newspaper article.
+Intuitively, we can imagine a scenario in which a reader is beginning a newspaper article.
 While they may have some knowledge about the article’s topic (perhaps from the title), they may not have complete knowledge of its contents, including the persons or events involved.
@@
 
-We consider such a scenario in simulation two -- analyzing efficient *use* of ambiguity over discourse in a reference game setting.
-Unlike our first simulation, we now assume a single language $\ell$, which contains both ambiguous and unambiguous utterances.
+We consider consider a single language $\ell$, which contains both ambiguous and unambiguous utterances.
 The ambiguous utterances are less costly than the unambiguous utterances.
@@
 Over the course of a discourse, $D$ the listener tries to infer both the current topic, $c_{current}$, as well as the particular meaning $m$ of a given utterance $u$.
-That is we consider a modified listener model $L(m, c|u,D;\ell)$ and speaker model $S(u|m,c,D;\ell)$.
+That is, we consider updated listener and speakers ($L(m, c|u,D;\ell)$, $S(u|m,c,D;\ell)$).
 Note that both the speaker and listener can track the history of previous utterances $D$.
-Importantly, a listener can attempt to infer the current topic of conversation $c_{current}$ using the discourse history $D$.
-Section 3 of the supplementary materials provides more detail on the speaker and listener models.\par
+Importantly, an agent can attempt to infer the current topic of conversation $c_{current}$ using the discourse history $D$ (see SI section 3 for more detail).\par
 
-```{r plot-optimal-use, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Left facet (A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position.
-As the discourse proceeds, only our Full Pragmatic speaker is more likely to use an ambiguous term.
-Plot B tracks the Cross-Entropy objective under our three speaker models.
-The objective declines in all cases as the listener infers the correct topic distribution.
-However the pragmatic speaker arrives as the lowest value, do to reduction in both speaker and listener effort.
-Plot C tracks speaker effort across the three models.
-Only the Full Pragmatic speaker shows declines in effort across the discourse as they increasinlgy use ambiguous items as the listener becomes more sure of the topic of conversation."}
+```{r plot-optimal-use, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=2.5, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "(A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position.
+(B) shows speaker effort across the three models.
+(C) shows the Cross-Entropy objective under our three speaker models."}
 img <- png::readPNG("figs/discourse_grid_plot.png")
@@
 We conduct $N=600$ simulations, generating discourses of length $|D|=30$ utterances with three different speaker models ($n=200$ each).
-We consider a single language $\ell$ with $|U|=6$ and $|M|=4$ specified by the boolean matrix:
-
-\begin{center}
-\begin{tabular}{ c | c c c c} 
-& $m_1$ & $m_2$ & $m_3$ & $m_4$ \\
-\hline
-$u_1$ & 1 & 0 & 0 & 0 \\
-$u_2$ & 0 & 1 & 0 & 0 \\
-$u_3$ & 0 & 0 & 1 & 0 \\
-$u_4$ & 0 & 0 & 0 & 1 \\
-$u_5$ & 1 & 1 & 0 & 0 \\
-$u_6$ & 0 & 0 & 1 & 1 \\
-\end{tabular}
-\end{center}
-
-We assume that $p(u_5) = p(u_6) > p(u_{1})=\dots=p(u_{4})$. That is, the two ambiguous utterances ($u_5$ and $u_6$) are less costly than the non-ambiguous utterances.\par
+We consider a single language $\ell$\footnote{See SI for the matrix notation of this langauge.} with $|U|=6$ and $|M|=4$ in which two of the utterances $u_5$ and $u_6$ are ambiguous and that $p(u_5) = p(u_6) > p(u_{1})=\dots=p(u_{4})$.
+That is, the two ambiguous utterances ($u_5$ and $u_6$) are less costly than the non-ambiguous utterances.
+(Note that use of this particular language is not essential -- the results are broadly generalizable languages that contain ambiguity.)\par
 
@@
 The first, a \textit{Partial pragmatics} baseline describes a speaker who reasons about a listener, but assumes they have no access to discourse history.
-The second is a \textit{No pragmatics} speaker who does not consider a listener at all, but produces utterances according to the underlying langauge semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$).
-We provide detailed descriptions of each model in the supplementary materials Section 3.\par
+The second is a \textit{No pragmatics} speaker who does not consider a listener at all, but produces utterances according to the underlying langauge semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$) (see SI section 3 for details).\par
 
 ## Hypotheses
-Under Piantadosi's claim it is efficient to use low-cost, ambiguous material when context is disambiguating.
-In the current setting, this corresponds to speaker behavior in which ambiguouous forms are avoided early in the discourse, but prefered later in the discourse once the listener is confident of the topic.
+
+We assess whether it is efficient to use low-cost, ambiguous material when context is disambiguating.
+In the current setting, this corresponds to speaker behavior in which ambiguous forms are avoided early in the discourse, but prefered later in the discourse once the listener is confident of the topic.
 We should expect this to be reflected in our \textit{Full pragmatics} model.
@@
 
-## Results
+## Results and Discussion
 
-Figure 2, plot A shows the empirical probability that a speaker uses an ambiguous utterance as a function of discourse position.
-Note that the \text{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse, the \text{Partial pragmatic} baseline avoids ambiguous utterances entirely (reasonining that the utterances cannout be disambiguated), but the \text{Full pragmatic} model avoids ambiguous material at first, but employing them gradually increasingly as the discourse proceeds.
-Figure 2, plot B tracks the speaker-listener Cross-Entropy objective.
+Figure 2, (A) shows the empirical probability that a speaker uses an ambiguous utterance as a function of discourse position.
+The \text{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse and the \text{Partial pragmatic} baseline avoids ambiguous utterances entirely (reasoning that the utterances cannot be disambiguated).
+But, the \text{Full pragmatic} model avoids ambiguous material at first, but employing them gradually increasingly as the discourse proceeds.
+(B) shows the speaker-listener Cross-Entropy objective.
 Note that the objective decreases for all three models.
@@
 Of particular importance is the value of the objective at the end of the discourse (position $i=30$).
-The difference in the Cross-Entropy between the \text{Full Pragmatic} and \text{Partial Pragmatic} models at the end of the discourse is driven by the reduction in speaker costs described in Figure 2, plot C.
+The difference in the Cross-Entropy between the \text{Full Pragmatic} and \text{Partial Pragmatic} models at the end of the discourse is driven by the reduction in speaker costs described in (C).
 That is, while speaker effort remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines, effort declines for the \text{Full Pragmatic} model as they are increasingly able to use ambiguous material later in the discourse.\par
 
-## Summary
-
-These simulations explored the efficient *use* of ambiguity over a discourse, in an alternative instantiation of the argument outlined by Piantadosi et al.
-[-@Piantadosi2011a].
-In the first set of simulations we varied the set of languages, examining the properties of languages which minimized our cross-entropy objective under different context conditions.
-In the current simulation we assumed that the listener did not know the current context (topic) \textit{a priori}, but could infer it from the discourse.
+In Simulation 2 we explored how ambiguity could be employed efficiently during discourse.
+We assumed that the listener did not know the current context (topic) \textit{a priori}, but could infer it from the discourse history.
 We examined speaker production behavior, tracking use of low-cost, but ambiguous material over the discourse.
-We found that early in the discourse our \textit{Full pragmatic} speaker avoided using ambiguous material, only using the material later when the context was known to the listener.
-This is consistent with Piantaodi's claim -- it is useful to use low-cost material, but only when context is disambiguating.
+Early in the discourse the \textit{Full pragmatic} speaker avoided using ambiguous material, only using the material later when the context was known to the listener -- it is useful to use low-cost material, but only when context is disambiguating.
 
@@
 Zipf [-@Zipf1949a] proposed that the asymmetry between speaker and listener costs gives rise to a range of properties at the level of the lexicon.
-Subsequent functionalist approaches to language science have provided additional evidence that aspects of language structure and language use appear to be optimized for efficiency.
-In Zipf’s famous example of an optimal speaker and listener language he appears indicate that such optimization occurs in in-the-moment, local interactions between speakers and listeners.
-We explore this idea in our current project, following theoretical work by Horn [-@Horn1984a] who proposed the speaker-listener pressures identified by Zipf can also be interpreted in terms of the maxims governing rational conversation proposed by [@Grice1975a].
-To so, we explored the interactions of rational pragmatic agents as a framework for understanding efficient language structure and use.
-We focused on an argument on the communicative function of ambiguity put forth by Piantadosi et al.
-[-@Piantadosi2011a].
+We explored the interactions of rational pragmatic agents as a framework for understanding efficient language structure and use.
+We focused on an argument on the communicative function of ambiguity [@Piantadosi2011a].
 We derive a novel speaker-listener cross-entropy objective for measuring the efficiency of a language in a reference game setting showing that optimal languages are more likely to contain ambiguous material when context is informative.
-Further we show that rational pragmatic agents will *use* ambiguous material efficiently, only when such use is supported by context.
-The intent is that this project provides a demonstration for the types of functionalist theories that can be explored within this framework.
-Among other language properties open to this type of analysis are work examining semantic typology [@RegierKempKay2015a], the emergence of compositionality [@SmithKirbyBrighton2003a] and convention formation, as well as efficient examples of language use including the efficient use of reduction and redundancy [@LevyJaeger2007a; @GenzelCharniak2002a].
+Further we show that rational pragmatic agents will *use* ambiguous material efficiently, only when such use is supported by context.\par
+
+The framework employed here -- the analysis of the dynamics of repeated reference games between rational pragmatic agents -- provides a tool for the analysis of functionalist theories, allowing linkage across timescales of analysis from conversation [@LevyJaeger2007a; @GenzelCharniak2002a] to typology [@RegierKempKay2015a].
+In theory, this framework provides extensive opportunity to vary properties of the languages and agents, however we have focused on only a small subset here.
+For example, examining how increasing or decreasing \textit{rationality} (instantiated through the RSA $\alpha$ parameter) or \textit{pragmatics} (instantiated through the RSA recursion depth) impacts these results is a clear avenue for future work.
+An important limitation of the current work is the broad assumption that language can be described through such reference games.
+The degree to which properties of non-referential language can be captured in such a framework remains an important target for future work.
+
+\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering All experiments, data, and analysis code are available in the public repository for the project: (link will be available upon acceptance)}} \vspace{1em}
+\noindent
 

### 2019-01-29 paper/paper.Rmd
updates to plots and add figs to repo for reproducibilityy

@@
 Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only consideration."}
-
 img <- png::readPNG("figs/fig1.png")
@@
 (C) shows the Cross-Entropy objective under our three speaker models."}
+
 img <- png::readPNG("figs/discourse_grid_plot.png")

### 2019-01-29 ambiguity/simulation-runners/context-ambiguity-plots.Rmd
move simluation run and plotting to simulation-runners. update README.

---
title: "ambiguity_context_analysis"
author: "Ben"
date: "11/12/2018"
output: html_document
---

```{r libraries}
rm(list = ls())  # clear workspace
library(cowplot)
library(tidyverse)
library(ggthemes)
```

```{r read-files}
fp <- '/Users/benpeloquin/Data/zipf_principles/ambiguity/runs-20190121/uniform_uniform_1_1_500.csv'
df <- read.csv(fp)
```

## Claim 1 - the more contexts the more useful ambiguity is.
```{r plot1-a}
p1 <- df %>%
  select(sim_id, idxs, n_contexts, speaker_k,
         utterance_order_fn_name, is_min_ce,
         is_min_listener_entropy, is_min_speaker_entropy, 
         contains_ambiguity) %>%
  gather(objective, val, c(is_min_ce, is_min_listener_entropy, is_min_speaker_entropy)) %>%
  # filter(objective=='base_speaker_listener_ce') %>%
  mutate(val=as.logical(val),
         contains_ambiguity=as.logical(contains_ambiguity)) %>%
  filter(val) %>%  # only look at 'optimal languages'
  group_by(objective, n_contexts) %>%
  summarise(n=n(),  # number of optimal languages (there can be ties)
            prop_ambiguous=mean(contains_ambiguity),
            # Binomial CIs
            ci_high = prop_ambiguous + qnorm(0.975)*sqrt((prop_ambiguous*(1 - prop_ambiguous))/n),
            ci_low=prop_ambiguous - qnorm(0.975)*(sqrt((prop_ambiguous*(1-prop_ambiguous))/n))) %>%
  ungroup() %>%
  # Normal approximation isn't great when we have proportions so 
  # close to 1. so just rounding down slightly. This is a little
  # loosey-goosey, but does not impact the overall narrative
  # in a significant way
  # mutate(ci_high=ifelse(ci_high > 1, 1, ci_high)) %>%
  mutate(objective=ifelse(objective=='is_min_ce', 'Cross-entropy',
                          ifelse(objective=='is_min_listener_entropy', 'Optimal Listener', 'Optimal Speaker'))) %>%
  rename(Objective=objective) %>%
  ggplot(aes(x=n_contexts, y=prop_ambiguous, col=Objective)) +
    geom_line(size=1.5) +
    geom_errorbar(aes(ymin=ci_low, ymax=ci_high), width=0.05, size=1, alpha=0.7) +
    ylab("Percent optimal languages \n containing ambiguity") +
    xlab("Number of contexts (|C|)") +
    theme_few()
    # In case we want to embed legend in the plot, but not really necessary...
    # theme(legend.position = c(0.8, 0.4),
    #       legend.background = element_rect(colour = 'grey', fill = 'white', linetype='solid'))
```

```{r plot1-b}
p2 <- df %>%
  select(sim_id, idxs, n_contexts, is_min_ce, ce, base_speaker_listener_ce, contains_ambiguity) %>%
  filter(as.logical(is_min_ce)) %>%
  mutate(ce_minus_base=ce-base_speaker_listener_ce) %>%
  ungroup %>%
  group_by(n_contexts) %>%
  summarise(n=n(),
            diff=mean(ce_minus_base),
            ci_high=diff + 1.96*sd(ce_minus_base)/sqrt(n),
            ci_low=diff - 1.96*sd(ce_minus_base)/sqrt(n)) %>%
  ggplot(aes(x=n_contexts, y=diff)) +
    geom_bar(stat='identity', position='dodge', alpha=0.8) +
    geom_errorbar(aes(ymin=ci_low, ymax=ci_high), width=0.1, col='black', alpha=0.8) +
    theme_few() +
    ylab("Difference in cross-entropy\n aware vs unaware of context") +
    xlab("Number of contexts (|C|)")
```

Saves single fig plot.
```{r saving}
# plot_grid(p1, p2, labels = c("A", "B"), align = "h")
# ggsave(file.path(getwd(), "../../paper/figs/fig1.png"), width=10, height=3)
ggsave(file.path(getwd(), "../../paper/figs/fig1.png"), plot=p1, width=7, height=3)
```

Save grid plot
```{r saving}
# plot_grid(p1, p2, labels = c("A", "B"), align = "h")
# ggsave(file.path(getwd(), "../../paper/figs/fig1.png"), width=10, height=3)
```

### 2019-01-29 ambiguity/simulation-runners/discourse-ambiguity-plots.Rmd
move simluation run and plotting to simulation-runners. update README.

---
title: "Discourse-ambiguity-plots"
author: "Ben"
date: "1/24/2019"
output: html_document
---

Plotting functionality for Simulation 2 plots in "The interactions of rational, pragmatic agentslead to efficient language structure and use."

```{r libraries, message=FALSE, warning=FALSE}
rm(list = ls())  # clear workspace

library(cowplot)
library(doParallel)
library(entropy)
library(ggthemes)
library(tidyverse)
library(zoo)

source(file.path(getwd(), "../utils.R"))
```

```{r load-data}
# Note you should point to the correct locations on your local machine
# reach out to bpeloqui@stanfor.edu for this data.
fpPreprocessed <- "/Users/benpeloquin/Data/zipf_principles/discourse_ambiguity/df_preprocessed_20190124.csv"
fpSims <- "/Users/benpeloquin/Data/zipf_principles/discourse_ambiguity/df_simsCmbd_20190124.csv"

df_preProcessedSims <- read.csv(fpPreprocessed)
df_simCmbd <- read.csv(fpSims)
```

Empirical probability of ambiguous utterance.
```{r plot-ambiguity-use}
# Get legend
p1_temp <- df_preProcessedSims %>%
  rename(model=resultType) %>%
  mutate(model=ifelse(model=='baseline', "Non-Pragmatic",
                      ifelse(model=="discourseUnaware", "Partial Pragmatic", "Full Pragmatic"))) %>%
  mutate(Model=factor(model, level=c("Non-Pragmatic", "Partial Pragmatic", "Full Pragmatic"))) %>%
  ggplot(aes(x=utteranceNum, y=probAmb, col=Model)) +
    geom_line(aes(col=Model)) +
    geom_errorbar(aes(ymin=probAmb-2*stdAmb, ymax=probAmb+2*stdAmb), alpha=0.5, width=0.4, col='black') +
    geom_point(size=2, alpha=0.8) +
    ylab("Probability \nof ambiguous utterance") +
    xlab("Utterance position in discourse") +
    theme_few() +
    theme(legend.background = element_rect(colour = 'grey', fill = 'white', linetype='solid'))
legend <- get_legend(p1_temp)

p1 <- df_preProcessedSims %>%
  rename(model=resultType) %>%
  mutate(model=ifelse(model=='baseline', "Non-Pragmatic",
                      ifelse(model=="discourseUnaware", "Partial Pragmatic", "Full Pragmatic"))) %>%
  mutate(Model=factor(model, level=c("Non-Pragmatic", "Partial Pragmatic", "Full Pragmatic"))) %>%
  ggplot(aes(x=utteranceNum, y=probAmb, col=Model)) +
    geom_line(aes(col=Model)) +
    geom_errorbar(aes(ymin=probAmb-2*stdAmb, ymax=probAmb+2*stdAmb), alpha=0.5, width=0.4, col='black') +
    geom_point(size=1.5, alpha=0.8) +
    ylab("Proportion \n ambiguous") +
    xlab("Utterance position in discourse") +
    theme_few() +
    theme(legend.position="None",
          axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          plot.margin = unit(c(0, 0, 0, 0), "cm"))
```

Cross-entropy measure
```{r cross-entropy}
p2 <- df_simCmbd %>%
  rename(model=speakerType) %>%
  mutate(model=ifelse(model=='baseline', "Non-Pragmatic",
                      ifelse(model=="discourseUnaware", "Partial Pragmatic", "Full Pragmatic"))) %>%
  mutate(Model=factor(model, level=c("Non-Pragmatic", "Partial Pragmatic", "Full Pragmatic"))) %>%
  mutate(CE=-CE) %>%
  group_by(Model, utteranceNum) %>%
  summarise(n=n(),
            meanCE= mean(CE),
            ci_high=meanCE + qnorm(.975)*(sd(CE)/sqrt(n)),
            ci_low=meanCE - qnorm(.975)*(sd(CE)/sqrt(n))) %>%
  ggplot(aes(x=utteranceNum, y=meanCE, col=Model)) +
    geom_line() +
    geom_errorbar(aes(ymin=ci_low, ymax=ci_high), alpha=0.5, width=0.4, col='black') +
    geom_point(size=1.5, alpha=0.8) +
    theme_few() +
    ylab('Average S/L cross-entropy') +
    xlab("Utterance position in discourse") +
    theme_few() +
    theme(legend.position="None",
          plot.margin = unit(c(0, 0, 0, 0.5), "cm"))
```

Speaker cost.
```{r speaker-costs}
p3 <- df_simCmbd %>%
  rename(model=speakerType) %>%
  mutate(model=ifelse(model=='baseline', "Non-Pragmatic",
                      ifelse(model=="discourseUnaware", "Partial Pragmatic", "Full Pragmatic"))) %>%
  mutate(Model=factor(model, level=c("Non-Pragmatic", "Partial Pragmatic", "Full Pragmatic"))) %>%
  mutate(speakerCosts=-speakerCosts) %>%
  group_by(Model, utteranceNum) %>%
  summarise(n=n(),
            mean_speakerCosts=mean(speakerCosts),
            ci_high=mean_speakerCosts + qnorm(.975)*(sd(speakerCosts)/sqrt(n)),
            ci_low=mean_speakerCosts - qnorm(.975)*(sd(speakerCosts)/sqrt(n))) %>%
  ggplot(aes(x=utteranceNum, y=mean_speakerCosts, col=Model)) +
    geom_line() +
    geom_errorbar(aes(ymin=ci_low, ymax=ci_high), alpha=0.5, width=0.4, col='black') +
    geom_point(size=1.5, alpha=0.8) +
    theme_few() +
    ylab('Average \n Speaker cost') +
    xlab("Utterance position in discourse") +
    theme_few() +
    theme(legend.position="None",
          plot.margin=unit(c(0, 0, 0, 0), "cm"))
```

Utterance position 3 delta in CE objective
```{r}
p_embed <- df_simCmbd %>%
  rename(model=speakerType) %>%
  mutate(model=ifelse(model=='baseline', "Non-Pragmatic",
                      ifelse(model=="discourseUnaware", "Partial Pragmatic", "Full Pragmatic"))) %>%
  mutate(Model=factor(model, level=c("Non-Pragmatic", "Partial Pragmatic", "Full Pragmatic"))) %>%
  mutate(CE=-CE) %>%
  group_by(Model, utteranceNum) %>%
  summarise(n=n(),
            meanCE=mean(CE),
            ci_high=meanCE + qnorm(.975)*(sd(CE)/sqrt(n)),
            ci_low=meanCE - qnorm(.975)*(sd(CE)/sqrt(n))) %>%
  filter(utteranceNum == 29) %>%
  ggplot(aes(x=Model, y=meanCE, fill=Model)) +
    geom_bar(stat='identity', position=position_dodge()) +
    # geom_line() +
    geom_errorbar(aes(ymin=ci_low, ymax=ci_high), position=position_dodge(), alpha=0.8, width=0.1) +
    ylab('Average S/L cross-entropy') +
    xlab("CE at last utterance in discourse") +
    theme_few() +
    theme(legend.position="None")
```

Paper figure.
```{r}
# Embed legend
p2a <- ggdraw() +
  draw_plot(p2) +
  draw_plot(legend, 0.55, 0.6, 0.4, 0.4)
p1a <- plot_grid(p1, NULL, p3, align='hv', ncol=1, labels=c("A", "", "B"), rel_heights=c(1, -0.14, 1))
# Embed legend
p_grid <- plot_grid(p1a, p2a, ncol=2, labels=c('', 'C'))
ggsave(file.path(getwd(), "../../paper/figs/discourse_grid_plot.png"), plot=p_grid, width=7.8, height=3.2)
```

Additional plots

Listener belief update.
```{r}
p4 <- df_preProcessedSims %>%
  mutate(ci_hi=avgT1+qnorm(0.975)*(sdT1/sqrt(n)),
         ci_lo=avgT1-qnorm(0.975)*(sdT1/sqrt(n))) %>%
  rename(model=resultType) %>%
  ggplot(aes(x=utteranceNum, y=avgT1)) +
    geom_smooth(method='loess', alpha=0.5, col='red') +
    geom_errorbar(aes(ymin=ci_lo, ymax=ci_hi), width=0.4, alpha=0.6) +
    geom_point(size=2, alpha=0.75) +
    ylab("Probability of\n current topic P(C=c|D)") +
    xlab("Utterance position in discourse") +
    theme_few()  +
    facet_grid(~model)
```

### 2019-01-29 ambiguity/simulation-runners/discourse-ambiguity.Rmd
move simluation run and plotting to simulation-runners. update README.

---
title: "Discourse-ambiguity"
author: "Ben"
date: "11/9/2018"
output: html_document
---

Simulation 2 run code for "The interactions of rational, pragmatic agentslead to efficient language structure and use." Note that we make use of the `rwebppl` package to interface with webppl. Modeling and discourse generation are contained in the "ambiguity/webppl" folder.

```{r libraries, message=FALSE, warning=FALSE}
rm(list = ls())  # clear workspace

library(cowplot)
library(doParallel)
library(entropy)
library(ggthemes)
library(tidyverse)
library(zoo)

source(file.path(getwd(), "../utils.R"))
```

Load model.
```{r getModel}
fPath <- file.path(getwd(), '../webppl/discourse_ambiguity_model.wppl')
m <- getModelFile(fPath)
```

Setup.
```{r runFn}
# Single run function
runAmbiguity <- createRunFn(m)
runAmbiguityFn <- function(i, targetDistr, nUtterances, resultType, alpha, recursionLevel) {
  dTemp <- data.frame(targetDistr=targetDistr, 
                      nUtterances=nUtterances, 
                      resultType=resultType, 
                      alpha=alpha, 
                      recursionLevel=recursionLevel)
  df <- runAmbiguity(dTemp) %>%
    mutate(runNum=i, targetDistr=targetDistr, alpha=alpha, resultType=resultType, recursionLevel=recursionLevel)
  df
}

# Multiple runs function
runAmibguityCorpus <- function(topic, alpha, resultType, recursionLevel, nUtterances, nSims) {
  # Parallelization setup
  no_cores <- detectCores() - 1
  cl <- makeCluster(no_cores, type='FORK')
  registerDoParallel(cl)

  # Run sims
  ptm <- proc.time()
  sims <- foreach(i=seq(1, nSims), .packages=c('dplyr', 'rwebppl'), .combine=rbind) %dopar% 
    runAmbiguityFn(i,
                   targetDistr=topic,
                   nUtterances=nUtterances,
                   resultType=resultType,
                   alpha=alpha,
                   recursionLevel=recursionLevel)
  stopCluster(cl)
  etm <- proc.time() - ptm

  # Preprocess for entropy calculations.
  binSize <- 1
  df_binned <- addUtteranceBins(sims, nUtterances=nUtterances, binSize=binSize)
  df_filled <- fillUtteranceProportions(df_binned, binSize) %>%
    mutate(recursionLevel=recursionLevel)
  list(sims, df_filled)
}
```

# Run sims
```{r}
# Note to use cached results you must hace access to the data. Data is available on request bpeloqui@stanford.edu.
fpPreprocssed <- "/Users/benpeloquin/Data/zipf_principles/discourse_ambiguity/df_preprocessed_20190124.csv"
fpSims <- "/Users/benpeloquin/Data/zipf_principles/discourse_ambiguity/df_simsCmbd_20190124.csv"

topic <- 'T1'
nUtterances <- 30
alpha=8
nSims <- 100

simsBaseline <- runAmibguityCorpus(
  topic=topic,
  alpha=8,
  resultType='baseline',
  recursionLevel=1,
  nUtterances=nUtterances,
  nSims=nSims)

simsUnaware <- runAmibguityCorpus(
  topic=topic,
  alpha=8,
  resultType='discourseUnaware',
  recursionLevel=1,
  nUtterances=nUtterances,
  nSims=nSims)

simsAware <- runAmibguityCorpus(
  topic=topic,
  alpha=8,
  resultType='discourseAware',
  recursionLevel=1,
  nUtterances=nUtterances,
  nSims=nSims)

simsAmbiguityCorpusUnaware <- simsUnaware[[1]]
dfAmbiguityCorpusUnaware <- simsUnaware[[2]]
simsAmbiguityCorpusAware <- simsAware[[1]]
dfAmbiguityCorpusAware <- simsAware[[2]]
simsAmbiguityCorpusBaseline <- simsBaseline[[1]]
dfAmbiguityCorpusBaseline <- simsBaseline[[2]]

# Combine data
simsCmbd <- rbind(simsAmbiguityCorpusUnaware,
                  simsAmbiguityCorpusAware,
                  simsAmbiguityCorpusBaseline)
# Write raw sims data
write.csv(simsCmbd, fpSims)

dfAmbiguityCorpusCmbd <- rbind(dfAmbiguityCorpusUnaware,
                               dfAmbiguityCorpusAware,
                               dfAmbiguityCorpusBaseline)

# Pre-process
preProcessedSims <- simsCmbd %>% 
  group_by(resultType, alpha) %>%
  mutate(isAmbiguous=(utterance=='x' | utterance=='y'),
         isX=utterance=='x',
         isA=utterance=='a',
         isB=utterance=='b') %>%
  group_by(resultType, alpha, utteranceNum) %>%
  summarise(n=n(),
            numAmb=sum(isAmbiguous),
            probAmb=numAmb/n,
            stdAmb=sqrt(probAmb*(1-probAmb)/n),
            numX=sum(isX),
            probX=numX/n,
            stdX=sqrt(probX*(1-probX)/n),
            numA=sum(isA),
            probA=numA/n,
            stdA=sqrt(probA*(1-probA)/n),
            numB=sum(isB),
            probB=numB/n,
            stdB=sqrt(probB*(1-probB)/n),
            avgT1=mean(T1),
            sdT1=sd(T1),
            avgT2=mean(T2),
            sdT2=sd(T2),
            avgT3=mean(T3),
            sdT3=sd(T3),
            avgT4=mean(T4),
            sdT4=sd(T4))
# Cache data
write.csv(preProcessedSims, fpPreprocssed)
df_preProcessedSims <- preProcessedSims
```

### 2019-01-29 paper/paper.tex
add .tex helpers to knitting

% Template for Cogsci submission with R Markdown

% Stuff changed from original Markdown PLOS Template
\documentclass[10pt, letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}

% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{apacite}

% KM added 1/4/18 to allow control of blind submission

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% % Text layout
% \topmargin 0.0cm
% \oddsidemargin 0.5cm
% \evensidemargin 0.5cm
% \textwidth 16cm
% \textheight 21cm

\title{The interactions of rational, pragmatic agents\\
lead to efficient language structure and use}

\author{{\large \bf Benjamin N. Peloquin} \\ \texttt{bpeloqui@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Noah D. Goodman} \\ \texttt{ngoodman@stanford.edu} \\ Department of Computer Science \\ Stanford University \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@university.edu} \\ Department of Psychology \\ Stanford University}

\begin{document}

\maketitle

\begin{abstract}
Languages display a diverse set of distributional regularities such as
the relation between a word's frequency and rank in a corpus, the
distribution of dependency lengths, or the presence of lexical
properties such as ambiguity. We discuss a framework for studying how
these properties emerge from in-the-moment interactions of rational,
pragmatic speakers and listeners. Our work takes Zipfian notions of
lexicon-level efficiency as a starting point, connecting these ideas to
Gricean notions of conversational-level efficiency. To do so, we derive
an objective function for measuring the communicative efficiency of
linguistic systems and then examining the behavior of this objective in
a series of simulations focusing on the communicative function of
ambiguity in language. These simulations suggest that rational pragmatic
agents will produce communicatively efficient systems and that
interactions between such agents provide a framework for examining
efficient properties of language more broadly.

\textbf{Keywords:}
Communicative efficiency, Rational Speech Act theory, computational
modeling, information theory, agent-based simulation
\end{abstract}

\section{Introduction}\label{introduction}

Why do languages look the way they do? Zipf (1949) proposed that
distributional properties found in natural language were evidence of
speaker-listener effort minimization. In his own words, ``we are arguing
that people do in fact act with a maximum economy of effort, and that
therefore in the process of speaking-listening they will automatically
minimize the expenditure of effort.'' Evidence for this claim has been
largely derived at the level of the lexicon. Zipf argued that the
particular relationship between a word's frequency and its rank, length,
and denotation size could be explained as an emergent property of
speaker-listener effort minimization. \par

Zipf articulated what is now considered a \emph{functionalist} approach
to language science -- analyzing language structure and use in terms of
efficiency. Such an approach might reframe our opening question as
follows -- how does having \textbf{property x} make using language
\(\ell\) more or less useful for communication? This reframing provides
an opportunity to study both language \emph{structure} (as Zipf
primarily did) or \emph{use} in terms of efficiency. For example, Regier
et al. (2015) showed that languages appear to organize semantic domains
(form-meaning mappings) in maximally efficient ways. Likewise,
Piantadosi et al. (2011) argued that lexical ambiguity is an efficient
property of any communication system when communication is
contextualized.\par

Zipf's original work and subsequent functionalist projects describe a
varied set of linguistic properties and human behaviors. Common to many
is a characterization of a fundamental effort-asymmetry underlying
everyday communication. Simply put, what is ``hard'' for a speaker is
likely different from what is ``hard'' for a listener. Zipf
characterized this as follows -- purely from the standpoint of speaker
effort, an optimal language \(\ell_{speaker}^*\) would tend toward a
vocabulary of just a single, low-cost word. Given such a language, the
full set of potential meanings would be conveyed using only that word,
i.e. \(\ell_{speaker}^*\) would be fully ambiguous and all possible
meanings would need to be disambiguated by a listener. From the
standpoint of listener effort, an optimal language \(\ell_{listener}^*\)
would map all possible meanings to distinct words, removing a listener's
need to disambiguate. In this example, speaker effort is related to
\emph{production cost} and listener effort to \emph{understanding or
disambiguation cost}. Clearly natural languages fall between the two
extremes of \(\ell_{speaker}^*\) and \(\ell_{listener}^*\). Zipf
proposed that the particular lexicon-level properties he observed were a
result of these competing forces -- the pressure to jointly minimize
speaker and listener effort.\par

But how does such optimization take place? The example given by Zipf
(1949), describes local, communicative interactions in terms of a
\textit{reference game}. Speakers intend to refer to some object in the
world \(m\). They choose some utterance \(u\) to transmit this intended
meaning, \(u \rightarrow m\). The listener attempts to reconstruct this
intended meaning given the transmitted utterance, \(m \rightarrow u\).
Other functionalist projects have also assumed this basic reference game
setting (Piantadosi et al., 2011; Regier et al., 2015) and this
simplification of the communicative act has proven productive in
theoretical (Ferrer-i-Cancho, 2018), simulation-based (Kirby, Griffiths,
\& Smith, 2014) and empirical explorations (Hawkins, Franke, Smith, \&
Goodman, 2018) of efficient language structure and use.\par

Zipf's conception of speaker and listener effort may be connected to
speakers' pragmatic reasoning in the moment. Under a Gricean treatment
of pragmatics, speakers and listeners are assumed to follow a set of
conversational maxims in which they cooperate to transfer information
(Grice, 1975). These maxims appear to emerge from efficiency concerns,
however (Horn, 1984).\par

In this work, we present a step toward formalizing the framing presented
above -- connecting Zipfian notions of efficient language structure to
Gricean notions of rational conversation (Grice, 1975). To do so
minimally requires three basic ingredients -- (1) a language property
we'd like to explain, (2) a framework for describing in-the-moment
interactions of speaker-listeners, and (3) some measure of linguistic
efficiency. For the current project we focus on the ``communicative
function of ambiguity'' for our property -- translating the
functionalist theory pursued by Piantadosi, et al. (2011) into our
framework. To model local, conversational interactions (2) we adopt the
Rational Speech-act framework (RSA) (Frank \& Goodman, 2012; Goodman \&
Frank, 2016). Finally, (3) we derive an objective function for measuring
language efficiency in the reference game setting.\par

The current work examines the question, ``how does having
\textit{lexical ambiguity} make using language \(\ell\) more or less
useful for communication''? We begin with a high-level introduction to
the modelling framework introducing the basic ingredients we will need
to represent language as repeated reference games. We derive anobjective
function for measuring the efficiency of linguistic systems in this
setting. We move on to two case-studies exploring the questions posed
above -- examining the efficiency of ambiguity from a language design
and language use perspective. The contributions of this work is to show
how the reference game framework, in combination with formal modles of
communication, can be used to study questions at a range of different
timescales, from language use in-the-moment to language change across
historical time.\par

\section{Exploring efficient language design and use in rational
pragmatic
agents}\label{exploring-efficient-language-design-and-use-in-rational-pragmatic-agents}

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/plot-reference-game-1}

}

\caption[An example reference game with associated literal semantics (in our terminology a ``language'')]{An example reference game with associated literal semantics (in our terminology a ``language'').}\label{fig:plot-reference-game}
\end{figure}
\end{CodeChunk}

\subsubsection{Reference games}\label{reference-games}

Zipf's example of optimal speaker- and listener-languages took the form
of a reference game. We adopt that formulation here, assuming these
communication games as our basic unit of analysis. In this setting,
speakers and listeners are aware of a set of objects \(M\), which we
will refer to as \emph{meanings} and are knowledgeable about the set of
possible signals \(U\) (\emph{utterances}) that can be used to refer to
a given meaning (see Figure 1). Utterances may have different relative
costs, operationalized via a prior over utterances \(P(U)\). Similarly,
meanings differ in the relative degree to which they need to be talked
about, operationalized as a prior over meanings
\(P(M)\)\footnote{The prior over meanings are analogous to the \textit{need probabilities} assumed in previous work Regier, Kemp \& Kay (2015).}.
We consider a set of contexts \(C\) with an associated prior \(P(C)\).
Contexts describe different orderings of the underlying meaning prior,
formalized as different conditional distributions over meanings e.g.
\(p(M|C=c)\). Finally, we consider a set of communicative events
\(e \in E\) where \(<u, m> = e\) is a tuple of utterance, meaning
pairs.\par

\subsubsection{Languages}\label{languages}

A language \(\ell\) defines the set of semantic mappings between
utterance and meanings. For example, Figure 1 contains four utterances
\(U = \{\text{blue}, \text{green}, \text{square}, \text{circle}\} \text{ and three meanings }M = \{\text{green-square}, \text{blue-square}, \text{green-circle}\}\).
The boolean matrix describes the literal semantics of the language
\(\ell\). Note that will consider language to ``contain ambiguous
items'' if there is at least one utterance \(u \in U\) which can apply
to multiple meanings (i.e.
\(|[[u_i]]| > 1\))\footnote{We use double brackets $[[\dots]]$ to represent denotation}.
For example, in Figure 1 both the words ``blue'' and ``green'' are
ambigous so we would say that \(\ell\) contains ambiguity.

\subsubsection{Speakers and listeners}\label{speakers-and-listeners}

The Rational Speech-act framework (RSA) is computational-level theory of
pragmatic language use. RSA can be understood as a formalization of
essential Gricean pragmatic principles -- agents reason about one
another and their shared context. For this reason, we adopt RSA as our
representational framework to operationalize Gricean (rational and
pragmatic) speaker-listeners (Grice, 1975). We refer the reader to
Section 1 of our supplementary materials for a more detailed description
of basic RSA speaker-listener definitions, and to Goodman \& Frank
(2016) for an overview of its appliations.\par

An RSA \emph{speaker agent} defines a conditional distribution over
utterances, mapping from intended meanings \(M\) to utterances \(U\)
using \(\ell\). That is, a speaker defines \(P_{speaker}(u|m;\ell)\). We
will use \(S(u|m;\ell)\) as short-hand throughout. A \emph{listener
agent} defines a conditional distribution over meanings, mapping from
utterances \(U\) to meanings \(M\) using \(\ell\). We will use
\(L(m|u; \ell)\) as shorthand. Note that both speakers and listeners can
induce joint distributions over the set of all events \(E\), although,
importantly, these distributions may differ:
\[P_{speaker}(u, m; \ell) = S(u|m; \ell)p(m)\]
\[P_{listener}(u, m; \ell) = L(m|u; \ell)p(u)\]

\section{Zipfian objective for linguistic system
efficiency}\label{zipfian-objective-for-linguistic-system-efficiency}

Zipf (1949) proposed that the particular distributional properties found
in natural language emerge from competing speaker and listener
pressures. We operationalize this objective in equation (1) -- the
efficiency of a linguistic system \(\ell\) being used by speaker and
listener agents \(S\) and \(L\) is the sum of the expected speaker and
listener effort to communicate over all possible communicative events
\(E\).\par

\begin{equation}
\begin{split}
  \text{Efficiency}(S, L, \ell) = \mathbb{E}_{e \sim P(E)}[\text{speaker effort}] \\+ \mathbb{E}_{e \sim P(E)}[\text{listener effort}]
\end{split}
\end{equation}

We assume that speaker effort is the negative log probability of an
utterance (surprisal) -- intuitively, the number of bits needed to
encode the utterance \(u\). This particular formalization of
speaker-cost is general enough to accommodate a range of instantiations
that might in theory be related to production difficult via articulation
effort, cognitive effort related to lexical access, etc (Bennett \&
Goodman, 2018).\par

\[\text{speaker effort} = -log_2(p(u))\]

We assume listener effort is the negative log probability of the
listener's conditional \(-log_2(L(M|U))\). This operationalization of
listener effort is intuitively related to existing work in sentence
processing in which word difficulty during comprehension is proportional
to surprisal (Hale, 2001; R. Levy, 2008).

\[\text{listener effort} = -log_2(L(m|u; \ell))\]

Finally, we assume that events \(e = <u, m>\) are sampled according the
to following generative model -- some object occurs in the world with
probability \(p(M=m)\), the speaker attempts to refer to that object by
sampling from her conditional distribution \(S(u|m)\) (i.e.
\(e \sim p(m)S(u|m)\)). Using these basic ingredients we arrive at the
following objective:

\begin{equation}
\begin{split}
  = \sum_{u, m}P_{speaker}(u, m; \ell)[-log_2(P_{listener}(u, m; \ell))]
\end{split}
\end{equation}

This equation is the simply the cross-entropy between the speaker and
listener joint distributions.

\begin{equation}
\begin{split}
  = \mathbb{E}_{P_{speaker}}[-log_2(P_{listener})] \\
  = H_{cross}(P_{speaker}, P_{listener}; \ell)
\end{split}
\end{equation}

From an information-theoretic perspective this objective is intuitive:
cross-entropy gives us a measure of dissimilarity between two
distributions -- the average number of bits required to communicate
under one distribution, given that the ``true'' distribution differs. In
our case, this is the difference between the joint distribution assumed
by the speaker \(P_{speaker}\) and listener \(P_{listener}\). In other
words, an ``efficient'' language \(\ell\) minimizes the distance between
what speakers and listener think. For a complete derivation and its
components we refer the reader to Section 2.1 of our supplementary
materials.\par

\section{Simulating the communicative function of
ambiguity}\label{simulating-the-communicative-function-of-ambiguity}

The task of understanding language is marked by a frequent need to
handle various forms of ambiguity; lexical, syntactic, among others
(Wasow, Perfors, \& Beaver, 2005). The ubiquity of this property,
however, has been argued to provide evidence that langauges are not been
optimized for
communication\footnote{Chomsky (2002) famously claimed "If you want to make sure that we never misunderstand one another, for that purpose language is not well designed, because you have such properties as ambiguity." Note that this analysis is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$ we described earlier and define in terms of equation 2 in SI 2.2 .}
(Chomsky, 2002).

Piantadosi et al. (2011) argue just the opposite, claiming that
ambiguity is an \emph{efficient} property of any communication system in
which \emph{communication is contextualized}. Simply put, it is useful
to have a language that re-uses low-cost material (has ambiguity) so
long as the cost of disambiguating the material is low. In particular,
context (or common ground) can provide useful information for
disambiguation.\par

We provide a toy example of their argument here. Say we have two objects
in the world \(m_1\) and \(m_2\), and two languages \(\ell_1\) and
\(\ell_2\). In language \(\ell_1\), the low-cost utterance \(u_1\) can
be used to refer to both \(m_1\) and \(m_2\) (i.e.
\([[u_1]]_{\ell_1} = \{m_1, m_2 \}\)). With language \(\ell_2\),
however, the low cost utterance \(u_1\) can only refer to \(m_1\) and a
higher-cost utterance \(u_2\) can refer to \(m_2\) (i.e.
\([[u_1]]_{\ell_2} = \{m_1\}\) and \([[u_2]]_{\ell_2} = \{m_2 \}\)).
While it is cheaper for a speaker to use \(\ell_1\) (because she can
always use a low-cost utterance), it is more difficult for a listener
(because \(u_1\) is ambiguous). If context isn't disambiguating then we
might prefer \(\ell_2\) to \(\ell_1\). But, if context is disambiguating
then the speaker can use \(u_1\) to refer to either \(m_1\) or \(m_2\)
and \(\ell_1\) should be preferred.\par

In the following experiments we explore two aspects of Piantadosi's
claim. First, we examine the efficient language \emph{structure} aspect
of their claim, exploring when the optimal linguistic system \(\ell^*\)
is most likely to contain ambiguous lexical items. In the second
experiment, we explore an efficient language \emph{use} aspect of the
claim -- at what point in a conversation is it useful for a speaker to
use ambiguous lexical material?\par

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/plot-optimal-langs-1}

}

\caption[Optimal languages are more likely to contain ambiguous items as the amount of contextual information increases]{Optimal languages are more likely to contain ambiguous items as the amount of contextual information increases. Vertical axis shows the proportion of optimal languages that contain ambiguity. Horizontal axis shows the number of contexts in each condition (1-4). Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only consideration.}\label{fig:plot-optimal-langs}
\end{figure*}
\end{CodeChunk}

\section{Simulation 1: Optimal languages contain ambiguity when context
is
informative}\label{simulation-1-optimal-languages-contain-ambiguity-when-context-is-informative}

We adpot the framework above, and (following Piantadosi et al. 2011),
add \textit{context} (\(C\)). The authors argue that mapping two
meanings \(m_1\) and \(m_2\) to a single utterance \(u_1\) is useful
when they can be disambiguated \emph{in context}. In our case we
consider a context \(c_i \in C\) to specify a permutation of the meaning
probabilities. That is, \(p(m|c_1) \neq p(m|c_2)\), when there are two
contexts \(|C| = \{c_1, c_2\}\).This in turn, leads to an update to our
linguistic efficiency objective as we now would like to consider the
average speaker-listener effort over all contextualized communicative
events (see section 3 SI).\par

We proceed by generating languages with different amounts of contextual
support (varying the size of \(|C|\)). We search the space of languages,
asessing if the ones which minimize our objective contain ambiguity. If
context leads to more efficient commnication, then optimal languages
should be more likley to be ambiguous as the amount of context
increases.\par

\subsection{Simulation set-up}\label{simulation-set-up}

We conduct \(N=2000\) simulations. For each simulation we enumerate the
set of \emph{valid} languages in which \(|U|=|M|=4\) (recall that \(U\)
is our set of utterances and \(M\) our set of meanings). Note that a
language \(\ell \in L\) is ``valid'' so long as each possible meaning in
\(m \in M\) can be referred to by at least one form \(u \in U\) (every
column of \(\ell\) has some non-zero assignment) and each form maps to
at least one meaning (every column has some non-zero
assignment)\footnote{For the current set of simulations we consider languages in which there are four or five true values -- in a "four" language ambiguity is not possible, but in a "five" language there are ambiguous mappings.}.
For a given simulation the goal is to find the language \(\ell^*\) which
minimizes our cross-entropy objective and then check to see if that
language contains ambiguity.\par

Recall that language efficiency is both a function of the particular
semantic mappings induced by that language, the speaker and listener
agents (\(S\) and \(L\)), as well as the utterance (\(P(U)\)), meaning
(\(P(M)\)), and context priors (\(P(C)\)). Rather than assume particular
structure for our utterance, meaning, and context prior distributions,
for each simulation we generate \(P(U) \sim \text{Dir}(1, |U|)\),
\(P(M) \sim \text{Dir}(1, |M|)\), and \(P(C) \sim \text{Dir}(1, |C|)\)
where \(\text{Dir}(1, k)\) specifies the uniform Dirichlet distribution
over a \(k\)-dimensional probability vector.\par

\subsubsection{Context}\label{context}

Following the argument given by Piantadosi et al. (2011), we want to
assess the impact of \emph{context} on our objective. To do so we
consider four conditions with \(n=500\) simulations each (that is, 500
unique sets of \(\{P(U), P(M), P(C)\}\). Our first is a
\emph{single-context} condition (\(|C|=1\)) -- there is a only a single
context describing \(P(M)\). Our second condition contains two-contexts
(\(|C| = 2\)) -- we consider efficiency under both \(P(M|C=c_1)\) as
well as \(P(M|C=c_2)\). The third and fourth condition correspond
accordingly with \(|C|=3\) and \(|C| = 4\), respectively.\par

\subsubsection{Baselines}\label{baselines}

For comparison, we also examine properties of optimal languages under
two additional objectives. Zipf (1949) proposed that the optimal speaker
language \(\ell_{speaker}^*\) would only optimize speaker effort and the
optimal listener language \(\ell_{listener}^*\) would only optimize
listener effort. We define these objectives using the first and second
half of equation 1 (see SI section 2.2. for more detail).\par

\subsection{Results and Discussion}\label{results-and-discussion}

Figure 1. plots the proportion of optimal languages under each
objective, as a function of context condition. The red line shows that
as the number of contexts increases, so does the probability that the
optimal language \(\ell^*_{cross}\) under our speaker-listener
cross-entropy objective contains ambiguity. We also plot the proportion
of optimal speaker-only \(\ell^*_{speaker}\) and listener-only
\(\ell^*_{listener}\) languages that contain ambiguity. In line with
Zipf's predictions, if languages are designed only to minimize speaker
effort then optimal languages will assign lower-cost forms multiple
meanings (all of the speaker-optimal languages contained ambiguity).
Likewise, if languages are designed only to minimize listener effort
then ambiguity should always be avoided (none of the listener-optimal
languages contained ambiguity).\par

We explored the degree to which ambiguity improved efficiency during
communication between rational, pragmatic agents. We showed that as
contextual information increased, the optimal languages under our
Cross-Entropy objective were more likely to have ambiguous lexical
items. Importantly, we did not see this effect under alternative
objectives which only emphasized speaker- or listener effort alone.\par

While our results indicate that ambiguity is an efficient property of
contextualized language these simulations assumed that speaker-listener
agents could always disambiguate items in context. That is, in our four
conditions both agents had perfect knowledge of the relevant conditional
distributions (\(P(M|C)\)). This assumption may be too strong for
describing much of day-to-day communication -- we seldom begin a
conversation with perfect knowledge of the current context of a
conversation. We explore a case in which the listener has imperfect
knowledge of context at the start of the conversation in Simulation
2.\par

\section{Simulation 2: Rational, pragmatic speakers use ambiguity
efficiently}\label{simulation-2-rational-pragmatic-speakers-use-ambiguity-efficiently}

We shift time-scales in Simulation 2 -- instead of examining the
objective landscape of a space of languages, we now consider a single
language that contains ambiguity and examine how that property is used
over a discourse. If context is not informative with respect to meaning
early in a discourse (position \(i\)), but is at later positions at
position (position \(i+k\)), then the speaker should avoid using
ambigous material early \(i\), only using it \(k\) steps later in the
conversation. In this toy example we can consider ``context'' as
analogous to a ``topic'' of conversation. Intuitively, we can imagine a
scenario in which a reader is beginning a newspaper article. While they
may have some knowledge about the article's topic (perhaps from the
title), they may not have complete knowledge of its contents, including
the persons or events involved. In this setting, using an ambiguous
pronoun early (position \(i\)) in the article can lead to
misunderstanding without sufficient context to support disambiguation.
But if by position \(i+k\) there is sufficient contextual information it
may be efficient for the writer to use ambiguous material.\par

\subsection{Simulation set-up}\label{simulation-set-up-1}

We consider consider a single language \(\ell\), which contains both
ambiguous and unambiguous utterances. The ambiguous utterances are less
costly than the unambiguous utterances. Crucially, we do not assume that
the listener knows the particular topic (\(c_{current}\)) of the
conversation \textit{a priori}. Rather, we assume that the listener has
knowledge of the set of possible topics \(C = \{c_1, \dots, c_k\}\), but
\textit{does not know which one is currently being used by the speaker}.
Formally, this means the listener does not have access to the correct
conditional distribution over meanings \(P(M|C=c_{current})\) at the
start of the discourse. In this setting, a speaker minimizing production
effort would like to use the low-cost material as frequently as
possible. However, this will incur a disambiguation cost to the listener
if they are uncertain of the current context \(c_{current}\) (topic of
conversation).\par

Over the course of a discourse, \(D\) the listener tries to infer both
the current topic, \(c_{current}\), as well as the particular meaning
\(m\) of a given utterance \(u\). That is, we consider updated listener
and speakers (\(L(m, c|u,D;\ell)\), \(S(u|m,c,D;\ell)\)). Note that both
the speaker and listener can track the history of previous utterances
\(D\). Importantly, an agent can attempt to infer the current topic of
conversation \(c_{current}\) using the discourse history \(D\) (see SI
section 3 for more detail).\par

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/plot-optimal-use-1}

}

\caption[(A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position]{(A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position. (B) shows speaker effort across the three models. (C) shows the Cross-Entropy objective under our three speaker models.}\label{fig:plot-optimal-use}
\end{figure*}
\end{CodeChunk}

We conduct \(N=600\) simulations, generating discourses of length
\(|D|=30\) utterances with three different speaker models (\(n=200\)
each). We consider a single language
\(\ell\)\footnote{See SI for the matrix notation of this langauge.} with
\(|U|=6\) and \(|M|=4\) in which two of the utterances \(u_5\) and
\(u_6\) are ambiguous and that
\(p(u_5) = p(u_6) > p(u_{1})=\dots=p(u_{4})\). That is, the two
ambiguous utterances (\(u_5\) and \(u_6\)) are less costly than the
non-ambiguous utterances. (Note that use of this particular language is
not essential -- the results are broadly generalizable languages that
contain ambiguity.)\par

\subsection{Speaker agents}\label{speaker-agents}

We consider three types of speaker models. In our
\textit{Full pragmatics} model, we consider a speaker agent who reasons
about her listener and also has complete recall of the set of utterances
in the discourse \(D\). This speaker believes that the listener may not
know the current topic \(c_{current}\) at the start of the discrouse,
but can infer it with enough data. In this way, contextual information
shared by the speaker and listener may increase over the discourse (as
the listeners estimate of the current topic (\(P(C=c|D)\)) comes in line
with the actual topic known to the speaker). We compare two baselines
models. The first, a \textit{Partial pragmatics} baseline describes a
speaker who reasons about a listener, but assumes they have no access to
discourse history. The second is a \textit{No pragmatics} speaker who
does not consider a listener at all, but produces utterances according
to the underlying langauge semantics (\(\ell\)) and topic probabilities
(\(p(M|C=c_{current}\)) (see SI section 3 for details).\par

\subsection{Hypotheses}\label{hypotheses}

We assess whether it is efficient to use low-cost, ambiguous material
when context is disambiguating. In the current setting, this corresponds
to speaker behavior in which ambiguous forms are avoided early in the
discourse, but prefered later in the discourse once the listener is
confident of the topic. We should expect this to be reflected in our
\textit{Full pragmatics} model. By contrast, our
\textit{Partial pragmatics} speaker does not believe that there is
sufficient contextual information for the listener to disambiguate her
utterances. This should lead to total avoidance of ambiguous material
over the discourse. Finally, our \textit{No pragmatics} speaker should
use low-cost material greedily since she does not consider a listener's
need to disambiguate whatsoever.\par

\subsection{Results and Discussion}\label{results-and-discussion-1}

Figure 2, (A) shows the empirical probability that a speaker uses an
ambiguous utterance as a function of discourse position. The
\text{No pragmatic} baseline uses ambiguous utterances frequently and at
a constant rate over the discourse and the \text{Partial pragmatic}
baseline avoids ambiguous utterances entirely (reasoning that the
utterances cannot be disambiguated). But, the \text{Full pragmatic}
model avoids ambiguous material at first, but employing them gradually
increasingly as the discourse proceeds. (B) shows the speaker-listener
Cross-Entropy objective. Note that the objective decreases for all three
models. This decrease in all three models is primarily driven by the
listener updating his belief about the actual topic
(\(P(C=c_{current}|D)\)). Of particular importance is the value of the
objective at the end of the discourse (position \(i=30\)). The
difference in the Cross-Entropy between the \text{Full Pragmatic} and
\text{Partial Pragmatic} models at the end of the discourse is driven by
the reduction in speaker costs described in (C). That is, while speaker
effort remains constant in both \textit{No pragmatic} and
\textit{Partial pragmatic} baselines, effort declines for the
\text{Full Pragmatic} model as they are increasingly able to use
ambiguous material later in the discourse.\par

In Simulation 2 we explored how ambiguity could be employed efficiently
during discourse. We assumed that the listener did not know the current
context (topic) \textit{a priori}, but could infer it from the discourse
history. We examined speaker production behavior, tracking use of
low-cost, but ambiguous material over the discourse. Early in the
discourse the \textit{Full pragmatic} speaker avoided using ambiguous
material, only using the material later when the context was known to
the listener -- it is useful to use low-cost material, but only when
context is disambiguating.

\section{General Discussion}\label{general-discussion}

How do the competing pressures of speakers and listeners give rise to
the distributional forms found in natural language? Zipf (1949) proposed
that the asymmetry between speaker and listener costs gives rise to a
range of properties at the level of the lexicon. We explored the
interactions of rational pragmatic agents as a framework for
understanding efficient language structure and use. We focused on an
argument on the communicative function of ambiguity (Piantadosi et al.,
2011). We derive a novel speaker-listener cross-entropy objective for
measuring the efficiency of a language in a reference game setting
showing that optimal languages are more likely to contain ambiguous
material when context is informative. Further we show that rational
pragmatic agents will \emph{use} ambiguous material efficiently, only
when such use is supported by context.\par

The framework employed here -- the analysis of the dynamics of repeated
reference games between rational pragmatic agents -- provides a tool for
the analysis of functionalist theories, allowing linkage across
timescales of analysis from conversation (Genzel \& Charniak, 2002; \&.
J. Levy R., 2007) to typology (Regier et al., 2015). In theory, this
framework provides extensive opportunity to vary properties of the
languages and agents, however we have focused on only a small subset
here. For example, examining how increasing or decreasing
\textit{rationality} (instantiated through the RSA \(\alpha\) parameter)
or \textit{pragmatics} (instantiated through the RSA recursion depth)
impacts these results is a clear avenue for future work. An important
limitation of the current work is the broad assumption that language can
be described through such reference games. The degree to which
properties of non-referential language can be captured in such a
framework remains an important target for future work.

\vspace{1em}

\fbox{\parbox[b][][c]{7.3cm}{\centering All experiments, data, and analysis code are available in the public repository for the project: (link will be available upon acceptance)}}
\vspace{1em} \noindent

\section{References}\label{references}

\setlength{\parindent}{-0.1in} \setlength{\leftskip}{0.125in} \noindent

\hypertarget{refs}{}
\hypertarget{ref-BennettGoodman2015a}{}
Bennett, E., \& Goodman, N. (2018). Extremely costly intensifiers are
stronger than quite costly ones. \emph{Cognition}.

\hypertarget{ref-Chomsky2002a}{}
Chomsky, N. (2002). An interview on minimilism. In \emph{N chomsky, on
nature and language}.

\hypertarget{ref-FerreriCancho2018a}{}
Ferrer-i-Cancho, R. (2018). Optimization models of natural
communication. \emph{Journal of Quantitative Linguistics}, \emph{25
(3)}, 207--237.

\hypertarget{ref-FrankGoodman2012a}{}
Frank, M., \& Goodman, N. (2012). Predicting pragmatic reasoning in
language games. \emph{Science}, \emph{336}, 998.

\hypertarget{ref-GenzelCharniak2002a}{}
Genzel, D., \& Charniak, E. (2002). Entropy rate constancy in text. In
\emph{Proceedings of the 40th annual meeting on association for
computational linguistics}.

\hypertarget{ref-GoodmanFrank2016a}{}
Goodman, N., \& Frank, M. (2016). Pragmatic language interpreation as
probabilistic inference. \emph{Trends in Cognitive Sciences},
\emph{20(11)}, 818--829.

\hypertarget{ref-Grice1975a}{}
Grice, P. H. (1975). Logic and conversation.

\hypertarget{ref-Hale2001a}{}
Hale, J. (2001). A probabilistic earley parser as a psycholinguistic
model. In \emph{Proceedings of the naacl}. 159-166.

\hypertarget{ref-HawkinsFrankeSmithGoodman2018a}{}
Hawkins, R., Franke, M., Smith, K., \& Goodman, N. (2018). Emerging
abstractions: Lexical conventions are shaped by communicative context.
In \emph{Proceedings of the 40th annual conference of the cognitive
science society}.

\hypertarget{ref-KirbyGriffithsSmith2014a}{}
Kirby, S., Griffiths, T., \& Smith, K. (2014). Iterated learnign and the
evoluation of language. \emph{Current Opinion in Neurobiology},
\emph{28}, 108--114.

\hypertarget{ref-LevyJaeger2007a}{}
Levy, \&. J., R. (2007). Speakers optimize information density through
syntactic reduction. In \emph{Proceedings of the twentieth annual
conference on neural information processing systems}.

\hypertarget{ref-Levy2008a}{}
Levy, R. (2008). Expectation-based syntactic comprehension.
\emph{Cognition}, \emph{106(3)}, 1126--1177.

\hypertarget{ref-Piantadosi2011a}{}
Piantadosi, S., Tily, H., \& Gibson, E. (2011). The communicative
function of ambiguity in language. \emph{Cognition}, \emph{122},
280--291.

\hypertarget{ref-RegierKempKay2015a}{}
Regier, T., Kemp, C., \& Kay, P. (2015). Word meanings across languages
support efficient communication. In B. M. \& W. O'Grady (Ed.), \emph{The
ahndbook of language emergence (pp. 237-263)}. Hoboken, NJ:
Wiley-Blackwell.

\hypertarget{ref-WasowPerforsBeaver2005a}{}
Wasow, T., Perfors, A., \& Beaver, D. (2005). The puzzle of ambiguity.
In CSLI (Ed.), \emph{Morphology and the web of grammar: Essays in memory
of steven g lapointe}. Stanford, CA: McGraw-Hill.

\hypertarget{ref-Zipf1949a}{}
Zipf, G. (1949). \emph{Human behavior and the principle of least
effort}. New York, NY: Prentice-Hall.

\bibliographystyle{apacite}

\end{document}

### 2019-01-30 paper/paper.Rmd
update with some (not all) of NDGs comments

@@
 We discuss a framework for studying how these properties emerge from in-the-moment interactions of rational, pragmatic speakers and listeners.
-Our work takes Zipfian notions of lexicon-level efficiency as a starting point, connecting these ideas to Gricean notions of conversational-level efficiency.
+Our work takes notions of lexicon-level efficiency as a starting point, connecting these ideas to notions of conversational-level efficiency.
 To do so, we derive an objective function for measuring the communicative efficiency of linguistic systems and then examining the behavior of this objective in a series of simulations focusing on the communicative function of ambiguity in language.
@@
 The listener attempts to reconstruct this intended meaning given the transmitted utterance, $m \rightarrow u$.
-Other functionalist projects have also assumed this basic reference game setting [@RegierKempKay2015a; @Piantadosi2011a] and this simplification of the communicative act has proven productive in theoretical [@FerreriCancho2018a], simulation-based [@KirbyGriffithsSmith2014a] and empirical explorations [@HawkinsFrankeSmithGoodman2018a] of efficient language structure and use.\par
+Other functionalist projects have assumed this basic reference game setting [@RegierKempKay2015a; @Piantadosi2011a] and this simplification of the communicative act has proven productive in theoretical [@FerreriCancho2018a], simulation-based [@KirbyGriffithsSmith2014a] and empirical explorations [@HawkinsFrankeSmithGoodman2018a] of efficient language structure and use.\par
 
@@
 Under a Gricean treatment of pragmatics, speakers and listeners are assumed to follow a set of conversational maxims in which they cooperate to transfer information (Grice, 1975).
-These maxims appear to emerge from efficiency concerns, however (Horn, 1984).\par
+These maxims appear to emerge from efficiency concerns, however (Horn, 1984).
+A primary target of the current work is to formalize this connection, by providing a framework for deriving system-level efficiencies from local, interaction behavior.
+Our claim is that to understand an "efficient" property of a system we need to consider *how* that property is *used* efficiently.
+Additionally, any quantitative description of "efficiency" should be a function of both the system and users.\par
 
-In this work, we present a step toward formalizing the framing presented above  -- connecting Zipfian notions of efficient language structure to Gricean notions of rational conversation [@Grice1975a].
-To do so minimally requires three basic ingredients -- (1) a language property we’d like to explain, (2) a framework for describing in-the-moment interactions of speaker-listeners, and (3) some measure of linguistic efficiency.
-For the current project we focus on the “communicative function of ambiguity” for our property -- translating the functionalist theory pursued by Piantadosi, et al.
-[-@Piantadosi2011a] into our framework.
-To model local, conversational interactions (2) we adopt the Rational Speech-act framework (RSA) [@FrankGoodman2012a; @GoodmanFrank2016a].
-Finally, (3) we derive an objective function for measuring language efficiency in the reference game setting.\par
-
-The current work examines the question, “how does having \textit{lexical ambiguity} make using language $\ell$ more or less useful for communication”?
-We begin with a high-level introduction to the modelling framework introducing the basic ingredients we will need to represent language as repeated reference games.
-We derive anobjective function for measuring the efficiency of linguistic systems in this setting.
-We move on to two case-studies exploring the questions posed above -- examining the efficiency of ambiguity from a language design and language use perspective.
-The contributions of this work is to show how the reference game framework, in combination with formal modles of communication, can be used to study questions at a range of different timescales, from language use in-the-moment to language change across historical time.\par
+In the current work we provide a case-study for the kind of framing presented above.
+We choose a property of languages that could, in principle, vary freely, but shows strong regularities across languages.
+The explanadum is why this regularity holds.
+We examine lexical ambiguity as our property, extending ideas by Piantadosi et al.
+[-@Piantadosi2011a].
+We define a novel measure of efficiency that depends on the interactional behavior of interlocutors.
+For our model of speakers and listeners we adopt the Rational Speech Act (RSA) framework, which is nicely attested in experimental data on pragmatic interaction.
+With these basic ingredients we have a measure of efficiency given language and speaker-listener agents.
+Finally, we show that the property of interest is prevalent in languages that optimize this measure of efficiency and also how this property is used efficiently during local interaction.
+The primary contribution of this work is to show how the reference game frameowrk, in combination with formal models of communication, can be used to connect ideas about system-level efficiencies to in-the-moment language use.\par
 
@@
 We consider a set of contexts $C$ with an associated prior $P(C)$.
-Contexts describe different orderings of the underlying meaning prior, formalized as different conditional distributions over meanings e.g.
-$p(M|C=c)$.
-Finally, we consider a set of communicative events $e \in E$ where $<u, m> = e$ is a tuple of utterance, meaning pairs.\par
+Contexts encode different meaning priors, formalized as separate conditional distributions over meanings e.g.
+$p(M|C=c_i) \neq p(M|C=c_j)$.
+Finally, we consider a set of communicative events $e \in E$ where $<u, m> = e$ is a tuple of utterance, meaning pairs.
+Note that in the case of single context $|C|=1$ we simply have a prior over meanings $p(M)$, a simplification we will use in introducing our objective function.\par
 
@@
 
-An RSA *speaker agent* defines a conditional distribution over utterances, mapping from intended meanings $M$ to utterances $U$ using $\ell$.
-That is, a speaker defines $P_{speaker}(u|m;\ell)$.
-We will use $S(u|m;\ell)$ as short-hand throughout.
-A *listener agent* defines a conditional distribution over meanings, mapping from utterances $U$ to meanings $M$ using $\ell$.
+An RSA *speaker agent* defines a conditional distribution over utterances, mapping from intended meanings $M$ to utterances $U$ using $\ell$ in a given context $c$.
+That is, a speaker defines $P_{speaker}(u|m; \ell)$.
+We will use $S(u|m; \ell)$ as short-hand throughout.
+A *listener agent* defines a conditional distribution over meanings, mapping from utterances $U$ to meanings $M$ using $\ell$ in a given context $c$.
 We will use $L(m|u; \ell)$ as shorthand.
@@
 Zipf (1949) proposed that the particular distributional properties found in natural language emerge from competing speaker and listener pressures.
-We operationalize this objective in equation (1) -- the efficiency of a linguistic system $\ell$ being used by speaker and listener agents $S$ and $L$ is the sum of the expected speaker and listener effort to communicate over all possible communicative events $E$.\par
+We operationalize this objective in equation (1) -- the efficiency of a linguistic system $\ell$ being used by speaker and listener agents $S$ and $L$ is the sum of the expected speaker and listener effort to communicate over all possible communicative events $E$.
+Note that to simplify the introduction of this objective we assume there is just a single context ($|C|=1$).
+In this case $p(C=c)=1$ and there is only distribution over meanings ($p(m|c) = p(m)$).
+We provide a more detailed description of this derivation in the supplementary materials, both with and without context $C$.\par
 
@@
 
-We assume listener effort is the negative log probability of the listener's conditional $-log_2(L(M|U))$.
+We assume listener effort is the negative log probability of the listener's conditional $-log_2(L(m|u))$.
 This operationalization of listener effort is intuitively related to existing work in sentence processing in which word difficulty during comprehension is proportional to surprisal [@Hale2001a; @Levy2008a].
@@
 
-Finally, we assume that events $e = <u, m>$ are sampled according the to following generative model -- some object occurs in the world with probability $p(M=m)$, the speaker attempts to refer to that object by sampling from her conditional distribution $S(u|m)$ (i.e.
+Importantly, we assume that events $e = <u, m>$ are sampled according the to following generative model -- some object occurs in the world with probability $p(M=m)$, the speaker attempts to refer to that object by sampling from her conditional distribution $S(u|m)$ (i.e.
 $e \sim p(m)S(u|m)$).
-Using these basic ingredients we arrive at the following objective:
+From the ingredients highlighted above, it is possible to derive the following objective between the speaker and listener joint distributions (see SI 2.2 for a complete derivation).
 
@@
 \begin{split}
-  = \sum_{u, m}P_{speaker}(u, m; \ell)[-log_2(P_{listener}(u, m; \ell))]
-\end{split}
-\end{equation}
-This equation is the simply the cross-entropy between the speaker and listener joint distributions.
-\begin{equation}
-\begin{split}
-  = \mathbb{E}_{P_{speaker}}[-log_2(P_{listener})] \\
+  = -\mathbb{E}_{P_{speaker}}[log_2(P_{listener})] \\
   = H_{cross}(P_{speaker}, P_{listener}; \ell)
@@
 In our case, this is the difference between the joint distribution assumed by the speaker $P_{speaker}$ and listener $P_{listener}$.
-In other words, an "efficient" language $\ell$ minimizes the distance between what speakers and listener think.
-For a complete derivation and its components we refer the reader to Section 2.1 of our supplementary materials.\par
+In other words, an "efficient" language $\ell$ minimizes the distance between what speakers and listeners think.
 
@@
 In language $\ell_1$, the low-cost utterance $u_1$ can be used to refer to both $m_1$ and $m_2$ (i.e.
-$[[u_1]]_{\ell_1} = \{m_1, m_2 \}$).
+$[[u_1]]_{\ell_1} = \{m_1, m_2 \}$ and $[[u_2]]_{\ell_1} = \emptyset$).
 With language $\ell_2$, however, the low cost utterance $u_1$ can only refer to $m_1$ and a higher-cost utterance $u_2$ can refer to $m_2$ (i.e.
@@
 Horizontal axis shows the number of contexts in each condition (1-4).
-Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only consideration."}
+Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only objectives."}
 img <- png::readPNG("figs/fig1.png")
@@
 
-We adpot the framework above, and (following Piantadosi et al.
+We adopt the framework above, and (following Piantadosi et al.
 2011), add \textit{context} ($C$).
@@
 Recall that language efficiency is both a function of the particular semantic mappings induced by that language, the speaker and listener agents ($S$ and $L$), as well as the utterance ($P(U)$), meaning ($P(M)$), and context priors ($P(C)$).
-Rather than assume particular structure for our utterance, meaning, and context prior distributions, for each simulation we generate $P(U) \sim \text{Dir}(1, |U|)$, $P(M) \sim \text{Dir}(1, |M|)$, and $P(C) \sim \text{Dir}(1, |C|)$ where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.\par
+Rather than assume particular structure for our utterance, meaning, and context prior distributions, for each simulation we generate $P(U) \sim \text{Dir}(1, |U|)$, $P(M|C=c) \sim \text{Dir}(1, |M|)$ (a separete conditional distribution over meanings for each context $c$), and $P(C) \sim \text{Dir}(1, |C|)$ where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.\par
 
@@
 [-@Piantadosi2011a], we want to assess the impact of *context* on our objective.
-To do so we consider four conditions with $n=500$ simulations each (that is, 500 unique sets of $\{P(U), P(M), P(C)\}$.
+To do so we consider four conditions with $n=500$ simulations each (that is, 500 unique sets of $\{P(U), P(M|C), P(C)\}$.
 Our first is a *single-context* condition ($|C|=1$) -- there is a only a single context describing $P(M)$.
@@
 
-\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering All experiments, data, and analysis code are available in the public repository for the project: (link will be available upon acceptance)}} \vspace{1em}
+\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering All simulation and analysis code are available in the public repository for the project: (link will be available upon acceptance)}} \vspace{1em}
 \noindent

### 2019-01-30 paper/paper.tex
update with some (not all) of NDGs comments

@@
 these properties emerge from in-the-moment interactions of rational,
-pragmatic speakers and listeners. Our work takes Zipfian notions of
+pragmatic speakers and listeners. Our work takes notions of
 lexicon-level efficiency as a starting point, connecting these ideas to
-Gricean notions of conversational-level efficiency. To do so, we derive
-an objective function for measuring the communicative efficiency of
+notions of conversational-level efficiency. To do so, we derive an
+objective function for measuring the communicative efficiency of
 linguistic systems and then examining the behavior of this objective in
@@
 intended meaning given the transmitted utterance, \(m \rightarrow u\).
-Other functionalist projects have also assumed this basic reference game
+Other functionalist projects have assumed this basic reference game
 setting (Piantadosi et al., 2011; Regier et al., 2015) and this
@@
 (Grice, 1975). These maxims appear to emerge from efficiency concerns,
-however (Horn, 1984).\par
+however (Horn, 1984). A primary target of the current work is to
+formalize this connection, by providing a framework for deriving
+system-level efficiencies from local, interaction behavior. Our claim is
+that to understand an ``efficient'' property of a system we need to
+consider \emph{how} that property is \emph{used} efficiently.
+Additionally, any quantitative description of ``efficiency'' should be a
+function of both the system and users.\par
 
-In this work, we present a step toward formalizing the framing presented
-above -- connecting Zipfian notions of efficient language structure to
-Gricean notions of rational conversation (Grice, 1975). To do so
-minimally requires three basic ingredients -- (1) a language property
-we'd like to explain, (2) a framework for describing in-the-moment
-interactions of speaker-listeners, and (3) some measure of linguistic
-efficiency. For the current project we focus on the ``communicative
-function of ambiguity'' for our property -- translating the
-functionalist theory pursued by Piantadosi, et al. (2011) into our
-framework. To model local, conversational interactions (2) we adopt the
-Rational Speech-act framework (RSA) (Frank \& Goodman, 2012; Goodman \&
-Frank, 2016). Finally, (3) we derive an objective function for measuring
-language efficiency in the reference game setting.\par
-
-The current work examines the question, ``how does having
-\textit{lexical ambiguity} make using language \(\ell\) more or less
-useful for communication''? We begin with a high-level introduction to
-the modelling framework introducing the basic ingredients we will need
-to represent language as repeated reference games. We derive anobjective
-function for measuring the efficiency of linguistic systems in this
-setting. We move on to two case-studies exploring the questions posed
-above -- examining the efficiency of ambiguity from a language design
-and language use perspective. The contributions of this work is to show
-how the reference game framework, in combination with formal modles of
-communication, can be used to study questions at a range of different
-timescales, from language use in-the-moment to language change across
-historical time.\par
+In the current work we provide a case-study for the kind of framing
+presented above. We choose a property of languages that could, in
+principle, vary freely, but shows strong regularities across languages.
+The explanadum is why this regularity holds. We examine lexical
+ambiguity as our property, extending ideas by Piantadosi et al. (2011).
+We define a novel measure of efficiency that depends on the
+interactional behavior of interlocutors. For our model of speakers and
+listeners we adopt the Rational Speech Act (RSA) framework, which is
+nicely attested in experimental data on pragmatic interaction. With
+these basic ingredients we have a measure of efficiency given language
+and speaker-listener agents. Finally, we show that the property of
+interest is prevalent in languages that optimize this measure of
+efficiency and also how this property is used efficiently during local
+interaction. The primary contribution of this work is to show how the
+reference game frameowrk, in combination with formal models of
+communication, can be used to connect ideas about system-level
+efficiencies to in-the-moment language use.\par
 
@@
 We consider a set of contexts \(C\) with an associated prior \(P(C)\).
-Contexts describe different orderings of the underlying meaning prior,
-formalized as different conditional distributions over meanings e.g.
-\(p(M|C=c)\). Finally, we consider a set of communicative events
-\(e \in E\) where \(<u, m> = e\) is a tuple of utterance, meaning
-pairs.\par
+Contexts encode different meaning priors, formalized as separate
+conditional distributions over meanings e.g.
+\(p(M|C=c_i) \neq p(M|C=c_j)\). Finally, we consider a set of
+communicative events \(e \in E\) where \(<u, m> = e\) is a tuple of
+utterance, meaning pairs. Note that in the case of single context
+\(|C|=1\) we simply have a prior over meanings \(p(M)\), a
+simplification we will use in introducing our objective function.\par
 
@@
 utterances, mapping from intended meanings \(M\) to utterances \(U\)
-using \(\ell\). That is, a speaker defines \(P_{speaker}(u|m;\ell)\). We
-will use \(S(u|m;\ell)\) as short-hand throughout. A \emph{listener
-agent} defines a conditional distribution over meanings, mapping from
-utterances \(U\) to meanings \(M\) using \(\ell\). We will use
-\(L(m|u; \ell)\) as shorthand. Note that both speakers and listeners can
-induce joint distributions over the set of all events \(E\), although,
-importantly, these distributions may differ:
+using \(\ell\) in a given context \(c\). That is, a speaker defines
+\(P_{speaker}(u|m; \ell)\). We will use \(S(u|m; \ell)\) as short-hand
+throughout. A \emph{listener agent} defines a conditional distribution
+over meanings, mapping from utterances \(U\) to meanings \(M\) using
+\(\ell\) in a given context \(c\). We will use \(L(m|u; \ell)\) as
+shorthand. Note that both speakers and listeners can induce joint
+distributions over the set of all events \(E\), although, importantly,
+these distributions may differ:
 \[P_{speaker}(u, m; \ell) = S(u|m; \ell)p(m)\]
@@
 listener effort to communicate over all possible communicative events
-\(E\).\par
+\(E\). Note that to simplify the introduction of this objective we
+assume there is just a single context (\(|C|=1\)). In this case
+\(p(C=c)=1\) and there is only distribution over meanings
+(\(p(m|c) = p(m)\)). We provide a more detailed description of this
+derivation in the supplementary materials, both with and without context
+\(C\).\par
 
@@
 We assume listener effort is the negative log probability of the
-listener's conditional \(-log_2(L(M|U))\). This operationalization of
+listener's conditional \(-log_2(L(m|u))\). This operationalization of
 listener effort is intuitively related to existing work in sentence
@@
 
-Finally, we assume that events \(e = <u, m>\) are sampled according the
-to following generative model -- some object occurs in the world with
-probability \(p(M=m)\), the speaker attempts to refer to that object by
-sampling from her conditional distribution \(S(u|m)\) (i.e.
-\(e \sim p(m)S(u|m)\)). Using these basic ingredients we arrive at the
-following objective:
+Importantly, we assume that events \(e = <u, m>\) are sampled according
+the to following generative model -- some object occurs in the world
+with probability \(p(M=m)\), the speaker attempts to refer to that
+object by sampling from her conditional distribution \(S(u|m)\) (i.e.
+\(e \sim p(m)S(u|m)\)). From the ingredients highlighted above, it is
+possible to derive the following objective between the speaker and
+listener joint distributions (see SI 2.2 for a complete derivation).
 
@@
 \begin{split}
-  = \sum_{u, m}P_{speaker}(u, m; \ell)[-log_2(P_{listener}(u, m; \ell))]
-\end{split}
-\end{equation}
-
-This equation is the simply the cross-entropy between the speaker and
-listener joint distributions.
-
-\begin{equation}
-\begin{split}
-  = \mathbb{E}_{P_{speaker}}[-log_2(P_{listener})] \\
+  = -\mathbb{E}_{P_{speaker}}[log_2(P_{listener})] \\
   = H_{cross}(P_{speaker}, P_{listener}; \ell)
@@
 words, an ``efficient'' language \(\ell\) minimizes the distance between
-what speakers and listener think. For a complete derivation and its
-components we refer the reader to Section 2.1 of our supplementary
-materials.\par
+what speakers and listeners think.
 
@@
 be used to refer to both \(m_1\) and \(m_2\) (i.e.
-\([[u_1]]_{\ell_1} = \{m_1, m_2 \}\)). With language \(\ell_2\),
-however, the low cost utterance \(u_1\) can only refer to \(m_1\) and a
+\([[u_1]]_{\ell_1} = \{m_1, m_2 \}\) and
+\([[u_2]]_{\ell_1} = \emptyset\)). With language \(\ell_2\), however,
+the low cost utterance \(u_1\) can only refer to \(m_1\) and a
 higher-cost utterance \(u_2\) can refer to \(m_2\) (i.e.
@@
 Horizontal axis shows the number of contexts in each condition (1-4).
-Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only consideration.}\label{fig:plot-optimal-langs}
+Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only objectives.}\label{fig:plot-optimal-langs}
 \end{figure*}
@@
 
-We adpot the framework above, and (following Piantadosi et al. 2011),
+We adopt the framework above, and (following Piantadosi et al. 2011),
 add \textit{context} (\(C\)). The authors argue that mapping two
@@
 for each simulation we generate \(P(U) \sim \text{Dir}(1, |U|)\),
-\(P(M) \sim \text{Dir}(1, |M|)\), and \(P(C) \sim \text{Dir}(1, |C|)\)
-where \(\text{Dir}(1, k)\) specifies the uniform Dirichlet distribution
-over a \(k\)-dimensional probability vector.\par
+\(P(M|C=c) \sim \text{Dir}(1, |M|)\) (a separete conditional
+distribution over meanings for each context \(c\)), and
+\(P(C) \sim \text{Dir}(1, |C|)\) where \(\text{Dir}(1, k)\) specifies
+the uniform Dirichlet distribution over a \(k\)-dimensional probability
+vector.\par
 
@@
 consider four conditions with \(n=500\) simulations each (that is, 500
-unique sets of \(\{P(U), P(M), P(C)\}\). Our first is a
+unique sets of \(\{P(U), P(M|C), P(C)\}\). Our first is a
 \emph{single-context} condition (\(|C|=1\)) -- there is a only a single
@@
 \vspace{1em}
-
-\fbox{\parbox[b][][c]{7.3cm}{\centering All experiments, data, and analysis code are available in the public repository for the project: (link will be available upon acceptance)}}
+\fbox{\parbox[b][][c]{7.3cm}{\centering All simulation and analysis code are available in the public repository for the project: (link will be available upon acceptance)}}
 \vspace{1em} \noindent
@@
 (3)}, 207--237.
-
-\hypertarget{ref-FrankGoodman2012a}{}
-Frank, M., \& Goodman, N. (2012). Predicting pragmatic reasoning in
-language games. \emph{Science}, \emph{336}, 998.
 

### 2019-01-30 paper/paper.Rmd
update context plot with new n=500 run

@@
 Horizontal axis shows the number of contexts in each condition (1-4).
-Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only objectives."}
+Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only objectives.
+Error bars represent 95% confidence intervals."}
 img <- png::readPNG("figs/fig1.png")
@@
 (B) shows speaker effort across the three models.
-(C) shows the Cross-Entropy objective under our three speaker models."}
+(C) shows the Cross-Entropy objective under our three speaker models.
+Error bars represent 95% confidence intervals."}
 

### 2019-01-30 paper/paper.tex
update context plot with new n=500 run

@@
 
-\begin{CodeChunk}
-\begin{figure*}[h]
+\textbackslash{}begin\{CodeChunk\}
+\textbackslash{}begin\{figure*\}{[}h{]}
 
-{\centering \includegraphics{figs/plot-optimal-langs-1}
+\{\centering \includegraphics{figs/plot-optimal-langs-1}
 
-}
+\}
 
-\caption[Optimal languages are more likely to contain ambiguous items as the amount of contextual information increases]{Optimal languages are more likely to contain ambiguous items as the amount of contextual information increases.
-Vertical axis shows the proportion of optimal languages that contain ambiguity.
+\textbackslash{}caption{[}Optimal languages are more likely to contain
+ambiguous items as the amount of contextual information
+increases{]}\{Optimal languages are more likely to contain ambiguous
+items as the amount of contextual information increases. Vertical axis
+shows the proportion of optimal languages that contain ambiguity.
 Horizontal axis shows the number of contexts in each condition (1-4).
-Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only objectives.}\label{fig:plot-optimal-langs}
-\end{figure*}
-\end{CodeChunk}
+Red-line represents the optimal language under our Zipfian cross-entropy
+objective while the blue and red lines show optimal languages under
+speaker- and listener-only objectives. Error bars represent 95\%
+confidence intervals.\}\label{fig:plot-optimal-langs}
+\textbackslash{}end\{figure*\} \textbackslash{}end\{CodeChunk\}
 
@@
 
-\begin{CodeChunk}
-\begin{figure*}[h]
+\textbackslash{}begin\{CodeChunk\}
+\textbackslash{}begin\{figure*\}{[}h{]}
 
-{\centering \includegraphics{figs/plot-optimal-use-1}
+\{\centering \includegraphics{figs/plot-optimal-use-1}
 
-}
+\}
 
-\caption[(A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position]{(A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position.
-(B) shows speaker effort across the three models.
-(C) shows the Cross-Entropy objective under our three speaker models.}\label{fig:plot-optimal-use}
-\end{figure*}
-\end{CodeChunk}
+\textbackslash{}caption{[}(A) shows the empirical probability that our
+speaker used an ambiguous utterance as a function of discourse
+position{]}\{(A) shows the empirical probability that our speaker used
+an ambiguous utterance as a function of discourse position. (B) shows
+speaker effort across the three models. (C) shows the Cross-Entropy
+objective under our three speaker models. Error bars represent 95\%
+confidence intervals.\}\label{fig:plot-optimal-use}
+\textbackslash{}end\{figure*\} \textbackslash{}end\{CodeChunk\}
 

### 2019-01-31 ambiguity/simulations.py
remove p(c) multiplier after talking with NDG

@@ -86,10 +86,13 @@ class OrderedSimulation(Simulation):
             contexts = [p_meanings]
 
             for i in range(1, n_contexts):
-                # Currently not using...
+                # Sample new context from Dirichlet
                 # new_p_m populates a totally new prior over meanings.
-                # new_p_m = np.random.dirichlet([meaning_order_fn(k) for k in range(N_MEANINGS)])
-                contexts.append(np.roll(contexts[0], i+1))
+                new_p_m = np.random.dirichlet([meaning_order_fn(k) for k in range(N_MEANINGS)])
+                # or
+                # New context is simply a permutation
+                # contexts.append(np.roll(contexts[0], i+1))
+                contexts.append(new_p_m)
             # Baseline model
             # (context isn't informative so we take)
             # p(m) = \sum_c p(m|c)p(c)
@@ -121,23 +124,18 @@ class OrderedSimulation(Simulation):
                                                        listener_alpha,
                                                        listener_k),
                                               p_utterances)
-
                     # p(u) term
                     # in \sum_{u, m}P_s(u, m)*log(p(u))
                     m_simple_speaker_costs = row_multiply(m, p_utterances)
 
-                    # L(m|u) term
+                    # L(m|u, c)p(c) term
                     # in \sum_{u, m}P_s(u, m)*log(L(m|u))
-                    m_simple_listener_costs = listener(m, p_utterances,
-                                                       p_meanings_,
-                                                       listener_alpha,
-                                                       listener_k)
+                    m_simple_listener_costs =  \
+                        listener(m, p_utterances,
+                                 p_meanings_,
+                                 listener_alpha,
+                                 listener_k)
 
-                    # Require m_speaker and m_listener are valid
-                    # joint distributions
-                    # (although expect some rounding error).
-                    assert(np.abs(np.sum(m_speaker) - 1) < 1e-05)
-                    assert (np.abs(np.sum(m_listener) - 1) < 1e-05)
 
                     # Safe-keeping for clean-use below
                     args = [('ce', m_speaker, m_listener, ce),

### 2019-01-31 ambiguity/simulation-runners/context-ambiguity-plots.Rmd
updates to plots

@@
 fp <- '/Users/benpeloquin/Data/zipf_principles/ambiguity/runs-20190121/uniform_uniform_1_1_500.csv'
+fp <- '/Users/benpeloquin/Desktop/Projects/zipf_principles/test-20190130/uniform_uniform_1_1_500.csv'
+fp <- '/Users/benpeloquin/Data/zipf_principles/ambiguity/runs-20190130/uniform_uniform_1_1_500.csv'
 df <- read.csv(fp)
@@
   ggplot(aes(x=n_contexts, y=prop_ambiguous, col=Objective)) +
-    geom_line(size=1.5) +
-    geom_errorbar(aes(ymin=ci_low, ymax=ci_high), width=0.05, size=1, alpha=0.7) +
+    geom_line(size=1, alpha=0.9) +
+    geom_errorbar(aes(ymin=ci_low, ymax=ci_high), width=0.05, alpha=0.7, col='black') +
     ylab("Percent optimal languages \n containing ambiguity") +

### 2019-01-31 ambiguity/config.py

@@ -17,6 +17,7 @@ CONTEXT_RUN_PARAMS = {
         'verbose': False
     },
     # fp pattern below
+    # ----------------
     # utterance_order_fn_name-meaning_order_fn_name-speaker_k-listener_k
     'out_file_name': 'uniform_uniform_1_1_500.csv'
 }

### 2019-01-31 paper/paper.Rmd
update objective function -- integrate comments from NDG about emphasis on novelty of objective and our framing.

@@
 abstract: 
-    Languages display a diverse set of distributional regularities such as the relation between a word's frequency and rank in a corpus, the distribution of dependency lengths, or the presence of lexical properties such as ambiguity.
-We discuss a framework for studying how these properties emerge from in-the-moment interactions of rational, pragmatic speakers and listeners.
-Our work takes notions of lexicon-level efficiency as a starting point, connecting these ideas to notions of conversational-level efficiency.
-To do so, we derive an objective function for measuring the communicative efficiency of linguistic systems and then examining the behavior of this objective in a series of simulations focusing on the communicative function of ambiguity in language.
-These simulations suggest that rational pragmatic agents will produce communicatively efficient systems and that interactions between such agents provide a framework for examining efficient properties of language more broadly.
+    Languages display a diverse set of distributional regularities such as the relation between a word's frequency and rank in a corpus, the distribution of syntactic dependency lengths across languages, or the presence of properties such as ambiguity.
+We discuss a framework for studying how these system-level properties emerge from local, in-the-moment interactions of rational, pragmatic speakers and listeners.
+To do so, we derive a novel objective function for measuring the communicative efficiency of linguistic systems in terms of the interactions of speakers and listeners.
+We examine the behavior of this objective in a series of simulations focusing on the communicative function of ambiguity in language.
+These simulations suggest that rational pragmatic agents will produce communicatively efficient systems and that interactions between such agents provide a framework for examining efficient properties of language structure and use more broadly.
 
@@
 Zipf articulated what is now considered a *functionalist* approach to language science -- analyzing language structure and use in terms of efficiency.
-Such an approach might reframe our opening question as follows -- how does having \textbf{property x} make using language $\ell$ more or less useful for communication?
-This reframing provides an opportunity to study both language *structure* (as Zipf primarily did) or *use* in terms of efficiency.
-For example, Regier et al.
-[-@RegierKempKay2015a] showed that languages appear to organize semantic domains (form-meaning mappings) in maximally efficient ways.
-Likewise, Piantadosi et al.
-[-@Piantadosi2011a] argued that lexical ambiguity is an efficient property of any communication system when communication is contextualized.\par
+Such an approach might reframe our opening question as follows: how does having property \textit{x} make using language $\ell$ more or less useful for communication?
+Among others, this efficiency-based framing has produced a rich set of theoretical and empirical targets exploring semantic typology [@RegierKempKay2015a], properties such as ambiguity [@Piantadosi2011a] and compositionality [@KirbyGriffithsSmith2014a], and the efficient use of reduction and redundancy in production [@LevyJaeger2007a; @GenzelCharniak2002a].
+These projects typically employ information-theoretic principles to characterize efficiency in natural or simulated language corpora.
+We adopt a different approach in this work, deriving a measure of efficiency from first principles of \textit{efficient language use} and then showing how this can lead to communicatively efficient systems.\par
 
-Zipf’s original work and subsequent functionalist projects describe a varied set of linguistic properties and human behaviors.
-Common to many is a characterization of a fundamental effort-asymmetry underlying everyday communication.
-Simply put, what is “hard” for a speaker is likely different from what is “hard” for a listener.
-Zipf characterized this as follows -- purely from the standpoint of speaker effort, an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of just a single, low-cost word.
+Functionalist theories commonly frame language efficiency in terms of a fundamental effort-assymetry underlying everyday communication: what is “hard” for a speaker is likely different than what is “hard” for a listener.
+Zipf gave the following intution for how this idea -- purely from the standpoint of speaker effort, an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of a single, low-cost word.
 Given such a language, the full set of potential meanings would be conveyed using only that word, i.e.
@@
 In this example, speaker effort is related to *production cost* and listener effort to *understanding or disambiguation cost*.
-Clearly natural languages fall between the two extremes of $\ell_{speaker}^*$ and $\ell_{listener}^*$.
+Clearly, natural languages fall between the two extremes of $\ell_{speaker}^*$ and $\ell_{listener}^*$.
 Zipf proposed that the particular lexicon-level properties he observed were a result of these competing forces -- the pressure to jointly minimize speaker and listener effort.\par
 
-But how does such optimization take place?
+But how does this optimization take place?
 The example given by Zipf [-@Zipf1949a], describes local, communicative interactions in terms of a \textit{reference game}.
@@
 The listener attempts to reconstruct this intended meaning given the transmitted utterance, $m \rightarrow u$.
-Other functionalist projects have assumed this basic reference game setting [@RegierKempKay2015a; @Piantadosi2011a] and this simplification of the communicative act has proven productive in theoretical [@FerreriCancho2018a], simulation-based [@KirbyGriffithsSmith2014a] and empirical explorations [@HawkinsFrankeSmithGoodman2018a] of efficient language structure and use.\par
+Other projects have assumed this basic reference game setting [@RegierKempKay2015a; @Piantadosi2011a] and this simplification of the communicative act has proven productive in theoretical [@FerreriCancho2018a], simulation-based [@KirbyGriffithsSmith2014a] and empirical explorations [@HawkinsFrankeSmithGoodman2018a] of efficient language structure and use.\par
 
 Zipf's conception of speaker and listener effort may be connected to speakers' pragmatic reasoning in the moment.
-Under a Gricean treatment of pragmatics, speakers and listeners are assumed to follow a set of conversational maxims in which they cooperate to transfer information (Grice, 1975).
+Under a Gricean treatment of pragmatics, speakers and listeners follow a set of conversational maxims in which they cooperate to transfer information [@Grice1975a].
 These maxims appear to emerge from efficiency concerns, however (Horn, 1984).
-A primary target of the current work is to formalize this connection, by providing a framework for deriving system-level efficiencies from local, interaction behavior.
-Our claim is that to understand an "efficient" property of a system we need to consider *how* that property is *used* efficiently.
-Additionally, any quantitative description of "efficiency" should be a function of both the system and users.\par
+We formalize this connection -- showing how system-level efficiencies can emerge from local interaction behavior.
+Intuitively, our claim is that to understand an "efficient" property of a system it is essential that we consider how that property is *used* efficiently.
+It follows that any quantitative measure of "efficiency" must be a function of both the system and its users.\par
 
-In the current work we provide a case-study for the kind of framing presented above.
+In the current work we provide a case-study for the framing presented above.
 We choose a property of languages that could, in principle, vary freely, but shows strong regularities across languages.
 The explanadum is why this regularity holds.
-We examine lexical ambiguity as our property, extending ideas by Piantadosi et al.
+We examine ambiguity as our property, extending ideas by Piantadosi et al.
 [-@Piantadosi2011a].
-We define a novel measure of efficiency that depends on the interactional behavior of interlocutors.
-For our model of speakers and listeners we adopt the Rational Speech Act (RSA) framework, which is nicely attested in experimental data on pragmatic interaction.
-With these basic ingredients we have a measure of efficiency given language and speaker-listener agents.
-Finally, we show that the property of interest is prevalent in languages that optimize this measure of efficiency and also how this property is used efficiently during local interaction.
-The primary contribution of this work is to show how the reference game frameowrk, in combination with formal models of communication, can be used to connect ideas about system-level efficiencies to in-the-moment language use.\par
+We define a novel measure of efficiency that depends on the interactional behavior of speaker and listener agents.
+We adopt the reference game as our primary unit of interaction and model language users using the Rational Speech Act (RSA) framework -- a computional model of language use, which is nicely attested by experimental data on interaction.
+Using these basic ingredients, we show that the property of interest (ambiguity) is prevalent in languages that optimize our measure of efficiency.
+Further, we show how ambiguity is \textit{used} efficiently during local, in-the-moment interactions.
+The contributions of this work are twofold -- we derive a novel objective for measuring linguistic efficiency and also show how the reference game framework, in combination with formal models of communication, can be used to connect ideas about system-level efficiencies to in-the-moment language use.\par
 
@@
 We adopt that formulation here, assuming these communication games as our basic unit of analysis.
-In this setting, speakers and listeners are aware of a set of objects $M$, which we will refer to as *meanings* and are knowledgeable about the set of possible signals $U$ (*utterances*) that can be used to refer to a given meaning (see Figure 1).
+In this framework, speakers and listeners are aware of a set of objects $M$ (*meanings*) and are knowledgeable about the set of possible signals $U$ (*utterances*) that can be used to refer to a given meaning (see Figure 1).
 Utterances may have different relative costs, operationalized via a prior over utterances $P(U)$.
-Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$\footnote{The prior over meanings are analogous to the \textit{need probabilities} assumed in previous work Regier, Kemp \& Kay (2015).}.
-We consider a set of contexts $C$ with an associated prior $P(C)$.
-Contexts encode different meaning priors, formalized as separate conditional distributions over meanings e.g.
+Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$\footnote{The prior over meanings are analogous to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}.
+We consider a set of contexts $c\in C$ with an associated prior $P(C)$.
+Each context $c_i$ describes a different distribution over meanings e.g.
 $p(M|C=c_i) \neq p(M|C=c_j)$.
-Finally, we consider a set of communicative events $e \in E$ where $<u, m> = e$ is a tuple of utterance, meaning pairs.
-Note that in the case of single context $|C|=1$ we simply have a prior over meanings $p(M)$, a simplification we will use in introducing our objective function.\par
+Finally, we consider a set of communicative events $e \in E$ where $<u, m, c> = e$ is an utterance, meaning and context triple.\par
 
@@
 For example, Figure 1 contains four utterances $U = \{\text{blue}, \text{green}, \text{square}, \text{circle}\} \text{ and three meanings }M = \{\text{green-square}, \text{blue-square}, \text{green-circle}\}$.
-The boolean matrix describes the literal semantics of the language $\ell$.
-Note that will consider language to "contain ambiguous items" if there is at least one utterance $u \in U$ which can apply to multiple meanings (i.e.
-$|[[u_i]]| > 1$)\footnote{We use double brackets $[[\dots]]$ to represent denotation}.
-For example, in Figure 1 both the words "blue" and "green" are ambigous so we would say that $\ell$ contains ambiguity.
+The boolean matrix describes the literal semantics of the language.
+We define a language as "ambiguous" if there is some utterance $u \in U$ which can apply to multiple meanings (i.e.
+$|[[u_i]]| > 1$)\footnote{We use double brackets $[[\dots]]$ to represent denotation.}.
+In Figure 1 both the words "square" and "green" are ambigous so we would say that $\ell$ contains ambiguity.
 
 ### Speakers and listeners
-The Rational Speech-act framework (RSA) is computational-level theory of pragmatic language use.
-RSA can be understood as a formalization of essential Gricean pragmatic principles -- agents reason about one another and their shared context.
-For this reason, we adopt RSA as our representational framework to operationalize Gricean (rational and pragmatic) speaker-listeners [@Grice1975a].
-We refer the reader to Section 1 of our supplementary materials for a more detailed description of basic RSA speaker-listener definitions, and to Goodman & Frank [-@GoodmanFrank2016a] for an overview of its appliations.\par
+The Rational Speech Act framework (RSA) is computational-level theory of pragmatic language use, which has been well attested in experimental work examining both speaker and listener communication behavior [@FrankGoodman2012a; @GoodmanFrank2016a].
+RSA is a formalization of essential Gricean pragmatic principles -- agents reason about one another and their shared context [@Grice1975a].
+We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speaker-listeners in the reference game setting\footnote{See SI section 1 for a more detailed description of basic RSA speaker-listener defintions and to Goodman \& Frank [-@GoodmanFrank2016a] for an overview of its applications.}.\par
 
 An RSA *speaker agent* defines a conditional distribution over utterances, mapping from intended meanings $M$ to utterances $U$ using $\ell$ in a given context $c$.
-That is, a speaker defines $P_{speaker}(u|m; \ell)$.
-We will use $S(u|m; \ell)$ as short-hand throughout.
-A *listener agent* defines a conditional distribution over meanings, mapping from utterances $U$ to meanings $M$ using $\ell$ in a given context $c$.
-We will use $L(m|u; \ell)$ as shorthand.
-Note that both speakers and listeners can induce joint distributions over the set of all events $E$, although, importantly, these distributions may differ:
-$$P_{speaker}(u, m; \ell) = S(u|m; \ell)p(m)$$
-$$P_{listener}(u, m; \ell) = L(m|u; \ell)p(u)$$
+That is, a speaker defines $P_{speaker}(u|m, c; \ell)$.
+We will use $S(u|m, c; \ell)$ as short-hand throughout.
+A *listener agent* defines a conditional distribution over meanings, mapping from utterances $U$ to meanings $M$ using $\ell$ in a given context $c$ (i.e.
+$L(m|u, c;\ell)$).
+Note that both speakers and listeners can induce joint distributions over utterace-meaning pairs, although, importantly, these distributions may differ:
+$$P_{speaker}(u, m | c; \ell) = S(u|m, c; \ell)p(m|c)$$
+$$P_{listener}(u, m| c; \ell) = L(m|u, c; \ell)p(u|c)$$
 
@@
 Zipf (1949) proposed that the particular distributional properties found in natural language emerge from competing speaker and listener pressures.
-We operationalize this objective in equation (1) -- the efficiency of a linguistic system $\ell$ being used by speaker and listener agents $S$ and $L$ is the sum of the expected speaker and listener effort to communicate over all possible communicative events $E$.
-Note that to simplify the introduction of this objective we assume there is just a single context ($|C|=1$).
-In this case $p(C=c)=1$ and there is only distribution over meanings ($p(m|c) = p(m)$).
-We provide a more detailed description of this derivation in the supplementary materials, both with and without context $C$.\par
+We operationalize this objective in equation (1) -- the efficiency of a linguistic system $\ell$ being used by speaker and listener agents $S$ and $L$ is the sum of the expected speaker and listener effort to communicate over all possible communicative events $e \in E$.\par
 
@@
 
-We assume that speaker effort is the negative log probability of an utterance (surprisal) -- intuitively, the number of bits needed to encode the utterance $u$.
-This particular formalization of speaker-cost is general enough to accommodate a range of instantiations that might in theory be related to production difficult via articulation effort, cognitive effort related to lexical access, etc [@BennettGoodman2015a].\par
+We assume that speaker effort is related to the suprisal of an utterance in a particular context\footnote{In the current set of simulations we consider utterances costs as independent from context (ie.
+$p(u|c)p(c)=p(u)p(c)$).} -- intuitively, the number of bits needed to encode the utterance $u$.
+This particular formalization of speaker-cost is general enough to accommodate a range of cost instantiations, such as production difficult via articulation effort, cognitive effort related to lexical access, or others [@BennettGoodman2015a].\par
 
-$$\text{speaker effort} = -log_2(p(u))$$
+$$\text{speaker effort} = -log_2(p(u|c))$$
 
-We assume listener effort is the negative log probability of the listener's conditional $-log_2(L(m|u))$.
-This operationalization of listener effort is intuitively related to existing work in sentence processing in which word difficulty during comprehension is proportional to surprisal [@Hale2001a; @Levy2008a].
+We assume listener effort is the semantic suprisal of a meaning given an utterance.
+This operationalization of listener effort is intuitively related to existing work in sentence processing in which word comprehension difficulty is proportional to surprisal [@Hale2001a; @Levy2008a].
 
-$$\text{listener effort} = -log_2(L(m|u; \ell))$$
+$$\text{listener effort} = -log_2(L(m|u, c; \ell))$$
 
-Importantly, we assume that events $e = <u, m>$ are sampled according the to following generative model -- some object occurs in the world with probability $p(M=m)$, the speaker attempts to refer to that object by sampling from her conditional distribution $S(u|m)$ (i.e.
-$e \sim p(m)S(u|m)$).
-From the ingredients highlighted above, it is possible to derive the following objective between the speaker and listener joint distributions (see SI 2.2 for a complete derivation).
+Importantly, we assume that events $e = <u, m, c>$ are sampled according the to following generative model -- some context occurs in the world with probability $P(C=c)$.
+Within this context, an object $m$ occurs with probability $p(m|c)$.
+The speaker attempts to refer to that object by sampling from her conditional distribution $S(u|m, c; \ell)$ (i.e.
+$e \sim p(c)p(m|c)S(u|m, c; \ell)$).
+From the ingredients described above, it is possible to derive the following objective between the speaker and listener joint distributions (see SI 2.2 for a complete derivation).
 
@@
 \begin{split}
-  = -\mathbb{E}_{P_{speaker}}[log_2(P_{listener})] \\
-  = H_{cross}(P_{speaker}, P_{listener}; \ell)
+  = \mathbb{E}_{c \sim P(C)}[H_{cross}(P_{speaker}, P_{listener} | c; \ell)]\\
 \end{split}
 \end{equation}
-From an information-theoretic perspective this objective is intuitive: cross-entropy gives us a measure of dissimilarity between two distributions -- the average number of bits required to communicate under one distribution, given that the “true” distribution differs.
-In our case, this is the difference between the joint distribution assumed by the speaker $P_{speaker}$ and listener $P_{listener}$.
+From an information-theoretic perspective this objective is intuitive: The Cross-Entropy (CE) gives us a measure of dissimilarity between two distributions -- the average number of bits required to communicate under one distribution, given that the “true” distribution differs.
+In our case, we have an expectation over this term -- the expected difference between the distributions assumed by the speaker $P_{speaker}$ and listener $P_{listener}$ given a set of contexts $C$.
 In other words, an "efficient" language $\ell$ minimizes the distance between what speakers and listeners think.
@@
 
-The task of understanding language is marked by a frequent need to handle various forms of ambiguity; lexical, syntactic, among others [@WasowPerforsBeaver2005a].
-The ubiquity of this property, however, has been argued to provide evidence that langauges are not been optimized for communication\footnote{Chomsky (2002) famously claimed "If you want to make sure that we never misunderstand one another, for that purpose language is not well designed, because you have such properties as ambiguity." Note that this analysis is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$ we described earlier and define in terms of equation 2 in SI 2.2 .} [@Chomsky2002a].
+The task of understanding language is marked by a frequent need to handle various forms of ambiguity: lexical, syntactic, among others[@WasowPerforsBeaver2005a].
+The ubiquity of this property, however, has been argued to provide evidence that langauges are not been optimized for communication\footnote{Chomsky's (2002) claim that "...
+language is not well designed, because you have such properties as ambiguity" is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$.} [@Chomsky2002a].
 
@@
 
-We provide a toy example of their argument here.
-Say we have two objects in the world $m_1$ and $m_2$, and two languages $\ell_1$ and $\ell_2$.
-In language $\ell_1$, the low-cost utterance $u_1$ can be used to refer to both $m_1$ and $m_2$ (i.e.
-$[[u_1]]_{\ell_1} = \{m_1, m_2 \}$ and $[[u_2]]_{\ell_1} = \emptyset$).
-With language $\ell_2$, however, the low cost utterance $u_1$ can only refer to $m_1$ and a higher-cost utterance $u_2$ can refer to $m_2$ (i.e.
-$[[u_1]]_{\ell_2} = \{m_1\}$ and $[[u_2]]_{\ell_2} = \{m_2 \}$).
-While it is cheaper for a speaker to use $\ell_1$ (because she can always use a low-cost utterance), it is more difficult for a listener (because $u_1$ is ambiguous).
-If context isn't disambiguating then we might prefer $\ell_2$ to $\ell_1$.
-But, if context is disambiguating then the speaker can use $u_1$ to refer to either $m_1$ or $m_2$ and $\ell_1$ should be preferred.\par
+We re-iterate their argument here.
+Say we have two objects ($m_1$ and $m_2$), two utterances ($u_1$ and $u_2$), which differ in cost, and two languages ($\ell_1$ and $\ell_2$) which describe different utterance-meaning mappings.
+In language $\ell_1$, the low-cost $u_1$ can be used to refer to both $m_1$ and $m_2$ ($[[u_1]]_{\ell_1} = \{m_1, m_2 \}$), but the high-cost $u_2$ cannot be used at all ($[[u_2]]_{\ell_1} = \emptyset$).
+By contrast, in language $\ell_2$, $u_1$ can only refer to $m_1$ and $u_2$ can only refer to $m_2$ ($[[u_1]]_{\ell_2} = \{m_1\}$ and $[[u_2]]_{\ell_2} = \{m_2 \}$).
+While it is cheaper for a speaker to use $\ell_1$ (because speaking is always low-cost), it is more difficult for a listener (because $u_1$ is ambiguous).
+Crucially, however, if context is disambiguating then the speaker can use $u_1$ to refer to either $m_1$ or $m_2$ and $\ell_1$ should be preferred to $\ell_2$.\par
 
@@
 ```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=8, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Optimal languages are more likely to contain ambiguous items as the amount of contextual information increases.
-Vertical axis shows the proportion of optimal languages that contain ambiguity.
-Horizontal axis shows the number of contexts in each condition (1-4).
+Vertical axis shows the proportion of optimal languages containing ambiguity.
+Horizontal axis shows the number of context-sizes (1-4).
 Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only objectives.
-Error bars represent 95% confidence intervals."}
+Error bars represent 95 percent confidence intervals."}
 img <- png::readPNG("figs/fig1.png")
@@
 
-We adopt the framework above, and (following Piantadosi et al.
-2011), add \textit{context} ($C$).
-The authors argue that mapping two meanings $m_1$ and $m_2$ to a single utterance $u_1$ is useful when they can be disambiguated *in context*.
-In our case we consider a context $c_i \in C$ to specify a permutation of the meaning probabilities.
-That is, $p(m|c_1) \neq p(m|c_2)$, when there are two contexts $|C| = \{c_1, c_2\}$.This in turn, leads to an update to our linguistic efficiency objective as we now would like to consider the average speaker-listener effort over all contextualized communicative events (see section 3 SI).\par
-
+We adopt the hypothesis presented by Piantadosi, formalizing their argument in our current framework -- showing that ambiguity is an efficient property under our CE objective in the reference game setting.
 We proceed by generating languages with different amounts of contextual support (varying the size of $|C|$).
-We search the space of languages, asessing if the ones which minimize our objective contain ambiguity.
-If context leads to more efficient commnication, then optimal languages should be more likley to be ambiguous as the amount of context increases.\par
+We search the space of languages, examining whether ones which minimize our objective contain ambiguity.
+If context leads to more efficient commnication, then optimal languages under our objective should be more likley to be ambiguous as the amount of context increases.\par
 
@@
 For each simulation we enumerate the set of *valid* languages in which $|U|=|M|=4$ (recall that $U$ is our set of utterances and $M$ our set of meanings).
-Note that a language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment) and each form maps to at least one meaning (every column has some non-zero assignment)\footnote{For the current set of simulations we consider languages in which there are four or five true values -- in a "four" language ambiguity is not possible, but in a "five" language there are ambiguous mappings.}.
-For a given simulation the goal is to find the language $\ell^*$ which minimizes our cross-entropy objective and then check to see if that language contains ambiguity.\par
+Note that a language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment) and each form maps to at least one meaning (every column has some non-zero assignment).
+For a given simulation the goal is to find the language $\ell^*$ which minimizes our CE objective and then check to see if that language contains ambiguity.\par
 
-Recall that language efficiency is both a function of the particular semantic mappings induced by that language, the speaker and listener agents ($S$ and $L$), as well as the utterance ($P(U)$), meaning ($P(M)$), and context priors ($P(C)$).
-Rather than assume particular structure for our utterance, meaning, and context prior distributions, for each simulation we generate $P(U) \sim \text{Dir}(1, |U|)$, $P(M|C=c) \sim \text{Dir}(1, |M|)$ (a separete conditional distribution over meanings for each context $c$), and $P(C) \sim \text{Dir}(1, |C|)$ where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.\par
+We define language efficiency as a function of the particular semantic mappings induced by that language, the speaker and listener agents ($S$ and $L$), as well as the utterance ($P(U)$), meaning ($P(M)$), and context priors ($P(C)$).
+Rather than assume particular structure, for each simulation we generate $P(U) \sim \text{Dir}(1, |U|)$, $P(M|C=c) \sim \text{Dir}(1, |M|)$ (a separate conditional distribution over meanings for each context $c$), and $P(C) \sim \text{Dir}(1, |C|)$, where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.\par
 
@@
 Following the argument given by Piantadosi et al.
-[-@Piantadosi2011a], we want to assess the impact of *context* on our objective.
+[-@Piantadosi2011a], we want to assess the impact of *context* on the presence of ambiguity in optimal languages.
 To do so we consider four conditions with $n=500$ simulations each (that is, 500 unique sets of $\{P(U), P(M|C), P(C)\}$.
-Our first is a *single-context* condition ($|C|=1$) -- there is a only a single context describing $P(M)$.
-Our second condition contains two-contexts  ($|C| = 2$) -- we consider efficiency under both $P(M|C=c_1)$ as well as $P(M|C=c_2)$.
-The third and fourth condition correspond accordingly with $|C|=3$ and $|C| = 4$, respectively.\par
+Our first is a *one-context* condition ($|C|=1$) -- only a single distribution over meanings $P(M)$.
+In our two-context condition ($|C| = 2$), we consider efficiency under both $P(M|C=c_1)$ as well as $P(M|C=c_2)$.
+Three- and four-context conditions corresponding accordingly.\par
 
@@
 For comparison, we also examine properties of optimal languages under two additional objectives.
-Zipf [-@Zipf1949a] proposed that the optimal speaker language $\ell_{speaker}^*$ would only optimize speaker effort and the optimal listener language $\ell_{listener}^*$ would only optimize listener effort.
+Zipf [-@Zipf1949a] proposed that the speaker-optimal language $\ell_{speaker}^*$ would minimize speaker effort and the listener-optimal language $\ell_{listener}^*$ would minimize listener effort.
 We define these objectives using the first and second half of equation 1 (see SI section 2.2.
@@
 Figure 1.
-plots the proportion of optimal languages under each objective, as a function of context condition.
-The red line shows that as the number of contexts increases, so does the probability that the optimal language $\ell^*_{cross}$ under our speaker-listener cross-entropy objective contains ambiguity.
-We also plot the proportion of optimal speaker-only $\ell^*_{speaker}$ and listener-only $\ell^*_{listener}$ languages that contain ambiguity.
-In line with Zipf's predictions, if languages are designed only to minimize speaker effort then optimal languages will assign lower-cost forms multiple meanings (all of the speaker-optimal languages contained ambiguity).
-Likewise, if languages are designed only to minimize listener effort then ambiguity should always be avoided (none of the listener-optimal languages contained ambiguity).\par
+plots the proportion of optimal languages under each objective as a function of context size.
+The red line shows that as the number of contexts increases, so does the probability that an optimal language $\ell^*_{cross}$ contains ambiguity under our CE objective.
+For comparison we also plot the proportion of speaker-optimal $\ell^*_{speaker}$ (blue line) and listener-optimal $\ell^*_{listener}$ (green line) languages that contain ambiguity.
+In line with Zipf's predictions, if languages are designed only to minimize speaker effort then optimal languages always contain ambiguity.
+If languages are designed to minimize listener effort then ambiguity is always avoided.\par
 
-We explored the degree to which ambiguity improved efficiency during communication between rational, pragmatic agents.
-We showed that as contextual information increased, the optimal languages under our Cross-Entropy objective were more likely to have ambiguous lexical items.
+In simulation 1 we explored the degree to which ambiguity was an efficient property when communication was contextualized.
+We showed that as contextual information increased, the optimal languages under our CE objective were more likely to have ambiguous lexical items.
 Importantly, we did not see this effect under alternative objectives which only emphasized speaker- or listener effort alone.\par
 
-While our results indicate that ambiguity is an efficient property of contextualized language these simulations assumed that speaker-listener agents could always disambiguate items in context.
+While our results indicate that ambiguity is an efficient property of contextualized language, these simulations assumed that agents could always disambiguate items in context.
 That is, in our four conditions both agents had perfect knowledge of the relevant conditional distributions ($P(M|C)$).
-This assumption may be too strong for describing much of day-to-day communication -- we seldom begin a conversation with perfect knowledge of the current context of a conversation.
-We explore a case in which the listener has imperfect knowledge of context at the start of the conversation in Simulation 2.\par
+This assumption may be too strong for describing much of day-to-day communication -- we seldom interact with others with perfect knowledge of the current context (or topic) of conversation.
+To explore how ambiguity may be \textit{used} efficiencly, we next explore a case in which the listener has imperfect knowledge of context at the start of the conversation, but may infer it from a discourse history in Simulation 2.\par
 
@@
 
-We shift time-scales in Simulation 2 -- instead of examining the objective landscape of a space of languages, we now consider a single language that contains ambiguity and examine how that property is used over a discourse.
-If context is not informative with respect to meaning early in a discourse (position $i$), but is at later positions at position (position $i+k$), then the speaker should avoid using ambigous material early $i$, only using it $k$ steps later in the conversation.
+We shift focus in Simulation 2 -- instead of examining the landscape of a space of languages under our CE objective, we consider a single language containing ambiguity, examining how that property is used over a discourse.
+If context is not informative with respect to meaning early in a discourse, but is at later positions, then an efficient speaking strategy would avoid ambiguous material early, saving it for when contextual information is disambiguating.
 In this toy example we can consider “context” as analogous to a “topic” of conversation.
-Intuitively, we can imagine a scenario in which a reader is beginning a newspaper article.
+As an example, imagine a scenario in which a reader is beginning a magazine article.
 While they may have some knowledge about the article’s topic (perhaps from the title), they may not have complete knowledge of its contents, including the persons or events involved.
-In this setting, using an ambiguous pronoun early (position $i$) in the article can lead to misunderstanding without sufficient context to support disambiguation.
-But if by position $i+k$ there is sufficient contextual information it may be efficient for the writer to use ambiguous material.\par
+In this setting, using an potential low-cost, but ambiguous pronoun early may lead to misunderstanding if context is sufficiently informative.
+But, if by a later position enough contextual information has accumulated, it may be efficient for the writer to use the ambiguous material.\par
 
@@
 We consider consider a single language $\ell$, which contains both ambiguous and unambiguous utterances.
-The ambiguous utterances are less costly than the unambiguous utterances.
+We assume ambiguous utterances are lower cost.
 Crucially, we do not assume that the listener knows the particular topic ($c_{current}$) of the conversation \textit{a priori}.
-Rather, we assume that the listener has knowledge of the set of possible topics $C = \{c_1, \dots, c_k\}$, but \textit{does not know which one is currently being used by the speaker}.
-Formally, this means the listener does not have access to the correct conditional distribution over meanings $P(M|C=c_{current})$ at the start of the discourse.
-In this setting, a speaker minimizing production effort would like to use the low-cost material as frequently as possible.
-However, this will incur a disambiguation cost to the listener if they are uncertain of the current context $c_{current}$ (topic of conversation).\par
+Rather, that the listener has knowledge of the set of possible topics $C = \{c_1, \dots, c_k\}$, but \textit{does not know which one is currently being used by the speaker}.
+Formally, this means the listener does not have access to the correct conditional distribution over meanings $P(M|C=c_{current})$ at the start of the discourse.\par
 
-Over the course of a discourse, $D$ the listener tries to infer both the current topic, $c_{current}$, as well as the particular meaning $m$ of a given utterance $u$.
-That is, we consider updated listener and speakers ($L(m, c|u,D;\ell)$, $S(u|m,c,D;\ell)$).
-Note that both the speaker and listener can track the history of previous utterances $D$.
+Over the course of a discourse $D$, the listener tries to infer both the current topic, $c_{current}$, as well as the particular meaning $m$ of a given utterance $u$.
+That is, we consider updated listener and speakers ($L(m, c|u,D;\ell)$, $S(u|m,c,D;\ell)$) in which both agents can track the history of previous utterances $D$.
 Importantly, an agent can attempt to infer the current topic of conversation $c_{current}$ using the discourse history $D$ (see SI section 3 for more detail).\par
 
-```{r plot-optimal-use, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=2.5, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "(A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position.
+```{r plot-optimal-use, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "(A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position.
 (B) shows speaker effort across the three models.
 (C) shows the Cross-Entropy objective under our three speaker models.
-Error bars represent 95% confidence intervals."}
+Error bars represent 95 percent confidence intervals."}
 
@@
 
-We conduct $N=600$ simulations, generating discourses of length $|D|=30$ utterances with three different speaker models ($n=200$ each).
-We consider a single language $\ell$\footnote{See SI for the matrix notation of this langauge.} with $|U|=6$ and $|M|=4$ in which two of the utterances $u_5$ and $u_6$ are ambiguous and that $p(u_5) = p(u_6) > p(u_{1})=\dots=p(u_{4})$.
-That is, the two ambiguous utterances ($u_5$ and $u_6$) are less costly than the non-ambiguous utterances.
-(Note that use of this particular language is not essential -- the results are broadly generalizable languages that contain ambiguity.)\par
+We conduct $N=600$ simulations, generating discourses of length $|D|=30$ utterances, comparing three speaker models ($n=200$ each).
+We consider a single language $\ell$\footnote{See SI for the matrix notation of this langauge.} with $|U|=6$ and $|M|=4$ in which two of the utterances are ambiguous and lower cost than the unambiguous utterances.
+(Note that use of this particular language is not essential -- the results are broadly generalizable to languages that contain ambiguity.)\par
 
@@
 We consider three types of speaker models.
-In our \textit{Full pragmatics} model, we consider a speaker agent who reasons about her listener and also has complete recall of the set of utterances in the discourse $D$.
-This speaker believes that the listener may not know the current topic $c_{current}$ at the start of the discrouse, but can infer it with enough data.
-In this way, contextual information shared by the speaker and listener may increase over the discourse (as the listeners estimate of the current topic ($P(C=c|D)$) comes in line with the actual topic known to the speaker).
+Our \textit{Full pragmatics} agent, models a speaker who reasons about her listener and also has complete recall of the set of utterances in the discourse $D$.
+This speaker believes that the listener may not know the current topic $c_{current}$ at the start of the discrouse, but should infer it over the discourse.
 We compare two baselines models.
 The first, a \textit{Partial pragmatics} baseline describes a speaker who reasons about a listener, but assumes they have no access to discourse history.
-The second is a \textit{No pragmatics} speaker who does not consider a listener at all, but produces utterances according to the underlying langauge semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$) (see SI section 3 for details).\par
+The second, a \textit{No pragmatics} speaker who does not consider a listener at all, but produces utterances according to the underlying langauge semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$) (see SI section 3 for details).\par
 
@@
 
-We assess whether it is efficient to use low-cost, ambiguous material when context is disambiguating.
-In the current setting, this corresponds to speaker behavior in which ambiguous forms are avoided early in the discourse, but prefered later in the discourse once the listener is confident of the topic.
-We should expect this to be reflected in our \textit{Full pragmatics} model.
-By contrast, our \textit{Partial pragmatics} speaker does not believe that there is sufficient contextual information for the listener to disambiguate her utterances.
-This should lead to total  avoidance of ambiguous material over the discourse.
-Finally, our \textit{No pragmatics} speaker should use low-cost material greedily since she does not consider a listener's need to disambiguate whatsoever.\par
+We are interested in \textit{when} it is efficient to use low-cost, ambiguous material in this discourse setting.
+A speaker strategy that is mutually efficient for both agents should avoid ambiguity until sufficient contextual information has accumulated.
+We should expect this to be reflected in our \textit{Full pragmatics} model who reasons about the listener and discourse history.
+By contrast, a speaker-optimal model who does not consider the listener should greedily use ambiguous utterances (\textit{No Pragmatics} model), while a listener-optimal model should avoid ambiguity entirely (\textit{Partial pragmatics} model).\par
 
@@
 Figure 2, (A) shows the empirical probability that a speaker uses an ambiguous utterance as a function of discourse position.
-The \text{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse and the \text{Partial pragmatic} baseline avoids ambiguous utterances entirely (reasoning that the utterances cannot be disambiguated).
-But, the \text{Full pragmatic} model avoids ambiguous material at first, but employing them gradually increasingly as the discourse proceeds.
-(B) shows the speaker-listener Cross-Entropy objective.
-Note that the objective decreases for all three models.
-This decrease in all three models is primarily driven by the listener updating his belief about the actual topic ($P(C=c_{current}|D)$).
-Of particular importance is the value of the objective at the end of the discourse (position $i=30$).
-The difference in the Cross-Entropy between the \text{Full Pragmatic} and \text{Partial Pragmatic} models at the end of the discourse is driven by the reduction in speaker costs described in (C).
-That is, while speaker effort remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines, effort declines for the \text{Full Pragmatic} model as they are increasingly able to use ambiguous material later in the discourse.\par
+The \textit{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse and the \textit{Partial pragmatic} baseline avoids ambiguous utterances entirely.
+But, the \textit{Full pragmatic} model avoids ambiguous material only at the start of the discourse, employing it increasingly as the discourse proceeds.
+(C) tracks our CE objective for each model over the discourse.
+Note that the objective decreases for all three models, primarily driven by the listener updating his belief about the actual topic ($P(C=c_{current}|D)$).
+However, the objective declines more quickly under the \textit{Full} and \textit{Partial} pragmatic speakers as listener agents are better able to infer the correct context.
+Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatic} models at the end of the discourse is driven by the reduction in speaker costs described in (B).
+While speaker effort remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines, effort declines in the \text{Full Pragmatic} model as she increasingly relies on ambiguous material later in the discourse.\par
 
-In Simulation 2 we explored how ambiguity could be employed efficiently during discourse.
+In Simulation 2 we explored how ambiguity could be employed efficiently during language use.
 We assumed that the listener did not know the current context (topic) \textit{a priori}, but could infer it from the discourse history.
 We examined speaker production behavior, tracking use of low-cost, but ambiguous material over the discourse.
-Early in the discourse the \textit{Full pragmatic} speaker avoided using ambiguous material, only using the material later when the context was known to the listener -- it is useful to use low-cost material, but only when context is disambiguating.
+Early in the discourse the \textit{Full pragmatic} speaker avoided using ambiguous material, only using the material later when the current topic was known to the listener.
 
@@
 We explored the interactions of rational pragmatic agents as a framework for understanding efficient language structure and use.
-We focused on an argument on the communicative function of ambiguity [@Piantadosi2011a].
-We derive a novel speaker-listener cross-entropy objective for measuring the efficiency of a language in a reference game setting showing that optimal languages are more likely to contain ambiguous material when context is informative.
-Further we show that rational pragmatic agents will *use* ambiguous material efficiently, only when such use is supported by context.\par
+We focused on an argument on the communicative function of ambiguity [@Piantadosi2011a], deriving a novel speaker-listener Cross-Entropy objective for measuring the efficiency linguistic systems from first principles of efficient language use.
+In a series of simulations we showed that optimal languages are more likely to contain ambiguous material when context is informative and that rational pragmatic agents will *use* ambiguous material efficiently.\par
 
-The framework employed here -- the analysis of the dynamics of repeated reference games between rational pragmatic agents -- provides a tool for the analysis of functionalist theories, allowing linkage across timescales of analysis from conversation [@LevyJaeger2007a; @GenzelCharniak2002a] to typology [@RegierKempKay2015a].
-In theory, this framework provides extensive opportunity to vary properties of the languages and agents, however we have focused on only a small subset here.
-For example, examining how increasing or decreasing \textit{rationality} (instantiated through the RSA $\alpha$ parameter) or \textit{pragmatics} (instantiated through the RSA recursion depth) impacts these results is a clear avenue for future work.
-An important limitation of the current work is the broad assumption that language can be described through such reference games.
-The degree to which properties of non-referential language can be captured in such a framework remains an important target for future work.
-
-\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering All simulation and analysis code are available in the public repository for the project: (link will be available upon acceptance)}} \vspace{1em}
-\noindent
+A limitation of the current work is an analysis of exactly how this objective compares to existing measures.
+For example, previous work has described competing speaker-listener pressures in terms of a trade-off of \textit{simplicity} and \textit{informativeness} [@KempRegier2012a] or \textit{expressivity} and \textit{compressibility} [@SmithTamarizKirby2013a].
+Future work should assess the degree to which we can derive the same properties as previous studies using our current framework.
+In line with this idea, a primary focus of the current work is to provide a case-study in the productivity of assessing functionalist theories in repeated reference games between rational pragmatic agents.
+In theory, this framework can serve as a domain general tool to assess the range of functionalist theories examining efficient language-structure and use.\par
 

### 2019-01-31 paper/paper.tex
update objective function -- integrate comments from NDG about emphasis on novelty of objective and our framing.

--- 2019-01-30 paper/paper.tex
@@
 the relation between a word's frequency and rank in a corpus, the
-distribution of dependency lengths, or the presence of lexical
-properties such as ambiguity. We discuss a framework for studying how
-these properties emerge from in-the-moment interactions of rational,
-pragmatic speakers and listeners. Our work takes notions of
-lexicon-level efficiency as a starting point, connecting these ideas to
-notions of conversational-level efficiency. To do so, we derive an
-objective function for measuring the communicative efficiency of
-linguistic systems and then examining the behavior of this objective in
-a series of simulations focusing on the communicative function of
-ambiguity in language. These simulations suggest that rational pragmatic
-agents will produce communicatively efficient systems and that
-interactions between such agents provide a framework for examining
-efficient properties of language more broadly.
+distribution of syntactic dependency lengths across languages, or the
+presence of properties such as ambiguity. We discuss a framework for
+studying how these system-level properties emerge from local,
+in-the-moment interactions of rational, pragmatic speakers and
+listeners. To do so, we derive a novel objective function for measuring
+the communicative efficiency of linguistic systems in terms of the
+interactions of speakers and listeners. We examine the behavior of this
+objective in a series of simulations focusing on the communicative
+function of ambiguity in language. These simulations suggest that
+rational pragmatic agents will produce communicatively efficient systems
+and that interactions between such agents provide a framework for
+examining efficient properties of language structure and use more
+broadly.
 
@@
 efficiency. Such an approach might reframe our opening question as
-follows -- how does having \textbf{property x} make using language
-\(\ell\) more or less useful for communication? This reframing provides
-an opportunity to study both language \emph{structure} (as Zipf
-primarily did) or \emph{use} in terms of efficiency. For example, Regier
-et al. (2015) showed that languages appear to organize semantic domains
-(form-meaning mappings) in maximally efficient ways. Likewise,
-Piantadosi et al. (2011) argued that lexical ambiguity is an efficient
-property of any communication system when communication is
-contextualized.\par
+follows: how does having property \textit{x} make using language
+\(\ell\) more or less useful for communication? Among others, this
+efficiency-based framing has produced a rich set of theoretical and
+empirical targets exploring semantic typology (Regier, Kemp, \& Kay,
+2015), properties such as ambiguity (Piantadosi, Tily, \& Gibson, 2011)
+and compositionality (Kirby, Griffiths, \& Smith, 2014), and the
+efficient use of reduction and redundancy in production (Genzel \&
+Charniak, 2002; Levy \& Jaeger, 2007). These projects typically employ
+information-theoretic principles to characterize efficiency in natural
+or simulated language corpora. We adopt a different approach in this
+work, deriving a measure of efficiency from first principles of
+\textit{efficient language use} and then showing how this can lead to
+communicatively efficient systems.\par
 
-Zipf's original work and subsequent functionalist projects describe a
-varied set of linguistic properties and human behaviors. Common to many
-is a characterization of a fundamental effort-asymmetry underlying
-everyday communication. Simply put, what is ``hard'' for a speaker is
-likely different from what is ``hard'' for a listener. Zipf
-characterized this as follows -- purely from the standpoint of speaker
-effort, an optimal language \(\ell_{speaker}^*\) would tend toward a
-vocabulary of just a single, low-cost word. Given such a language, the
-full set of potential meanings would be conveyed using only that word,
-i.e. \(\ell_{speaker}^*\) would be fully ambiguous and all possible
-meanings would need to be disambiguated by a listener. From the
-standpoint of listener effort, an optimal language \(\ell_{listener}^*\)
-would map all possible meanings to distinct words, removing a listener's
-need to disambiguate. In this example, speaker effort is related to
-\emph{production cost} and listener effort to \emph{understanding or
-disambiguation cost}. Clearly natural languages fall between the two
-extremes of \(\ell_{speaker}^*\) and \(\ell_{listener}^*\). Zipf
-proposed that the particular lexicon-level properties he observed were a
-result of these competing forces -- the pressure to jointly minimize
-speaker and listener effort.\par
+Functionalist theories commonly frame language efficiency in terms of a
+fundamental effort-assymetry underlying everyday communication: what is
+``hard'' for a speaker is likely different than what is ``hard'' for a
+listener. Zipf gave the following intution for how this idea -- purely
+from the standpoint of speaker effort, an optimal language
+\(\ell_{speaker}^*\) would tend toward a vocabulary of a single,
+low-cost word. Given such a language, the full set of potential meanings
+would be conveyed using only that word, i.e. \(\ell_{speaker}^*\) would
+be fully ambiguous and all possible meanings would need to be
+disambiguated by a listener. From the standpoint of listener effort, an
+optimal language \(\ell_{listener}^*\) would map all possible meanings
+to distinct words, removing a listener's need to disambiguate. In this
+example, speaker effort is related to \emph{production cost} and
+listener effort to \emph{understanding or disambiguation cost}. Clearly,
+natural languages fall between the two extremes of \(\ell_{speaker}^*\)
+and \(\ell_{listener}^*\). Zipf proposed that the particular
+lexicon-level properties he observed were a result of these competing
+forces -- the pressure to jointly minimize speaker and listener
+effort.\par
 
-But how does such optimization take place? The example given by Zipf
+But how does this optimization take place? The example given by Zipf
 (1949), describes local, communicative interactions in terms of a
@@
 intended meaning given the transmitted utterance, \(m \rightarrow u\).
-Other functionalist projects have assumed this basic reference game
-setting (Piantadosi et al., 2011; Regier et al., 2015) and this
-simplification of the communicative act has proven productive in
-theoretical (Ferrer-i-Cancho, 2018), simulation-based (Kirby, Griffiths,
-\& Smith, 2014) and empirical explorations (Hawkins, Franke, Smith, \&
-Goodman, 2018) of efficient language structure and use.\par
+Other projects have assumed this basic reference game setting
+(Piantadosi et al., 2011; Regier et al., 2015) and this simplification
+of the communicative act has proven productive in theoretical
+(Ferrer-i-Cancho, 2018), simulation-based (Kirby et al., 2014) and
+empirical explorations (Hawkins, Franke, Smith, \& Goodman, 2018) of
+efficient language structure and use.\par
 
@@
 speakers' pragmatic reasoning in the moment. Under a Gricean treatment
-of pragmatics, speakers and listeners are assumed to follow a set of
-conversational maxims in which they cooperate to transfer information
-(Grice, 1975). These maxims appear to emerge from efficiency concerns,
-however (Horn, 1984). A primary target of the current work is to
-formalize this connection, by providing a framework for deriving
-system-level efficiencies from local, interaction behavior. Our claim is
-that to understand an ``efficient'' property of a system we need to
-consider \emph{how} that property is \emph{used} efficiently.
-Additionally, any quantitative description of ``efficiency'' should be a
-function of both the system and users.\par
+of pragmatics, speakers and listeners follow a set of conversational
+maxims in which they cooperate to transfer information (Grice, 1975).
+These maxims appear to emerge from efficiency concerns, however (Horn,
+1984). We formalize this connection -- showing how system-level
+efficiencies can emerge from local interaction behavior. Intuitively,
+our claim is that to understand an ``efficient'' property of a system it
+is essential that we consider how that property is \emph{used}
+efficiently. It follows that any quantitative measure of ``efficiency''
+must be a function of both the system and its users.\par
 
-In the current work we provide a case-study for the kind of framing
-presented above. We choose a property of languages that could, in
-principle, vary freely, but shows strong regularities across languages.
-The explanadum is why this regularity holds. We examine lexical
-ambiguity as our property, extending ideas by Piantadosi et al. (2011).
-We define a novel measure of efficiency that depends on the
-interactional behavior of interlocutors. For our model of speakers and
-listeners we adopt the Rational Speech Act (RSA) framework, which is
-nicely attested in experimental data on pragmatic interaction. With
-these basic ingredients we have a measure of efficiency given language
-and speaker-listener agents. Finally, we show that the property of
-interest is prevalent in languages that optimize this measure of
-efficiency and also how this property is used efficiently during local
-interaction. The primary contribution of this work is to show how the
-reference game frameowrk, in combination with formal models of
-communication, can be used to connect ideas about system-level
+In the current work we provide a case-study for the framing presented
+above. We choose a property of languages that could, in principle, vary
+freely, but shows strong regularities across languages. The explanadum
+is why this regularity holds. We examine ambiguity as our property,
+extending ideas by Piantadosi et al. (2011). We define a novel measure
+of efficiency that depends on the interactional behavior of speaker and
+listener agents. We adopt the reference game as our primary unit of
+interaction and model language users using the Rational Speech Act (RSA)
+framework -- a computional model of language use, which is nicely
+attested by experimental data on interaction. Using these basic
+ingredients, we show that the property of interest (ambiguity) is
+prevalent in languages that optimize our measure of efficiency. Further,
+we show how ambiguity is \textit{used} efficiently during local,
+in-the-moment interactions. The contributions of this work are twofold
+-- we derive a novel objective for measuring linguistic efficiency and
+also show how the reference game framework, in combination with formal
+models of communication, can be used to connect ideas about system-level
 efficiencies to in-the-moment language use.\par
@@
 of a reference game. We adopt that formulation here, assuming these
-communication games as our basic unit of analysis. In this setting,
-speakers and listeners are aware of a set of objects \(M\), which we
-will refer to as \emph{meanings} and are knowledgeable about the set of
-possible signals \(U\) (\emph{utterances}) that can be used to refer to
-a given meaning (see Figure 1). Utterances may have different relative
-costs, operationalized via a prior over utterances \(P(U)\). Similarly,
+communication games as our basic unit of analysis. In this framework,
+speakers and listeners are aware of a set of objects \(M\)
+(\emph{meanings}) and are knowledgeable about the set of possible
+signals \(U\) (\emph{utterances}) that can be used to refer to a given
+meaning (see Figure 1). Utterances may have different relative costs,
+operationalized via a prior over utterances \(P(U)\). Similarly,
 meanings differ in the relative degree to which they need to be talked
 about, operationalized as a prior over meanings
-\(P(M)\)\footnote{The prior over meanings are analogous to the \textit{need probabilities} assumed in previous work Regier, Kemp \& Kay (2015).}.
-We consider a set of contexts \(C\) with an associated prior \(P(C)\).
-Contexts encode different meaning priors, formalized as separate
-conditional distributions over meanings e.g.
-\(p(M|C=c_i) \neq p(M|C=c_j)\). Finally, we consider a set of
-communicative events \(e \in E\) where \(<u, m> = e\) is a tuple of
-utterance, meaning pairs. Note that in the case of single context
-\(|C|=1\) we simply have a prior over meanings \(p(M)\), a
-simplification we will use in introducing our objective function.\par
+\(P(M)\)\footnote{The prior over meanings are analogous to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}.
+We consider a set of contexts \(c\in C\) with an associated prior
+\(P(C)\). Each context \(c_i\) describes a different distribution over
+meanings e.g. \(p(M|C=c_i) \neq p(M|C=c_j)\). Finally, we consider a set
+of communicative events \(e \in E\) where \(<u, m, c> = e\) is an
+utterance, meaning and context triple.\par
 
@@
 \(U = \{\text{blue}, \text{green}, \text{square}, \text{circle}\} \text{ and three meanings }M = \{\text{green-square}, \text{blue-square}, \text{green-circle}\}\).
-The boolean matrix describes the literal semantics of the language
-\(\ell\). Note that will consider language to ``contain ambiguous
-items'' if there is at least one utterance \(u \in U\) which can apply
-to multiple meanings (i.e.
-\(|[[u_i]]| > 1\))\footnote{We use double brackets $[[\dots]]$ to represent denotation}.
-For example, in Figure 1 both the words ``blue'' and ``green'' are
-ambigous so we would say that \(\ell\) contains ambiguity.
+The boolean matrix describes the literal semantics of the language. We
+define a language as ``ambiguous'' if there is some utterance
+\(u \in U\) which can apply to multiple meanings (i.e.
+\(|[[u_i]]| > 1\))\footnote{We use double brackets $[[\dots]]$ to represent denotation.}.
+In Figure 1 both the words ``square'' and ``green'' are ambigous so we
+would say that \(\ell\) contains ambiguity.
 
@@
 
-The Rational Speech-act framework (RSA) is computational-level theory of
-pragmatic language use. RSA can be understood as a formalization of
+The Rational Speech Act framework (RSA) is computational-level theory of
+pragmatic language use, which has been well attested in experimental
+work examining both speaker and listener communication behavior (Frank
+\& Goodman, 2012; Goodman \& Frank, 2016). RSA is a formalization of
 essential Gricean pragmatic principles -- agents reason about one
-another and their shared context. For this reason, we adopt RSA as our
-representational framework to operationalize Gricean (rational and
-pragmatic) speaker-listeners (Grice, 1975). We refer the reader to
-Section 1 of our supplementary materials for a more detailed description
-of basic RSA speaker-listener definitions, and to Goodman \& Frank
-(2016) for an overview of its appliations.\par
+another and their shared context (Grice, 1975). We adopt RSA as our
+representational framework to model Gricean (rational and pragmatic)
+speaker-listeners in the reference game
+setting\footnote{See SI section 1 for a more detailed description of basic RSA speaker-listener defintions and to Goodman \& Frank [-@GoodmanFrank2016a] for an overview of its applications.}.\par
 
@@
 using \(\ell\) in a given context \(c\). That is, a speaker defines
-\(P_{speaker}(u|m; \ell)\). We will use \(S(u|m; \ell)\) as short-hand
-throughout. A \emph{listener agent} defines a conditional distribution
-over meanings, mapping from utterances \(U\) to meanings \(M\) using
-\(\ell\) in a given context \(c\). We will use \(L(m|u; \ell)\) as
-shorthand. Note that both speakers and listeners can induce joint
-distributions over the set of all events \(E\), although, importantly,
-these distributions may differ:
-\[P_{speaker}(u, m; \ell) = S(u|m; \ell)p(m)\]
-\[P_{listener}(u, m; \ell) = L(m|u; \ell)p(u)\]
+\(P_{speaker}(u|m, c; \ell)\). We will use \(S(u|m, c; \ell)\) as
+short-hand throughout. A \emph{listener agent} defines a conditional
+distribution over meanings, mapping from utterances \(U\) to meanings
+\(M\) using \(\ell\) in a given context \(c\) (i.e. \(L(m|u, c;\ell)\)).
+Note that both speakers and listeners can induce joint distributions
+over utterace-meaning pairs, although, importantly, these distributions
+may differ: \[P_{speaker}(u, m | c; \ell) = S(u|m, c; \ell)p(m|c)\]
+\[P_{listener}(u, m| c; \ell) = L(m|u, c; \ell)p(u|c)\]
 
@@
 listener effort to communicate over all possible communicative events
-\(E\). Note that to simplify the introduction of this objective we
-assume there is just a single context (\(|C|=1\)). In this case
-\(p(C=c)=1\) and there is only distribution over meanings
-(\(p(m|c) = p(m)\)). We provide a more detailed description of this
-derivation in the supplementary materials, both with and without context
-\(C\).\par
+\(e \in E\).\par
 
@@
 
-We assume that speaker effort is the negative log probability of an
-utterance (surprisal) -- intuitively, the number of bits needed to
-encode the utterance \(u\). This particular formalization of
-speaker-cost is general enough to accommodate a range of instantiations
-that might in theory be related to production difficult via articulation
-effort, cognitive effort related to lexical access, etc (Bennett \&
-Goodman, 2018).\par
+We assume that speaker effort is related to the suprisal of an utterance
+in a particular
+context\footnote{In the current set of simulations we consider utterances costs as independent from context (ie. $p(u|c)p(c)=p(u)p(c)$).}
+-- intuitively, the number of bits needed to encode the utterance \(u\).
+This particular formalization of speaker-cost is general enough to
+accommodate a range of cost instantiations, such as production difficult
+via articulation effort, cognitive effort related to lexical access, or
+others (Bennett \& Goodman, 2018).\par
 
-\[\text{speaker effort} = -log_2(p(u))\]
+\[\text{speaker effort} = -log_2(p(u|c))\]
 
-We assume listener effort is the negative log probability of the
-listener's conditional \(-log_2(L(m|u))\). This operationalization of
-listener effort is intuitively related to existing work in sentence
-processing in which word difficulty during comprehension is proportional
-to surprisal (Hale, 2001; R. Levy, 2008).
+We assume listener effort is the semantic suprisal of a meaning given an
+utterance. This operationalization of listener effort is intuitively
+related to existing work in sentence processing in which word
+comprehension difficulty is proportional to surprisal (Hale, 2001; Levy,
+2008).
 
-\[\text{listener effort} = -log_2(L(m|u; \ell))\]
+\[\text{listener effort} = -log_2(L(m|u, c; \ell))\]
 
-Importantly, we assume that events \(e = <u, m>\) are sampled according
-the to following generative model -- some object occurs in the world
-with probability \(p(M=m)\), the speaker attempts to refer to that
-object by sampling from her conditional distribution \(S(u|m)\) (i.e.
-\(e \sim p(m)S(u|m)\)). From the ingredients highlighted above, it is
-possible to derive the following objective between the speaker and
-listener joint distributions (see SI 2.2 for a complete derivation).
+Importantly, we assume that events \(e = <u, m, c>\) are sampled
+according the to following generative model -- some context occurs in
+the world with probability \(P(C=c)\). Within this context, an object
+\(m\) occurs with probability \(p(m|c)\). The speaker attempts to refer
+to that object by sampling from her conditional distribution
+\(S(u|m, c; \ell)\) (i.e. \(e \sim p(c)p(m|c)S(u|m, c; \ell)\)). From
+the ingredients described above, it is possible to derive the following
+objective between the speaker and listener joint distributions (see SI
+2.2 for a complete derivation).
 
@@
 \begin{split}
-  = -\mathbb{E}_{P_{speaker}}[log_2(P_{listener})] \\
-  = H_{cross}(P_{speaker}, P_{listener}; \ell)
+  = \mathbb{E}_{c \sim P(C)}[H_{cross}(P_{speaker}, P_{listener} | c; \ell)]\\
 \end{split}
@@
 From an information-theoretic perspective this objective is intuitive:
-cross-entropy gives us a measure of dissimilarity between two
+The Cross-Entropy (CE) gives us a measure of dissimilarity between two
 distributions -- the average number of bits required to communicate
 under one distribution, given that the ``true'' distribution differs. In
-our case, this is the difference between the joint distribution assumed
-by the speaker \(P_{speaker}\) and listener \(P_{listener}\). In other
-words, an ``efficient'' language \(\ell\) minimizes the distance between
-what speakers and listeners think.
+our case, we have an expectation over this term -- the expected
+difference between the distributions assumed by the speaker
+\(P_{speaker}\) and listener \(P_{listener}\) given a set of contexts
+\(C\). In other words, an ``efficient'' language \(\ell\) minimizes the
+distance between what speakers and listeners think.
 
@@
 The task of understanding language is marked by a frequent need to
-handle various forms of ambiguity; lexical, syntactic, among others
-(Wasow, Perfors, \& Beaver, 2005). The ubiquity of this property,
+handle various forms of ambiguity: lexical, syntactic, among
+others(Wasow, Perfors, \& Beaver, 2005). The ubiquity of this property,
 however, has been argued to provide evidence that langauges are not been
 optimized for
-communication\footnote{Chomsky (2002) famously claimed "If you want to make sure that we never misunderstand one another, for that purpose language is not well designed, because you have such properties as ambiguity." Note that this analysis is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$ we described earlier and define in terms of equation 2 in SI 2.2 .}
+communication\footnote{Chomsky's (2002) claim that "...
+language is not well designed, because you have such properties as ambiguity" is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$.}
 (Chomsky, 2002).
@@
 
-We provide a toy example of their argument here. Say we have two objects
-in the world \(m_1\) and \(m_2\), and two languages \(\ell_1\) and
-\(\ell_2\). In language \(\ell_1\), the low-cost utterance \(u_1\) can
-be used to refer to both \(m_1\) and \(m_2\) (i.e.
-\([[u_1]]_{\ell_1} = \{m_1, m_2 \}\) and
-\([[u_2]]_{\ell_1} = \emptyset\)). With language \(\ell_2\), however,
-the low cost utterance \(u_1\) can only refer to \(m_1\) and a
-higher-cost utterance \(u_2\) can refer to \(m_2\) (i.e.
-\([[u_1]]_{\ell_2} = \{m_1\}\) and \([[u_2]]_{\ell_2} = \{m_2 \}\)).
-While it is cheaper for a speaker to use \(\ell_1\) (because she can
-always use a low-cost utterance), it is more difficult for a listener
-(because \(u_1\) is ambiguous). If context isn't disambiguating then we
-might prefer \(\ell_2\) to \(\ell_1\). But, if context is disambiguating
-then the speaker can use \(u_1\) to refer to either \(m_1\) or \(m_2\)
-and \(\ell_1\) should be preferred.\par
+We re-iterate their argument here. Say we have two objects (\(m_1\) and
+\(m_2\)), two utterances (\(u_1\) and \(u_2\)), which differ in cost,
+and two languages (\(\ell_1\) and \(\ell_2\)) which describe different
+utterance-meaning mappings. In language \(\ell_1\), the low-cost \(u_1\)
+can be used to refer to both \(m_1\) and \(m_2\)
+(\([[u_1]]_{\ell_1} = \{m_1, m_2 \}\)), but the high-cost \(u_2\) cannot
+be used at all (\([[u_2]]_{\ell_1} = \emptyset\)). By contrast, in
+language \(\ell_2\), \(u_1\) can only refer to \(m_1\) and \(u_2\) can
+only refer to \(m_2\) (\([[u_1]]_{\ell_2} = \{m_1\}\) and
+\([[u_2]]_{\ell_2} = \{m_2 \}\)). While it is cheaper for a speaker to
+use \(\ell_1\) (because speaking is always low-cost), it is more
+difficult for a listener (because \(u_1\) is ambiguous). Crucially,
+however, if context is disambiguating then the speaker can use \(u_1\)
+to refer to either \(m_1\) or \(m_2\) and \(\ell_1\) should be preferred
+to \(\ell_2\).\par
 
@@
 \caption[Optimal languages are more likely to contain ambiguous items as the amount of contextual information increases]{Optimal languages are more likely to contain ambiguous items as the amount of contextual information increases.
-Vertical axis shows the proportion of optimal languages that contain ambiguity.
-Horizontal axis shows the number of contexts in each condition (1-4).
-Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only objectives.}\label{fig:plot-optimal-langs}
+Vertical axis shows the proportion of optimal languages containing ambiguity.
+Horizontal axis shows the number of context-sizes (1-4).
+Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only objectives.
+Error bars represent 95 percent confidence intervals.}\label{fig:plot-optimal-langs}
 \end{figure*}
@@
 
-We adopt the framework above, and (following Piantadosi et al. 2011),
-add \textit{context} (\(C\)). The authors argue that mapping two
-meanings \(m_1\) and \(m_2\) to a single utterance \(u_1\) is useful
-when they can be disambiguated \emph{in context}. In our case we
-consider a context \(c_i \in C\) to specify a permutation of the meaning
-probabilities. That is, \(p(m|c_1) \neq p(m|c_2)\), when there are two
-contexts \(|C| = \{c_1, c_2\}\).This in turn, leads to an update to our
-linguistic efficiency objective as we now would like to consider the
-average speaker-listener effort over all contextualized communicative
-events (see section 3 SI).\par
-
+We adopt the hypothesis presented by Piantadosi, formalizing their
+argument in our current framework -- showing that ambiguity is an
+efficient property under our CE objective in the reference game setting.
 We proceed by generating languages with different amounts of contextual
 support (varying the size of \(|C|\)). We search the space of languages,
-asessing if the ones which minimize our objective contain ambiguity. If
-context leads to more efficient commnication, then optimal languages
-should be more likley to be ambiguous as the amount of context
-increases.\par
+examining whether ones which minimize our objective contain ambiguity.
+If context leads to more efficient commnication, then optimal languages
+under our objective should be more likley to be ambiguous as the amount
+of context increases.\par
 
@@
 column of \(\ell\) has some non-zero assignment) and each form maps to
-at least one meaning (every column has some non-zero
-assignment)\footnote{For the current set of simulations we consider languages in which there are four or five true values -- in a "four" language ambiguity is not possible, but in a "five" language there are ambiguous mappings.}.
-For a given simulation the goal is to find the language \(\ell^*\) which
-minimizes our cross-entropy objective and then check to see if that
-language contains ambiguity.\par
+at least one meaning (every column has some non-zero assignment). For a
+given simulation the goal is to find the language \(\ell^*\) which
+minimizes our CE objective and then check to see if that language
+contains ambiguity.\par
 
-Recall that language efficiency is both a function of the particular
-semantic mappings induced by that language, the speaker and listener
-agents (\(S\) and \(L\)), as well as the utterance (\(P(U)\)), meaning
+We define language efficiency as a function of the particular semantic
+mappings induced by that language, the speaker and listener agents
+(\(S\) and \(L\)), as well as the utterance (\(P(U)\)), meaning
 (\(P(M)\)), and context priors (\(P(C)\)). Rather than assume particular
-structure for our utterance, meaning, and context prior distributions,
-for each simulation we generate \(P(U) \sim \text{Dir}(1, |U|)\),
-\(P(M|C=c) \sim \text{Dir}(1, |M|)\) (a separete conditional
-distribution over meanings for each context \(c\)), and
-\(P(C) \sim \text{Dir}(1, |C|)\) where \(\text{Dir}(1, k)\) specifies
-the uniform Dirichlet distribution over a \(k\)-dimensional probability
-vector.\par
+structure, for each simulation we generate
+\(P(U) \sim \text{Dir}(1, |U|)\), \(P(M|C=c) \sim \text{Dir}(1, |M|)\)
+(a separate conditional distribution over meanings for each context
+\(c\)), and \(P(C) \sim \text{Dir}(1, |C|)\), where \(\text{Dir}(1, k)\)
+specifies the uniform Dirichlet distribution over a \(k\)-dimensional
+probability vector.\par
 
@@
 Following the argument given by Piantadosi et al. (2011), we want to
-assess the impact of \emph{context} on our objective. To do so we
-consider four conditions with \(n=500\) simulations each (that is, 500
-unique sets of \(\{P(U), P(M|C), P(C)\}\). Our first is a
-\emph{single-context} condition (\(|C|=1\)) -- there is a only a single
-context describing \(P(M)\). Our second condition contains two-contexts
-(\(|C| = 2\)) -- we consider efficiency under both \(P(M|C=c_1)\) as
-well as \(P(M|C=c_2)\). The third and fourth condition correspond
-accordingly with \(|C|=3\) and \(|C| = 4\), respectively.\par
+assess the impact of \emph{context} on the presence of ambiguity in
+optimal languages. To do so we consider four conditions with \(n=500\)
+simulations each (that is, 500 unique sets of
+\(\{P(U), P(M|C), P(C)\}\). Our first is a \emph{one-context} condition
+(\(|C|=1\)) -- only a single distribution over meanings \(P(M)\). In our
+two-context condition (\(|C| = 2\)), we consider efficiency under both
+\(P(M|C=c_1)\) as well as \(P(M|C=c_2)\). Three- and four-context
+conditions corresponding accordingly.\par
 
@@
 For comparison, we also examine properties of optimal languages under
-two additional objectives. Zipf (1949) proposed that the optimal speaker
-language \(\ell_{speaker}^*\) would only optimize speaker effort and the
-optimal listener language \(\ell_{listener}^*\) would only optimize
-listener effort. We define these objectives using the first and second
-half of equation 1 (see SI section 2.2. for more detail).\par
+two additional objectives. Zipf (1949) proposed that the speaker-optimal
+language \(\ell_{speaker}^*\) would minimize speaker effort and the
+listener-optimal language \(\ell_{listener}^*\) would minimize listener
+effort. We define these objectives using the first and second half of
+equation 1 (see SI section 2.2. for more detail).\par
 
@@
 
-Figure 1. plots the proportion of optimal languages under each
-objective, as a function of context condition. The red line shows that
-as the number of contexts increases, so does the probability that the
-optimal language \(\ell^*_{cross}\) under our speaker-listener
-cross-entropy objective contains ambiguity. We also plot the proportion
-of optimal speaker-only \(\ell^*_{speaker}\) and listener-only
-\(\ell^*_{listener}\) languages that contain ambiguity. In line with
-Zipf's predictions, if languages are designed only to minimize speaker
-effort then optimal languages will assign lower-cost forms multiple
-meanings (all of the speaker-optimal languages contained ambiguity).
-Likewise, if languages are designed only to minimize listener effort
-then ambiguity should always be avoided (none of the listener-optimal
-languages contained ambiguity).\par
+Figure 1. plots the proportion of optimal languages under each objective
+as a function of context size. The red line shows that as the number of
+contexts increases, so does the probability that an optimal language
+\(\ell^*_{cross}\) contains ambiguity under our CE objective. For
+comparison we also plot the proportion of speaker-optimal
+\(\ell^*_{speaker}\) (blue line) and listener-optimal
+\(\ell^*_{listener}\) (green line) languages that contain ambiguity. In
+line with Zipf's predictions, if languages are designed only to minimize
+speaker effort then optimal languages always contain ambiguity. If
+languages are designed to minimize listener effort then ambiguity is
+always avoided.\par
 
-We explored the degree to which ambiguity improved efficiency during
-communication between rational, pragmatic agents. We showed that as
-contextual information increased, the optimal languages under our
-Cross-Entropy objective were more likely to have ambiguous lexical
-items. Importantly, we did not see this effect under alternative
-objectives which only emphasized speaker- or listener effort alone.\par
+In simulation 1 we explored the degree to which ambiguity was an
+efficient property when communication was contextualized. We showed that
+as contextual information increased, the optimal languages under our CE
+objective were more likely to have ambiguous lexical items. Importantly,
+we did not see this effect under alternative objectives which only
+emphasized speaker- or listener effort alone.\par
 
 While our results indicate that ambiguity is an efficient property of
-contextualized language these simulations assumed that speaker-listener
-agents could always disambiguate items in context. That is, in our four
-conditions both agents had perfect knowledge of the relevant conditional
+contextualized language, these simulations assumed that agents could
+always disambiguate items in context. That is, in our four conditions
+both agents had perfect knowledge of the relevant conditional
 distributions (\(P(M|C)\)). This assumption may be too strong for
-describing much of day-to-day communication -- we seldom begin a
-conversation with perfect knowledge of the current context of a
-conversation. We explore a case in which the listener has imperfect
-knowledge of context at the start of the conversation in Simulation
-2.\par
+describing much of day-to-day communication -- we seldom interact with
+others with perfect knowledge of the current context (or topic) of
+conversation. To explore how ambiguity may be \textit{used} efficiencly,
+we next explore a case in which the listener has imperfect knowledge of
+context at the start of the conversation, but may infer it from a
+discourse history in Simulation 2.\par
 
@@
 
-We shift time-scales in Simulation 2 -- instead of examining the
-objective landscape of a space of languages, we now consider a single
-language that contains ambiguity and examine how that property is used
-over a discourse. If context is not informative with respect to meaning
-early in a discourse (position \(i\)), but is at later positions at
-position (position \(i+k\)), then the speaker should avoid using
-ambigous material early \(i\), only using it \(k\) steps later in the
-conversation. In this toy example we can consider ``context'' as
-analogous to a ``topic'' of conversation. Intuitively, we can imagine a
-scenario in which a reader is beginning a newspaper article. While they
-may have some knowledge about the article's topic (perhaps from the
-title), they may not have complete knowledge of its contents, including
-the persons or events involved. In this setting, using an ambiguous
-pronoun early (position \(i\)) in the article can lead to
-misunderstanding without sufficient context to support disambiguation.
-But if by position \(i+k\) there is sufficient contextual information it
-may be efficient for the writer to use ambiguous material.\par
+We shift focus in Simulation 2 -- instead of examining the landscape of
+a space of languages under our CE objective, we consider a single
+language containing ambiguity, examining how that property is used over
+a discourse. If context is not informative with respect to meaning early
+in a discourse, but is at later positions, then an efficient speaking
+strategy would avoid ambiguous material early, saving it for when
+contextual information is disambiguating. In this toy example we can
+consider ``context'' as analogous to a ``topic'' of conversation. As an
+example, imagine a scenario in which a reader is beginning a magazine
+article. While they may have some knowledge about the article's topic
+(perhaps from the title), they may not have complete knowledge of its
+contents, including the persons or events involved. In this setting,
+using an potential low-cost, but ambiguous pronoun early may lead to
+misunderstanding if context is sufficiently informative. But, if by a
+later position enough contextual information has accumulated, it may be
+efficient for the writer to use the ambiguous material.\par
 
@@
 We consider consider a single language \(\ell\), which contains both
-ambiguous and unambiguous utterances. The ambiguous utterances are less
-costly than the unambiguous utterances. Crucially, we do not assume that
-the listener knows the particular topic (\(c_{current}\)) of the
-conversation \textit{a priori}. Rather, we assume that the listener has
-knowledge of the set of possible topics \(C = \{c_1, \dots, c_k\}\), but
+ambiguous and unambiguous utterances. We assume ambiguous utterances are
+lower cost. Crucially, we do not assume that the listener knows the
+particular topic (\(c_{current}\)) of the conversation
+\textit{a priori}. Rather, that the listener has knowledge of the set of
+possible topics \(C = \{c_1, \dots, c_k\}\), but
 \textit{does not know which one is currently being used by the speaker}.
@@
 conditional distribution over meanings \(P(M|C=c_{current})\) at the
-start of the discourse. In this setting, a speaker minimizing production
-effort would like to use the low-cost material as frequently as
-possible. However, this will incur a disambiguation cost to the listener
-if they are uncertain of the current context \(c_{current}\) (topic of
-conversation).\par
+start of the discourse.\par
 
-Over the course of a discourse, \(D\) the listener tries to infer both
+Over the course of a discourse \(D\), the listener tries to infer both
 the current topic, \(c_{current}\), as well as the particular meaning
 \(m\) of a given utterance \(u\). That is, we consider updated listener
-and speakers (\(L(m, c|u,D;\ell)\), \(S(u|m,c,D;\ell)\)). Note that both
-the speaker and listener can track the history of previous utterances
-\(D\). Importantly, an agent can attempt to infer the current topic of
-conversation \(c_{current}\) using the discourse history \(D\) (see SI
-section 3 for more detail).\par
+and speakers (\(L(m, c|u,D;\ell)\), \(S(u|m,c,D;\ell)\)) in which both
+agents can track the history of previous utterances \(D\). Importantly,
+an agent can attempt to infer the current topic of conversation
+\(c_{current}\) using the discourse history \(D\) (see SI section 3 for
+more detail).\par
 
@@
 (B) shows speaker effort across the three models.
-(C) shows the Cross-Entropy objective under our three speaker models.}\label{fig:plot-optimal-use}
+(C) shows the Cross-Entropy objective under our three speaker models.
+Error bars represent 95 percent confidence intervals.}\label{fig:plot-optimal-use}
 \end{figure*}
@@
 We conduct \(N=600\) simulations, generating discourses of length
-\(|D|=30\) utterances with three different speaker models (\(n=200\)
-each). We consider a single language
+\(|D|=30\) utterances, comparing three speaker models (\(n=200\) each).
+We consider a single language
 \(\ell\)\footnote{See SI for the matrix notation of this langauge.} with
-\(|U|=6\) and \(|M|=4\) in which two of the utterances \(u_5\) and
-\(u_6\) are ambiguous and that
-\(p(u_5) = p(u_6) > p(u_{1})=\dots=p(u_{4})\). That is, the two
-ambiguous utterances (\(u_5\) and \(u_6\)) are less costly than the
-non-ambiguous utterances. (Note that use of this particular language is
-not essential -- the results are broadly generalizable languages that
-contain ambiguity.)\par
+\(|U|=6\) and \(|M|=4\) in which two of the utterances are ambiguous and
+lower cost than the unambiguous utterances. (Note that use of this
+particular language is not essential -- the results are broadly
+generalizable to languages that contain ambiguity.)\par
 
@@
 
-We consider three types of speaker models. In our
-\textit{Full pragmatics} model, we consider a speaker agent who reasons
-about her listener and also has complete recall of the set of utterances
-in the discourse \(D\). This speaker believes that the listener may not
-know the current topic \(c_{current}\) at the start of the discrouse,
-but can infer it with enough data. In this way, contextual information
-shared by the speaker and listener may increase over the discourse (as
-the listeners estimate of the current topic (\(P(C=c|D)\)) comes in line
-with the actual topic known to the speaker). We compare two baselines
-models. The first, a \textit{Partial pragmatics} baseline describes a
-speaker who reasons about a listener, but assumes they have no access to
-discourse history. The second is a \textit{No pragmatics} speaker who
-does not consider a listener at all, but produces utterances according
-to the underlying langauge semantics (\(\ell\)) and topic probabilities
+We consider three types of speaker models. Our \textit{Full pragmatics}
+agent, models a speaker who reasons about her listener and also has
+complete recall of the set of utterances in the discourse \(D\). This
+speaker believes that the listener may not know the current topic
+\(c_{current}\) at the start of the discrouse, but should infer it over
+the discourse. We compare two baselines models. The first, a
+\textit{Partial pragmatics} baseline describes a speaker who reasons
+about a listener, but assumes they have no access to discourse history.
+The second, a \textit{No pragmatics} speaker who does not consider a
+listener at all, but produces utterances according to the underlying
+langauge semantics (\(\ell\)) and topic probabilities
 (\(p(M|C=c_{current}\)) (see SI section 3 for details).\par
@@
 
-We assess whether it is efficient to use low-cost, ambiguous material
-when context is disambiguating. In the current setting, this corresponds
-to speaker behavior in which ambiguous forms are avoided early in the
-discourse, but prefered later in the discourse once the listener is
-confident of the topic. We should expect this to be reflected in our
-\textit{Full pragmatics} model. By contrast, our
-\textit{Partial pragmatics} speaker does not believe that there is
-sufficient contextual information for the listener to disambiguate her
-utterances. This should lead to total avoidance of ambiguous material
-over the discourse. Finally, our \textit{No pragmatics} speaker should
-use low-cost material greedily since she does not consider a listener's
-need to disambiguate whatsoever.\par
+We are interested in \textit{when} it is efficient to use low-cost,
+ambiguous material in this discourse setting. A speaker strategy that is
+mutually efficient for both agents should avoid ambiguity until
+sufficient contextual information has accumulated. We should expect this
+to be reflected in our \textit{Full pragmatics} model who reasons about
+the listener and discourse history. By contrast, a speaker-optimal model
+who does not consider the listener should greedily use ambiguous
+utterances (\textit{No Pragmatics} model), while a listener-optimal
+model should avoid ambiguity entirely (\textit{Partial pragmatics}
+model).\par
 
@@
 ambiguous utterance as a function of discourse position. The
-\text{No pragmatic} baseline uses ambiguous utterances frequently and at
-a constant rate over the discourse and the \text{Partial pragmatic}
-baseline avoids ambiguous utterances entirely (reasoning that the
-utterances cannot be disambiguated). But, the \text{Full pragmatic}
-model avoids ambiguous material at first, but employing them gradually
-increasingly as the discourse proceeds. (B) shows the speaker-listener
-Cross-Entropy objective. Note that the objective decreases for all three
-models. This decrease in all three models is primarily driven by the
-listener updating his belief about the actual topic
-(\(P(C=c_{current}|D)\)). Of particular importance is the value of the
-objective at the end of the discourse (position \(i=30\)). The
-difference in the Cross-Entropy between the \text{Full Pragmatic} and
-\text{Partial Pragmatic} models at the end of the discourse is driven by
-the reduction in speaker costs described in (C). That is, while speaker
-effort remains constant in both \textit{No pragmatic} and
-\textit{Partial pragmatic} baselines, effort declines for the
-\text{Full Pragmatic} model as they are increasingly able to use
-ambiguous material later in the discourse.\par
+\textit{No pragmatic} baseline uses ambiguous utterances frequently and
+at a constant rate over the discourse and the \textit{Partial pragmatic}
+baseline avoids ambiguous utterances entirely. But, the
+\textit{Full pragmatic} model avoids ambiguous material only at the
+start of the discourse, employing it increasingly as the discourse
+proceeds. (C) tracks our CE objective for each model over the discourse.
+Note that the objective decreases for all three models, primarily driven
+by the listener updating his belief about the actual topic
+(\(P(C=c_{current}|D)\)). However, the objective declines more quickly
+under the \textit{Full} and \textit{Partial} pragmatic speakers as
+listener agents are better able to infer the correct context.
+Additionally, the difference in CE between the \textit{Full}- and
+\textit{Partial pragmatic} models at the end of the discourse is driven
+by the reduction in speaker costs described in (B). While speaker effort
+remains constant in both \textit{No pragmatic} and
+\textit{Partial pragmatic} baselines, effort declines in the
+\text{Full Pragmatic} model as she increasingly relies on ambiguous
+material later in the discourse.\par
 
 In Simulation 2 we explored how ambiguity could be employed efficiently
-during discourse. We assumed that the listener did not know the current
-context (topic) \textit{a priori}, but could infer it from the discourse
-history. We examined speaker production behavior, tracking use of
-low-cost, but ambiguous material over the discourse. Early in the
+during language use. We assumed that the listener did not know the
+current context (topic) \textit{a priori}, but could infer it from the
+discourse history. We examined speaker production behavior, tracking use
+of low-cost, but ambiguous material over the discourse. Early in the
 discourse the \textit{Full pragmatic} speaker avoided using ambiguous
-material, only using the material later when the context was known to
-the listener -- it is useful to use low-cost material, but only when
-context is disambiguating.
+material, only using the material later when the current topic was known
+to the listener.
 
@@
 argument on the communicative function of ambiguity (Piantadosi et al.,
-2011). We derive a novel speaker-listener cross-entropy objective for
-measuring the efficiency of a language in a reference game setting
-showing that optimal languages are more likely to contain ambiguous
-material when context is informative. Further we show that rational
-pragmatic agents will \emph{use} ambiguous material efficiently, only
-when such use is supported by context.\par
+2011), deriving a novel speaker-listener Cross-Entropy objective for
+measuring the efficiency linguistic systems from first principles of
+efficient language use. In a series of simulations we showed that
+optimal languages are more likely to contain ambiguous material when
+context is informative and that rational pragmatic agents will
+\emph{use} ambiguous material efficiently.\par
 
-The framework employed here -- the analysis of the dynamics of repeated
-reference games between rational pragmatic agents -- provides a tool for
-the analysis of functionalist theories, allowing linkage across
-timescales of analysis from conversation (Genzel \& Charniak, 2002; \&.
-J. Levy R., 2007) to typology (Regier et al., 2015). In theory, this
-framework provides extensive opportunity to vary properties of the
-languages and agents, however we have focused on only a small subset
-here. For example, examining how increasing or decreasing
-\textit{rationality} (instantiated through the RSA \(\alpha\) parameter)
-or \textit{pragmatics} (instantiated through the RSA recursion depth)
-impacts these results is a clear avenue for future work. An important
-limitation of the current work is the broad assumption that language can
-be described through such reference games. The degree to which
-properties of non-referential language can be captured in such a
-framework remains an important target for future work.
-
-\vspace{1em}
-\fbox{\parbox[b][][c]{7.3cm}{\centering All simulation and analysis code are available in the public repository for the project: (link will be available upon acceptance)}}
-\vspace{1em} \noindent
+A limitation of the current work is an analysis of exactly how this
+objective compares to existing measures. For example, previous work has
+described competing speaker-listener pressures in terms of a trade-off
+of \textit{simplicity} and \textit{informativeness} (Kemp \& Regier,
+2012) or \textit{expressivity} and \textit{compressibility} (Smith,
+Tamariz, \& Kirby, 2013). Future work should assess the degree to which
+we can derive the same properties as previous studies using our current
+framework. In line with this idea, a primary focus of the current work
+is to provide a case-study in the productivity of assessing
+functionalist theories in repeated reference games between rational
+pragmatic agents. In theory, this framework can serve as a domain
+general tool to assess the range of functionalist theories examining
+efficient language-structure and use.\par
 
@@
 (3)}, 207--237.
+
+\hypertarget{ref-FrankGoodman2012a}{}
+Frank, M., \& Goodman, N. (2012). Predicting pragmatic reasoning in
+language games. \emph{Science}, \emph{336}, 998.
 
@@
 
+\hypertarget{ref-KempRegier2012a}{}
+Kemp, C., \& Regier, T. (2012). Kinship categories across languages
+reflect general communicative principles. \emph{Science}, \emph{336},
+1049--1054.
+
 \hypertarget{ref-KirbyGriffithsSmith2014a}{}
@@
 
-\hypertarget{ref-LevyJaeger2007a}{}
-Levy, \&. J., R. (2007). Speakers optimize information density through
-syntactic reduction. In \emph{Proceedings of the twentieth annual
-conference on neural information processing systems}.
-
 \hypertarget{ref-Levy2008a}{}
@@
 \emph{Cognition}, \emph{106(3)}, 1126--1177.
+
+\hypertarget{ref-LevyJaeger2007a}{}
+Levy, R., \& Jaeger, T. (2007). Speakers optimize information density
+through syntactic reduction. In \emph{Proceedings of the twentieth
+annual conference on neural information processing systems}.
 
@@
 Wiley-Blackwell.
+
+\hypertarget{ref-SmithTamarizKirby2013a}{}
+Smith, K., Tamariz, M., \& Kirby, S. (2013). Linguistic structure is an
+evolutionary trade-off between simplicity and expressivity. In
+\emph{Proceedings of the 36th annual conference of the cognitive science
+society}.
 

### 2019-01-31 paper/paper.Rmd
updates Thurs morning.

@@
 These projects typically employ information-theoretic principles to characterize efficiency in natural or simulated language corpora.
-We adopt a different approach in this work, deriving a measure of efficiency from first principles of \textit{efficient language use} and then showing how this can lead to communicatively efficient systems.\par
+We adopt a different approach in this work, deriving a measure of efficiency from first principles of \textit{efficient language use} and show how this can lead to communicatively efficient systems.\par
 
 Functionalist theories commonly frame language efficiency in terms of a fundamental effort-assymetry underlying everyday communication: what is “hard” for a speaker is likely different than what is “hard” for a listener.
-Zipf gave the following intution for how this idea -- purely from the standpoint of speaker effort, an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of a single, low-cost word.
+Zipf described this as follows -- purely from the standpoint of speaker effort, an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of a single, low-cost word.
 Given such a language, the full set of potential meanings would be conveyed using only that word, i.e.
@@
 We define a novel measure of efficiency that depends on the interactional behavior of speaker and listener agents.
-We adopt the reference game as our primary unit of interaction and model language users using the Rational Speech Act (RSA) framework -- a computional model of language use, which is nicely attested by experimental data on interaction.
+We adopt the reference game as our primary unit of interaction and model language users with the Rational Speech Act (RSA) framework -- a computional model of language use, which is nicely attested by experimental data on interaction.
 Using these basic ingredients, we show that the property of interest (ambiguity) is prevalent in languages that optimize our measure of efficiency.
@@
 Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$\footnote{The prior over meanings are analogous to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}.
-We consider a set of contexts $c\in C$ with an associated prior $P(C)$.
-Each context $c_i$ describes a different distribution over meanings e.g.
+We consider a set of contexts $C$ with an associated prior $P(C)$.
+Each context $c\in C$ describes a different distribution over meanings e.g.
 $p(M|C=c_i) \neq p(M|C=c_j)$.
-Finally, we consider a set of communicative events $e \in E$ where $<u, m, c> = e$ is an utterance, meaning and context triple.\par
+Finally, we consider a set of communicative events $e \in E$ where $<u, m, c> = e$ is an utterance-meaning-context triple.\par
 
@@
 ### Speakers and listeners
-The Rational Speech Act framework (RSA) is computational-level theory of pragmatic language use, which has been well attested in experimental work examining both speaker and listener communication behavior [@FrankGoodman2012a; @GoodmanFrank2016a].
+The Rational Speech Act framework (RSA) is a computational-level theory of pragmatic language use, which has been well attested in experimental work examining both speaker and listener communication behavior [@FrankGoodman2012a; @GoodmanFrank2016a].
 RSA is a formalization of essential Gricean pragmatic principles -- agents reason about one another and their shared context [@Grice1975a].
-We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speaker-listeners in the reference game setting\footnote{See SI section 1 for a more detailed description of basic RSA speaker-listener defintions and to Goodman \& Frank [-@GoodmanFrank2016a] for an overview of its applications.}.\par
+We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speaker-listeners in the reference game setting\footnote{See SI section 1 for a more detailed description of basic RSA speaker-listener defintions and to Goodman \& Frank (2016) for an overview of its applications.}.\par
 
@@
 \end{equation}
+
 From an information-theoretic perspective this objective is intuitive: The Cross-Entropy (CE) gives us a measure of dissimilarity between two distributions -- the average number of bits required to communicate under one distribution, given that the “true” distribution differs.
-In our case, we have an expectation over this term -- the expected difference between the distributions assumed by the speaker $P_{speaker}$ and listener $P_{listener}$ given a set of contexts $C$.
+In our case, we have an expectation over this term -- the expected difference between the distributions assumed by the speaker $P_{speaker}$ and listener $P_{listener}$ given a set of contexts $C$\footnote{Note that in the single context case $|C|=1$ this objective is simply the speaker-listener Cross-Entropy over utterance-meaning pairs $H_{cross}[P_{speaker}(u, m; \ell), P_{listener}(u, m; \ell)]$}.
 In other words, an "efficient" language $\ell$ minimizes the distance between what speakers and listeners think.
@@
 
-The task of understanding language is marked by a frequent need to handle various forms of ambiguity: lexical, syntactic, among others[@WasowPerforsBeaver2005a].
-The ubiquity of this property, however, has been argued to provide evidence that langauges are not been optimized for communication\footnote{Chomsky's (2002) claim that "...
+The task of understanding language is marked by a frequent need to handle various forms of ambiguity: lexical, syntactic, among others [@WasowPerforsBeaver2005a].
+The ubiquity of this property, however, has been argued to provide evidence that languages have not been optimized for communication\footnote{Chomsky's (2002) claim that "...
 language is not well designed, because you have such properties as ambiguity" is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$.} [@Chomsky2002a].
@@
 
-We re-iterate their argument here.
-Say we have two objects ($m_1$ and $m_2$), two utterances ($u_1$ and $u_2$), which differ in cost, and two languages ($\ell_1$ and $\ell_2$) which describe different utterance-meaning mappings.
+As an example, say we have two objects ($m_1$ and $m_2$), two utterances ($u_1$ and $u_2$), which differ in cost, and two languages ($\ell_1$ and $\ell_2$), which describe different utterance-meaning mappings.
 In language $\ell_1$, the low-cost $u_1$ can be used to refer to both $m_1$ and $m_2$ ($[[u_1]]_{\ell_1} = \{m_1, m_2 \}$), but the high-cost $u_2$ cannot be used at all ($[[u_2]]_{\ell_1} = \emptyset$).
@@
 In the following experiments we explore two aspects of Piantadosi’s claim.
-First, we examine the efficient language *structure* aspect of their claim, exploring when the optimal linguistic system $\ell^*$ is most likely to contain ambiguous lexical items.
-In the second experiment, we explore an efficient language *use* aspect of the claim -- at what point in a conversation is it useful for a speaker to use ambiguous lexical material?\par
+In Simulation 1, we examine the efficient language *structure* aspect of their claim, exploring when the optimal linguistic system $\ell^*$ is most likely to contain ambiguous lexical items.
+In Simulation 2, we explore an efficient language *use* aspect of the claim -- at what point in a conversation is it useful for a speaker to use ambiguous lexical material?\par
 
@@
 We search the space of languages, examining whether ones which minimize our objective contain ambiguity.
-If context leads to more efficient commnication, then optimal languages under our objective should be more likley to be ambiguous as the amount of context increases.\par
+If context leads to more efficient communication, then optimal languages under our objective should be more likley to be ambiguous as the amount of context increases.\par
 
@@
 Note that a language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment) and each form maps to at least one meaning (every column has some non-zero assignment).
-For a given simulation the goal is to find the language $\ell^*$ which minimizes our CE objective and then check to see if that language contains ambiguity.\par
+For a given simulation, the goal is to find the language $\ell^*$ which minimizes our objective and then check to see if that language contains ambiguity.\par
 
@@
 To do so we consider four conditions with $n=500$ simulations each (that is, 500 unique sets of $\{P(U), P(M|C), P(C)\}$.
-Our first is a *one-context* condition ($|C|=1$) -- only a single distribution over meanings $P(M)$.
-In our two-context condition ($|C| = 2$), we consider efficiency under both $P(M|C=c_1)$ as well as $P(M|C=c_2)$.
-Three- and four-context conditions corresponding accordingly.\par
+Our first is a \textit{one-context} condition ($|C|=1$) -- only a single distribution over meanings $P(M)$.
+In our \textit{two-context} condition ($|C| = 2$), we consider efficiency under both $P(M|C=c_1)$ as well as $P(M|C=c_2)$.
+\textit{Three-} and \textit{four-context} conditions corresponding accordingly.\par
 
@@
 
+In Simulation 1 we explored the degree to which ambiguity is an efficient property of languages when communication is contextualized.
 Figure 1.
@@
 
-In simulation 1 we explored the degree to which ambiguity was an efficient property when communication was contextualized.
-We showed that as contextual information increased, the optimal languages under our CE objective were more likely to have ambiguous lexical items.
-Importantly, we did not see this effect under alternative objectives which only emphasized speaker- or listener effort alone.\par
-
 While our results indicate that ambiguity is an efficient property of contextualized language, these simulations assumed that agents could always disambiguate items in context.
 That is, in our four conditions both agents had perfect knowledge of the relevant conditional distributions ($P(M|C)$).
-This assumption may be too strong for describing much of day-to-day communication -- we seldom interact with others with perfect knowledge of the current context (or topic) of conversation.
-To explore how ambiguity may be \textit{used} efficiencly, we next explore a case in which the listener has imperfect knowledge of context at the start of the conversation, but may infer it from a discourse history in Simulation 2.\par
+This assumption may be too strong for describing much of day-to-day communication -- we seldom interact with others with perfect knowledge of the current context (or topic) at the start of a conversation.
+To explore how ambiguity may be \textit{used} efficiently in our framework, we next examine a case in which the listener has imperfect knowledge of context at the start of the conversation, but may infer it from the contents of discourse.\par
 
@@
 If context is not informative with respect to meaning early in a discourse, but is at later positions, then an efficient speaking strategy would avoid ambiguous material early, saving it for when contextual information is disambiguating.
-In this toy example we can consider “context” as analogous to a “topic” of conversation.
+We can consider “context” as analogous to a “topic” of conversation.
 As an example, imagine a scenario in which a reader is beginning a magazine article.
 While they may have some knowledge about the article’s topic (perhaps from the title), they may not have complete knowledge of its contents, including the persons or events involved.
-In this setting, using an potential low-cost, but ambiguous pronoun early may lead to misunderstanding if context is sufficiently informative.
-But, if by a later position enough contextual information has accumulated, it may be efficient for the writer to use the ambiguous material.\par
+In this setting, using a low-cost, but ambiguous pronoun early may lead to misunderstanding if context is not informative.
+But, if by a later position enough contextual information has accumulated, it may be efficient to use the ambiguous material.\par
 
@@
 That is, we consider updated listener and speakers ($L(m, c|u,D;\ell)$, $S(u|m,c,D;\ell)$) in which both agents can track the history of previous utterances $D$.
-Importantly, an agent can attempt to infer the current topic of conversation $c_{current}$ using the discourse history $D$ (see SI section 3 for more detail).\par
+Importantly, an agent can attempt to infer the current topic of conversation $c_{current}$ using the discourse history $D$.\par
 
@@
 Our \textit{Full pragmatics} agent, models a speaker who reasons about her listener and also has complete recall of the set of utterances in the discourse $D$.
-This speaker believes that the listener may not know the current topic $c_{current}$ at the start of the discrouse, but should infer it over the discourse.
-We compare two baselines models.
+This speaker believes that the listener may not know the current topic $c_{current}$ at the start of the discrouse, but can infer it over the discourse.
+We compare two baseline models.
 The first, a \textit{Partial pragmatics} baseline describes a speaker who reasons about a listener, but assumes they have no access to discourse history.
@@
 Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatic} models at the end of the discourse is driven by the reduction in speaker costs described in (B).
-While speaker effort remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines, effort declines in the \text{Full Pragmatic} model as she increasingly relies on ambiguous material later in the discourse.\par
+While speaker effort remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines, effort declines in the \textit{Full pragmatic} model as she increasingly relies on ambiguous material later in the discourse.\par
 

### 2019-01-31 paper/paper.tex
updates Thurs morning.

@@
 work, deriving a measure of efficiency from first principles of
-\textit{efficient language use} and then showing how this can lead to
+\textit{efficient language use} and show how this can lead to
 communicatively efficient systems.\par
@@
 ``hard'' for a speaker is likely different than what is ``hard'' for a
-listener. Zipf gave the following intution for how this idea -- purely
-from the standpoint of speaker effort, an optimal language
-\(\ell_{speaker}^*\) would tend toward a vocabulary of a single,
-low-cost word. Given such a language, the full set of potential meanings
-would be conveyed using only that word, i.e. \(\ell_{speaker}^*\) would
-be fully ambiguous and all possible meanings would need to be
-disambiguated by a listener. From the standpoint of listener effort, an
-optimal language \(\ell_{listener}^*\) would map all possible meanings
-to distinct words, removing a listener's need to disambiguate. In this
-example, speaker effort is related to \emph{production cost} and
-listener effort to \emph{understanding or disambiguation cost}. Clearly,
-natural languages fall between the two extremes of \(\ell_{speaker}^*\)
-and \(\ell_{listener}^*\). Zipf proposed that the particular
-lexicon-level properties he observed were a result of these competing
-forces -- the pressure to jointly minimize speaker and listener
-effort.\par
+listener. Zipf described this as follows -- purely from the standpoint
+of speaker effort, an optimal language \(\ell_{speaker}^*\) would tend
+toward a vocabulary of a single, low-cost word. Given such a language,
+the full set of potential meanings would be conveyed using only that
+word, i.e. \(\ell_{speaker}^*\) would be fully ambiguous and all
+possible meanings would need to be disambiguated by a listener. From the
+standpoint of listener effort, an optimal language \(\ell_{listener}^*\)
+would map all possible meanings to distinct words, removing a listener's
+need to disambiguate. In this example, speaker effort is related to
+\emph{production cost} and listener effort to \emph{understanding or
+disambiguation cost}. Clearly, natural languages fall between the two
+extremes of \(\ell_{speaker}^*\) and \(\ell_{listener}^*\). Zipf
+proposed that the particular lexicon-level properties he observed were a
+result of these competing forces -- the pressure to jointly minimize
+speaker and listener effort.\par
 
@@
 listener agents. We adopt the reference game as our primary unit of
-interaction and model language users using the Rational Speech Act (RSA)
+interaction and model language users with the Rational Speech Act (RSA)
 framework -- a computional model of language use, which is nicely
@@
 \(P(M)\)\footnote{The prior over meanings are analogous to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}.
-We consider a set of contexts \(c\in C\) with an associated prior
-\(P(C)\). Each context \(c_i\) describes a different distribution over
-meanings e.g. \(p(M|C=c_i) \neq p(M|C=c_j)\). Finally, we consider a set
-of communicative events \(e \in E\) where \(<u, m, c> = e\) is an
-utterance, meaning and context triple.\par
+We consider a set of contexts \(C\) with an associated prior \(P(C)\).
+Each context \(c\in C\) describes a different distribution over meanings
+e.g. \(p(M|C=c_i) \neq p(M|C=c_j)\). Finally, we consider a set of
+communicative events \(e \in E\) where \(<u, m, c> = e\) is an
+utterance-meaning-context triple.\par
 
@@
 
-The Rational Speech Act framework (RSA) is computational-level theory of
-pragmatic language use, which has been well attested in experimental
+The Rational Speech Act framework (RSA) is a computational-level theory
+of pragmatic language use, which has been well attested in experimental
 work examining both speaker and listener communication behavior (Frank
@@
 speaker-listeners in the reference game
-setting\footnote{See SI section 1 for a more detailed description of basic RSA speaker-listener defintions and to Goodman \& Frank [-@GoodmanFrank2016a] for an overview of its applications.}.\par
+setting\footnote{See SI section 1 for a more detailed description of basic RSA speaker-listener defintions and to Goodman \& Frank (2016) for an overview of its applications.}.\par
 
@@
 \(P_{speaker}\) and listener \(P_{listener}\) given a set of contexts
-\(C\). In other words, an ``efficient'' language \(\ell\) minimizes the
+\(C\)\footnote{Note that in the single context case $|C|=1$ this objective is simply the speaker-listener Cross-Entropy over utterance-meaning pairs $H_{cross}[P_{speaker}(u, m; \ell), P_{listener}(u, m; \ell)]$}.
+In other words, an ``efficient'' language \(\ell\) minimizes the
 distance between what speakers and listeners think.
@@
 The task of understanding language is marked by a frequent need to
-handle various forms of ambiguity: lexical, syntactic, among
-others(Wasow, Perfors, \& Beaver, 2005). The ubiquity of this property,
-however, has been argued to provide evidence that langauges are not been
-optimized for
+handle various forms of ambiguity: lexical, syntactic, among others
+(Wasow, Perfors, \& Beaver, 2005). The ubiquity of this property,
+however, has been argued to provide evidence that languages have not
+been optimized for
 communication\footnote{Chomsky's (2002) claim that "...
@@
 
-We re-iterate their argument here. Say we have two objects (\(m_1\) and
-\(m_2\)), two utterances (\(u_1\) and \(u_2\)), which differ in cost,
-and two languages (\(\ell_1\) and \(\ell_2\)) which describe different
+As an example, say we have two objects (\(m_1\) and \(m_2\)), two
+utterances (\(u_1\) and \(u_2\)), which differ in cost, and two
+languages (\(\ell_1\) and \(\ell_2\)), which describe different
 utterance-meaning mappings. In language \(\ell_1\), the low-cost \(u_1\)
@@
 In the following experiments we explore two aspects of Piantadosi's
-claim. First, we examine the efficient language \emph{structure} aspect
-of their claim, exploring when the optimal linguistic system \(\ell^*\)
-is most likely to contain ambiguous lexical items. In the second
-experiment, we explore an efficient language \emph{use} aspect of the
-claim -- at what point in a conversation is it useful for a speaker to
-use ambiguous lexical material?\par
+claim. In Simulation 1, we examine the efficient language
+\emph{structure} aspect of their claim, exploring when the optimal
+linguistic system \(\ell^*\) is most likely to contain ambiguous lexical
+items. In Simulation 2, we explore an efficient language \emph{use}
+aspect of the claim -- at what point in a conversation is it useful for
+a speaker to use ambiguous lexical material?\par
 
@@
 examining whether ones which minimize our objective contain ambiguity.
-If context leads to more efficient commnication, then optimal languages
+If context leads to more efficient communication, then optimal languages
 under our objective should be more likley to be ambiguous as the amount
@@
 at least one meaning (every column has some non-zero assignment). For a
-given simulation the goal is to find the language \(\ell^*\) which
-minimizes our CE objective and then check to see if that language
-contains ambiguity.\par
+given simulation, the goal is to find the language \(\ell^*\) which
+minimizes our objective and then check to see if that language contains
+ambiguity.\par
 
@@
 simulations each (that is, 500 unique sets of
-\(\{P(U), P(M|C), P(C)\}\). Our first is a \emph{one-context} condition
-(\(|C|=1\)) -- only a single distribution over meanings \(P(M)\). In our
-two-context condition (\(|C| = 2\)), we consider efficiency under both
-\(P(M|C=c_1)\) as well as \(P(M|C=c_2)\). Three- and four-context
-conditions corresponding accordingly.\par
+\(\{P(U), P(M|C), P(C)\}\). Our first is a \textit{one-context}
+condition (\(|C|=1\)) -- only a single distribution over meanings
+\(P(M)\). In our \textit{two-context} condition (\(|C| = 2\)), we
+consider efficiency under both \(P(M|C=c_1)\) as well as \(P(M|C=c_2)\).
+\textit{Three-} and \textit{four-context} conditions corresponding
+accordingly.\par
 
@@
 
+In Simulation 1 we explored the degree to which ambiguity is an
+efficient property of languages when communication is contextualized.
 Figure 1. plots the proportion of optimal languages under each objective
@@
 
-In simulation 1 we explored the degree to which ambiguity was an
-efficient property when communication was contextualized. We showed that
-as contextual information increased, the optimal languages under our CE
-objective were more likely to have ambiguous lexical items. Importantly,
-we did not see this effect under alternative objectives which only
-emphasized speaker- or listener effort alone.\par
-
 While our results indicate that ambiguity is an efficient property of
@@
 describing much of day-to-day communication -- we seldom interact with
-others with perfect knowledge of the current context (or topic) of
-conversation. To explore how ambiguity may be \textit{used} efficiencly,
-we next explore a case in which the listener has imperfect knowledge of
-context at the start of the conversation, but may infer it from a
-discourse history in Simulation 2.\par
+others with perfect knowledge of the current context (or topic) at the
+start of a conversation. To explore how ambiguity may be \textit{used}
+efficiently in our framework, we next examine a case in which the
+listener has imperfect knowledge of context at the start of the
+conversation, but may infer it from the contents of discourse.\par
 
@@
 strategy would avoid ambiguous material early, saving it for when
-contextual information is disambiguating. In this toy example we can
-consider ``context'' as analogous to a ``topic'' of conversation. As an
-example, imagine a scenario in which a reader is beginning a magazine
-article. While they may have some knowledge about the article's topic
-(perhaps from the title), they may not have complete knowledge of its
-contents, including the persons or events involved. In this setting,
-using an potential low-cost, but ambiguous pronoun early may lead to
-misunderstanding if context is sufficiently informative. But, if by a
-later position enough contextual information has accumulated, it may be
-efficient for the writer to use the ambiguous material.\par
+contextual information is disambiguating. We can consider ``context'' as
+analogous to a ``topic'' of conversation. As an example, imagine a
+scenario in which a reader is beginning a magazine article. While they
+may have some knowledge about the article's topic (perhaps from the
+title), they may not have complete knowledge of its contents, including
+the persons or events involved. In this setting, using a low-cost, but
+ambiguous pronoun early may lead to misunderstanding if context is not
+informative. But, if by a later position enough contextual information
+has accumulated, it may be efficient to use the ambiguous material.\par
 
@@
 an agent can attempt to infer the current topic of conversation
-\(c_{current}\) using the discourse history \(D\) (see SI section 3 for
-more detail).\par
+\(c_{current}\) using the discourse history \(D\).\par
 
@@
 speaker believes that the listener may not know the current topic
-\(c_{current}\) at the start of the discrouse, but should infer it over
-the discourse. We compare two baselines models. The first, a
+\(c_{current}\) at the start of the discrouse, but can infer it over the
+discourse. We compare two baseline models. The first, a
 \textit{Partial pragmatics} baseline describes a speaker who reasons
@@
 \textit{Partial pragmatic} baselines, effort declines in the
-\text{Full Pragmatic} model as she increasingly relies on ambiguous
+\textit{Full pragmatic} model as she increasingly relies on ambiguous
 material later in the discourse.\par

### 2019-01-31 paper/paper.Rmd
fix typos

@@
 
-We assume that speaker effort is related to the suprisal of an utterance in a particular context\footnote{In the current set of simulations we consider utterances costs as independent from context (ie.
+We assume that speaker effort is related to the surprisal of an utterance in a particular context\footnote{In the current set of simulations we consider utterances costs as independent from context (ie.
 $p(u|c)p(c)=p(u)p(c)$).} -- intuitively, the number of bits needed to encode the utterance $u$.
@@
 
-We assume listener effort is the semantic suprisal of a meaning given an utterance.
+We assume listener effort is the semantic surprisal of a meaning given an utterance.
 This operationalization of listener effort is intuitively related to existing work in sentence processing in which word comprehension difficulty is proportional to surprisal [@Hale2001a; @Levy2008a].

### 2019-01-31 paper/paper.tex
fix typos

@@
 
-We assume that speaker effort is related to the suprisal of an utterance
-in a particular
+We assume that speaker effort is related to the surprisal of an
+utterance in a particular
 context\footnote{In the current set of simulations we consider utterances costs as independent from context (ie. $p(u|c)p(c)=p(u)p(c)$).}
@@
 
-We assume listener effort is the semantic suprisal of a meaning given an
-utterance. This operationalization of listener effort is intuitively
+We assume listener effort is the semantic surprisal of a meaning given
+an utterance. This operationalization of listener effort is intuitively
 related to existing work in sentence processing in which word

### 2019-01-31 paper/paper.Rmd
updates from MF / BP edits

@@
 abstract: 
-    Languages display a diverse set of distributional regularities such as the relation between a word's frequency and rank in a corpus, the distribution of syntactic dependency lengths across languages, or the presence of properties such as ambiguity.
+    Despite their diversity, languages around the world share a consistent set of properties and distributional regularities.
+For example, the distribution of word frequencies, the distribution of syntactic dependency lengths, and the presence of ambiguity are all remarkably consistent across languages.
 We discuss a framework for studying how these system-level properties emerge from local, in-the-moment interactions of rational, pragmatic speakers and listeners.
@@
 But how does this optimization take place?
-The example given by Zipf [-@Zipf1949a], describes local, communicative interactions in terms of a \textit{reference game}.
+The example given by Zipf [-@Zipf1949a] describes local, communicative interactions in terms of a \textit{reference game}.
 Speakers intend to refer to some object in the world $m$.
@@
 
-Zipf's conception of speaker and listener effort may be connected to speakers' pragmatic reasoning in the moment.
+Adopting reference games as a basic unit of analysis suggests that optimization may take place at the level of conversation.
+Importantly, Zipf's conception of speaker and listener effort may be connected to how language is used; in particular, whether interlocutors engage in pragmatic reasoning during conversation.
 Under a Gricean treatment of pragmatics, speakers and listeners follow a set of conversational maxims in which they cooperate to transfer information [@Grice1975a].
 These maxims appear to emerge from efficiency concerns, however (Horn, 1984).
-We formalize this connection -- showing how system-level efficiencies can emerge from local interaction behavior.
-Intuitively, our claim is that to understand an "efficient" property of a system it is essential that we consider how that property is *used* efficiently.
-It follows that any quantitative measure of "efficiency" must be a function of both the system and its users.\par
+We formalize this connection -- showing how system-level efficiencies can emerge from local interaction behavior of pragmatic agents.
+Intuitively, our claim is that to understand an "efficient" property of a system it is essential that we consider how that property is *used* efficiently.\par
 
-In the current work we provide a case-study for the framing presented above.
+We provide a case study for this approach, in which functionalist regularities emerge from the dynamics of pragmatic communication.
 We choose a property of languages that could, in principle, vary freely, but shows strong regularities across languages.
@@
 ### Speakers and listeners
-The Rational Speech Act framework (RSA) is a computational-level theory of pragmatic language use, which has been well attested in experimental work examining both speaker and listener communication behavior [@FrankGoodman2012a; @GoodmanFrank2016a].
+The Rational Speech Act framework (RSA) is a computational-level theory of pragmatic language use, which has produced good fit to human communication behavior across a range of language phenomena [@FrankGoodman2012a; @GoodmanFrank2016a].
 RSA is a formalization of essential Gricean pragmatic principles -- agents reason about one another and their shared context [@Grice1975a].
-We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speaker-listeners in the reference game setting\footnote{See SI section 1 for a more detailed description of basic RSA speaker-listener defintions and to Goodman \& Frank (2016) for an overview of its applications.}.\par
+We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speaker-listeners in the reference game setting (see SI and @GoodmanFrank2016a for review).\par
 
@@
 $p(u|c)p(c)=p(u)p(c)$).} -- intuitively, the number of bits needed to encode the utterance $u$.
-This particular formalization of speaker-cost is general enough to accommodate a range of cost instantiations, such as production difficult via articulation effort, cognitive effort related to lexical access, or others [@BennettGoodman2015a].\par
+This particular formalization of speaker-cost is general enough to accommodate a range of cost instantiations, such as production difficulty via articulation effort, cognitive effort related to lexical access, or others [@BennettGoodman2015a].\par
 
@@
 
-From an information-theoretic perspective this objective is intuitive: The Cross-Entropy (CE) gives us a measure of dissimilarity between two distributions -- the average number of bits required to communicate under one distribution, given that the “true” distribution differs.
+From an information-theoretic perspective this objective is intuitive: $H_{cross}$ denotes the Cross-Entropy (CE), a measure of dissimilarity between two distributions -- the average number of bits required to communicate under one distribution, given that the “true” distribution differs.
 In our case, we have an expectation over this term -- the expected difference between the distributions assumed by the speaker $P_{speaker}$ and listener $P_{listener}$ given a set of contexts $C$\footnote{Note that in the single context case $|C|=1$ this objective is simply the speaker-listener Cross-Entropy over utterance-meaning pairs $H_{cross}[P_{speaker}(u, m; \ell), P_{listener}(u, m; \ell)]$}.
@@
 
-In the following experiments we explore two aspects of Piantadosi’s claim.
+In the following experiments we explore two aspects of Piantadosi’s et al.'s claim.
 In Simulation 1, we examine the efficient language *structure* aspect of their claim, exploring when the optimal linguistic system $\ell^*$ is most likely to contain ambiguous lexical items.
@@
 
-We adopt the hypothesis presented by Piantadosi, formalizing their argument in our current framework -- showing that ambiguity is an efficient property under our CE objective in the reference game setting.
+We show that ambiguity is an efficient property under our CE objective in the reference game setting.
 We proceed by generating languages with different amounts of contextual support (varying the size of $|C|$).
 We search the space of languages, examining whether ones which minimize our objective contain ambiguity.
-If context leads to more efficient communication, then optimal languages under our objective should be more likley to be ambiguous as the amount of context increases.\par
+If context leads to more efficient communication, then optimal languages under our objective should be more likely to be ambiguous as the amount of context increases.\par
 
@@
 
+```{r plot-optimal-use, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "(A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position.
+(B) shows speaker effort across the three models.
+(C) shows the Cross-Entropy objective under our three speaker models.
+Error bars represent 95 percent confidence intervals."}
+
+img <- png::readPNG("figs/discourse_grid_plot.png")
+grid::grid.raster(img)
+```
+
 We define language efficiency as a function of the particular semantic mappings induced by that language, the speaker and listener agents ($S$ and $L$), as well as the utterance ($P(U)$), meaning ($P(M)$), and context priors ($P(C)$).
@@
 
-Following the argument given by Piantadosi et al.
-[-@Piantadosi2011a], we want to assess the impact of *context* on the presence of ambiguity in optimal languages.
+We want to assess the impact of *context* on the presence of ambiguity in optimal languages.
 To do so we consider four conditions with $n=500$ simulations each (that is, 500 unique sets of $\{P(U), P(M|C), P(C)\}$.
@@
 In Simulation 1 we explored the degree to which ambiguity is an efficient property of languages when communication is contextualized.
-Figure 1.
-plots the proportion of optimal languages under each objective as a function of context size.
+Figure 1 plots the proportion of optimal languages under each objective as a function of context size.
 The red line shows that as the number of contexts increases, so does the probability that an optimal language $\ell^*_{cross}$ contains ambiguity under our CE objective.
@@
 This assumption may be too strong for describing much of day-to-day communication -- we seldom interact with others with perfect knowledge of the current context (or topic) at the start of a conversation.
-To explore how ambiguity may be \textit{used} efficiently in our framework, we next examine a case in which the listener has imperfect knowledge of context at the start of the conversation, but may infer it from the contents of discourse.\par
+To explore how ambiguity may be \textit{used} efficiently in our framework, we next examine a case in which the listener has imperfect knowledge of context at the start of the conversation, but may infer it from the discourse history.\par
 
@@
 While they may have some knowledge about the article’s topic (perhaps from the title), they may not have complete knowledge of its contents, including the persons or events involved.
-In this setting, using a low-cost, but ambiguous pronoun early may lead to misunderstanding if context is not informative.
-But, if by a later position enough contextual information has accumulated, it may be efficient to use the ambiguous material.\par
+In this setting, using a low-cost, but ambiguous referring expression (say a pronoun like "he") early may lead to misunderstanding if context is not informative.
+But, if by a later position enough contextual information has accumulated, it may be efficient to use the ambiguous expression.\par
 
@@
 Crucially, we do not assume that the listener knows the particular topic ($c_{current}$) of the conversation \textit{a priori}.
-Rather, that the listener has knowledge of the set of possible topics $C = \{c_1, \dots, c_k\}$, but \textit{does not know which one is currently being used by the speaker}.
+Rather, that the listener has knowledge of the set of possible topics $C = \{c_1, \dots, c_k\}$, but does not know which one is currently being used by the speaker.
 Formally, this means the listener does not have access to the correct conditional distribution over meanings $P(M|C=c_{current})$ at the start of the discourse.\par
@@
 
-```{r plot-optimal-use, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "(A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position.
-(B) shows speaker effort across the three models.
-(C) shows the Cross-Entropy objective under our three speaker models.
-Error bars represent 95 percent confidence intervals."}
-
-img <- png::readPNG("figs/discourse_grid_plot.png")
-grid::grid.raster(img)
-```
-
 We conduct $N=600$ simulations, generating discourses of length $|D|=30$ utterances, comparing three speaker models ($n=200$ each).
 We consider a single language $\ell$\footnote{See SI for the matrix notation of this langauge.} with $|U|=6$ and $|M|=4$ in which two of the utterances are ambiguous and lower cost than the unambiguous utterances.
-(Note that use of this particular language is not essential -- the results are broadly generalizable to languages that contain ambiguity.)\par
+(Note that use of this particular language is not essential -- the results are broadly generalizable to languages that contain ambiguity, but exploring this space is computationally expensive.)\par
 
@@
 The first, a \textit{Partial pragmatics} baseline describes a speaker who reasons about a listener, but assumes they have no access to discourse history.
-The second, a \textit{No pragmatics} speaker who does not consider a listener at all, but produces utterances according to the underlying langauge semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$) (see SI section 3 for details).\par
+The second, a \textit{No pragmatics} speaker who does not consider a listener at all, but produces utterances according to the underlying langauge semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$) (see SI section 3).\par
 
@@
 
-We are interested in \textit{when} it is efficient to use low-cost, ambiguous material in this discourse setting.
+We are interested in when it is efficient to use low-cost, ambiguous material in this discourse setting.
 A speaker strategy that is mutually efficient for both agents should avoid ambiguity until sufficient contextual information has accumulated.
@@
 
-Figure 2, (A) shows the empirical probability that a speaker uses an ambiguous utterance as a function of discourse position.
+Figure 2 (A) shows the empirical probability that a speaker uses an ambiguous utterance as a function of discourse position.
 The \textit{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse and the \textit{Partial pragmatic} baseline avoids ambiguous utterances entirely.
 But, the \textit{Full pragmatic} model avoids ambiguous material only at the start of the discourse, employing it increasingly as the discourse proceeds.
-(C) tracks our CE objective for each model over the discourse.
+Panel (C) tracks our CE objective for each model over the discourse.
 Note that the objective decreases for all three models, primarily driven by the listener updating his belief about the actual topic ($P(C=c_{current}|D)$).
 However, the objective declines more quickly under the \textit{Full} and \textit{Partial} pragmatic speakers as listener agents are better able to infer the correct context.
-Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatic} models at the end of the discourse is driven by the reduction in speaker costs described in (B).
+Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatic} models at the end of the discourse is driven by the reduction in speaker costs (see panel B).
 While speaker effort remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines, effort declines in the \textit{Full pragmatic} model as she increasingly relies on ambiguous material later in the discourse.\par
@@
 
-How do the competing pressures of speakers and listeners give rise to the distributional forms found in natural language?
+How do the competing pressures of speakers and listeners give rise to the distributional regularities found in natural language?
 Zipf [-@Zipf1949a] proposed that the asymmetry between speaker and listener costs gives rise to a range of properties at the level of the lexicon.
@@
 We focused on an argument on the communicative function of ambiguity [@Piantadosi2011a], deriving a novel speaker-listener Cross-Entropy objective for measuring the efficiency linguistic systems from first principles of efficient language use.
-In a series of simulations we showed that optimal languages are more likely to contain ambiguous material when context is informative and that rational pragmatic agents will *use* ambiguous material efficiently.\par
+In a series of simulations we showed that optimal languages are more likely to contain ambiguous material when context is informative and that rational pragmatic agents will use ambiguous material efficiently.\par
 
 A limitation of the current work is an analysis of exactly how this objective compares to existing measures.
-For example, previous work has described competing speaker-listener pressures in terms of a trade-off of \textit{simplicity} and \textit{informativeness} [@KempRegier2012a] or \textit{expressivity} and \textit{compressibility} [@SmithTamarizKirby2013a].
+For example, previous work has described competing speaker-listener pressures in terms of a trade-off of simplicity and informativeness [@KempRegier2012a] or expressivity and compressibility [@SmithTamarizKirby2013a].
 Future work should assess the degree to which we can derive the same properties as previous studies using our current framework.
-In line with this idea, a primary focus of the current work is to provide a case-study in the productivity of assessing functionalist theories in repeated reference games between rational pragmatic agents.
-In theory, this framework can serve as a domain general tool to assess the range of functionalist theories examining efficient language-structure and use.\par
+In line with this idea, a primary focus of the current work is to provide a case-study on the productivity of assessing functionalist theories in repeated reference games between rational pragmatic agents.
+We hope this framework can serve as a domain general tool to assess the range of functionalist theories examining efficient language-structure and use.\par
 

### 2019-01-31 paper/paper.tex
updates from MF / BP edits

@@
 \begin{abstract}
-Languages display a diverse set of distributional regularities such as
-the relation between a word's frequency and rank in a corpus, the
-distribution of syntactic dependency lengths across languages, or the
-presence of properties such as ambiguity. We discuss a framework for
-studying how these system-level properties emerge from local,
-in-the-moment interactions of rational, pragmatic speakers and
-listeners. To do so, we derive a novel objective function for measuring
-the communicative efficiency of linguistic systems in terms of the
-interactions of speakers and listeners. We examine the behavior of this
-objective in a series of simulations focusing on the communicative
-function of ambiguity in language. These simulations suggest that
-rational pragmatic agents will produce communicatively efficient systems
-and that interactions between such agents provide a framework for
-examining efficient properties of language structure and use more
-broadly.
+Despite their diversity, languages around the world share a consistent
+set of properties and distributional regularities. For example, the
+distribution of word frequencies, the distribution of syntactic
+dependency lengths, and the presence of ambiguity are all remarkably
+consistent across languages. We discuss a framework for studying how
+these system-level properties emerge from local, in-the-moment
+interactions of rational, pragmatic speakers and listeners. To do so, we
+derive a novel objective function for measuring the communicative
+efficiency of linguistic systems in terms of the interactions of
+speakers and listeners. We examine the behavior of this objective in a
+series of simulations focusing on the communicative function of
+ambiguity in language. These simulations suggest that rational pragmatic
+agents will produce communicatively efficient systems and that
+interactions between such agents provide a framework for examining
+efficient properties of language structure and use more broadly.
 
@@
 But how does this optimization take place? The example given by Zipf
-(1949), describes local, communicative interactions in terms of a
+(1949) describes local, communicative interactions in terms of a
 \textit{reference game}. Speakers intend to refer to some object in the
@@
 
-Zipf's conception of speaker and listener effort may be connected to
-speakers' pragmatic reasoning in the moment. Under a Gricean treatment
-of pragmatics, speakers and listeners follow a set of conversational
-maxims in which they cooperate to transfer information (Grice, 1975).
-These maxims appear to emerge from efficiency concerns, however (Horn,
-1984). We formalize this connection -- showing how system-level
-efficiencies can emerge from local interaction behavior. Intuitively,
-our claim is that to understand an ``efficient'' property of a system it
-is essential that we consider how that property is \emph{used}
-efficiently. It follows that any quantitative measure of ``efficiency''
-must be a function of both the system and its users.\par
+Adopting reference games as a basic unit of analysis suggests that
+optimization may take place at the level of conversation. Importantly,
+Zipf's conception of speaker and listener effort may be connected to how
+language is used; in particular, whether interlocutors engage in
+pragmatic reasoning during conversation. Under a Gricean treatment of
+pragmatics, speakers and listeners follow a set of conversational maxims
+in which they cooperate to transfer information (Grice, 1975). These
+maxims appear to emerge from efficiency concerns, however (Horn, 1984).
+We formalize this connection -- showing how system-level efficiencies
+can emerge from local interaction behavior of pragmatic agents.
+Intuitively, our claim is that to understand an ``efficient'' property
+of a system it is essential that we consider how that property is
+\emph{used} efficiently.\par
 
-In the current work we provide a case-study for the framing presented
-above. We choose a property of languages that could, in principle, vary
-freely, but shows strong regularities across languages. The explanadum
-is why this regularity holds. We examine ambiguity as our property,
-extending ideas by Piantadosi et al. (2011). We define a novel measure
-of efficiency that depends on the interactional behavior of speaker and
+We provide a case study for this approach, in which functionalist
+regularities emerge from the dynamics of pragmatic communication. We
+choose a property of languages that could, in principle, vary freely,
+but shows strong regularities across languages. The explanadum is why
+this regularity holds. We examine ambiguity as our property, extending
+ideas by Piantadosi et al. (2011). We define a novel measure of
+efficiency that depends on the interactional behavior of speaker and
 listener agents. We adopt the reference game as our primary unit of
@@
 The Rational Speech Act framework (RSA) is a computational-level theory
-of pragmatic language use, which has been well attested in experimental
-work examining both speaker and listener communication behavior (Frank
-\& Goodman, 2012; Goodman \& Frank, 2016). RSA is a formalization of
+of pragmatic language use, which has produced good fit to human
+communication behavior across a range of language phenomena (Frank \&
+Goodman, 2012; Goodman \& Frank, 2016). RSA is a formalization of
 essential Gricean pragmatic principles -- agents reason about one
@@
 representational framework to model Gricean (rational and pragmatic)
-speaker-listeners in the reference game
-setting\footnote{See SI section 1 for a more detailed description of basic RSA speaker-listener defintions and to Goodman \& Frank (2016) for an overview of its applications.}.\par
+speaker-listeners in the reference game setting (see SI and Goodman \&
+Frank (2016) for review).\par
 
@@
 From an information-theoretic perspective this objective is intuitive:
-The Cross-Entropy (CE) gives us a measure of dissimilarity between two
-distributions -- the average number of bits required to communicate
-under one distribution, given that the ``true'' distribution differs. In
-our case, we have an expectation over this term -- the expected
-difference between the distributions assumed by the speaker
+\(H_{cross}\) denotes the Cross-Entropy (CE), a measure of dissimilarity
+between two distributions -- the average number of bits required to
+communicate under one distribution, given that the ``true'' distribution
+differs. In our case, we have an expectation over this term -- the
+expected difference between the distributions assumed by the speaker
 \(P_{speaker}\) and listener \(P_{listener}\) given a set of contexts
@@
 
-In the following experiments we explore two aspects of Piantadosi's
-claim. In Simulation 1, we examine the efficient language
+In the following experiments we explore two aspects of Piantadosi's et
+al.'s claim. In Simulation 1, we examine the efficient language
 \emph{structure} aspect of their claim, exploring when the optimal
@@
 
-We adopt the hypothesis presented by Piantadosi, formalizing their
-argument in our current framework -- showing that ambiguity is an
-efficient property under our CE objective in the reference game setting.
-We proceed by generating languages with different amounts of contextual
-support (varying the size of \(|C|\)). We search the space of languages,
-examining whether ones which minimize our objective contain ambiguity.
-If context leads to more efficient communication, then optimal languages
-under our objective should be more likley to be ambiguous as the amount
-of context increases.\par
+We show that ambiguity is an efficient property under our CE objective
+in the reference game setting. We proceed by generating languages with
+different amounts of contextual support (varying the size of \(|C|\)).
+We search the space of languages, examining whether ones which minimize
+our objective contain ambiguity. If context leads to more efficient
+communication, then optimal languages under our objective should be more
+likely to be ambiguous as the amount of context increases.\par
 
@@
 
+\begin{CodeChunk}
+\begin{figure*}[h]
+
+{\centering \includegraphics{figs/plot-optimal-use-1}
+
+}
+
+\caption[(A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position]{(A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position.
+(B) shows speaker effort across the three models.
+(C) shows the Cross-Entropy objective under our three speaker models.
+Error bars represent 95 percent confidence intervals.}\label{fig:plot-optimal-use}
+\end{figure*}
+\end{CodeChunk}
+
 We define language efficiency as a function of the particular semantic
@@
 
-Following the argument given by Piantadosi et al. (2011), we want to
-assess the impact of \emph{context} on the presence of ambiguity in
-optimal languages. To do so we consider four conditions with \(n=500\)
-simulations each (that is, 500 unique sets of
+We want to assess the impact of \emph{context} on the presence of
+ambiguity in optimal languages. To do so we consider four conditions
+with \(n=500\) simulations each (that is, 500 unique sets of
 \(\{P(U), P(M|C), P(C)\}\). Our first is a \textit{one-context}
@@
 efficient property of languages when communication is contextualized.
-Figure 1. plots the proportion of optimal languages under each objective
+Figure 1 plots the proportion of optimal languages under each objective
 as a function of context size. The red line shows that as the number of
@@
 listener has imperfect knowledge of context at the start of the
-conversation, but may infer it from the contents of discourse.\par
+conversation, but may infer it from the discourse history.\par
 
@@
 the persons or events involved. In this setting, using a low-cost, but
-ambiguous pronoun early may lead to misunderstanding if context is not
-informative. But, if by a later position enough contextual information
-has accumulated, it may be efficient to use the ambiguous material.\par
+ambiguous referring expression (say a pronoun like ``he'') early may
+lead to misunderstanding if context is not informative. But, if by a
+later position enough contextual information has accumulated, it may be
+efficient to use the ambiguous expression.\par
 
@@
 \textit{a priori}. Rather, that the listener has knowledge of the set of
-possible topics \(C = \{c_1, \dots, c_k\}\), but
-\textit{does not know which one is currently being used by the speaker}.
-Formally, this means the listener does not have access to the correct
-conditional distribution over meanings \(P(M|C=c_{current})\) at the
-start of the discourse.\par
+possible topics \(C = \{c_1, \dots, c_k\}\), but does not know which one
+is currently being used by the speaker. Formally, this means the
+listener does not have access to the correct conditional distribution
+over meanings \(P(M|C=c_{current})\) at the start of the discourse.\par
 
@@
 
-\begin{CodeChunk}
-\begin{figure*}[h]
-
-{\centering \includegraphics{figs/plot-optimal-use-1}
-
-}
-
-\caption[(A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position]{(A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position.
-(B) shows speaker effort across the three models.
-(C) shows the Cross-Entropy objective under our three speaker models.
-Error bars represent 95 percent confidence intervals.}\label{fig:plot-optimal-use}
-\end{figure*}
-\end{CodeChunk}
-
 We conduct \(N=600\) simulations, generating discourses of length
@@
 particular language is not essential -- the results are broadly
-generalizable to languages that contain ambiguity.)\par
+generalizable to languages that contain ambiguity, but exploring this
+space is computationally expensive.)\par
 
@@
 langauge semantics (\(\ell\)) and topic probabilities
-(\(p(M|C=c_{current}\)) (see SI section 3 for details).\par
+(\(p(M|C=c_{current}\)) (see SI section 3).\par
 
@@
 
-We are interested in \textit{when} it is efficient to use low-cost,
-ambiguous material in this discourse setting. A speaker strategy that is
-mutually efficient for both agents should avoid ambiguity until
-sufficient contextual information has accumulated. We should expect this
-to be reflected in our \textit{Full pragmatics} model who reasons about
-the listener and discourse history. By contrast, a speaker-optimal model
-who does not consider the listener should greedily use ambiguous
-utterances (\textit{No Pragmatics} model), while a listener-optimal
-model should avoid ambiguity entirely (\textit{Partial pragmatics}
-model).\par
+We are interested in when it is efficient to use low-cost, ambiguous
+material in this discourse setting. A speaker strategy that is mutually
+efficient for both agents should avoid ambiguity until sufficient
+contextual information has accumulated. We should expect this to be
+reflected in our \textit{Full pragmatics} model who reasons about the
+listener and discourse history. By contrast, a speaker-optimal model who
+does not consider the listener should greedily use ambiguous utterances
+(\textit{No Pragmatics} model), while a listener-optimal model should
+avoid ambiguity entirely (\textit{Partial pragmatics} model).\par
 
@@
 
-Figure 2, (A) shows the empirical probability that a speaker uses an
+Figure 2 (A) shows the empirical probability that a speaker uses an
 ambiguous utterance as a function of discourse position. The
@@
 start of the discourse, employing it increasingly as the discourse
-proceeds. (C) tracks our CE objective for each model over the discourse.
-Note that the objective decreases for all three models, primarily driven
-by the listener updating his belief about the actual topic
-(\(P(C=c_{current}|D)\)). However, the objective declines more quickly
-under the \textit{Full} and \textit{Partial} pragmatic speakers as
-listener agents are better able to infer the correct context.
+proceeds. Panel (C) tracks our CE objective for each model over the
+discourse. Note that the objective decreases for all three models,
+primarily driven by the listener updating his belief about the actual
+topic (\(P(C=c_{current}|D)\)). However, the objective declines more
+quickly under the \textit{Full} and \textit{Partial} pragmatic speakers
+as listener agents are better able to infer the correct context.
 Additionally, the difference in CE between the \textit{Full}- and
 \textit{Partial pragmatic} models at the end of the discourse is driven
-by the reduction in speaker costs described in (B). While speaker effort
+by the reduction in speaker costs (see panel B). While speaker effort
 remains constant in both \textit{No pragmatic} and
@@
 How do the competing pressures of speakers and listeners give rise to
-the distributional forms found in natural language? Zipf (1949) proposed
-that the asymmetry between speaker and listener costs gives rise to a
-range of properties at the level of the lexicon. We explored the
-interactions of rational pragmatic agents as a framework for
+the distributional regularities found in natural language? Zipf (1949)
+proposed that the asymmetry between speaker and listener costs gives
+rise to a range of properties at the level of the lexicon. We explored
+the interactions of rational pragmatic agents as a framework for
 understanding efficient language structure and use. We focused on an
@@
 optimal languages are more likely to contain ambiguous material when
-context is informative and that rational pragmatic agents will
-\emph{use} ambiguous material efficiently.\par
+context is informative and that rational pragmatic agents will use
+ambiguous material efficiently.\par
 
@@
 described competing speaker-listener pressures in terms of a trade-off
-of \textit{simplicity} and \textit{informativeness} (Kemp \& Regier,
-2012) or \textit{expressivity} and \textit{compressibility} (Smith,
-Tamariz, \& Kirby, 2013). Future work should assess the degree to which
-we can derive the same properties as previous studies using our current
-framework. In line with this idea, a primary focus of the current work
-is to provide a case-study in the productivity of assessing
-functionalist theories in repeated reference games between rational
-pragmatic agents. In theory, this framework can serve as a domain
+of simplicity and informativeness (Kemp \& Regier, 2012) or expressivity
+and compressibility (Smith, Tamariz, \& Kirby, 2013). Future work should
+assess the degree to which we can derive the same properties as previous
+studies using our current framework. In line with this idea, a primary
+focus of the current work is to provide a case-study on the productivity
+of assessing functionalist theories in repeated reference games between
+rational pragmatic agents. We hope this framework can serve as a domain
 general tool to assess the range of functionalist theories examining

### 2019-02-01 ambiguity/simulation-runners/context-ambiguity-plots.Rmd
initial ndg changes to simulation 2 intro and matrix added to plot 1. just pushing in case i mess all the formatting up...

@@
     xlab("Number of contexts (|C|)") +
-    theme_few()
+    theme_few() +
     # In case we want to embed legend in the plot, but not really necessary...
-    # theme(legend.position = c(0.8, 0.4),
-    #       legend.background = element_rect(colour = 'grey', fill = 'white', linetype='solid'))
+    theme(legend.position = c(0.7, 0.3),
+          legend.background = element_rect(colour = 'grey', fill = 'white', linetype='solid'),
+          plot.margin = unit(c(0, 0, 0, 0), "cm"))
 ```
@@
 
+Saves plot grid with example unambiguous and ambiguous language.
+```{r}
+p2 <- ggdraw() +
+  draw_image("../../paper/figs/ambiguous_lang.png", scale=0.75)
+p3 <- ggdraw() +
+  draw_image("../../paper/figs/unambiguous_lang.png", scale=0.75, clip="on")
+
+g1 <- plot_grid(p2, p3, align='vh', ncol=1, labels=c("B", "C")) +
+  theme(plot.margin = unit(c(0, 0, 0, 0.4), "cm"))
+g2 <- plot_grid(p1, g1, ncol=2, labels=c("A", "", ""))
+
+ggsave(file.path(getwd(), "../../paper/figs/fig1.png"), plot=g2, width=7, height=3)
+```
+
 Saves single fig plot.
@@
 ```
+```{r}
+df %>%
+  filter(sim_id==1, n_contexts==1) %>%
+  select(idxs, is_min_ce, is_min_listener_entropy, is_min_speaker_entropy) %>%
+  gather(typ, val, c(is_min_ce, is_min_listener_entropy, is_min_speaker_entropy)) %>%
+  filter(val=='True')
+
+```

### 2019-02-01 paper/paper.Rmd
initial ndg changes to simulation 2 intro and matrix added to plot 1. just pushing in case i mess all the formatting up...

@@
 We conduct $N=2000$ simulations.
-For each simulation we enumerate the set of *valid* languages in which $|U|=|M|=4$ (recall that $U$ is our set of utterances and $M$ our set of meanings).
-Note that a language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment) and each form maps to at least one meaning (every column has some non-zero assignment).
+For each simulation we enumerate the set of *valid* languages in which $|U|=|M|=4$ ($U$ is our set of utterances and $M$ our set of meanings).
+Recall that languages are boolean matrices and a language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment) and each form maps to at least one meaning (every row has some non-zero assignment).
 For a given simulation, the goal is to find the language $\ell^*$ which minimizes our objective and then check to see if that language contains ambiguity.\par
@@
 In Simulation 1 we explored the degree to which ambiguity is an efficient property of languages when communication is contextualized.
-Figure 1 plots the proportion of optimal languages under each objective as a function of context size.
-The red line shows that as the number of contexts increases, so does the probability that an optimal language $\ell^*_{cross}$ contains ambiguity under our CE objective.
+Figure 1 plots the proportion of optimal languages under each objective as a function of number of contexts.
+The red line shows that as the number of contexts increases, so does the probability that an optimal language $\ell^*_{cross}$ contains ambiguity (at least one utterance maps to two meanings) under our CE objective.
 For comparison we also plot the proportion of speaker-optimal $\ell^*_{speaker}$ (blue line) and listener-optimal $\ell^*_{listener}$ (green line) languages that contain ambiguity.
@@
 
-While our results indicate that ambiguity is an efficient property of contextualized language, these simulations assumed that agents could always disambiguate items in context.
-That is, in our four conditions both agents had perfect knowledge of the relevant conditional distributions ($P(M|C)$).
+While our results indicate that ambiguity is an efficient property of contextualized language, these simulations assumed that agents had perfect knowledge of the relevant conditional distributions ($P(M|C)$).
 This assumption may be too strong for describing much of day-to-day communication -- we seldom interact with others with perfect knowledge of the current context (or topic) at the start of a conversation.
@@
 
-We shift focus in Simulation 2 -- instead of examining the landscape of a space of languages under our CE objective, we consider a single language containing ambiguity, examining how that property is used over a discourse.
-If context is not informative with respect to meaning early in a discourse, but is at later positions, then an efficient speaking strategy would avoid ambiguous material early, saving it for when contextual information is disambiguating.
-We can consider “context” as analogous to a “topic” of conversation.
-As an example, imagine a scenario in which a reader is beginning a magazine article.
+In Simulation 1 we showed that efficiency defined in terms of pragmatic agents leads to a preference for languages that contain ambiguity.
+In Simulation 2 we assume a single fixed language $\ell$, which contains ambiguity, and instead vary the types of agents using $\ell$.
+We will show that efficient *use* of ambiguity depends on an agent's ability to use context for disambiguation.
+This instantiates an on-line perspective on the communicative function of ambiguity theory.
+More generally, it demonstrates how the reference game framework can be used to asses questions of efficient language use (current simulation) as well as language design (Simulation 1).\par
+
+Imagine a scenario in which a reader is beginning a news article.
 While they may have some knowledge about the article’s topic (perhaps from the title), they may not have complete knowledge of its contents, including the persons or events involved.
 In this setting, using a low-cost, but ambiguous referring expression (say a pronoun like "he") early may lead to misunderstanding if context is not informative.
-But, if by a later position enough contextual information has accumulated, it may be efficient to use the ambiguous expression.\par
+But, if by a later position enough contextual information has accumulated, it may be efficient to use the ambiguous expression.
+We pursue this general framework in Simulation 2 -- examining when in a discourse using ambigity is efficient.
+Note that we will consider “context” as analogous to a “topic” of conversation.
+\par
 
@@
 Over the course of a discourse $D$, the listener tries to infer both the current topic, $c_{current}$, as well as the particular meaning $m$ of a given utterance $u$.
-That is, we consider updated listener and speakers ($L(m, c|u,D;\ell)$, $S(u|m,c,D;\ell)$) in which both agents can track the history of previous utterances $D$.
+That is, we consider listener and speaker agents who can track the history of previous utterances $D$.
 Importantly, an agent can attempt to infer the current topic of conversation $c_{current}$ using the discourse history $D$.\par
@@
 ## Speaker agents
+We vary the degree to which agents can use context for disambiguation.
 We consider three types of speaker models.
 Our \textit{Full pragmatics} agent, models a speaker who reasons about her listener and also has complete recall of the set of utterances in the discourse $D$.
-This speaker believes that the listener may not know the current topic $c_{current}$ at the start of the discrouse, but can infer it over the discourse.
+This speaker believes that the listener may not know the current topic $c_{current}$ at the start of the discourse, but can infer it over the discourse.
 We compare two baseline models.
-The first, a \textit{Partial pragmatics} baseline describes a speaker who reasons about a listener, but assumes they have no access to discourse history.
-The second, a \textit{No pragmatics} speaker who does not consider a listener at all, but produces utterances according to the underlying langauge semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$) (see SI section 3).\par
+The first, a \textit{Partial pragmatics} baseline describes a speaker who reasons about a listener, but assumes they have no access to the discourse history.
+The second, a \textit{No pragmatics} speaker does not consider a listener at all, but produces utterances according to the underlying langauge semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$) (see SI section 3).\par
 
@@
 
-We are interested in when it is efficient to use low-cost, ambiguous material in this discourse setting.
+We are interested in how each speaker-model uses ambiguity over the discourse.
 A speaker strategy that is mutually efficient for both agents should avoid ambiguity until sufficient contextual information has accumulated.
@@
 While speaker effort remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines, effort declines in the \textit{Full pragmatic} model as she increasingly relies on ambiguous material later in the discourse.\par
-
-In Simulation 2 we explored how ambiguity could be employed efficiently during language use.
-We assumed that the listener did not know the current context (topic) \textit{a priori}, but could infer it from the discourse history.
-We examined speaker production behavior, tracking use of low-cost, but ambiguous material over the discourse.
-Early in the discourse the \textit{Full pragmatic} speaker avoided using ambiguous material, only using the material later when the current topic was known to the listener.
 

### 2019-02-01 paper/paper.tex
initial ndg changes to simulation 2 intro and matrix added to plot 1. just pushing in case i mess all the formatting up...

@@
 This particular formalization of speaker-cost is general enough to
-accommodate a range of cost instantiations, such as production difficult
-via articulation effort, cognitive effort related to lexical access, or
-others (Bennett \& Goodman, 2018).\par
+accommodate a range of cost instantiations, such as production
+difficulty via articulation effort, cognitive effort related to lexical
+access, or others (Bennett \& Goodman, 2018).\par
 
@@
 We conduct \(N=2000\) simulations. For each simulation we enumerate the
-set of \emph{valid} languages in which \(|U|=|M|=4\) (recall that \(U\)
-is our set of utterances and \(M\) our set of meanings). Note that a
-language \(\ell \in L\) is ``valid'' so long as each possible meaning in
-\(m \in M\) can be referred to by at least one form \(u \in U\) (every
-column of \(\ell\) has some non-zero assignment) and each form maps to
-at least one meaning (every column has some non-zero assignment). For a
-given simulation, the goal is to find the language \(\ell^*\) which
-minimizes our objective and then check to see if that language contains
-ambiguity.\par
+set of \emph{valid} languages in which \(|U|=|M|=4\) (\(U\) is our set
+of utterances and \(M\) our set of meanings). Recall that languages are
+boolean matrices and a language \(\ell \in L\) is ``valid'' so long as
+each possible meaning in \(m \in M\) can be referred to by at least one
+form \(u \in U\) (every column of \(\ell\) has some non-zero assignment)
+and each form maps to at least one meaning (every row has some non-zero
+assignment). For a given simulation, the goal is to find the language
+\(\ell^*\) which minimizes our objective and then check to see if that
+language contains ambiguity.\par
 
@@
 Figure 1 plots the proportion of optimal languages under each objective
-as a function of context size. The red line shows that as the number of
-contexts increases, so does the probability that an optimal language
-\(\ell^*_{cross}\) contains ambiguity under our CE objective. For
-comparison we also plot the proportion of speaker-optimal
-\(\ell^*_{speaker}\) (blue line) and listener-optimal
-\(\ell^*_{listener}\) (green line) languages that contain ambiguity. In
-line with Zipf's predictions, if languages are designed only to minimize
-speaker effort then optimal languages always contain ambiguity. If
-languages are designed to minimize listener effort then ambiguity is
-always avoided.\par
+as a function of number of contexts. The red line shows that as the
+number of contexts increases, so does the probability that an optimal
+language \(\ell^*_{cross}\) contains ambiguity (at least one utterance
+maps to two meanings) under our CE objective. For comparison we also
+plot the proportion of speaker-optimal \(\ell^*_{speaker}\) (blue line)
+and listener-optimal \(\ell^*_{listener}\) (green line) languages that
+contain ambiguity. In line with Zipf's predictions, if languages are
+designed only to minimize speaker effort then optimal languages always
+contain ambiguity. If languages are designed to minimize listener effort
+then ambiguity is always avoided.\par
 
 While our results indicate that ambiguity is an efficient property of
-contextualized language, these simulations assumed that agents could
-always disambiguate items in context. That is, in our four conditions
-both agents had perfect knowledge of the relevant conditional
-distributions (\(P(M|C)\)). This assumption may be too strong for
-describing much of day-to-day communication -- we seldom interact with
-others with perfect knowledge of the current context (or topic) at the
-start of a conversation. To explore how ambiguity may be \textit{used}
-efficiently in our framework, we next examine a case in which the
-listener has imperfect knowledge of context at the start of the
-conversation, but may infer it from the discourse history.\par
+contextualized language, these simulations assumed that agents had
+perfect knowledge of the relevant conditional distributions
+(\(P(M|C)\)). This assumption may be too strong for describing much of
+day-to-day communication -- we seldom interact with others with perfect
+knowledge of the current context (or topic) at the start of a
+conversation. To explore how ambiguity may be \textit{used} efficiently
+in our framework, we next examine a case in which the listener has
+imperfect knowledge of context at the start of the conversation, but may
+infer it from the discourse history.\par
 
@@
 
-We shift focus in Simulation 2 -- instead of examining the landscape of
-a space of languages under our CE objective, we consider a single
-language containing ambiguity, examining how that property is used over
-a discourse. If context is not informative with respect to meaning early
-in a discourse, but is at later positions, then an efficient speaking
-strategy would avoid ambiguous material early, saving it for when
-contextual information is disambiguating. We can consider ``context'' as
-analogous to a ``topic'' of conversation. As an example, imagine a
-scenario in which a reader is beginning a magazine article. While they
-may have some knowledge about the article's topic (perhaps from the
+In Simulation 1 we showed that efficiency defined in terms of pragmatic
+agents leads to a preference for languages that contain ambiguity. In
+Simulation 2 we assume a single fixed language \(\ell\), which contains
+ambiguity, and instead vary the types of agents using \(\ell\). We will
+show that efficient \emph{use} of ambiguity depends on an agent's
+ability to use context for disambiguation. This instantiates an on-line
+perspective on the communicative function of ambiguity theory. More
+generally, it demonstrates how the reference game framework can be used
+to asses questions of efficient language use (current simulation) as
+well as language design (Simulation 1).\par
+
+Imagine a scenario in which a reader is beginning a news article. While
+they may have some knowledge about the article's topic (perhaps from the
 title), they may not have complete knowledge of its contents, including
@@
 later position enough contextual information has accumulated, it may be
-efficient to use the ambiguous expression.\par
+efficient to use the ambiguous expression. We pursue this general
+framework in Simulation 2 -- examining when in a discourse using
+ambigity is efficient. Note that we will consider ``context'' as
+analogous to a ``topic'' of conversation. \par
 
@@
 the current topic, \(c_{current}\), as well as the particular meaning
-\(m\) of a given utterance \(u\). That is, we consider updated listener
-and speakers (\(L(m, c|u,D;\ell)\), \(S(u|m,c,D;\ell)\)) in which both
-agents can track the history of previous utterances \(D\). Importantly,
-an agent can attempt to infer the current topic of conversation
-\(c_{current}\) using the discourse history \(D\).\par
+\(m\) of a given utterance \(u\). That is, we consider listener and
+speaker agents who can track the history of previous utterances \(D\).
+Importantly, an agent can attempt to infer the current topic of
+conversation \(c_{current}\) using the discourse history \(D\).\par
 
@@
 
+We vary the degree to which agents can use context for disambiguation.
 We consider three types of speaker models. Our \textit{Full pragmatics}
@@
 speaker believes that the listener may not know the current topic
-\(c_{current}\) at the start of the discrouse, but can infer it over the
+\(c_{current}\) at the start of the discourse, but can infer it over the
 discourse. We compare two baseline models. The first, a
 \textit{Partial pragmatics} baseline describes a speaker who reasons
-about a listener, but assumes they have no access to discourse history.
-The second, a \textit{No pragmatics} speaker who does not consider a
-listener at all, but produces utterances according to the underlying
+about a listener, but assumes they have no access to the discourse
+history. The second, a \textit{No pragmatics} speaker does not consider
+a listener at all, but produces utterances according to the underlying
 langauge semantics (\(\ell\)) and topic probabilities
@@
 
-We are interested in when it is efficient to use low-cost, ambiguous
-material in this discourse setting. A speaker strategy that is mutually
-efficient for both agents should avoid ambiguity until sufficient
-contextual information has accumulated. We should expect this to be
-reflected in our \textit{Full pragmatics} model who reasons about the
-listener and discourse history. By contrast, a speaker-optimal model who
-does not consider the listener should greedily use ambiguous utterances
+We are interested in how each speaker-model uses ambiguity over the
+discourse. A speaker strategy that is mutually efficient for both agents
+should avoid ambiguity until sufficient contextual information has
+accumulated. We should expect this to be reflected in our
+\textit{Full pragmatics} model who reasons about the listener and
+discourse history. By contrast, a speaker-optimal model who does not
+consider the listener should greedily use ambiguous utterances
 (\textit{No Pragmatics} model), while a listener-optimal model should
@@
 material later in the discourse.\par
-
-In Simulation 2 we explored how ambiguity could be employed efficiently
-during language use. We assumed that the listener did not know the
-current context (topic) \textit{a priori}, but could infer it from the
-discourse history. We examined speaker production behavior, tracking use
-of low-cost, but ambiguous material over the discourse. Early in the
-discourse the \textit{Full pragmatic} speaker avoided using ambiguous
-material, only using the material later when the current topic was known
-to the listener.
 
@@
 \hypertarget{ref-GoodmanFrank2016a}{}
-Goodman, N., \& Frank, M. (2016). Pragmatic language interpreation as
+Goodman, N., \& Frank, M. (2016). Pragmatic language interpretation as
 probabilistic inference. \emph{Trends in Cognitive Sciences},

### 2019-02-01 paper/paper.Rmd
update with matrix in plots according to NDG reco

@@
 
-```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=8, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Optimal languages are more likely to contain ambiguous items as the amount of contextual information increases.
-Vertical axis shows the proportion of optimal languages containing ambiguity.
+```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Panel (A) Vertical axis shows the proportion of optimal languages containing ambiguity.
 Horizontal axis shows the number of context-sizes (1-4).
-Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only objectives.
-Error bars represent 95 percent confidence intervals."}
+Optimal language under CE objective (red).
+Speaker-optimal (blue).
+Listener-optimal (green).
+Error bars represent 95 percent confidence intervals.
+Panel (B), example CE-optimal language (ambiguous) under in a four-condition context.
+Panel (C), example CE-optimal language in a single-condition context."}
 img <- png::readPNG("figs/fig1.png")

### 2019-02-01 paper/paper.tex
update with matrix in plots according to NDG reco

@@
 
-\caption[Optimal languages are more likely to contain ambiguous items as the amount of contextual information increases]{Optimal languages are more likely to contain ambiguous items as the amount of contextual information increases.
-Vertical axis shows the proportion of optimal languages containing ambiguity.
+\caption[Panel (A) Vertical axis shows the proportion of optimal languages containing ambiguity]{Panel (A) Vertical axis shows the proportion of optimal languages containing ambiguity.
 Horizontal axis shows the number of context-sizes (1-4).
-Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only objectives.
-Error bars represent 95 percent confidence intervals.}\label{fig:plot-optimal-langs}
+Optimal language under CE objective (red).
+Speaker-optimal (blue).
+Listener-optimal (green).
+Error bars represent 95 percent confidence intervals.
+Panel (B), example CE-optimal language (ambiguous) under in a four-condition context.
+Panel (C), example CE-optimal language in a single-condition context.}\label{fig:plot-optimal-langs}
 \end{figure*}

### 2019-02-01 paper/paper.Rmd
fix typo, almost good to go

@@
 We explored the interactions of rational pragmatic agents as a framework for understanding efficient language structure and use.
-We focused on an argument on the communicative function of ambiguity [@Piantadosi2011a], deriving a novel speaker-listener Cross-Entropy objective for measuring the efficiency linguistic systems from first principles of efficient language use.
+We focused on an argument on the communicative function of ambiguity [@Piantadosi2011a], deriving a novel speaker-listener Cross-Entropy objective for measuring the efficiency of linguistic systems from first principles of efficient language use.
 In a series of simulations we showed that optimal languages are more likely to contain ambiguous material when context is informative and that rational pragmatic agents will use ambiguous material efficiently.\par

### 2019-02-01 paper/paper.tex
fix typo, almost good to go

@@
 2011), deriving a novel speaker-listener Cross-Entropy objective for
-measuring the efficiency linguistic systems from first principles of
+measuring the efficiency of linguistic systems from first principles of
 efficient language use. In a series of simulations we showed that

### 2019-02-01 ambiguity/simulation-runners/context-ambiguity-plots.Rmd
update fig1 blot aesthetics

@@
   ggplot(aes(x=n_contexts, y=prop_ambiguous, col=Objective)) +
-    geom_line(size=1, alpha=0.9) +
-    geom_errorbar(aes(ymin=ci_low, ymax=ci_high), width=0.05, alpha=0.7, col='black') +
+    geom_line() +
+    geom_errorbar(aes(ymin=ci_low, ymax=ci_high), alpha=0.5, width=0.1, col='black') +
+    geom_point(size=1.5, alpha=0.8) +
     ylab("Percent optimal languages \n containing ambiguity") +

### 2019-02-01 paper/paper.Rmd
1st round FD updates

@@
 Among others, this efficiency-based framing has produced a rich set of theoretical and empirical targets exploring semantic typology [@RegierKempKay2015a], properties such as ambiguity [@Piantadosi2011a] and compositionality [@KirbyGriffithsSmith2014a], and the efficient use of reduction and redundancy in production [@LevyJaeger2007a; @GenzelCharniak2002a].
-These projects typically employ information-theoretic principles to characterize efficiency in natural or simulated language corpora.
-We adopt a different approach in this work, deriving a measure of efficiency from first principles of \textit{efficient language use} and show how this can lead to communicatively efficient systems.\par
+These projects typically employ information-theoretic principles to characterize efficiency in natural or simulated language corpora.\par
+
+We adopt a different approach in this work in two ways.
+First we derive a measure of efficiency from first principles of \textit{efficient language use} and show how this can lead to communicatively efficient systems.
+Second we demonstrate how assumptions about language-users impacts whether language properties are used efficiently.
+In this way, we bridge questions of design and use in a single framework.\par
 
@@
 We formalize this connection -- showing how system-level efficiencies can emerge from local interaction behavior of pragmatic agents.
-Intuitively, our claim is that to understand an "efficient" property of a system it is essential that we consider how that property is *used* efficiently.\par
+Our claim is that to understand an "efficient" property of a system it is essential that we consider how that property is *used* efficiently.\par
 
@@
 RSA is a formalization of essential Gricean pragmatic principles -- agents reason about one another and their shared context [@Grice1975a].
-We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speaker-listeners in the reference game setting (see SI and @GoodmanFrank2016a for review).\par
+We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speaker-listeners in the reference game setting (see \href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI} and @GoodmanFrank2016a for a review).\par
 
@@
 $e \sim p(c)p(m|c)S(u|m, c; \ell)$).
-From the ingredients described above, it is possible to derive the following objective between the speaker and listener joint distributions (see SI 2.2 for a complete derivation).
+From the ingredients described above, it is possible to derive the following objective between the speaker and listener distributions (see \href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 2.1} for a complete derivation).
 
@@
 From an information-theoretic perspective this objective is intuitive: $H_{cross}$ denotes the Cross-Entropy (CE), a measure of dissimilarity between two distributions -- the average number of bits required to communicate under one distribution, given that the “true” distribution differs.
-In our case, we have an expectation over this term -- the expected difference between the distributions assumed by the speaker $P_{speaker}$ and listener $P_{listener}$ given a set of contexts $C$\footnote{Note that in the single context case $|C|=1$ this objective is simply the speaker-listener Cross-Entropy over utterance-meaning pairs $H_{cross}[P_{speaker}(u, m; \ell), P_{listener}(u, m; \ell)]$}.
+In our case, we have an expectation over this term -- the expected difference between the distributions assumed by the speaker $P_{speaker}$ and listener $P_{listener}$ given a set of contexts $C$\footnote{Note that in the single context case $|C|=1$ this objective is simply the speaker-listener Cross-Entropy.}.
 In other words, an "efficient" language $\ell$ minimizes the distance between what speakers and listeners think.
@@
 The task of understanding language is marked by a frequent need to handle various forms of ambiguity: lexical, syntactic, among others [@WasowPerforsBeaver2005a].
-The ubiquity of this property, however, has been argued to provide evidence that languages have not been optimized for communication\footnote{Chomsky's (2002) claim that "...
-language is not well designed, because you have such properties as ambiguity" is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$.} [@Chomsky2002a].
+The ubiquity of this property, however, has been argued to provide evidence that languages have not been optimized for communication [@Chomsky2002a].
 
@@
 In the following experiments we explore two aspects of Piantadosi’s et al.'s claim.
-In Simulation 1, we examine the efficient language *structure* aspect of their claim, exploring when the optimal linguistic system $\ell^*$ is most likely to contain ambiguous lexical items.
-In Simulation 2, we explore an efficient language *use* aspect of the claim -- at what point in a conversation is it useful for a speaker to use ambiguous lexical material?\par
+In Simulation 1, we examine the efficient language *structure* aspect of their claim, exploring when the optimal linguistic system $\ell^*$ is most likely to contain ambiguous expressions.
+In Simulation 2, we explore an efficient language *use* aspect of the claim -- at what point in a conversation is it useful for a speaker to use ambiguous expressions?\par
 
-```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Panel (A) Vertical axis shows the proportion of optimal languages containing ambiguity.
+```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=2.8, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Panel (A) Vertical axis shows the proportion of optimal languages containing ambiguity.
 Horizontal axis shows the number of context-sizes (1-4).
@@
 Panel (B), example CE-optimal language (ambiguous) under in a four-condition context.
-Panel (C), example CE-optimal language in a single-condition context."}
+Panel (C), example CE-optimal language (unambiguous) in a single-condition context."}
 img <- png::readPNG("figs/fig1.png")
@@
 We search the space of languages, examining whether ones which minimize our objective contain ambiguity.
-If context leads to more efficient communication, then optimal languages under our objective should be more likely to be ambiguous as the amount of context increases.\par
+If context leads to more efficient communication, then optimal languages should be more likely to be ambiguous as the amount of context increases.\par
 
@@
 For a given simulation, the goal is to find the language $\ell^*$ which minimizes our objective and then check to see if that language contains ambiguity.\par
+
+We define language efficiency as a function of the particular semantic mappings induced by that language, the speaker and listener agents ($S$ and $L$), as well as the utterance ($P(U)$), meaning ($P(M)$), and context priors ($P(C)$).
+Rather than assume particular structure, for each simulation we generate $P(U) \sim \text{Dir}(1, |U|)$, $P(M|C=c) \sim \text{Dir}(1, |M|)$ (a separate conditional distribution over meanings for each context $c$), and $P(C) \sim \text{Dir}(1, |C|)$, where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.\par
 
@@
 ```
-
-We define language efficiency as a function of the particular semantic mappings induced by that language, the speaker and listener agents ($S$ and $L$), as well as the utterance ($P(U)$), meaning ($P(M)$), and context priors ($P(C)$).
-Rather than assume particular structure, for each simulation we generate $P(U) \sim \text{Dir}(1, |U|)$, $P(M|C=c) \sim \text{Dir}(1, |M|)$ (a separate conditional distribution over meanings for each context $c$), and $P(C) \sim \text{Dir}(1, |C|)$, where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.\par
 
@@
 Zipf [-@Zipf1949a] proposed that the speaker-optimal language $\ell_{speaker}^*$ would minimize speaker effort and the listener-optimal language $\ell_{listener}^*$ would minimize listener effort.
-We define these objectives using the first and second half of equation 1 (see SI section 2.2.
-for more detail).\par
+We define these objectives using the first and second half of equation 1 (see \href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 2.2.}).\par
 
@@
 In Simulation 1 we explored the degree to which ambiguity is an efficient property of languages when communication is contextualized.
-Figure 1 plots the proportion of optimal languages under each objective as a function of number of contexts.
+Figure 2, panel (A) plots the proportion of optimal languages under each objective as a function of number of contexts.
 The red line shows that as the number of contexts increases, so does the probability that an optimal language $\ell^*_{cross}$ contains ambiguity (at least one utterance maps to two meanings) under our CE objective.
@@
 
-We consider consider a single language $\ell$, which contains both ambiguous and unambiguous utterances.
+We consider a single language $\ell$, which contains both ambiguous and unambiguous utterances.
 We assume ambiguous utterances are lower cost.
@@
 Over the course of a discourse $D$, the listener tries to infer both the current topic, $c_{current}$, as well as the particular meaning $m$ of a given utterance $u$.
-That is, we consider listener and speaker agents who can track the history of previous utterances $D$.
+That is, we consider agents who can track the history of previous utterances $D$.
 Importantly, an agent can attempt to infer the current topic of conversation $c_{current}$ using the discourse history $D$.\par
@@
 The first, a \textit{Partial pragmatics} baseline describes a speaker who reasons about a listener, but assumes they have no access to the discourse history.
-The second, a \textit{No pragmatics} speaker does not consider a listener at all, but produces utterances according to the underlying langauge semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$) (see SI section 3).\par
+The second, a \textit{No pragmatics} baseline speaker does not consider a listener at all, but produces utterances according to the underlying langauge semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$) (see \href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 3}).\par
 
@@
 Figure 2 (A) shows the empirical probability that a speaker uses an ambiguous utterance as a function of discourse position.
-The \textit{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse and the \textit{Partial pragmatic} baseline avoids ambiguous utterances entirely.
+The \textit{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse.
+The \textit{Partial pragmatic} baseline avoids ambiguous utterances entirely.
 But, the \textit{Full pragmatic} model avoids ambiguous material only at the start of the discourse, employing it increasingly as the discourse proceeds.
@@
 We focused on an argument on the communicative function of ambiguity [@Piantadosi2011a], deriving a novel speaker-listener Cross-Entropy objective for measuring the efficiency of linguistic systems from first principles of efficient language use.
-In a series of simulations we showed that optimal languages are more likely to contain ambiguous material when context is informative and that rational pragmatic agents will use ambiguous material efficiently.\par
+In Simulation 1 we showed that optimal languages are more likely to contain ambiguous material when context is informative.
+In Simulation 2 we showed how rational pragmatic agents use ambiguous material efficiently in conversation.\par
 
-A limitation of the current work is an analysis of exactly how this objective compares to existing measures.
-For example, previous work has described competing speaker-listener pressures in terms of a trade-off of simplicity and informativeness [@KempRegier2012a] or expressivity and compressibility [@SmithTamarizKirby2013a].
+A limitation of the current work is an analysis of exactly how the CE objective compares to existing measures.
+For example, previous work has described competing speaker-listener pressures in terms of a trade-off of simplicity and informativeness [@KempRegier2012a] or expressivity and compressibility [@SmithTamarizKirby2013a] to explain linguistic regularities.
 Future work should assess the degree to which we can derive the same properties as previous studies using our current framework.
@@
 We hope this framework can serve as a domain general tool to assess the range of functionalist theories examining efficient language-structure and use.\par
+
+\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI are available at this \href{https://github.com/benpeloquin7/zipf_principles/}{link}.}}} \vspace{1em}
+\noindent
 

### 2019-02-01 paper/paper.tex
1st round FD updates

@@
 information-theoretic principles to characterize efficiency in natural
-or simulated language corpora. We adopt a different approach in this
-work, deriving a measure of efficiency from first principles of
+or simulated language corpora.\par
+
+We adopt a different approach in this work in two ways. First we derive
+a measure of efficiency from first principles of
 \textit{efficient language use} and show how this can lead to
-communicatively efficient systems.\par
+communicatively efficient systems. Second we demonstrate how assumptions
+about language-users impacts whether language properties are used
+efficiently. In this way, we bridge questions of design and use in a
+single framework.\par
 
@@
 We formalize this connection -- showing how system-level efficiencies
-can emerge from local interaction behavior of pragmatic agents.
-Intuitively, our claim is that to understand an ``efficient'' property
-of a system it is essential that we consider how that property is
-\emph{used} efficiently.\par
+can emerge from local interaction behavior of pragmatic agents. Our
+claim is that to understand an ``efficient'' property of a system it is
+essential that we consider how that property is \emph{used}
+efficiently.\par
 
@@
 representational framework to model Gricean (rational and pragmatic)
-speaker-listeners in the reference game setting (see SI and Goodman \&
-Frank (2016) for review).\par
+speaker-listeners in the reference game setting (see
+\href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI}
+and Goodman \& Frank (2016) for a review).\par
 
@@
 the ingredients described above, it is possible to derive the following
-objective between the speaker and listener joint distributions (see SI
-2.2 for a complete derivation).
+objective between the speaker and listener distributions (see
+\href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 2.1}
+for a complete derivation).
 
@@
 \(P_{speaker}\) and listener \(P_{listener}\) given a set of contexts
-\(C\)\footnote{Note that in the single context case $|C|=1$ this objective is simply the speaker-listener Cross-Entropy over utterance-meaning pairs $H_{cross}[P_{speaker}(u, m; \ell), P_{listener}(u, m; \ell)]$}.
+\(C\)\footnote{Note that in the single context case $|C|=1$ this objective is simply the speaker-listener Cross-Entropy.}.
 In other words, an ``efficient'' language \(\ell\) minimizes the
@@
 however, has been argued to provide evidence that languages have not
-been optimized for
-communication\footnote{Chomsky's (2002) claim that "...
-language is not well designed, because you have such properties as ambiguity" is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$.}
-(Chomsky, 2002).
+been optimized for communication (Chomsky, 2002).
 
@@
 \emph{structure} aspect of their claim, exploring when the optimal
-linguistic system \(\ell^*\) is most likely to contain ambiguous lexical
-items. In Simulation 2, we explore an efficient language \emph{use}
-aspect of the claim -- at what point in a conversation is it useful for
-a speaker to use ambiguous lexical material?\par
+linguistic system \(\ell^*\) is most likely to contain ambiguous
+expressions. In Simulation 2, we explore an efficient language
+\emph{use} aspect of the claim -- at what point in a conversation is it
+useful for a speaker to use ambiguous expressions?\par
 
@@
 Panel (B), example CE-optimal language (ambiguous) under in a four-condition context.
-Panel (C), example CE-optimal language in a single-condition context.}\label{fig:plot-optimal-langs}
+Panel (C), example CE-optimal language (unambiguous) in a single-condition context.}\label{fig:plot-optimal-langs}
 \end{figure*}
@@
 our objective contain ambiguity. If context leads to more efficient
-communication, then optimal languages under our objective should be more
-likely to be ambiguous as the amount of context increases.\par
+communication, then optimal languages should be more likely to be
+ambiguous as the amount of context increases.\par
 
@@
 
+We define language efficiency as a function of the particular semantic
+mappings induced by that language, the speaker and listener agents
+(\(S\) and \(L\)), as well as the utterance (\(P(U)\)), meaning
+(\(P(M)\)), and context priors (\(P(C)\)). Rather than assume particular
+structure, for each simulation we generate
+\(P(U) \sim \text{Dir}(1, |U|)\), \(P(M|C=c) \sim \text{Dir}(1, |M|)\)
+(a separate conditional distribution over meanings for each context
+\(c\)), and \(P(C) \sim \text{Dir}(1, |C|)\), where \(\text{Dir}(1, k)\)
+specifies the uniform Dirichlet distribution over a \(k\)-dimensional
+probability vector.\par
+
 \begin{CodeChunk}
@@
 \end{CodeChunk}
-
-We define language efficiency as a function of the particular semantic
-mappings induced by that language, the speaker and listener agents
-(\(S\) and \(L\)), as well as the utterance (\(P(U)\)), meaning
-(\(P(M)\)), and context priors (\(P(C)\)). Rather than assume particular
-structure, for each simulation we generate
-\(P(U) \sim \text{Dir}(1, |U|)\), \(P(M|C=c) \sim \text{Dir}(1, |M|)\)
-(a separate conditional distribution over meanings for each context
-\(c\)), and \(P(C) \sim \text{Dir}(1, |C|)\), where \(\text{Dir}(1, k)\)
-specifies the uniform Dirichlet distribution over a \(k\)-dimensional
-probability vector.\par
 
@@
 effort. We define these objectives using the first and second half of
-equation 1 (see SI section 2.2. for more detail).\par
+equation 1 (see
+\href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 2.2.}).\par
 
@@
 efficient property of languages when communication is contextualized.
-Figure 1 plots the proportion of optimal languages under each objective
-as a function of number of contexts. The red line shows that as the
-number of contexts increases, so does the probability that an optimal
-language \(\ell^*_{cross}\) contains ambiguity (at least one utterance
-maps to two meanings) under our CE objective. For comparison we also
-plot the proportion of speaker-optimal \(\ell^*_{speaker}\) (blue line)
-and listener-optimal \(\ell^*_{listener}\) (green line) languages that
-contain ambiguity. In line with Zipf's predictions, if languages are
-designed only to minimize speaker effort then optimal languages always
-contain ambiguity. If languages are designed to minimize listener effort
-then ambiguity is always avoided.\par
+Figure 2, panel (A) plots the proportion of optimal languages under each
+objective as a function of number of contexts. The red line shows that
+as the number of contexts increases, so does the probability that an
+optimal language \(\ell^*_{cross}\) contains ambiguity (at least one
+utterance maps to two meanings) under our CE objective. For comparison
+we also plot the proportion of speaker-optimal \(\ell^*_{speaker}\)
+(blue line) and listener-optimal \(\ell^*_{listener}\) (green line)
+languages that contain ambiguity. In line with Zipf's predictions, if
+languages are designed only to minimize speaker effort then optimal
+languages always contain ambiguity. If languages are designed to
+minimize listener effort then ambiguity is always avoided.\par
 
@@
 
-We consider consider a single language \(\ell\), which contains both
-ambiguous and unambiguous utterances. We assume ambiguous utterances are
-lower cost. Crucially, we do not assume that the listener knows the
-particular topic (\(c_{current}\)) of the conversation
-\textit{a priori}. Rather, that the listener has knowledge of the set of
-possible topics \(C = \{c_1, \dots, c_k\}\), but does not know which one
-is currently being used by the speaker. Formally, this means the
-listener does not have access to the correct conditional distribution
-over meanings \(P(M|C=c_{current})\) at the start of the discourse.\par
+We consider a single language \(\ell\), which contains both ambiguous
+and unambiguous utterances. We assume ambiguous utterances are lower
+cost. Crucially, we do not assume that the listener knows the particular
+topic (\(c_{current}\)) of the conversation \textit{a priori}. Rather,
+that the listener has knowledge of the set of possible topics
+\(C = \{c_1, \dots, c_k\}\), but does not know which one is currently
+being used by the speaker. Formally, this means the listener does not
+have access to the correct conditional distribution over meanings
+\(P(M|C=c_{current})\) at the start of the discourse.\par
 
@@
 the current topic, \(c_{current}\), as well as the particular meaning
-\(m\) of a given utterance \(u\). That is, we consider listener and
-speaker agents who can track the history of previous utterances \(D\).
-Importantly, an agent can attempt to infer the current topic of
-conversation \(c_{current}\) using the discourse history \(D\).\par
+\(m\) of a given utterance \(u\). That is, we consider agents who can
+track the history of previous utterances \(D\). Importantly, an agent
+can attempt to infer the current topic of conversation \(c_{current}\)
+using the discourse history \(D\).\par
 
@@
 about a listener, but assumes they have no access to the discourse
-history. The second, a \textit{No pragmatics} speaker does not consider
-a listener at all, but produces utterances according to the underlying
-langauge semantics (\(\ell\)) and topic probabilities
-(\(p(M|C=c_{current}\)) (see SI section 3).\par
+history. The second, a \textit{No pragmatics} baseline speaker does not
+consider a listener at all, but produces utterances according to the
+underlying langauge semantics (\(\ell\)) and topic probabilities
+(\(p(M|C=c_{current}\)) (see
+\href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 3}).\par
 
@@
 \textit{No pragmatic} baseline uses ambiguous utterances frequently and
-at a constant rate over the discourse and the \textit{Partial pragmatic}
+at a constant rate over the discourse. The \textit{Partial pragmatic}
 baseline avoids ambiguous utterances entirely. But, the
@@
 measuring the efficiency of linguistic systems from first principles of
-efficient language use. In a series of simulations we showed that
-optimal languages are more likely to contain ambiguous material when
-context is informative and that rational pragmatic agents will use
-ambiguous material efficiently.\par
+efficient language use. In Simulation 1 we showed that optimal languages
+are more likely to contain ambiguous material when context is
+informative. In Simulation 2 we showed how rational pragmatic agents use
+ambiguous material efficiently in conversation.\par
 
-A limitation of the current work is an analysis of exactly how this
+A limitation of the current work is an analysis of exactly how the CE
 objective compares to existing measures. For example, previous work has
@@
 of simplicity and informativeness (Kemp \& Regier, 2012) or expressivity
-and compressibility (Smith, Tamariz, \& Kirby, 2013). Future work should
-assess the degree to which we can derive the same properties as previous
-studies using our current framework. In line with this idea, a primary
-focus of the current work is to provide a case-study on the productivity
-of assessing functionalist theories in repeated reference games between
-rational pragmatic agents. We hope this framework can serve as a domain
-general tool to assess the range of functionalist theories examining
-efficient language-structure and use.\par
+and compressibility (Smith, Tamariz, \& Kirby, 2013) to explain
+linguistic regularities. Future work should assess the degree to which
+we can derive the same properties as previous studies using our current
+framework. In line with this idea, a primary focus of the current work
+is to provide a case-study on the productivity of assessing
+functionalist theories in repeated reference games between rational
+pragmatic agents. We hope this framework can serve as a domain general
+tool to assess the range of functionalist theories examining efficient
+language-structure and use.\par
+
+\vspace{1em}
+
+\fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI are available at this \href{https://github.com/benpeloquin7/zipf_principles/}{link}.}}}
+\vspace{1em} \noindent
 

### 2019-02-01 paper/paper.Rmd
BP final read edits

@@
 
-We adopt a different approach in this work in two ways.
+To preview, we diverge from this approach in two ways.
 First we derive a measure of efficiency from first principles of \textit{efficient language use} and show how this can lead to communicatively efficient systems.
@@
 Using these basic ingredients, we show that the property of interest (ambiguity) is prevalent in languages that optimize our measure of efficiency.
-Further, we show how ambiguity is \textit{used} efficiently during local, in-the-moment interactions.
+Further, we show \textit{how} ambiguity is used efficiently during local, in-the-moment interactions.
 The contributions of this work are twofold -- we derive a novel objective for measuring linguistic efficiency and also show how the reference game framework, in combination with formal models of communication, can be used to connect ideas about system-level efficiencies to in-the-moment language use.\par
@@
 ```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=2.8, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Panel (A) Vertical axis shows the proportion of optimal languages containing ambiguity.
-Horizontal axis shows the number of context-sizes (1-4).
+Horizontal axis shows the context-size (1-4) in each condition.
 Optimal language under CE objective (red).
@@
 This instantiates an on-line perspective on the communicative function of ambiguity theory.
-More generally, it demonstrates how the reference game framework can be used to asses questions of efficient language use (current simulation) as well as language design (Simulation 1).\par
+More generally, it demonstrates how we can assess both questions of efficient use (current simulation) as well as design (previous simulation) in the same framework.\par
 
@@
 
-Figure 2 (A) shows the empirical probability that a speaker uses an ambiguous utterance as a function of discourse position.
+Figure 3, panel (A) shows the empirical probability that a speaker uses an ambiguous utterance as a function of discourse position.
 The \textit{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse.
@@
 
-\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI are available at this \href{https://github.com/benpeloquin7/zipf_principles/}{link}.}}} \vspace{1em}
+\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI are available at this \href{https://github.com/benpeloquin7/zipf_principles/}{repository}.}}} \vspace{1em}
 \noindent

### 2019-02-01 paper/paper.tex
BP final read edits

@@
 
-We adopt a different approach in this work in two ways. First we derive
-a measure of efficiency from first principles of
+To preview, we diverge from this approach in two ways. First we derive a
+measure of efficiency from first principles of
 \textit{efficient language use} and show how this can lead to
@@
 prevalent in languages that optimize our measure of efficiency. Further,
-we show how ambiguity is \textit{used} efficiently during local,
+we show \textit{how} ambiguity is used efficiently during local,
 in-the-moment interactions. The contributions of this work are twofold
@@
 \caption[Panel (A) Vertical axis shows the proportion of optimal languages containing ambiguity]{Panel (A) Vertical axis shows the proportion of optimal languages containing ambiguity.
-Horizontal axis shows the number of context-sizes (1-4).
+Horizontal axis shows the context-size (1-4) in each condition.
 Optimal language under CE objective (red).
@@
 perspective on the communicative function of ambiguity theory. More
-generally, it demonstrates how the reference game framework can be used
-to asses questions of efficient language use (current simulation) as
-well as language design (Simulation 1).\par
+generally, it demonstrates how we can assess both questions of efficient
+use (current simulation) as well as design (previous simulation) in the
+same framework.\par
 
@@
 
-Figure 2 (A) shows the empirical probability that a speaker uses an
-ambiguous utterance as a function of discourse position. The
+Figure 3, panel (A) shows the empirical probability that a speaker uses
+an ambiguous utterance as a function of discourse position. The
 \textit{No pragmatic} baseline uses ambiguous utterances frequently and
@@
 
-\fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI are available at this \href{https://github.com/benpeloquin7/zipf_principles/}{link}.}}}
+\fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI are available at this \href{https://github.com/benpeloquin7/zipf_principles/}{repository}.}}}
 \vspace{1em} \noindent

### 2019-02-01 paper/paper.Rmd
update with NDGs last comments and BP walk through

@@
 Such an approach might reframe our opening question as follows: how does having property \textit{x} make using language $\ell$ more or less useful for communication?
-Among others, this efficiency-based framing has produced a rich set of theoretical and empirical targets exploring semantic typology [@RegierKempKay2015a], properties such as ambiguity [@Piantadosi2011a] and compositionality [@KirbyGriffithsSmith2014a], and the efficient use of reduction and redundancy in production [@LevyJaeger2007a; @GenzelCharniak2002a].
-These projects typically employ information-theoretic principles to characterize efficiency in natural or simulated language corpora.\par
+Among others, this efficiency-based framing has produced a rich set of theoretical and empirical targets exploring semantic typology [@RegierKempKay2015a], properties such as ambiguity [@Piantadosi2011a] and compositionality [@KirbyGriffithsSmith2014a], and the efficient use of reduction and redundancy in production [@LevyJaeger2007a; @GenzelCharniak2002a].\par
 
-To preview, we diverge from this approach in two ways.
-First we derive a measure of efficiency from first principles of \textit{efficient language use} and show how this can lead to communicatively efficient systems.
-Second we demonstrate how assumptions about language-users impacts whether language properties are used efficiently.
+The approaches above typically posit efficiency measures motivated by information-theoretic principles, but not directly grounded in language use by interacting agents.
+In this work, we derive a measure of efficiency from first principles of \textit{efficient language use} and show how this can lead to communicatively efficient systems.
+We also demonstrate how assumptions about interlocutors impacts how language properties are used efficiently.
 In this way, we bridge questions of design and use in a single framework.\par
@@
 We adopt the reference game as our primary unit of interaction and model language users with the Rational Speech Act (RSA) framework -- a computional model of language use, which is nicely attested by experimental data on interaction.
-Using these basic ingredients, we show that the property of interest (ambiguity) is prevalent in languages that optimize our measure of efficiency.
-Further, we show \textit{how} ambiguity is used efficiently during local, in-the-moment interactions.
-The contributions of this work are twofold -- we derive a novel objective for measuring linguistic efficiency and also show how the reference game framework, in combination with formal models of communication, can be used to connect ideas about system-level efficiencies to in-the-moment language use.\par
+Using these ingredients, we show that the property of interest (ambiguity) is prevalent in languages that optimize our measure of efficiency (Simulation 1).
+Further, we show how ambiguity is used efficiently during local, in-the-moment interactions (Simulation 2).
+Put differently, these simulations examine efficiency from two angles -- in the first we vary languages, fixing agents, and search for efficient designs.
+In the second we vary agents, fixing language, and examine efficient use.
+\par
+
+The contributions of this work are twofold -- we derive a novel measure of linguistic efficiency and also show how the reference game framework, in combination with formal models of communication, can be used to connect ideas about system-level efficiencies to in-the-moment language use.\par
 
@@
 ### Languages
-A language $\ell$ defines the set of semantic mappings between utterance and meanings.
-For example, Figure 1 contains four utterances $U = \{\text{blue}, \text{green}, \text{square}, \text{circle}\} \text{ and three meanings }M = \{\text{green-square}, \text{blue-square}, \text{green-circle}\}$.
+A language $\ell$ defines the set of semantic mappings between utterances and meanings.
+For example, Figure 1 contains four utterances $U = \{\text{"blue"}, \text{"green"}, \text{"square"}, \text{"circle"}\} \text{ and three meanings }M = \{\text{green-square}, \text{blue-square}, \text{green-circle}\}$.
 The boolean matrix describes the literal semantics of the language.
@@
 $L(m|u, c;\ell)$).
-Note that both speakers and listeners can induce joint distributions over utterace-meaning pairs, although, importantly, these distributions may differ:
+Note that both speakers and listeners can induce joint distributions over utterance-meaning pairs, although, importantly, these distributions may differ:
 $$P_{speaker}(u, m | c; \ell) = S(u|m, c; \ell)p(m|c)$$
@@
 $e \sim p(c)p(m|c)S(u|m, c; \ell)$).
-From the ingredients described above, it is possible to derive the following objective between the speaker and listener distributions (see \href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 2.1} for a complete derivation).
-
+From these ingredients it is possible to derive the following objective between the speaker and listener distributions (see \href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 2.1} for complete derivation).
 \begin{equation}
@@
 \end{equation}
-
 From an information-theoretic perspective this objective is intuitive: $H_{cross}$ denotes the Cross-Entropy (CE), a measure of dissimilarity between two distributions -- the average number of bits required to communicate under one distribution, given that the “true” distribution differs.
@@
 
-As an example, say we have two objects ($m_1$ and $m_2$), two utterances ($u_1$ and $u_2$), which differ in cost, and two languages ($\ell_1$ and $\ell_2$), which describe different utterance-meaning mappings.
+As an example, say we have two objects ($m_1$ and $m_2$), two utterances ($u_1$ and $u_2$), differing in cost, and two languages ($\ell_1$, $\ell_2$), describing different utterance-meaning mappings.
 In language $\ell_1$, the low-cost $u_1$ can be used to refer to both $m_1$ and $m_2$ ($[[u_1]]_{\ell_1} = \{m_1, m_2 \}$), but the high-cost $u_2$ cannot be used at all ($[[u_2]]_{\ell_1} = \emptyset$).
@@
 
-In the following experiments we explore two aspects of Piantadosi’s et al.'s claim.
+In the following simulations we explore two aspects of Piantadosi’s et al.'s claim.
 In Simulation 1, we examine the efficient language *structure* aspect of their claim, exploring when the optimal linguistic system $\ell^*$ is most likely to contain ambiguous expressions.
-In Simulation 2, we explore an efficient language *use* aspect of the claim -- at what point in a conversation is it useful for a speaker to use ambiguous expressions?\par
+In Simulation 2, we explore an efficient language *use* aspect of the claim -- under what assumptions will agents use ambiguity efficiently in a conversation?\par
 
@@
 
-While our results indicate that ambiguity is an efficient property of contextualized language, these simulations assumed that agents had perfect knowledge of the relevant conditional distributions ($P(M|C)$).
+While our results indicate that ambiguity is an efficient property of contextualized language use, these simulations assumed that agents had perfect knowledge of the relevant conditional distributions ($P(M|C)$).
 This assumption may be too strong for describing much of day-to-day communication -- we seldom interact with others with perfect knowledge of the current context (or topic) at the start of a conversation.
@@
 We will show that efficient *use* of ambiguity depends on an agent's ability to use context for disambiguation.
-This instantiates an on-line perspective on the communicative function of ambiguity theory.
-More generally, it demonstrates how we can assess both questions of efficient use (current simulation) as well as design (previous simulation) in the same framework.\par
+More generally, Simulation 2 is intended to demonstrate how we can assess both questions of efficient use (current simulation) as well as design (previous simulation) in the same framework.\par
 
@@
 
-Figure 3, panel (A) shows the empirical probability that a speaker uses an ambiguous utterance as a function of discourse position.
+Figure 3, panel (A) shows the empirical probability a speaker uses an ambiguous utterance as a function of discourse position.
 The \textit{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse.
@@
 Note that the objective decreases for all three models, primarily driven by the listener updating his belief about the actual topic ($P(C=c_{current}|D)$).
-However, the objective declines more quickly under the \textit{Full} and \textit{Partial} pragmatic speakers as listener agents are better able to infer the correct context.
-Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatic} models at the end of the discourse is driven by the reduction in speaker costs (see panel B).
-While speaker effort remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines, effort declines in the \textit{Full pragmatic} model as she increasingly relies on ambiguous material later in the discourse.\par
+However, the objective declines more quickly under the \textit{Full} and \textit{Partial pragmatic} speakers as listener agents are better able to infer the correct context.
+Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatic} models at the end of the discourse is driven by the reduction in speaker costs.
+Panel (B) tracks speaker effort, which remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines.
+But, effort declines in the \textit{Full pragmatic} model as she increasingly relies on ambiguous material later in the discourse.\par
 
@@
 
-\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI are available at this \href{https://github.com/benpeloquin7/zipf_principles/}{repository}.}}} \vspace{1em}
+\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI: \href{https://bit.ly/2DOnjI4}{\url{https://bit.ly/2DOnjI4}}.}}} \vspace{1em}
 \noindent

### 2019-02-01 paper/paper.tex
update with NDGs last comments and BP walk through

@@
 efficient use of reduction and redundancy in production (Genzel \&
-Charniak, 2002; Levy \& Jaeger, 2007). These projects typically employ
-information-theoretic principles to characterize efficiency in natural
-or simulated language corpora.\par
+Charniak, 2002; Levy \& Jaeger, 2007).\par
 
-To preview, we diverge from this approach in two ways. First we derive a
-measure of efficiency from first principles of
-\textit{efficient language use} and show how this can lead to
-communicatively efficient systems. Second we demonstrate how assumptions
-about language-users impacts whether language properties are used
-efficiently. In this way, we bridge questions of design and use in a
-single framework.\par
+The approaches above typically posit efficiency measures motivated by
+information-theoretic principles, but not directly grounded in language
+use by interacting agents. In this work, we derive a measure of
+efficiency from first principles of \textit{efficient language use} and
+show how this can lead to communicatively efficient systems. We also
+demonstrate how assumptions about interlocutors impacts how language
+properties are used efficiently. In this way, we bridge questions of
+design and use in a single framework.\par
 
@@
 framework -- a computional model of language use, which is nicely
-attested by experimental data on interaction. Using these basic
-ingredients, we show that the property of interest (ambiguity) is
-prevalent in languages that optimize our measure of efficiency. Further,
-we show \textit{how} ambiguity is used efficiently during local,
-in-the-moment interactions. The contributions of this work are twofold
--- we derive a novel objective for measuring linguistic efficiency and
-also show how the reference game framework, in combination with formal
-models of communication, can be used to connect ideas about system-level
-efficiencies to in-the-moment language use.\par
+attested by experimental data on interaction. Using these ingredients,
+we show that the property of interest (ambiguity) is prevalent in
+languages that optimize our measure of efficiency (Simulation 1).
+Further, we show how ambiguity is used efficiently during local,
+in-the-moment interactions (Simulation 2). Put differently, these
+simulations examine efficiency from two angles -- in the first we vary
+languages, fixing agents, and search for efficient designs. In the
+second we vary agents, fixing language, and examine efficient use. \par
+
+The contributions of this work are twofold -- we derive a novel measure
+of linguistic efficiency and also show how the reference game framework,
+in combination with formal models of communication, can be used to
+connect ideas about system-level efficiencies to in-the-moment language
+use.\par
 
@@
 A language \(\ell\) defines the set of semantic mappings between
-utterance and meanings. For example, Figure 1 contains four utterances
-\(U = \{\text{blue}, \text{green}, \text{square}, \text{circle}\} \text{ and three meanings }M = \{\text{green-square}, \text{blue-square}, \text{green-circle}\}\).
+utterances and meanings. For example, Figure 1 contains four utterances
+\(U = \{\text{"blue"}, \text{"green"}, \text{"square"}, \text{"circle"}\} \text{ and three meanings }M = \{\text{green-square}, \text{blue-square}, \text{green-circle}\}\).
 The boolean matrix describes the literal semantics of the language. We
@@
 Note that both speakers and listeners can induce joint distributions
-over utterace-meaning pairs, although, importantly, these distributions
+over utterance-meaning pairs, although, importantly, these distributions
 may differ: \[P_{speaker}(u, m | c; \ell) = S(u|m, c; \ell)p(m|c)\]
@@
 \(S(u|m, c; \ell)\) (i.e. \(e \sim p(c)p(m|c)S(u|m, c; \ell)\)). From
-the ingredients described above, it is possible to derive the following
-objective between the speaker and listener distributions (see
+these ingredients it is possible to derive the following objective
+between the speaker and listener distributions (see
 \href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 2.1}
-for a complete derivation).
+for complete derivation).
 
@@
 As an example, say we have two objects (\(m_1\) and \(m_2\)), two
-utterances (\(u_1\) and \(u_2\)), which differ in cost, and two
-languages (\(\ell_1\) and \(\ell_2\)), which describe different
-utterance-meaning mappings. In language \(\ell_1\), the low-cost \(u_1\)
-can be used to refer to both \(m_1\) and \(m_2\)
+utterances (\(u_1\) and \(u_2\)), differing in cost, and two languages
+(\(\ell_1\), \(\ell_2\)), describing different utterance-meaning
+mappings. In language \(\ell_1\), the low-cost \(u_1\) can be used to
+refer to both \(m_1\) and \(m_2\)
 (\([[u_1]]_{\ell_1} = \{m_1, m_2 \}\)), but the high-cost \(u_2\) cannot
@@
 
-In the following experiments we explore two aspects of Piantadosi's et
+In the following simulations we explore two aspects of Piantadosi's et
 al.'s claim. In Simulation 1, we examine the efficient language
@@
 expressions. In Simulation 2, we explore an efficient language
-\emph{use} aspect of the claim -- at what point in a conversation is it
-useful for a speaker to use ambiguous expressions?\par
+\emph{use} aspect of the claim -- under what assumptions will agents use
+ambiguity efficiently in a conversation?\par
 
@@
 While our results indicate that ambiguity is an efficient property of
-contextualized language, these simulations assumed that agents had
+contextualized language use, these simulations assumed that agents had
 perfect knowledge of the relevant conditional distributions
@@
 show that efficient \emph{use} of ambiguity depends on an agent's
-ability to use context for disambiguation. This instantiates an on-line
-perspective on the communicative function of ambiguity theory. More
-generally, it demonstrates how we can assess both questions of efficient
-use (current simulation) as well as design (previous simulation) in the
-same framework.\par
+ability to use context for disambiguation. More generally, Simulation 2
+is intended to demonstrates how we can assess both questions of
+efficient use (current simulation) as well as design (previous
+simulation) in the same framework.\par
 
@@
 
-Figure 3, panel (A) shows the empirical probability that a speaker uses
-an ambiguous utterance as a function of discourse position. The
+Figure 3, panel (A) shows the empirical probability a speaker uses an
+ambiguous utterance as a function of discourse position. The
 \textit{No pragmatic} baseline uses ambiguous utterances frequently and
@@
 \textit{Partial pragmatic} models at the end of the discourse is driven
-by the reduction in speaker costs (see panel B). While speaker effort
-remains constant in both \textit{No pragmatic} and
-\textit{Partial pragmatic} baselines, effort declines in the
+by the reduction in speaker costs. Panel (B) tracks speaker effort,
+which remains constant in both \textit{No pragmatic} and
+\textit{Partial pragmatic} baselines. But, effort declines in the
 \textit{Full pragmatic} model as she increasingly relies on ambiguous
@@
 
-\fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI are available at this \href{https://github.com/benpeloquin7/zipf_principles/}{repository}.}}}
+\fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI: \href{https://bit.ly/2DOnjI4}{\url{https://bit.ly/2DOnjI4}}.}}}
 \vspace{1em} \noindent

### 2019-02-01 paper/paper.tex
more tweaks

@@
 show how this can lead to communicatively efficient systems. We also
-demonstrate how assumptions about interlocutors impacts how language
+demonstrate that assumptions about interlocutors impact how language
 properties are used efficiently. In this way, we bridge questions of
@@
 ability to use context for disambiguation. More generally, Simulation 2
-is intended to demonstrates how we can assess both questions of
-efficient use (current simulation) as well as design (previous
-simulation) in the same framework.\par
+is intended to demonstrate how we can assess both questions of efficient
+use (current simulation) as well as design (previous simulation) in the
+same framework.\par
 
@@
 topic (\(P(C=c_{current}|D)\)). However, the objective declines more
-quickly under the \textit{Full} and \textit{Partial} pragmatic speakers
+quickly under the \textit{Full} and \textit{Partial pragmatic} speakers
 as listener agents are better able to infer the correct context.

### 2019-02-02 paper/paper.Rmd
more tweaks
MF last changes and spell-checking.

@@
 Zipf articulated what is now considered a *functionalist* approach to language science -- analyzing language structure and use in terms of efficiency.
-Such an approach might reframe our opening question as follows: how does having property \textit{x} make using language $\ell$ more or less useful for communication?
-Among others, this efficiency-based framing has produced a rich set of theoretical and empirical targets exploring semantic typology [@RegierKempKay2015a], properties such as ambiguity [@Piantadosi2011a] and compositionality [@KirbyGriffithsSmith2014a], and the efficient use of reduction and redundancy in production [@LevyJaeger2007a; @GenzelCharniak2002a].\par
+Such an approach might re-frame our opening question as follows: how does having property \textit{x} make using language $\ell$ more or less useful for communication?
+This efficiency-based framing has produced a rich set of theoretical and empirical targets exploring semantic typology [@RegierKempKay2015a], properties such as ambiguity [@Piantadosi2011a] and compositionality [@KirbyGriffithsSmith2014a], and the efficient use of reduction and redundancy in production [@LevyJaeger2007a; @GenzelCharniak2002a].\par
 
-The approaches above typically posit efficiency measures motivated by information-theoretic principles, but not directly grounded in language use by interacting agents.
-In this work, we derive a measure of efficiency from first principles of \textit{efficient language use} and show how this can lead to communicatively efficient systems.
-We also demonstrate how assumptions about interlocutors impacts how language properties are used efficiently.
-In this way, we bridge questions of design and use in a single framework.\par
+The approaches above typically posit efficiency measures that are motivated by information-theoretic principles, but they typically do not ground out in language use by interacting agents.
+In this work, we derive a novel objective function from first principles of \textit{rational language use} and show how optimizing this objective can lead to communicatively efficient systems.
+We also demonstrate that assumptions about interlocutors impact whether language properties are used efficiently.
+In this way, we integrate questions of language design and language use in a single framework.\par
 
-Functionalist theories commonly frame language efficiency in terms of a fundamental effort-assymetry underlying everyday communication: what is “hard” for a speaker is likely different than what is “hard” for a listener.
-Zipf described this as follows -- purely from the standpoint of speaker effort, an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of a single, low-cost word.
+Functionalist theories commonly frame language efficiency in terms of a fundamental effort-asymmetry underlying everyday communication: what is “hard” for a speaker is likely different than what is “hard” for a listener.
+Zipf described this as follows: purely from the standpoint of speaker effort, an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of a single, low-cost word.
 Given such a language, the full set of potential meanings would be conveyed using only that word, i.e.
@@
 Clearly, natural languages fall between the two extremes of $\ell_{speaker}^*$ and $\ell_{listener}^*$.
-Zipf proposed that the particular lexicon-level properties he observed were a result of these competing forces -- the pressure to jointly minimize speaker and listener effort.\par
+Zipf proposed that the particular lexicon-level properties he observed were a result of optimization based on these competing forces -- the pressure to jointly minimize speaker and listener effort.\par
 
@@
 Adopting reference games as a basic unit of analysis suggests that optimization may take place at the level of conversation.
-Importantly, Zipf's conception of speaker and listener effort may be connected to how language is used; in particular, whether interlocutors engage in pragmatic reasoning during conversation.
+Importantly, Zipf's conception of speaker and listener effort should be connected to how language is used; in particular, whether interlocutors engage in pragmatic reasoning during conversation.
 Under a Gricean treatment of pragmatics, speakers and listeners follow a set of conversational maxims in which they cooperate to transfer information [@Grice1975a].
@@
 We choose a property of languages that could, in principle, vary freely, but shows strong regularities across languages.
-The explanadum is why this regularity holds.
+The explanandum is why this regularity holds.
 We examine ambiguity as our property, extending ideas by Piantadosi et al.
@@
 We define a novel measure of efficiency that depends on the interactional behavior of speaker and listener agents.
-We adopt the reference game as our primary unit of interaction and model language users with the Rational Speech Act (RSA) framework -- a computional model of language use, which is nicely attested by experimental data on interaction.
+We adopt the reference game as our primary unit of interaction and model language users with the Rational Speech Act (RSA) framework -- a computational model of language use, which is supported by experimental data on interaction.
 Using these ingredients, we show that the property of interest (ambiguity) is prevalent in languages that optimize our measure of efficiency (Simulation 1).
 Further, we show how ambiguity is used efficiently during local, in-the-moment interactions (Simulation 2).
-Put differently, these simulations examine efficiency from two angles -- in the first we vary languages, fixing agents, and search for efficient designs.
+Put differently, these simulations examine efficiency from two angles -- in the first we vary languages, fixing agents, and search for efficient language designs.
 In the second we vary agents, fixing language, and examine efficient use.
@@
 Utterances may have different relative costs, operationalized via a prior over utterances $P(U)$.
-Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$\footnote{The prior over meanings are analogous to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}.
+Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$\footnote{The prior over meanings is equivalent to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}.
 We consider a set of contexts $C$ with an associated prior $P(C)$.
@@
 $|[[u_i]]| > 1$)\footnote{We use double brackets $[[\dots]]$ to represent denotation.}.
-In Figure 1 both the words "square" and "green" are ambigous so we would say that $\ell$ contains ambiguity.
+In Figure 1 both the words "square" and "green" are ambiguous so we would say that $\ell$ contains ambiguity.
 
@@
 RSA is a formalization of essential Gricean pragmatic principles -- agents reason about one another and their shared context [@Grice1975a].
-We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speaker-listeners in the reference game setting (see \href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI} and @GoodmanFrank2016a for a review).\par
+We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speaker-listeners in the reference game setting (see \href{https://bit.ly/2RBSGcU}{SI}).\par
 
@@
 
-We assume that speaker effort is related to the surprisal of an utterance in a particular context\footnote{In the current set of simulations we consider utterances costs as independent from context (ie.
+We assume that speaker effort is related to the surprisal of an utterance in a particular context\footnote{In the current set of simulations we consider utterances costs as independent from context (i.e..
 $p(u|c)p(c)=p(u)p(c)$).} -- intuitively, the number of bits needed to encode the utterance $u$.
@@
 $e \sim p(c)p(m|c)S(u|m, c; \ell)$).
-From these ingredients it is possible to derive the following objective between the speaker and listener distributions (see \href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 2.1} for complete derivation).
+From these ingredients it is possible to derive the following objective between the speaker and listener distributions (see \href{https://bit.ly/2RBSGcU}{SI 2.1} for complete derivation).
 \begin{equation}
@@
 Error bars represent 95 percent confidence intervals.
-Panel (B), example CE-optimal language (ambiguous) under in a four-condition context.
-Panel (C), example CE-optimal language (unambiguous) in a single-condition context."}
+Panel (B), example CE-optimal language (ambiguous) from a four-context simulation.
+Panel (C), example CE-optimal language (unambiguous) from a single-context simulation."}
 img <- png::readPNG("figs/fig1.png")
@@
 
-For comparison, we also examine properties of optimal languages under two additional objectives.
+For comparison, we examine properties of optimal languages under two additional objectives.
 Zipf [-@Zipf1949a] proposed that the speaker-optimal language $\ell_{speaker}^*$ would minimize speaker effort and the listener-optimal language $\ell_{listener}^*$ would minimize listener effort.
-We define these objectives using the first and second half of equation 1 (see \href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 2.2.}).\par
+We define these objectives using the first and second half of equation 1 (see \href{https://bit.ly/2RBSGcU}{SI 2.2.}).\par
 
@@
 But, if by a later position enough contextual information has accumulated, it may be efficient to use the ambiguous expression.
-We pursue this general framework in Simulation 2 -- examining when in a discourse using ambigity is efficient.
-Note that we will consider “context” as analogous to a “topic” of conversation.
+We pursue this general framework in Simulation 2 -- examining when in a discourse using ambiguity is efficient.
+We will consider “context” as analogous to a “topic” of conversation.
 \par
@@
 We conduct $N=600$ simulations, generating discourses of length $|D|=30$ utterances, comparing three speaker models ($n=200$ each).
-We consider a single language $\ell$\footnote{See SI for the matrix notation of this langauge.} with $|U|=6$ and $|M|=4$ in which two of the utterances are ambiguous and lower cost than the unambiguous utterances.
+We consider a single language $\ell$\footnote{See SI for the matrix notation of this language.} with $|U|=6$ and $|M|=4$ in which two of the utterances are ambiguous and lower cost than the unambiguous utterances.
 (Note that use of this particular language is not essential -- the results are broadly generalizable to languages that contain ambiguity, but exploring this space is computationally expensive.)\par
@@
 The first, a \textit{Partial pragmatics} baseline describes a speaker who reasons about a listener, but assumes they have no access to the discourse history.
-The second, a \textit{No pragmatics} baseline speaker does not consider a listener at all, but produces utterances according to the underlying langauge semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$) (see \href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 3}).\par
+The second, a \textit{No pragmatics} baseline speaker does not consider a listener at all, but produces utterances according to the underlying language semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$) (see \href{https://bit.ly/2RBSGcU}{SI 3}).\par
 
@@
 Panel (C) tracks our CE objective for each model over the discourse.
-Note that the objective decreases for all three models, primarily driven by the listener updating his belief about the actual topic ($P(C=c_{current}|D)$).
+Note that the objective decreases for all three models, primarily driven by the listeners updating their beliefs about the actual topic ($P(C=c_{current}|D)$).
 However, the objective declines more quickly under the \textit{Full} and \textit{Partial pragmatic} speakers as listener agents are better able to infer the correct context.
@@
 Panel (B) tracks speaker effort, which remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines.
-But, effort declines in the \textit{Full pragmatic} model as she increasingly relies on ambiguous material later in the discourse.\par
+But, effort declines in the \textit{Full pragmatic} model as speakers increasingly rely on ambiguous material later in the discourse.\par
 
@@
 
-How do the competing pressures of speakers and listeners give rise to the distributional regularities found in natural language?
+How do the competing pressures imposed by speakers and listeners give rise to the distributional regularities found in natural language?
 Zipf [-@Zipf1949a] proposed that the asymmetry between speaker and listener costs gives rise to a range of properties at the level of the lexicon.
@@
 Future work should assess the degree to which we can derive the same properties as previous studies using our current framework.
-In line with this idea, a primary focus of the current work is to provide a case-study on the productivity of assessing functionalist theories in repeated reference games between rational pragmatic agents.
-We hope this framework can serve as a domain general tool to assess the range of functionalist theories examining efficient language-structure and use.\par
+More generally, we hope that this framework can serve as a domain general tool to assess the range of functionalist theories examining efficient language-structure and use.\par
 
-\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI: \href{https://bit.ly/2DOnjI4}{\url{https://bit.ly/2DOnjI4}}.}}} \vspace{1em}
+\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering {SI available at: \href{https://bit.ly/2RBSGcU}{\url{https://bit.ly/2RBSGcU}}. Code will be available upon acceptance.}}} \vspace{1em}
 \noindent

### 2019-02-02 paper/paper.tex
MF last changes and spell-checking.

@@
 to language science -- analyzing language structure and use in terms of
-efficiency. Such an approach might reframe our opening question as
+efficiency. Such an approach might re-frame our opening question as
 follows: how does having property \textit{x} make using language
-\(\ell\) more or less useful for communication? Among others, this
-efficiency-based framing has produced a rich set of theoretical and
-empirical targets exploring semantic typology (Regier, Kemp, \& Kay,
-2015), properties such as ambiguity (Piantadosi, Tily, \& Gibson, 2011)
-and compositionality (Kirby, Griffiths, \& Smith, 2014), and the
-efficient use of reduction and redundancy in production (Genzel \&
-Charniak, 2002; Levy \& Jaeger, 2007).\par
+\(\ell\) more or less useful for communication? This efficiency-based
+framing has produced a rich set of theoretical and empirical targets
+exploring semantic typology (Regier, Kemp, \& Kay, 2015), properties
+such as ambiguity (Piantadosi, Tily, \& Gibson, 2011) and
+compositionality (Kirby, Griffiths, \& Smith, 2014), and the efficient
+use of reduction and redundancy in production (Genzel \& Charniak, 2002;
+Levy \& Jaeger, 2007).\par
 
-The approaches above typically posit efficiency measures motivated by
-information-theoretic principles, but not directly grounded in language
-use by interacting agents. In this work, we derive a measure of
-efficiency from first principles of \textit{efficient language use} and
-show how this can lead to communicatively efficient systems. We also
-demonstrate that assumptions about interlocutors impact how language
-properties are used efficiently. In this way, we bridge questions of
-design and use in a single framework.\par
+The approaches above typically posit efficiency measures that are
+motivated by information-theoretic principles, but they typically do not
+ground out in language use by interacting agents. In this work, we
+derive an novel objective function from first principles of
+\textit{rational language use} and show how optimizing this objective
+can lead to communicatively efficient systems. We also demonstrate that
+assumptions about interlocutors impact whether language properties are
+used efficiently. In this way, we integrate questions of language design
+and language use in a single framework.\par
 
 Functionalist theories commonly frame language efficiency in terms of a
-fundamental effort-assymetry underlying everyday communication: what is
+fundamental effort-asymmetry underlying everyday communication: what is
 ``hard'' for a speaker is likely different than what is ``hard'' for a
-listener. Zipf described this as follows -- purely from the standpoint
-of speaker effort, an optimal language \(\ell_{speaker}^*\) would tend
+listener. Zipf described this as follows: purely from the standpoint of
+speaker effort, an optimal language \(\ell_{speaker}^*\) would tend
 toward a vocabulary of a single, low-cost word. Given such a language,
@@
 proposed that the particular lexicon-level properties he observed were a
-result of these competing forces -- the pressure to jointly minimize
-speaker and listener effort.\par
+result of optimization based on these competing forces -- the pressure
+to jointly minimize speaker and listener effort.\par
 
@@
 optimization may take place at the level of conversation. Importantly,
-Zipf's conception of speaker and listener effort may be connected to how
-language is used; in particular, whether interlocutors engage in
+Zipf's conception of speaker and listener effort should be connected to
+how language is used; in particular, whether interlocutors engage in
 pragmatic reasoning during conversation. Under a Gricean treatment of
@@
 choose a property of languages that could, in principle, vary freely,
-but shows strong regularities across languages. The explanadum is why
+but shows strong regularities across languages. The explanandum is why
 this regularity holds. We examine ambiguity as our property, extending
@@
 interaction and model language users with the Rational Speech Act (RSA)
-framework -- a computional model of language use, which is nicely
-attested by experimental data on interaction. Using these ingredients,
-we show that the property of interest (ambiguity) is prevalent in
-languages that optimize our measure of efficiency (Simulation 1).
-Further, we show how ambiguity is used efficiently during local,
-in-the-moment interactions (Simulation 2). Put differently, these
-simulations examine efficiency from two angles -- in the first we vary
-languages, fixing agents, and search for efficient designs. In the
-second we vary agents, fixing language, and examine efficient use. \par
+framework -- a computational model of language use, which is supported
+by experimental data on interaction. Using these ingredients, we show
+that the property of interest (ambiguity) is prevalent in languages that
+optimize our measure of efficiency (Simulation 1). Further, we show how
+ambiguity is used efficiently during local, in-the-moment interactions
+(Simulation 2). Put differently, these simulations examine efficiency
+from two angles -- in the first we vary languages, fixing agents, and
+search for efficient language designs. In the second we vary agents,
+fixing language, and examine efficient use. \par
 
@@
 about, operationalized as a prior over meanings
-\(P(M)\)\footnote{The prior over meanings are analogous to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}.
+\(P(M)\)\footnote{The prior over meanings is equivalent to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}.
 We consider a set of contexts \(C\) with an associated prior \(P(C)\).
@@
 \(|[[u_i]]| > 1\))\footnote{We use double brackets $[[\dots]]$ to represent denotation.}.
-In Figure 1 both the words ``square'' and ``green'' are ambigous so we
+In Figure 1 both the words ``square'' and ``green'' are ambiguous so we
 would say that \(\ell\) contains ambiguity.
@@
 speaker-listeners in the reference game setting (see
-\href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI}
-and Goodman \& Frank (2016) for a review).\par
+\href{https://bit.ly/2RBSGcU}{SI}).\par
 
@@
 utterance in a particular
-context\footnote{In the current set of simulations we consider utterances costs as independent from context (ie. $p(u|c)p(c)=p(u)p(c)$).}
+context\footnote{In the current set of simulations we consider utterances costs as independent from context (i.e.. $p(u|c)p(c)=p(u)p(c)$).}
 -- intuitively, the number of bits needed to encode the utterance \(u\).
@@
 between the speaker and listener distributions (see
-\href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 2.1}
-for complete derivation).
+\href{https://bit.ly/2RBSGcU}{SI 2.1} for complete derivation).
 
@@
 Error bars represent 95 percent confidence intervals.
-Panel (B), example CE-optimal language (ambiguous) under in a four-condition context.
-Panel (C), example CE-optimal language (unambiguous) in a single-condition context.}\label{fig:plot-optimal-langs}
+Panel (B), example CE-optimal language (ambiguous) from a four-context simulation.
+Panel (C), example CE-optimal language (unambiguous) from a single-context simulation.}\label{fig:plot-optimal-langs}
 \end{figure*}
@@
 
-For comparison, we also examine properties of optimal languages under
-two additional objectives. Zipf (1949) proposed that the speaker-optimal
+For comparison, we examine properties of optimal languages under two
+additional objectives. Zipf (1949) proposed that the speaker-optimal
 language \(\ell_{speaker}^*\) would minimize speaker effort and the
@@
 effort. We define these objectives using the first and second half of
-equation 1 (see
-\href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 2.2.}).\par
+equation 1 (see \href{https://bit.ly/2RBSGcU}{SI 2.2.}).\par
 
@@
 framework in Simulation 2 -- examining when in a discourse using
-ambigity is efficient. Note that we will consider ``context'' as
-analogous to a ``topic'' of conversation. \par
+ambiguity is efficient. We will consider ``context'' as analogous to a
+``topic'' of conversation. \par
 
@@
 We consider a single language
-\(\ell\)\footnote{See SI for the matrix notation of this langauge.} with
+\(\ell\)\footnote{See SI for the matrix notation of this language.} with
 \(|U|=6\) and \(|M|=4\) in which two of the utterances are ambiguous and
@@
 consider a listener at all, but produces utterances according to the
-underlying langauge semantics (\(\ell\)) and topic probabilities
-(\(p(M|C=c_{current}\)) (see
-\href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 3}).\par
+underlying language semantics (\(\ell\)) and topic probabilities
+(\(p(M|C=c_{current}\)) (see \href{https://bit.ly/2RBSGcU}{SI 3}).\par
 
@@
 discourse. Note that the objective decreases for all three models,
-primarily driven by the listener updating his belief about the actual
-topic (\(P(C=c_{current}|D)\)). However, the objective declines more
-quickly under the \textit{Full} and \textit{Partial pragmatic} speakers
-as listener agents are better able to infer the correct context.
-Additionally, the difference in CE between the \textit{Full}- and
-\textit{Partial pragmatic} models at the end of the discourse is driven
-by the reduction in speaker costs. Panel (B) tracks speaker effort,
-which remains constant in both \textit{No pragmatic} and
+primarily driven by the listeners updating their beliefs about the
+actual topic (\(P(C=c_{current}|D)\)). However, the objective declines
+more quickly under the \textit{Full} and \textit{Partial pragmatic}
+speakers as listener agents are better able to infer the correct
+context. Additionally, the difference in CE between the \textit{Full}-
+and \textit{Partial pragmatic} models at the end of the discourse is
+driven by the reduction in speaker costs. Panel (B) tracks speaker
+effort, which remains constant in both \textit{No pragmatic} and
 \textit{Partial pragmatic} baselines. But, effort declines in the
-\textit{Full pragmatic} model as she increasingly relies on ambiguous
+\textit{Full pragmatic} model as speakers increasingly rely on ambiguous
 material later in the discourse.\par
@@
 
-How do the competing pressures of speakers and listeners give rise to
-the distributional regularities found in natural language? Zipf (1949)
-proposed that the asymmetry between speaker and listener costs gives
-rise to a range of properties at the level of the lexicon. We explored
-the interactions of rational pragmatic agents as a framework for
-understanding efficient language structure and use. We focused on an
+How do the competing pressures imposed by speakers and listeners give
+rise to the distributional regularities found in natural language? Zipf
+(1949) proposed that the asymmetry between speaker and listener costs
+gives rise to a range of properties at the level of the lexicon. We
+explored the interactions of rational pragmatic agents as a framework
+for understanding efficient language structure and use. We focused on an
 argument on the communicative function of ambiguity (Piantadosi et al.,
@@
 we can derive the same properties as previous studies using our current
-framework. In line with this idea, a primary focus of the current work
-is to provide a case-study on the productivity of assessing
-functionalist theories in repeated reference games between rational
-pragmatic agents. We hope this framework can serve as a domain general
-tool to assess the range of functionalist theories examining efficient
-language-structure and use.\par
+framework. More generally, we hope that this framework can serve as a
+domain general tool to assess the range of functionalist theories
+examining efficient language-structure and use.\par
 
@@
 
-\fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI: \href{https://bit.ly/2DOnjI4}{\url{https://bit.ly/2DOnjI4}}.}}}
+\fbox{\parbox[b][][c]{7.3cm}{\centering {SI available at: \href{https://bit.ly/2RBSGcU}{\url{https://bit.ly/2RBSGcU}}. Code will be available upon acceptance.}}}
 \vspace{1em} \noindent
@@
 support efficient communication. In B. M. \& W. O'Grady (Ed.), \emph{The
-ahndbook of language emergence (pp. 237-263)}. Hoboken, NJ:
+handbook of language emergence (pp. 237-263)}. Hoboken, NJ:
 Wiley-Blackwell.

### 2019-02-02 paper/paper.Rmd
BP final read through, now submitting.

@@
 RSA is a formalization of essential Gricean pragmatic principles -- agents reason about one another and their shared context [@Grice1975a].
-We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speaker-listeners in the reference game setting (see \href{https://bit.ly/2RBSGcU}{SI}).\par
+We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speakers and listeners in the reference game setting (see \href{https://bit.ly/2RBSGcU}{SI}).\par
 
@@
 Piantadosi et al.
-[-@Piantadosi2011a] argue just the opposite, claiming that ambiguity is an *efficient* property of any communication system in which *communication is contextualized*.
+[-@Piantadosi2011a] argue just the opposite, claiming that ambiguity is an efficient property of any communication system in which *communication is contextualized*.
 Simply put, it is useful to have a language that re-uses low-cost material (has ambiguity) so long as the cost of disambiguating the material is low.
@@
 
-As an example, say we have two objects ($m_1$ and $m_2$), two utterances ($u_1$ and $u_2$), differing in cost, and two languages ($\ell_1$, $\ell_2$), describing different utterance-meaning mappings.
+As an example, say we have two objects ($m_1$ and $m_2$), two utterances ($u_1$ and $u_2$), differing in cost, and two languages ($\ell_1$ and $\ell_2$), describing different utterance-meaning mappings.
 In language $\ell_1$, the low-cost $u_1$ can be used to refer to both $m_1$ and $m_2$ ($[[u_1]]_{\ell_1} = \{m_1, m_2 \}$), but the high-cost $u_2$ cannot be used at all ($[[u_2]]_{\ell_1} = \emptyset$).
@@
 For each simulation we enumerate the set of *valid* languages in which $|U|=|M|=4$ ($U$ is our set of utterances and $M$ our set of meanings).
-Recall that languages are boolean matrices and a language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment) and each form maps to at least one meaning (every row has some non-zero assignment).
+Recall that languages are boolean matrices and a language $\ell \in L$ is "valid" so long as each possible meaning $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment) and each form maps to at least one meaning (every row has some non-zero assignment).
 For a given simulation, the goal is to find the language $\ell^*$ which minimizes our objective and then check to see if that language contains ambiguity.\par
@@
 We should expect this to be reflected in our \textit{Full pragmatics} model who reasons about the listener and discourse history.
-By contrast, a speaker-optimal model who does not consider the listener should greedily use ambiguous utterances (\textit{No Pragmatics} model), while a listener-optimal model should avoid ambiguity entirely (\textit{Partial pragmatics} model).\par
+By contrast, a speaker-optimal model who does not consider the listener should greedily use ambiguous utterances (\textit{No pragmatics} model), while a listener-optimal model should avoid ambiguity entirely (\textit{Partial pragmatics} model).\par
 
@@
 Figure 3, panel (A) shows the empirical probability a speaker uses an ambiguous utterance as a function of discourse position.
-The \textit{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse.
-The \textit{Partial pragmatic} baseline avoids ambiguous utterances entirely.
-But, the \textit{Full pragmatic} model avoids ambiguous material only at the start of the discourse, employing it increasingly as the discourse proceeds.
+The \textit{No pragmatics} baseline uses ambiguous utterances frequently and at a constant rate over the discourse.
+The \textit{Partial pragmatics} baseline avoids ambiguous utterances entirely.
+But, the \textit{Full pragmatics} model avoids ambiguous material only at the start of the discourse, employing it increasingly as the discourse proceeds.
 Panel (C) tracks our CE objective for each model over the discourse.
 Note that the objective decreases for all three models, primarily driven by the listeners updating their beliefs about the actual topic ($P(C=c_{current}|D)$).
-However, the objective declines more quickly under the \textit{Full} and \textit{Partial pragmatic} speakers as listener agents are better able to infer the correct context.
-Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatic} models at the end of the discourse is driven by the reduction in speaker costs.
-Panel (B) tracks speaker effort, which remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines.
-But, effort declines in the \textit{Full pragmatic} model as speakers increasingly rely on ambiguous material later in the discourse.\par
+However, the objective declines more quickly under the \textit{Full} and \textit{Partial pragmatics} speakers as listener agents are better able to infer the correct context.
+Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatics} models at the end of the discourse is driven by the reduction in speaker costs.
+Panel (B) tracks speaker effort, which remains constant in both \textit{No pragmatics} and \textit{Partial pragmatics} baselines.
+But, effort declines in the \textit{Full pragmatics} model as speakers increasingly rely on ambiguous material later in the discourse.\par
 

### 2019-02-02 paper/paper.tex
BP final read through, now submitting.

@@
 ground out in language use by interacting agents. In this work, we
-derive an novel objective function from first principles of
+derive a novel objective function from first principles of
 \textit{rational language use} and show how optimizing this objective
@@
 representational framework to model Gricean (rational and pragmatic)
-speaker-listeners in the reference game setting (see
+speakers and listeners in the reference game setting (see
 \href{https://bit.ly/2RBSGcU}{SI}).\par
@@
 Piantadosi et al. (2011) argue just the opposite, claiming that
-ambiguity is an \emph{efficient} property of any communication system in
-which \emph{communication is contextualized}. Simply put, it is useful
-to have a language that re-uses low-cost material (has ambiguity) so
-long as the cost of disambiguating the material is low. In particular,
-context (or common ground) can provide useful information for
-disambiguation.\par
+ambiguity is an efficient property of any communication system in which
+\emph{communication is contextualized}. Simply put, it is useful to have
+a language that re-uses low-cost material (has ambiguity) so long as the
+cost of disambiguating the material is low. In particular, context (or
+common ground) can provide useful information for disambiguation.\par
 
@@
 utterances (\(u_1\) and \(u_2\)), differing in cost, and two languages
-(\(\ell_1\), \(\ell_2\)), describing different utterance-meaning
+(\(\ell_1\) and \(\ell_2\)), describing different utterance-meaning
 mappings. In language \(\ell_1\), the low-cost \(u_1\) can be used to
@@
 boolean matrices and a language \(\ell \in L\) is ``valid'' so long as
-each possible meaning in \(m \in M\) can be referred to by at least one
+each possible meaning \(m \in M\) can be referred to by at least one
 form \(u \in U\) (every column of \(\ell\) has some non-zero assignment)
@@
 consider the listener should greedily use ambiguous utterances
-(\textit{No Pragmatics} model), while a listener-optimal model should
+(\textit{No pragmatics} model), while a listener-optimal model should
 avoid ambiguity entirely (\textit{Partial pragmatics} model).\par
@@
 ambiguous utterance as a function of discourse position. The
-\textit{No pragmatic} baseline uses ambiguous utterances frequently and
-at a constant rate over the discourse. The \textit{Partial pragmatic}
+\textit{No pragmatics} baseline uses ambiguous utterances frequently and
+at a constant rate over the discourse. The \textit{Partial pragmatics}
 baseline avoids ambiguous utterances entirely. But, the
-\textit{Full pragmatic} model avoids ambiguous material only at the
+\textit{Full pragmatics} model avoids ambiguous material only at the
 start of the discourse, employing it increasingly as the discourse
@@
 actual topic (\(P(C=c_{current}|D)\)). However, the objective declines
-more quickly under the \textit{Full} and \textit{Partial pragmatic}
+more quickly under the \textit{Full} and \textit{Partial pragmatics}
 speakers as listener agents are better able to infer the correct
 context. Additionally, the difference in CE between the \textit{Full}-
-and \textit{Partial pragmatic} models at the end of the discourse is
+and \textit{Partial pragmatics} models at the end of the discourse is
 driven by the reduction in speaker costs. Panel (B) tracks speaker
-effort, which remains constant in both \textit{No pragmatic} and
-\textit{Partial pragmatic} baselines. But, effort declines in the
-\textit{Full pragmatic} model as speakers increasingly rely on ambiguous
-material later in the discourse.\par
+effort, which remains constant in both \textit{No pragmatics} and
+\textit{Partial pragmatics} baselines. But, effort declines in the
+\textit{Full pragmatics} model as speakers increasingly rely on
+ambiguous material later in the discourse.\par
 

### 2019-05-11 paper/paper.Rmd
de-anonymized for psycharxiv
reviewer 3 edits

@@
 output: cogsci2016::cogsci_paper
+final-submission: \cogscifinalcopy
 ---
@@
 
-Importantly, we assume that events $e = <u, m, c>$ are sampled according the to following generative model -- some context occurs in the world with probability $P(C=c)$.
+Importantly, we assume that events $e = <u, m, c>$ are sampled according to the following generative model -- some context occurs in the world with probability $P(C=c)$.
 Within this context, an object $m$ occurs with probability $p(m|c)$.
@@
 By contrast, in language $\ell_2$, $u_1$ can only refer to $m_1$ and $u_2$ can only refer to $m_2$ ($[[u_1]]_{\ell_2} = \{m_1\}$ and $[[u_2]]_{\ell_2} = \{m_2 \}$).
-While it is cheaper for a speaker to use $\ell_1$ (because speaking is always low-cost), it is more difficult for a listener (because $u_1$ is ambiguous).
+While it is cheaper for a speaker to use $\ell_1$ (because speaking it is always lower cost), it is more difficult for a listener (because $u_1$ is ambiguous).
 Crucially, however, if context is disambiguating then the speaker can use $u_1$ to refer to either $m_1$ or $m_2$ and $\ell_1$ should be preferred to $\ell_2$.\par

### 2019-05-11 paper/paper.Rmd

@@
 Utterances may have different relative costs, operationalized via a prior over utterances $P(U)$.
-Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$\footnote{The prior over meanings is equivalent to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}.
+Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$\footnote{The prior over meanings is equivalent to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015)).}.
 We consider a set of contexts $C$ with an associated prior $P(C)$.
@@
 $L(m|u, c;\ell)$).
-Note that both speakers and listeners can induce joint distributions over utterance-meaning pairs, although, importantly, these distributions may differ:
+Speakers and listeners can induce joint distributions over utterance-meaning pairs, although, these distributions may differ:
 $$P_{speaker}(u, m | c; \ell) = S(u|m, c; \ell)p(m|c)$$
@@
 
-We assume that speaker effort is related to the surprisal of an utterance in a particular context\footnote{In the current set of simulations we consider utterances costs as independent from context (i.e..
+We assume that speaker effort is related to the surprisal of an utterance in a particular context\footnote{In the current set of simulations we consider utterance costs as independent from context (i.e..
 $p(u|c)p(c)=p(u)p(c)$).} -- intuitively, the number of bits needed to encode the utterance $u$.
@@
 While it is cheaper for a speaker to use $\ell_1$ (because speaking it is always lower cost), it is more difficult for a listener (because $u_1$ is ambiguous).
-Crucially, however, if context is disambiguating then the speaker can use $u_1$ to refer to either $m_1$ or $m_2$ and $\ell_1$ should be preferred to $\ell_2$.\par
+Crucially, if context is disambiguating then the speaker can use $u_1$ to refer to either $m_1$ or $m_2$ and $\ell_1$ should be preferred to $\ell_2$.\par
 
@@
 
-\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering {SI available at: \href{https://bit.ly/2RBSGcU}{\url{https://bit.ly/2RBSGcU}}. Code will be available upon acceptance.}}} \vspace{1em}
+\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering {SI and simulations: \href{https://bit.ly/2RBSGcU}{\url{https://bit.ly/2RBSGcU}}, \href{https://github.com/benpeloquin7/zipf_principles}{\url{https://github.com/benpeloquin7/zipf_principles}}}}} \vspace{1em}
 \noindent

### 2019-05-11 paper/paper.tex
de-anonymized for psycharxiv

@@
 % KM added 1/4/18 to allow control of blind submission
+\cogscifinalcopy
 
@@
 about, operationalized as a prior over meanings
-\(P(M)\)\footnote{The prior over meanings is equivalent to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}.
+\(P(M)\)\footnote{The prior over meanings is equivalent to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015)).}.
 We consider a set of contexts \(C\) with an associated prior \(P(C)\).
@@
 \(M\) using \(\ell\) in a given context \(c\) (i.e. \(L(m|u, c;\ell)\)).
-Note that both speakers and listeners can induce joint distributions
-over utterance-meaning pairs, although, importantly, these distributions
-may differ: \[P_{speaker}(u, m | c; \ell) = S(u|m, c; \ell)p(m|c)\]
+Speakers and listeners can induce joint distributions over
+utterance-meaning pairs, although, these distributions may differ:
+\[P_{speaker}(u, m | c; \ell) = S(u|m, c; \ell)p(m|c)\]
 \[P_{listener}(u, m| c; \ell) = L(m|u, c; \ell)p(u|c)\]
@@
 utterance in a particular
-context\footnote{In the current set of simulations we consider utterances costs as independent from context (i.e.. $p(u|c)p(c)=p(u)p(c)$).}
+context\footnote{In the current set of simulations we consider utterance costs as independent from context (i.e.. $p(u|c)p(c)=p(u)p(c)$).}
 -- intuitively, the number of bits needed to encode the utterance \(u\).
@@
 Importantly, we assume that events \(e = <u, m, c>\) are sampled
-according the to following generative model -- some context occurs in
+according to the following generative model -- some context occurs in
 the world with probability \(P(C=c)\). Within this context, an object
@@
 \([[u_2]]_{\ell_2} = \{m_2 \}\)). While it is cheaper for a speaker to
-use \(\ell_1\) (because speaking is always low-cost), it is more
-difficult for a listener (because \(u_1\) is ambiguous). Crucially,
-however, if context is disambiguating then the speaker can use \(u_1\)
-to refer to either \(m_1\) or \(m_2\) and \(\ell_1\) should be preferred
-to \(\ell_2\).\par
+use \(\ell_1\) (because speaking it is always lower cost), it is more
+difficult for a listener (because \(u_1\) is ambiguous). Crucially, if
+context is disambiguating then the speaker can use \(u_1\) to refer to
+either \(m_1\) or \(m_2\) and \(\ell_1\) should be preferred to
+\(\ell_2\).\par
 
@@
 
-\fbox{\parbox[b][][c]{7.3cm}{\centering {SI available at: \href{https://bit.ly/2RBSGcU}{\url{https://bit.ly/2RBSGcU}}. Code will be available upon acceptance.}}}
+\fbox{\parbox[b][][c]{7.3cm}{\centering {SI and simulations: \href{https://bit.ly/2RBSGcU}{\url{https://bit.ly/2RBSGcU}}, \href{https://github.com/benpeloquin7/zipf_principles}{\url{https://github.com/benpeloquin7/zipf_principles}}}}}
 \vspace{1em} \noindent

### 2019-11-26 paper/paper.Rmd
last push before port

@@
 # Zipfian objective for linguistic system efficiency
+
+Zipf (1949) proposed that the particular distributional properties found in natural language emerge from competing speaker and listener pressures.
+We operationalize this objective in equation (1) -- the efficiency of a linguistic system $\ell$ being used by speaker and listener agents $S$ and $L$ is the sum of the expected speaker and listener effort to communicate over all possible communicative events $e \in E$.\par
+
+\begin{equation}
+\begin{split}
+  \text{Efficiency}(S, L, \ell) = \mathbb{E}_{e \sim P(E)}[\text{speaker effort}] \\+ \mathbb{E}_{e \sim P(E)}[\text{listener effort}]
+\end{split}
+\end{equation}
+
+We assume that speaker effort is related to the surprisal of an utterance in a particular context -- intuitively, the number of bits needed to encode the utterance $u$.\footnote{In the current set of simulations we consider utterances costs as independent from context (i.e.
+$p(u|c)=p(u)$).
+Hence surprisal can be thought of more simply as cost.} This particular formalization of speaker-cost is general enough to accommodate a range of cost instantiations, such as production difficulty via articulation effort, cognitive effort related to lexical access, or others (Bennett \& Goodman, 2015).\par
+
+$$\text{speaker effort} = -log_2(p(u|c))$$
+
+We assume listener effort is the surprisal of a meaning given an utterance.
+This operationalization of listener effort is intuitively related to existing work in sentence processing in which word comprehension difficulty is proportional to surprisal (Hale, 2001; Levy, 2008).
+
+$$\text{listener effort} = -log_2(L(m|u, c; \ell))$$
+
+Rewriting (1) we have
+\begin{equation}
+\begin{split}
+\text{Efficiency}(S, L, C, \ell) = \mathbb{E}_{e \sim P(E)}[-log_2(p(u|c))] + 
+\mathbb{E}_{e \sim P(E)}[-log_2(L(m|u, c; \ell))]
+\end{split}
+\end{equation}
+
+We assume that the particular joint distribution over utterance-meaning-context triples $e = <u, m, c>$  follows from a simple generative model: First, some context is sampled with probability $p(c)$.
+Then some meaning is sampled with probability $p(m|c)$.
+Our speaker attempts to convey this intended meaning to a listener via an utterance $u$ by sampling from the speaker conditional distribution $S(u|m, c; \ell)$.
+This allows us to re-write the objective as:
+
+\begin{equation}
+\begin{split}
+  = -\sum_{u, m, c}P_{speaker}(u, m | c; \ell)p(c)[log_2(p(u|c))] - \\ 
+   \sum_{u, m, c}P_{speaker}(u, m | c; \ell)p(c)[log_2(L(m|u, c; \ell))]
+\end{split}
+\end{equation}
+
+\begin{equation}
+\begin{split}
+  = -\sum_{c\in C}p(c)(\sum_{u, m}P_{speaker}(u, m | c; \ell)[log_2(p(u|c))] + 
+  \sum_{u, m}P_{speaker}(u, m | c; \ell)[log_2(L(m|u, c; \ell))])
+\end{split}
+\end{equation}
+
+\begin{equation}
+\begin{split}
+ = -\sum_{c\in C}p(c)\sum_{u, m}P_{speaker}(u, m | c; \ell)[log_2(L(m|u, c; \ell)p(u|c)]
+ \end{split}
+\end{equation}
+
+Note that $P_{listener}(u, m | c; \ell) = L(m|u, c; \ell)p(u|c)$:
+
+\begin{equation}
+\begin{split}
+ = -\sum_{c\in C}p(c)\sum_{u, m}P_{speaker}(u, m | c; \ell)[log_2(P_{listener}(u, m | c; \ell))]
+ \end{split}
+\end{equation}
+
+The inner summation of (6) is the cross-entropy between speaker and listener conditional distributions over utterance-meaning pairs.
+
+\begin{equation}
+\begin{split}
+ = -\sum_{c\in C}p(c) H_{cross}(P_{speaker}(u, m |c; \ell), P_{listener}(u, m |c; \ell))
+ \end{split}
+\end{equation}
+
+This final form is simply an expectation of speaker-listener cross-entropy over contextualized language use.
+
+\begin{equation}
+\begin{split}
+ = \mathbb{E}_{c \sim P(C)}[H_{cross}(P_{speaker}, P_{listener}|c; \ell)]
+ \end{split}
+\end{equation}
+
+Note that in the case that $|C| =1$, our objective simplifies to a simple Cross-Entropy between speaker-listener joint distributions over utterance-meaning pairs.\par
+
+\begin{equation}
+\begin{split}
+ = H_{cross}(P_{speaker}, P_{listener}; \ell)
+ \end{split}
+\end{equation}
+
+From an information-theoretic perspective this objective is intuitive: $H_{cross}$ denotes the Cross-Entropy (CE), a measure of dissimilarity between two distributions -- the average number of bits required to communicate under one distribution, given that the “true” distribution differs.
+In our case, we have an expectation over this term -- the expected difference between the distributions assumed by the speaker $P_{speaker}$ and listener $P_{listener}$ given a set of contexts $C$.
+In other words, an "efficient" language $\ell$ minimizes the distance between what speakers and listeners think.\par
 

### 2019-11-26 ambiguity/config.py

@@ -1,11 +1,11 @@
 """config.py"""
 
-from simulations import VaryContextSimulations, RecursiveSimulation
+from .simulations import VaryContextSimulations, RecursiveSimulation
 
 CONTEXT_RUN_PARAMS = {
     'simulator': VaryContextSimulations,
     'run_params': {
-        'n_sims': 2,
+        'n_sims': 30,
         'utterance_order_fn': lambda x: 1.,
         'utterance_order_fn_name': 'uniform',
         'meaning_order_fn': lambda x: 1.,
@@ -15,7 +15,7 @@ CONTEXT_RUN_PARAMS = {
         'speaker_k': 1,
         'listener_k': 1,
         'verbose': False,
-        'matrix_density': 4
+        'matrix_density': 2
     },
     # fp pattern below
     # ----------------

### 2019-11-26 ambiguity/run.py

@@ -12,7 +12,7 @@ Example
 import logging
 import os
 
-from config import SIMULATIONS
+from .config import SIMULATIONS
 
 logging.getLogger().setLevel(logging.INFO)
 
